### 2017-09-17 article.Rnw
Add abstract

\documentclass[a4paper]{article}

\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{bm}
\DeclareRobustCommand{\rchi}{{\mathpalette\irchi\relax}}
\newcommand{\irchi}[2]{\raisebox{\depth}{$#1\chi$}}

\usepackage{hyperref}
\hypersetup{colorlinks=true, linkcolor=blue, citecolor=blue,
  urlcolor=blue}

\usepackage[bottom]{footmisc}
\usepackage{geometry}
\geometry{lmargin=3cm, rmargin=3cm, tmargin=3cm, bmargin=3cm}

% Algorithm
\usepackage[english, vlined, ruled]{algorithm2e}
\SetAlCapSty{}
% \SetAlCapFnt{\footnotesize}
% \SetAlCapNameFnt{\footnotesize}

\usepackage{natbib}
% \bibliographystyle{humannat}
\bibliographystyle{agsm}

\usepackage{lipsum}
\usepackage{float}
\usepackage{mathpazo}

\usepackage{booktabs}
\usepackage{tabularx}
\newcolumntype{C}{>{\centering\arraybackslash}X}

\makeatletter
\def\and{%
  \end{tabular}%
  \hskip 0em \@plus.17fil \protect$\bullet$
  \begin{tabular}[t]{c}}%
\makeatother

%-----------------------------------------------------------------------
\title{Reparametrization of COM-Poisson Regression Models with
  Applications in the Analysis of Experimental Data}

\author{
  Eduardo E. R. Junior\thanks{Corresponding author:
    \href{mailto:jreduardo@usp.br}{\nolinkurl{jreduardo@usp.br}}}
    \hspace*{.8mm}\thanks{Department of Exact Sciences, ESALQ-USP,
    Piracicaba, São Paulo, Brazil} \and
  Walmes M. Zeviani\thanks{Department of Statistics, UFPR, Curitiba,
    Paraná, Brazil} \and
  Wagner H. Bonat\footnotemark[3] \and
  Clarice G. B. Dem\'{e}trio\footnotemark[2]
}

\date{\today}

<<setup, include=FALSE>>=

library(knitr)
library(xtable)
options(digits = 3, OutDec = ".",
        xtable.caption.placement = "top",
        xtable.booktabs = TRUE,
        xtable.sanitize.text.function = identity)
opts_chunk$set(
    warning = FALSE,
    message = FALSE,
    echo = FALSE,
    results = "hide",
    fig.width = 7,
    fig.height = 5,
    out.width = "1\\textwidth",
    fig.align = "center",
    fig.pos = "H",
    # dev = "tikz"
    dev.args = list(family = "Palatino")
    )

source("./codes/lattice-panels.R")
source("./codes/functions.R")

# Colors for legends
cols <- trellis.par.get("superpose.line")$col

# Useful packages
library(bbmle)
library(multcomp)
library(plyr)
library(tidyr)
library(dplyr)

@

%-----------------------------------------------------------------------

\begin{document}
\maketitle

\begin{abstract}

  In the analysis of count data seldom the equidispersion assumption is
  suitable, hence the Poisson regression model is inappropriate in many
  situations. The COM-Poisson is a member of the exponential family, has
  the Poisson and geometric distributions as special cases and the
  Bernoulli distribution as a limiting case. It can deal with both under
  and overdispersed count data. The main disadvantage of COM-Poisson
  distribution is that the location parameter does not represent the
  average making difficult interpretation in a regression model. In this
  paper we present a very simple reparametrization based on mean
  approximation. The estimation and inference is based on likelihood and
  simulation study showed that the maximum likelihood estimators are
  unbiased and consistent for both regression and precision
  parameters. Furthermore we observe empirical correlation between
  regression and precision parameters close to 0 suggest the
  orthogonality for the proposed reparametrization. We presents
  applications of the proposed model by three case studies with over,
  under and equidispersed counts. Results show improvements over the
  conventional approaches. Therefore, its use in the analysis of count
  data is encouraged. The computational routines for fitting of original
  and reparametrized COM-Poisson models are available in the
  supplementary material.

\vspace{0.1cm}
\noindent\textbf{keywords:} COM-Poisson, Over/Under-dispersion, Count
data, Likelihood inference.
\end{abstract}

\pagebreak

\section{Introduction}
\label{introduction}

Count data are variables that assume non-negative integer values and
representing the number of times an event occurs in the observation
unit. This kind of data are common in experimental research such as
number of grains produced by a plant, number of fruits produced by a
tree, number of insects on a particular cell, among others. Since the
seminal paper by \citet{Nelder1972} introduced the class of generalized
linear models (GLM), analysis of count data often consider the Poisson
regression model. These models provide a suitable distribution for a
counting random variable and an efficient Newton scoring algorithm are
used to fit them. Furthermore, the theoretical background for GLMs is
well established in the class of dispersion models \citep{Jorgensen1997}
as a generalization of the exponential family of distributions.

In spite of the advantages of the GLM framework, the Poisson
distribution has only one parameter, which represents the mean, and
also, the variance. This relationship is referred as equidispersion.
However, this assumption can be unsuitable to many situations and when
the Poisson model is applied in these cases the standard errors for the
parameters estimates are inconsistent \citep{Winkelmann1995}.

In practice, overdispersion, when the variance greater than the mean, is
largely reported in the literature and may occur due to the absence of
relevant covariates, heterogeneity of sampling units, different domain
not considered and excess of zero \citep{Hinde1998}. Lesser reported,
but that has received attention from the statistical community, is a
underdispersion case, when the variance is less than the mean. The
process that reduce the variability are not as well known that lead to
extra variability. For the same reason, there a few approaches to deal
underdispersed cases. The explanatory mechanisms to underdispersion may
be related to Poisson process. When the time between events is not
exponential distributed, the number of events can be over or
underdispersed, this motivated the class of duration dependence models
\citep{Winkelmann1995}. Other possibity to explain underdispersion are
order statistics. Suppose $Y_i$ is Poisson distributed then maximum of
$Y_i$ is underdispersed.

The strategies for constructing alternative count distributions are
related with the causes of the non-equidispersion. Particular to
overdispersion case Poisson mixed models are widely applied. One popular
example of this approach is the negative-binomial model, where the
natural is assumed to follow a gamma distribution. However, other
distributions for the random effects may be considered, for example the
model Poisson-Tweedie \citep{Bonat2017} and
Poisson-Inverse-Gaussian. For underdispersion, the Gamma-Count
distribution assumes a gamma distribution for the time between events in
Poisson process, thus it can deal with dispersed counts
\citep{Zeviani2014}. The COM-Poisson distribution is given by a
generalization the Poisson distribution allowed for a non-linear
decrease in ratios of successive probabilities \citep{Shmueli2005}.

The COM-Poisson is a member of the exponential family, has the Poisson
and geometric distributions as special cases and the Bernoulli
distribution as a limiting case. It can deal with both under and
overdispersed count data. Some recently applications of the COM-Poisson
distribution are related by \citet{Lord2010} for analysis of traffic
crash data, \citet{Sellers2010} to model airfreight breakage and book
purchases and \citet{Huang2017} with analysis of attendance data,
takeover birds and cotton bolls count. The main disadvantage of
COM-Poisson distribution is that the location parameter does not
represent the average making difficult interpretation in a regression
structure. \citep{Huang2017} proposed a mean-parametrization COM-Poisson
distribution to avoid this issue, in this approach the mean parameter is
obtained by solving the root of an equation defined as a infinite sum.

The main goal of this article is to proposed a novel COM-Poisson
parametrization based on mean approximation given by
\citet{Shmueli2005}. In this parametrization the probability mass
function is wrote with $\mu$ and $\phi$ as function of original
parameters, where $\mu$ is the mean and $\phi$ is the precision or
dispersion parameter. Opposite the original parametrization, the
proposed parametrization leads to interpretability of the parameters in
a regression model and it is more simple than mean-parametrization gives
by \cite{Huang2017} because in this case it is not necessary find $\mu$
as the root of an equation involving an infinite sum. In simulation
study we show the properties about maximum likelihood estimators in
proposed and

This paper is organized as follows. In \autoref{background} we present
the COM-Poisson distribution and \cite{Huang2017} parametrization. The
proposed reparametrization and assessment of moments approximations is
considered in the \autoref{reparametrization}. In the
\autoref{estimation-and-inference} we show the estimation and inference
methods based on likelihood function whose properties are assess in
\autoref{simulation-study} by simulations. Applications of
reparametrized COM-Poisson regression model are illustrated in
\autoref{case-studies} with three case studies. The analysis are
performed using the software R \citep{Rcore2017}. The source code and
data sets are avaliable in supplementary material\footnote{Available on
  \texttt{\url{http://www.leg.ufpr.br/~eduardojr/papercompanions}}
  \label{papercompanion}.}.

\section{Background}
\label{background}

The COM-Poisson distribution generalizes the Poisson distribution in
terms of the ratio between the probabilities of two consecutive values,
with addition of a precision parameter \cite{Sellers2010}. Let $Y$ a
COM-Poisson random variable, then
$$\frac{\Pr(Y=y-1)}{\Pr(Y=y)} = \frac{y^\nu}{\lambda}$$ while for the
Poisson distribution this ratio is $\frac{y}{\lambda}$. This allows the
distribution for overdispersion or underdispersion. The probability mass
function for the COM-Poisson random variables takes the form
\begin{equation}
  \label{eqn:pmf-cmp}
  \Pr(Y=y \mid \lambda, \nu) = \frac{\lambda^y}{(y!)^\nu Z(\lambda, \nu)},
  \qquad y = 0, 1, 2, \ldots
\end{equation},
where $\lambda > 0$, $\nu \geq 0$ and $Z(\lambda, \nu) =
\sum_{j=0}^\infty \frac{\lambda^j}{(j!)^\nu}$ is a normalizing
constant.

The $Z(\lambda, \nu)$ series diverges theoretically only when $\nu=0$
and $\lambda \geq 1$, but numerically for small values of $\nu$ combined
with large of $\lambda$, the sum is so huge it causes
\textit{overflow}. \autoref{tab:convergenceZ} shows the calculated
normalizing constants using a thousand increments, that is,
$\sum_{j=0}^{1000}\lambda^j/(j!)^\nu$ for different of $\lambda$ and
$\phi$ \footnote{For $\lambda=1$ e $\nu=0$ we show Inf, although the
  numerically value is $1000$, since the series clearly diverges.}.

<<convergenceZ, echo=FALSE, results="asis">>=

#-----------------------------------------------------------------------
# Convergence of Z(lambda, nu) constant
computeZ <- function(lambda, nu, maxit = 1e4, tol = 1e-5) {
    z <- vector("numeric", maxit)
    j = 1
    z[j] <- exp(j * log(lambda) - nu * lfactorial(j))
    while (abs(z[j] - 0) > tol && j <= maxit) {
        j = j + 1
        z[j] <- exp(j * log(lambda) - nu * lfactorial(j))
    }
    if (lambda == 1 & nu == 0) z <- Inf
    return(sum(z, na.rm = TRUE) + 1)
}

grid <- expand.grid(
    lambda = c(0.5, 1, 5, 10, 30, 50),
    nu = seq(0, 1, length.out = 11))
grid$z <- apply(grid, 1, function(par) {
    computeZ(lambda = par[1], nu = par[2], maxit = 1e4)
})

xt <- xtabs(z ~ nu + lambda, data = grid)
caption <- paste(
    "Values for $Z(\\lambda, \\nu)$ constant (calculated numerically)",
    "for values of $\\lambda$ (0.5 to 50) and $\\phi$ (0 to 1)")

# Remove excessive scientific notation
xt2 <- gsub("e\\+00", "    ", format(xt))
xt2 <- gsub("\\|([0-9].{2}) \\|", "\\|**\\1** \\|", xt2)
print(xtable(xt2, digits = -2,
             caption = caption,
             align = "C|CCCCCC",
             # align = "ccccccc",
             label = "tab:convergenceZ"),
      include.colnames = FALSE,
      tabular.environment = "tabularx",
      width = "\\textwidth",
      add.to.row = list(
          pos = list(0, 0),
          command = c(
              " \\multicolumn{7}{c}{$\\bm{\\lambda}$} \\\\\n",
              "$\\bm{\\nu}$ & 0,5 & 1 & 5 & 10 & 30 & 50 \\\\\n")
      ))

@

An undesirable feature of COM-Poisson distribution is that the moments
cannot be solved in closed form. \citet{Shmueli2005} and
\citet{Sellers2010} using an asymptotic approximation for
$Z(\lambda,\nu)$, shows that an approximate forms can be closed by
\begin{equation}
  \label{eqn:mean-approx}
  E(Y) \approx \lambda^{1/\nu} - \frac{\nu - 1}{2\nu} \qquad
  \textrm{and} \qquad
  \textrm{var}(Y) \approx \frac{\lambda^{1/\nu}}{\nu}\,,
\end{equation} which is particularly accurate for  $\nu \leq 1$ or
$\lambda > 10$. The authors also shows that the mean-variance
relationship can be approximate by $\frac{1}{\nu}E(Y)$. In
\autoref{reparametrization} we assess the approximations.

A regression formulation for COM-Poisson models was proposed by
\citet{Sellers2010}, using the original parametrization. In this case,
the COM-Poisson regression is $\eta(E(Y_i)) = \log(\lambda_i)$ and the
relationship between $E(Y_i)$ and $\bm{x}_i$ is modelled
indirectly. \citet{Huang2017} shows how the means of COM-Poisson
distributions can be modelled directly. in the \autoref{eqn:pmf-cmp},
\Citeauthor{Huang2017} proposes that the parameter $\lambda$ is a
function of $\mu$ and $\nu$, given by the solution to
\begin{equation*}
  \sum_{j=0}^{\infty} (j - \mu) \frac{\lambda^j}{(y!)^\nu} = 0\,.
\end{equation*}
Thus the mean-parametrized COM-Poisson regression is $\eta(E(Y_i)) =
\log(\mu_i)$. In this article, we presented a alternative
mean-parametrization of COM-Poisson models to avoid the limitations of
original parametrization and complexity of \Citeauthor{Huang2017}
parametrization.

\section{Reparametrization}
\label{reparametrization}

The proposed reparametrization of COM-Poisson models is based on
mean approximation (\autoref{eqn:mean-approx}). We introduced new
parameter $\mu$, using the mean approximation
\begin{equation}
  \label{eqn:repar-cmp}
  \mu = h(\lambda, \nu) = \lambda^{1/\nu} - \frac{\nu - 1}{2\nu}
  \quad \Rightarrow \quad
  \lambda = h^{-1}(\mu, \nu) = \left (\mu +
    \frac{(\nu - 1)}{2\nu} \right )^\nu.
\end{equation}
Precision parameter is taken on the log scale to avoid restrictions on the
parameter space, $\phi = \log(\nu), \phi \in \mathbb{R}$. The
interpretation of $\phi$ are the same to $\nu$, but in other scale. For
$\phi<0$ we have overdispersed case and $\phi>0$ overdispersed, relatives
to Poisson distribution. $\phi=0$ represents the special case that the
COM-Poisson comes down to Poisson.

<<data-approx, include=FALSE, cache=TRUE>>=

#-----------------------------------------------------------------------
# Study the approximation

#-------------------------------------------
# Mean and variance relationship
aux <- expand.grid(
    mu = seq(2, 30, length.out = 50),
    phi = seq(log(0.3), log(2.5), length.out = 50))

moments <- mapply(FUN = calc_moments,
                  mu = aux$mu,
                  phi = aux$phi,
                  MoreArgs = list(sumto = 300),
                  SIMPLIFY = FALSE)
grid <- cbind(aux, t(do.call(cbind, moments)))
grid <- transform(grid, va = mu / exp(phi))

@

In order to assess the accuracy of the moments approximations (
\autoref{eqn:mean-approx}) we presented the quadratic errors of moments
approximations in \autoref{fig:approx-plot}(a) and
\ref{fig:approx-plot}(b). Quadratic errors are obtained by $[\mu -
E(Y)]^2$ for expectation and by $\left[ \frac{\mu}{\exp(\phi)} V(Y)
  \right ]^2$, where
$E(Y)$ and $V(Y)$ are numerically calculated using the moments
definition and dotted lines represents the regions $\nu \leq 1$ and
$\lambda > 10^\nu$ in the $(\mu\,,\phi)$ scale.

<<approx-plot, fig.height=3, fig.width=8, fig.cap="Quadratic errors for expectation (a) and variance (b) approximation, dotted lines representing the restriction for good approximations by \\cite{Shmueli2005}. Mean-variance relationship (c), dotted line represents $E(Y)=V(Y)$.">>=

#-------------------------------------------
# Errors in approximations for E[Y] and V[Y]
grid <- transform(grid,
                  emu = (mu - mean)^2,
                  eva = (va - var)^2)

myreg <- colorRampPalette(c("gray90",  "gray20"))
xy1 <- levelplot(emu ~ phi + mu, data = grid,
                 aspect = "fill",
                 col.regions = myreg(100),
                 xlab = expression(phi),
                 ylab = expression(mu),
                 sub = "(a)",
                 colorkey = list(space = "top"),
                 par.settings = ps2,
                 panel = function(x, y, z, ...) {
                     panel.levelplot(x, y, z, ...)
                     panel.curve(10 - (exp(x) - 1)/(2 * exp(x)),
                                 lty = 2)
                     panel.abline(v = 0, lty = 2)
                 })

xy2 <- levelplot(eva ~ phi + mu, data = grid,
                 aspect = "fill",
                 col.regions = myreg(100),
                 xlab = expression(phi),
                 ylab = expression(mu),
                 sub = "(b)",
                 colorkey = list(space = "top"),
                 par.settings = ps2,
                 panel = function(x, y, z, ...) {
                     panel.levelplot(x, y, z, ...)
                     panel.curve(10 - (exp(x) - 1)/(2 * exp(x)),
                                 lty = 2)
                     panel.abline(v = 0, lty = 2)
                 })

#-------------------------------------------
# Mean-variance relantionship
xy3 <- xyplot(var ~ mean,
              groups = phi,
              data = grid,
              type = "l",
              lwd = 2,
              axis = axis.grid,
              xlab = expression(E(X)),
              ylab = expression(V(X)),
              sub = "(c)",
              legend = list(
                  top = list(
                      fun = draw.colorkey,
                      args = list(
                          key = list(
                              space = "top",
                              col = myreg(length(unique(grid$phi))),
                              at = unique(grid$phi),
                              draw = FALSE)))),
              par.settings = modifyList(
                  ps2, list(superpose.line = list(
                                col = myreg(length(unique(grid$phi))))
                            )
              ),
              panel = function(x, y, ...) {
                  panel.xyplot(x, y, ...)
                  panel.curve(1*x, min(x), max(x), lty = 2)
              },
              page = function(...) {
                  grid.text(expression(phi),
                            just = "bottom",
                            x = unit(0.15, "npc"),
                            y = unit(0.83, "npc"))
              })

print(xy1, split = c(1, 1, 3, 1), more = TRUE)
print(xy2, split = c(2, 1, 3, 1), more = TRUE)
print(xy3, split = c(3, 1, 3, 1), more = TRUE)

@

The results in \autoref{fig:approx-plot} shows that the mean
approximation is accurate, see maximum quadratic errors is
\Sexpr{max(grid[["emu"]])}. For the variance approximation the maximum
is \Sexpr{max(grid[["eva"]])} and this occurs to negative values of
$\phi$.  Interestingly, the errors are bigger to negative values of
$\phi$ and appear to have no relation with $\mu$ against the regions
gives by \citet{Shmueli2005} ($\phi \leq 0$ and
$\mu > 10 - \frac{\exp(\phi) - 1}{2\exp(\phi)}$). The relationship
between mean and variance is linear with slope controlled by precision
parameter $\phi$ (\autoref{fig:approx-plot}(c)).

The results presented in \autoref{fig:approx-plot}(a) support the
proposed reparametrization. Replacing $\lambda$ and $\nu$ as function of
$\mu$ and $\phi$ in \autoref{eqn:pmf-cmp} the reparametrized
distribution takes the form
\begin{equation}
  \label{eqn:pmf-cmpmu}
  \Pr(Y=y \mid \mu, \phi) =
  \left ( \mu +\frac{ e^\phi-1}{2e^\phi} \right )^{ye^\phi}
  \frac{(y!)^{-e^\phi}}{Z(\mu, \phi)},
  \qquad y = 0, 1, 2, \ldots\,
\end{equation} where $\mu > 0$. We denotes this distribution as
COM-Poisson$_\mu$. In \autoref{fig:pmf-cmp} we show the shapes of
COM-Poisson$_\mu$ distribution.

<<pmf-cmp, fig.height=3, fig.width=7, fig.cap="Shapes of the COM-Poisson distribution for different parameter values.">>=

#-------------------------------------------
# COM-Poisson probabilities
parg <- expand.grid(mu = c(2, 8, 15), phi = c(-0.7, 0, 0.7))
y <- 0:30
py <- mapply(FUN = dcmp,
             mu = parg$mu,
             phi = parg$phi,
             MoreArgs = list(y = y, sumto = 100),
             SIMPLIFY = FALSE)
parg <- cbind(parg[rep(1:nrow(parg), each = length(y)), ],
              y = y, py = unlist(py))

leg_phi <- parse(
    text = paste("phi == \"",
                 formatC(unique(parg$phi), 1, format = "f"),
                 "\""))
barchart(py ~ y | factor(mu),
         groups = factor(phi),
         data = parg,
         horizontal = FALSE,
         layout = c(NA, 1),
         as.table = TRUE,
         axis = axis.grid,
         origin = 0,
         xlim = extendrange(y, f = 0.01),
         border = "transparent",
         scales = list(x = list(at = pretty(y))),
         ylab = expression(P(Y==y)),
         xlab = expression(y),
         par.settings = list(
             superpose.polygon = list(
                 col = c("gray40", "gray60", "gray10")),
             superpose.line = list(
                 col = c("gray40", "gray60", "gray10"),
                 lwd = 2)
         ),
         auto.key = list(
             columns = 3,
             rectangles = FALSE,
             lines = TRUE,
             text = leg_phi
         ),
         strip = strip.custom(
             strip.names = TRUE,
             var.name = expression(mu == ""),
             sep = ""))

@

\section{Estimation and Inference}
\label{estimation-and-inference}

In this section we shall describe the estimation and inference methods
adopted in this paper. The two forms of COM-Poisson models are fitted by
maximization of the log-likelihood function. The log-likelihood function
based on a set of independent observation $y_i$, $i=1,2,\ldots,n$ for
COM-Poisson$_\mu$ has the following form,
\begin{equation}
  \label{eqn:ll-rcmp}
  \ell = \ell(\bm{\beta}, \phi \mid \bm{y}) =
  e^\phi \left [
    \sum_{i=1}^n y_i
    \log \left( \mu_i + \frac{e^\phi-1}{2e^\phi} \right ) -
    \sum_{i=1}^n \log(y_i!) \right ] -
  \sum_{i=1}^n \log(Z_i(\mu_i, \phi)),
\end{equation}
where $\mu_i = \exp(\bm{x}_i^\top\bm{\beta})$, with
$\bm{x}_i^\top = (x_{i1},\, x_{i2},\, \ldots,\, x_{ip})$ the vector of
known covariates for the $i$-th observation, and $(\bm{\beta},\, \phi)
\in \mathbb{R}^{p+1}$. The normalizing constant $Z_i(\mu_i, \phi)$ is
given by
\begin{equation*}
  Z_i(\mu_i, \phi) = \sum_{j=0}^\infty \left [ \left (
    \mu_i + \frac{e^\phi - 1}{2e^\phi} \right )^{je^\phi}
  \frac{1}{(j!)^{e^\phi}} \right ]
\end{equation*}
To compute the log-likelihood function we need the calculation of an
infinite series for each observation, which makes the computation
expensive for parametric space regions that the convergence of sum is
slow.

Parameter estimation requires numerical maximization of
\autoref{eqn:ll-rcmp}. Since $\ell$ derivatives cannot be solved in
closed forms, we adopted the BFGS algorithm \cite{Nocedal1995} that uses
numerical estimate for the Hessian matrix
$\mathcal{H}(\bm{\theta})$. Standard errors for regression coefficients
are obtained based on the observed information matrix
$\mathcal{I}(\bm{\theta})$, where
$\mathcal{I}(\bm{\theta}) = -\mathcal{H}(\bm{\theta})$ numerically
calculated by BFGS algorithm. Confidence intervals for $\hat{\mu}_i$ are
obtained by using the delta method \citep[p. 89]{Pawitan2001}.

The parameter estimation for COM-Poisson model in the original
parametrization is analogous to presented for COM-Poisson$_\mu$, however
considering \autoref{eqn:ll-rcmp} in terms of $\lambda$. Even for the
standard COM-Poisson distribution, the precision parameter is taken on
the log scale, to avoid restriction of the parametric space.

Values of the maximized log-likelihood, the Akaike criterion (AIC) and
Bayesian criterion (BIC) are computed to compare the parametric
models. As a baseline model, in the applications we fitted the
quasi-Poisson model \citep{Wedderburn1974}. This approach is based on
second-moment assumption that allows more flexibility to model. In this
case the variance of the response variable is fixed by a additional
parameter $\sigma$, $\textrm{var}(Y_i)=\sigma \mu_i$. These models are
fitted in software R \citep{Rcore2017} by function \texttt{glm(...,
  family = quasipoisson)}.

\section{Simulation study}
\label{simulation-study}

In this section we conducted a simulation study to assess the properties
of the proposed model and estimators0. We considered average counts varying
from 3 to 27 according to a regression model structure with a continuous
variable and a categorical variable with three levels. With respect to
the precision parameter $\phi$, we fixed it to four values $-1.6$,
$-1.0$, $0.0$ and $1.8$, in order to have strong overdispersed, moderate
overdispersed, equidispersed and underdispersed counts respectively. For
each value of precision parameter, we considered four different sample
sizes, 50, 100, 300 and 1000, generating 1000 datasets in each case. The
\autoref{alg:simulation} shows the details of the steps in simulation
study. In \autoref{fig:bias-plot} we show the bias of the estimators for
each scenario (combination between values of the dispersion parameter
and samples sizes) together with a confidence interval calculated as
average bias plus and minus $\Phi(0.975)$ times the standard error. The
scales are standardized for each parameter by dividing the bias by the
standard error obtained for the sample of size 50.

\begin{algorithm}
\caption{Steps in simulation study.}
\label{alg:simulation}
  \Begin{
  $\bm{\beta} = \begin{bmatrix} \beta_0 & \beta_1 & \beta_{21} &
    \beta_{22} \end{bmatrix}^\top = \begin{bmatrix} 2.0 &
    0.5 & 0.8 & -0.8 \end{bmatrix}^\top$\;
  \For{$n \in \{50, 100, 300, 1000\}$}{
    set $\bm{x}_1$ as a sequence, with $n$ elements, between $0$ and
    $1$\;
    set $\bm{x}_2$ as a repetition, with $n$ elements, of three
    categories\;
    compute $\bm{\mu}$ using $\bm{\mu} = \exp(\beta_0 + \beta_1 \bm{x}_1
    +  \beta_{21} \bm{x}_{21} + \beta_{22} \bm{x}_{22})$, where
    $\bm{x}_{21}$ and $\bm{x}_{22}$ are \textit{dummy} variable for
    $\bm{x}_2$\;
    \For{$\phi \in \{-1.6, -1.0, 0.0, 1.8\}$}{
      \Repeat{$1000$ times}{
        simulate $\bm{y}$ from COM-Poisson distribution with $\bm{\mu}$
        and $\phi$ parameters\;
        fit COM-Poisson$_\mu$ regression model to $\bm{y}$ data with
        $\bm{X} = \begin{bmatrix} \bm{1} & \bm{x}_1 & \bm{x}_{21} &
          \bm{x}_{22} \end{bmatrix}$ design matrix\;
        get $\hat{\bm{\theta}} = \begin{bmatrix} \hat{\phi} & \hat{\beta}_0 &
          \hat{\beta}_1 & \hat{\beta}_{21} & \hat{\beta}_{22}
        \end{bmatrix}^\top$\;
        get confidence intervals for $\hat{\bm{\theta}}$ by quadratic
        approximation at the maximum likelihood estimate
        (assumes $\hat{\bm{\theta}} \sim \mathcal{N}(\hat{\bm{\theta}},
        \sqrt{-\bm{v}})$, where $\bm{v}$ is a diagonal of the inverse of
        the Hessian matrix)\;
      }
    }
  }
}
\end{algorithm}

<<load-simulation, cache=TRUE>>=

# Configuration
B <- 1000
beta <- c("b0" = 2, "b1" = 0.5, "b21" = 0.8, "b22" = -0.8)
phis <- c(0, -1.6, -1, 1.8)
names(phis) <- sprintf("phi=%s", phis)

sizes <- c(50, 100, 300, 1000)
names(sizes) <- sprintf("n=%s", sizes)

# Load results
results <- readRDS("./codes/simulation.rds")

@

<<bias-data, cache=TRUE>>=

#-----------------------------------------------------------------------
# Compute standardized bias
std <- lapply(results["n=50"], function(x) {
    ind <- names(x); names(ind) <- ind
    lapply(ind, function(y) {
        bhat <- t(vapply(x[[y]], "[[", double(5), "coef"))
        real <- matrix(c(phis[y], beta), byrow = TRUE,
                       nrow = B, ncol = 5)
        matrix(apply(bhat - real, 2, sd, na.rm = TRUE),
               byrow = TRUE, nrow = B, ncol = 5)
    })
})

aux <- ldply(lapply(results, function(x) {
    ind <- names(x); names(ind) <- ind
    out <- lapply(ind, function(y) {
        bhat <- t(vapply(x[[y]], "[[", double(5), "coef"))
        real <- matrix(c(phis[y], beta), byrow = TRUE,
                       nrow = B, ncol = 5)
        # (bhat - real)                 # raw bias
        (bhat - real) / std[[1]][[y]] # standardized bias
    })
    ldply(out, .id = "phi")
}), .id = "n")

# Organize the results
aux <- na.omit(aux)
bias <- gather(aux, param, bias, phi2:b22, factor_key = TRUE)
bias$phi <- ordered(bias$phi, c("phi=-1.6", "phi=-1", "phi=0",
                                "phi=1.8"))

@

The results in \autoref{fig:bias-plot} show that for all dispersed
levels, both the expected bias and standard error tend to 0 as the
sample size is increased. The estimators for the regression parameters
are unbiased, consistency and their empirical distributions are
symmetric. For the precision parameter, the estimator is asymptotically
unbiased, in small samples the parameter is overestimated and the
empirical distribution is right-skewed. \autoref{fig:coverage-plot}
presents the confidence interval coverage rate based on quadratic
approximation at the maximum likelihood estimate by sample size and
dispersed levels.

<<bias-plot, cache=TRUE, fig.height=5, fig.width=8, fig.cap="Distributions of standardized bias (gray points) and average with confidence intervals (black segments) by differents sample sizes and dispersion levels.">>=

# Distributions of bias and average with confidence intervals
ci <- function(x) {
    ci <- mean(x) + c(-1, 0, 1) * qnorm(0.975) * sd(x)
    names(ci) <- c("lwr", "fit", "upr")
    ci
}
cidata <- aggregate(bias ~ n + phi + param, data = bias, ci)

key <- list(
    type = "o",
    divide = 1,
    columns = 4,
    title = "Sample size",
    cex.title = 1.1,
    lines = list(pch = c(21:24), cex = 0.8),
            text = list(names(sizes))
)

fl <- parse(text = gsub("=", "==", levels(bias$phi)))
yl <- parse(text = c("hat(phi)",
                     paste0("hat(beta)[", c(0, 1, 21, 22), "]")))

xyplot(param ~ bias | phi, groups = n, data = bias,
       panel = panel.superpose,
       layout = c(4, 1),
       key = key,
       strip = strip.custom(factor.levels = fl),
       scales = list(y = list(labels = yl)),
       ylab = "",
       xlab = "Standardized Bias",
       jitter.y = TRUE,
       factor = 0.1,
       gap = 0.3,
       alpha = 0.6,
       col = "gray80",
       cex = 0.7,
       panel.groups = function(x, y, group.number,
                               subscripts, gap, ...){
           noise <- centfac(factor(levels(bias$n)), space = gap)
           noise <- sort(noise)
           panel.xyplot(x, y + noise[group.number], ...)
       }) +
    as.layer(
        segplot(param ~ bias[, "lwr"] + bias[, "upr"] | phi,
                centers = bias[, "fit"],
                data = cidata,
                draw = FALSE,
                horizontal = TRUE,
                layout = c(4, 1),
                groups = n, gap = 0.3,
                key = key,
                pch = 21:24,
                cex = 0.7,
                lwd = 2,
                panel = function(...) {
                    panel.groups.segplot(...)
                    panel.abline(v = 0, lty = 2)
                    panel.abline(h = 1:5, col = "lightgray", lty = 2)
                })
    )
@

<<coverage-data>>=

#-----------------------------------------------------------------------
# Compute coverage rate
aux <- ldply(lapply(results, function(x) {
    ind <- names(x); names(ind) <- ind
    out <- lapply(ind, function(y) {
        ind <- lapply(x[[y]], function(z) {
            real <- c(phis[[y]], beta)
            cint <- z[["cint"]]
            ind <- as.integer(cint[, 1] < real & cint[, 2] > real)
            names(ind) <- c("phi2", names(beta))
            ind
        })
        do.call("rbind", ind)
    })
    ldply(out, .id = "phi")
}), .id = "n")

# Organize the results
aux <- na.omit(aux)
coverage <- gather(aux, param, coverage, phi2:b22, factor_key = TRUE)
coverage$phi <- ordered(coverage$phi, c("phi=-1.6", "phi=-1", "phi=0",
                                        "phi=1.8"))
coverage$n <- as.numeric(gsub("n=([0-9]+)", "\\1", coverage$n))
covdata <- aggregate(coverage ~ n + phi + param, data = coverage,
                     function(x) sum(x == 1) / length(x))

@

The coverage rates in \autoref{fig:coverage-plot} show that for the
regression parameters the empirical coverage rates are close to the
nominal level of 95\% for sample sizes greater than 100 and all
dispersed levels. For the precision parameter the empirical coverage
rates are slightly lower than the nominal level, however, they become
closer for large samples. The worst scenario is when we have small
sample size and strong overdispersed counts.

<<coverage-plot, fig.height=4, fig.width=9, fig.cap="Coverage rate based on confidence intervals obtained by quadratic approximation for differents sample sizes and dispersion levels.">>=

fl <- parse(text = gsub("=", "==", levels(bias$phi)))
yl <- parse(text = c("hat(phi)",
                     paste0("hat(beta)[", c(0, 1, 21, 22), "]")))
xyplot(coverage ~ n | param,
       type = c("g", "p", "l"),
       groups = phi,
       data = covdata,
       pch = 19,
       lwd = 1.5,
       layout = c(NA, 1),
       xlab = "Sample size",
       ylab = "Coverage rate",
       par.settings = list(
           superpose.symbol = list(pch = 19),
           layout.heights = list(strip = 1.5)
       ),
       auto.key = list(
           column = 4,
           lines = TRUE,
           text = fl
       ),
       strip = strip.custom(
           factor.levels = yl
       ),
       panel = function(x, y, ...) {
           panel.xyplot(x, y, ...)
           panel.abline(h = 0.95, lty = 2)
       })

@

\section{Case studies}
\label{case-studies}

In this section, we report three illustrative examples of real count
data analysis obtained of the experimental research. We considered as
alternative models for the analysis the standard Poisson model, the
COM-Poisson model in two forms (original and mean parametrization) and
Quasi-Poisson model, based on second-moment assumptions. The data sets
and \texttt{R} codes for their analysis are available on on-line
supplements.

\subsection{Artificial defoliation in cotton phenology}
\label{case-cotton}

This example relates to cotton plants (\textit{Gossypium hirsutum})
submitted to five levels of artificial defoliation (\texttt{des}) and
observed on five growth stages (\texttt{est}). The purpose of this
research was to study the effect of defoliation levels at different
growth stages of cotton plants on the observed number of bolls
produced. The study was conducted in a greenhouse and the experimental
design was completely randomized with five replicates. This data set was
analysed in \citet{Zeviani2014} using the \textit{Gamma-Count}
distribution.

Following the results of \citet{Zeviani2014}, we considered the linear
predictor specified by
$$\log(\mu_{ij}) = \beta_0 + \beta_{1j} \textrm{def}_i + \beta_{2j}
\textrm{def}_i^2$$
where $\mu_{ij}$ is the expected number of cotton bolls for the $i$-th
defoliation level ($i=$ 1: 0\%, 2: 25\%, 3: 50\%, 4: 75\% e 5: 100\%)
and $j$-th growth stage ($j$ = 1: vegetative, 2: flower bud, 3: blossom,
4: boll, 5: boll open), that is, we have a second order effect of
defoliation in each growth stage. The parameters estimates and
goodness-of-fit measures for the Poisson, COM-Poisson, COM-Poisson$_\mu$
e quasi-Poisson models are presented in \autoref{tab:coef-cotton}.

<<fit-cotton, cache=TRUE, include=FALSE>>=

#-----------------------------------------------------------------------
# Load data
# data(cottonBolls, package = "cmpreg")
cottonBolls <- read.table("./data/cottonBolls.txt",
                          header = TRUE, sep = "\t")
cottonBolls$est <- ordered(
    cottonBolls$est,
    c("vegetative", "flower bud", "blossom", "boll", "boll open")
)

#-----------------------------------------------------------------------
# Fit models
mnames <- c("PO", "C1", "C2", "QP")

# Predictor, following Zeviani et al. (2014)
form1 <- ncap ~ est:(des + I(des^2))

m1PO <- glm(form1, data = cottonBolls, family = poisson)
time11 <- system.time(
    m1C1 <- fitcm(form1, data = cottonBolls, model = "CP", sumto = 50)
)
time12 <- system.time(
    m1C2 <- fitcm(form1, data = cottonBolls, model = "CP2", sumto = 50)
)
m1QP <- glm(form1, data = cottonBolls, family = quasipoisson)

models.ncap <- list(m1PO, m1C1, m1C2, m1QP)
names(models.ncap) <- mnames

# Numbers of calls to loglik and numerical gradient
c11 <- models.ncap$C1@details$counts
c12 <- models.ncap$C2@details$counts

# LRT between Poisson and COM-Poisson (test: phi == 0)
lrt.ncap <- getAnova(m1PO, m1C2)

@

\begin{table}[h]
\centering
\caption{Parameter estimates (Est) and ratio between estimate and
  standard error (SE) for the four model strategies for the cotton
  experiment.}
\label{tab:coef-cotton}
\begin{tabularx}{\textwidth}{lCCCCCCCC}
  \toprule
  & \multicolumn{2}{c}{Poisson} &
    \multicolumn{2}{c}{COM-Poisson} &
    \multicolumn{2}{c}{COM-Poisson$_\mu$} &
    \multicolumn{2}{c}{Quasi-Poisson} \\
\cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7} \cmidrule(lr){8-9}
<<results-cotton, results="asis">>=

#-----------------------------------------------------------------------
# Goodness of fit measures and estimate parameters

# GoF measures
measures.ncap <- lapply(models.ncap, function(x)
    c("LogLik" = logLik(x), "AIC" = AIC(x), "BIC" = BIC(x)))

# Get the estimates
co1 <- coef(m1C2)
est <- lapply(models.ncap, FUN = function(x) getCoefs(x))
est.ncap <- do.call(cbind, est)

# Organize in table
pnames <- c("\\phi\\,,\\,\\sigma", "\\beta_0",
            paste0("\\beta_{1", 1:5, "}"),
            paste0("\\beta_{2", 1:5, "}"))
rownames(est.ncap) <- paste0("$", pnames, "$")
meds <- apply(do.call(cbind, measures.ncap), 1, function(x) {
    x <- formatC(x, getOption("digits"), format = "f")
    x <- gsub("NA", "---", x)
    paste(paste0("\\multicolumn{2}{c}{", x, "}"),
          collapse = " & ")
})
text_gof <- paste(paste(names(meds), "&", meds),
                  collapse = "\\\\\n ")

append_gof <- list(
    pos = list(nrow(est.ncap)),
    command = paste("\\specialrule{0.01em}{0.3em}{0.3em} \n",
                    text_gof, "\\\\\n",
                    "\\bottomrule"))
print.xtable(xtable(est.ncap, digits = 4,
                    label = "tab:coef-cotton"),
             hline.after = 0,
             only.contents = TRUE,
             add.to.row = append_gof)

@
\end{tabularx}
\end{table}

The results presented in \autoref{tab:coef-cotton} show that the
goodness-of-fit measures (maximum log-likelihood, AIC and BIC) are quite
similar between COM-Poisson and COM-Poisson$_\mu$. This shows that the
reparametrization does not change the model fit, as expected. The Poisson
model is clearly unsuitable, being overly conservative. The difference
between the log-likelihoods of the Poisson and COM-Poisson$_\mu$ models
is \Sexpr{lrt.ncap[2, 4]}. Comparing this value with the chi-square
distribution, with one degree of freedom, we have these models are
significantly different ($\phi \neq 0$). The estimated value of
precision parameter, $\hat{\phi}$ is \Sexpr{co1[1]}, indicating
underdispersion.

\autoref{tab:coef-cotton} also shows the advantage of COM-Poisson$_\mu$
(reparametrized), since the estimates are quite similar to obtained by
the Poisson model, whereas estimates obtained by the COM-Poisson model
in original parametrization are on a non-interpretable scale. The ratios
between estimates and their respective standard deviations for the
COM-Poisson models are very close to ratios obtained by quasi-Poisson
model. However, it is important to note that COM-Poisson model is a
full parametric approach, that is, there is a probability distribution
associate to counts.

In the \autoref{fig:pred-cotton} are presented the observed values and
curves of fitted values with confidence intervals (95\%) as functions of
the defoliation level for each growth stage. The fitted values are same
for both models, but confidence intervals are bigger for Poisson model
because the equidispersion assumption. The results from COM-Poisson
models are consistent with those from the Gamma-Count model, fitted by
\cite{Zeviani2014}, Poisson-Tweedie model, fitted by \cite{Bonat2017}
and alternative COM-Poisson parametrization, fitted by \cite{Huang2017},
in that both methods indicate underdispersion and significant effects of
defoliation for the vegetative, blossom and boll growth stages.

<<pred-cotton, fig.height=3.5, fig.width=8.5, fig.cap="Scatterplots of the observed data and curves of fitted values with 95\\% confidence intervals as functions of the defoliation level for each growth stage.">>=

#-----------------------------------------------------------------------
# Prediction

# Data for prediction
pred <- with(cottonBolls,
             expand.grid(
                 est = levels(est),
                 des = seq(min(des), max(des), l = 20)
             ))
qn <- qnorm(0.975) * c(fit = 0, lwr = -1, upr = 1)

# Design matrix for prediction
X <- model.matrix(update(form1, NULL~.), pred)

# Considering Poisson
aux <- exp(confint(
    glht(m1PO, linfct = X), calpha = univariate_calpha())$confint)
colnames(aux) <- c("fit", "lwr", "upr")
aux <- data.frame(modelo = "Poisson", aux)
predPO.ncap <- cbind(pred, aux)

# Considering COM-Poisson
aux <- predictcm(m1C1, newdata = X)
aux <- data.frame(modelo = "COM-Poisson", aux)
predC1.ncap <- cbind(pred, aux)

# Considering COM-Poisson (mean parametrization)
aux <- predictcm(m1C2, newdata = X)
aux <- data.frame(modelo = "COM-Poisson2", aux)
predC2.ncap <- cbind(pred, aux)

# Considering Quasi-Poisson
aux <- exp(confint(
    glht(m1QP, linfct = X), calpha = univariate_calpha())$confint)
colnames(aux) <- c("fit", "lwr", "upr")
aux <- data.frame(modelo = "Quasi-Poisson", aux)
predQP.ncap <- cbind(pred, aux)

# Representing the confidence intervals
pred.ncap <- rbind(predPO.ncap, predC1.ncap, predC2.ncap, predQP.ncap)

# Legend
key <- list(columns = 4,
            cex = 0.9,
            lines = list(col = 1, lty = rev(1:4)),
            text = list(parse(
                text = c("'Poisson'", "'COM-Poisson'",
                         "'COM-Poisson'[mu]", "'Quasi-Poisson'"))
                ))

# Graph
xyplot(ncap ~ des | est,
       data = cottonBolls,
       layout = c(NA, 1),
       as.table = TRUE,
       grid = TRUE,
       type = "p",
       xlab = "Artificial defoliation level",
       ylab = "Number of bolls produced",
       spread = 0.05,
       key = key,
       alpha = 0.6,
       panel = panel.beeswarm) +
    as.layer(
        xyplot(fit ~ des | est,
               auto.key = TRUE,
               data = pred.ncap,
               groups = modelo,
               type = "l",
               layout = c(NA, 1),
               as.table = TRUE,
               col = 1,
               ly = pred.ncap$lwr,
               uy = pred.ncap$ upr,
               cty = "bands",
               fill = "gray80",
               alpha = 0.1,
               panel = panel.superpose,
               panel.groups = panel.cbH,
               prepanel = prepanel.cbH,
               lty = rev(1:4))
    )

@

In order to assess the orthogonality between $\bm{\mu}$ and $\phi$ in
the COM-Poisson$_\mu$ parametrization, \autoref{tab:corr-cotton}
presents the empirical correlations between precision and regression
parameters, obtained by variance and covariance matrix. The correlations
are practically null considering the COM-Poisson$_\mu$, whereas to the
original parametrization they are quite high. This reflects a better
performance of the maximization algorithm in mean-parametrization model,
in this case COM-Poisson$_\mu$ fit was \Sexpr{time12[3] * 100 /
  time11[3]}\% slower then COM-Poisson model.

<<corr-cotton, results="asis">>=

#-----------------------------------------------------------------------
# Correlation between estimates
corr.ncap <- do.call("rbind",
                     lapply(models.ncap[c("C1", "C2")],
                            function(x) cov2cor(vcov(x))[1, -1]))

# Organize on table
rownames(corr.ncap) <- paste0("COM-Poisson", c("", "$_\\mu$"))
colnames(corr.ncap) <- gsub("beta", "hat{\\\\beta}", pnames[-1])

caption <- paste("Empirical correlations between $\\hat{\\phi}$ and",
                 "$\\hat{\\bm{\\beta}}$ for the two parametrizations",
                 "of COM-Poisson model fit to underdispersed data.")
print(xtable(corr.ncap,
             align = c("lccccccccccc"),
             caption = caption,
             digits = 3,
             label = "tab:corr-cotton"),
      size = "small",
      sanitize.rownames.function = identity,
      sanitize.colnames.function = function(x) sprintf("$%s$", x))

@

\subsection{Soil moisture and potassium doses on soy bean culture}
\label{case-soybean}

The second example is a experiment completely randomized block with
treatments in a $5\times 3$ factorial arrangement about soy bean
culture. The aim of this study was to evaluate the effects of potassium
doses (\texttt{K}) applied to soil (0, 0.3, 0.6, 1.2 e 1.8 100mg
dm$^{-3}$) and soil humidity (\texttt{umid}) levels (37.5, 50, 62.5\%,
which represent few, ideal and abundance water respectively) on the
soy bean production. The experiment was carried out in a greenhouse, in
pots with two plants, and the count variable measured was the number of
grains per pot \citep{Serafim2012}. \autoref{fig:desc-soy} (left) shows
the number of grains recorded for each combination of potassium dose and
humidity level, it is important to note the indication of a quadratic
level of the potassium levels. Most points in the sample means and
variances dispersion diagram (right) are above the identity line,
suggesting overdispersion.

<<fit-soy, cache=TRUE, include=FALSE>>=

#-----------------------------------------------------------------------
# Load data
# data(soyaBeans, package = "cmpreg")
soyBeans <- read.table("./data/soyaBeans.txt",
                        header = TRUE, sep = "\t")
soyBeans$umid <- as.factor(soyBeans$umid)
soyBeans <- soyBeans[-74, ] # Incorrect observation
soyBeans <- transform(soyBeans, K = K / 100)

#-----------------------------------------------------------------------
# Fit models

# Predictor
form2 <-  ngra ~ bloc + umid * K + I(K^2)

m2PO <- glm(form2, data = soyBeans, family = poisson)
time21 <- system.time(
    m2C1 <- fitcm(form2, data = soyBeans, model = "CP", sumto = 700)
)
time22 <- system.time(
    m2C2 <- fitcm(form2, data = soyBeans, model = "CP2", sumto = 700)
)
m2QP <- glm(form2, data = soyBeans, family = quasipoisson)

models.ngra <- list(m2PO, m2C1, m2C2, m2QP)
names(models.ngra) <- mnames

# Numbers of calls to loglik and numerical gradient
c21 <- models.ngra$C1@details$counts
c22 <- models.ngra$C2@details$counts

# # Profile extra parameter
# profs.ngra <- lapply(list(c(m2C1, "phi"), c(m2C2, "phi2")),
#                      function(x) myprofile(x[[1]], x[[2]]))
# profs.ngra <- do.call("rbind", profs.ngra)

# LRT between Poisson and COM-Poisson (test: phi == 0)
lrt.ngra <- getAnova(m2PO, m2C2)

@

<<desc-soy, fig.pos="H", fig.height=4, fig.width=8, fig.cap="Number of grains per pot for each potassium dose and humidity level (left) and sample mean agains sample variance of the five replicates for each experimental treatment (right).">>=

#-----------------------------------------------------------------------
# Exploratory analysis

# Scatter plot
xy1 <- xyplot(ngra ~ K | umid,
              data = soyBeans,
              xlab = "Potassium fertilization level",
              ylab = "Number of grains per pot",
              type = c("p", "g", "smooth"),
              as.table =  TRUE,
              layout = c(2, 2),
              strip = strip.custom(
                  strip.names = TRUE, var.name = "humidity",
                  factor.levels = paste0(levels(soyBeans$umid), "%")))

# Sample variance vs sample mean (evidence in favor of the
# overdispersion).
mv <- soyBeans %>%
    group_by(K, umid) %>%
    summarise(mu = mean(ngra), va = var(ngra))
xlim <- ylim <- extendrange(c(mv$mu, mv$va), f = 0.05)

xy2 <- xyplot(va ~ mu,
              data = mv,
              type = c("p", "r", "g"),
              xlim = xlim,
              ylim = ylim,
              xlab = expression("Sample"~"mean"~(bar(y))),
              ylab = expression("Sample"~"variance"~(s^2)),
              panel = function(...) {
                  panel.xyplot(...)
                  panel.abline(a = 0, b = 1, lty = 2)
              })

print(xy1, split = c(1, 1, 2, 1), more = TRUE)
print(xy2, split = c(2, 1, 2, 1), more = FALSE)

@

For the analysis of this data set we proposed, based on descriptive
analysis \autoref{fig:desc-soy}, the linear predictor specified as a
linear doses for each humidity level and a global quadratic dose effect,
i.e
$$
\log(\mu_{ijk}) = \beta_0 + \gamma_i + \tau_j +
  \beta_{1}\texttt{K}_k + \beta_{2}\texttt{K}_k^2 +
  \beta_{3j}\texttt{K}_k
$$
with $i=$1: block II, 2: block III, 3: block IV e 4: block V; $j=$1:
50\% e 2: 62.5\%; and $k=$1: 0.0, 2: 0.3, 3: 0.6, 4: 1.2, 5: 1.8 100mg
dm$^{-3}$, where $\gamma_i$ is the effect of $i$-th block, $\tau_j$ is
the effect of $j$-th humidity level and $\beta_{3j}$ is the first order
potassium effect (\texttt{K}) for the $j$-th humidity level
(\texttt{umid}). \autoref{tab:coef-soy} presents the estimates, ratio
between estimate and standard deviation and goodness-of-fit measures for
the alternative models fitted.

The results in \autoref{tab:coef-soy} has similar interpretations to
results presented in \autoref{case-cotton}. The two parametrization of
COM-Poisson presented very similar goodness-of-fit measures and better
than Poisson model. The difference between the log-likelihoods of the
Poisson and COM-Poisson models is \Sexpr{lrt.ngra[2, 4]}, indicating
that $\phi$ is significantly different of zero. With respect to the
regression parameters, the similarities between models are analogous to
the previous section. Both models agree on the indication of effects,
however the Poisson model indicates the effects with greater
significance, because it does not fit the extra variability.

\begin{table}[ht]
\centering
\caption{Parameter estimates (Est) and ratio between estimate and
  standard error (SE) for the four model strategies for the soy bean
  experiment.}
\label{tab:coef-soy}
\begin{tabularx}{\textwidth}{lCCCCCCCC}
  \toprule
  & \multicolumn{2}{c}{Poisson} &
    \multicolumn{2}{c}{COM-Poisson} &
    \multicolumn{2}{c}{COM-Poisson$_\mu$} &
    \multicolumn{2}{c}{Quasi-Poisson} \\
\cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7} \cmidrule(lr){8-9}
<<results-soya, results="asis">>=

#-----------------------------------------------------------------------
# Goodness of fit measures and estimate parameters

# GoF measures
measures.ngra <- lapply(models.ngra, function(x)
    c("LogLik" = logLik(x), "AIC" = AIC(x), "BIC" = BIC(x)))

# Get the estimates
co2 <- coef(m2C2)
est <- lapply(models.ngra, FUN = function(x) getCoefs(x))
est.ngra <- do.call(cbind, est)

# Organize in table
pnames <- c("\\phi\\,,\\,\\sigma", "\\beta_0",
            paste0("\\gamma_{", 1:4, "}"),
            paste0("\\tau_{", 1:2, "}"),
            "\\beta_1", "\\beta_2",
            paste0("\\beta_{3", 1:2, "}"))

rownames(est.ngra) <- paste0("$", pnames, "$")
meds <- apply(do.call(cbind, measures.ngra), 1, function(x) {
    x <- formatC(x, getOption("digits"), format = "f")
    x <- gsub("NA", "---", x)
    paste(paste0("\\multicolumn{2}{c}{", x, "}"),
          collapse = " & ")
})
text_gof <- paste(paste(names(meds), "&", meds),
                  collapse = "\\\\\n ")

append_gof <- list(
    pos = list(nrow(est.ngra)),
    command = paste("\\specialrule{0.01em}{0.3em}{0.3em} \n",
                    text_gof, "\\\\\n",
                    "\\bottomrule"))
print.xtable(xtable(est.ngra, digits = 4,
                    label = "tab:coef-soy"),
             hline.after = 0,
             only.contents = TRUE,
             add.to.row = append_gof)

@
\end{tabularx}
\end{table}

The infinite sum $Z(\mu, \phi)$ in cases of oversidepersed counts needs
a greater upper bound for the convergence, making the computation of the
likelihood more expansive. In this case, the upper bound was 700. The
number of likelihood computations, on the BFGS algorithm, was
\Sexpr{c21[1]} and \Sexpr{c22[2]} for the original and
mean-parametrization and the time to fit was \Sexpr{time21[3]} and
\Sexpr{time22[3]} seconds respectively. This may be due the
well-behaviour of likelihood function for orthogonal parameters. In
\autoref{tab:corr-soy} we present the empirical correlations between
regression and precision parameter. The correlations are close to zero
for the COM-Poisson$_\mu$ parametrization, indicating the empirical
orthogonality between $\mu$ and $\phi$ on COM-Poisson$_\mu$
distribution.

<<corr-soy, results="asis">>=

##----------------------------------------------------------------------
## Correlation between estimates
corr.ngra <- do.call("rbind",
                     lapply(models.ngra[c("C1", "C2")],
                            function(x) cov2cor(vcov(x))[1, -1]))

## Organize on table
rownames(corr.ngra) <- paste0("COM-Poisson", c("", "$_\\mu$"))
colnames(corr.ngra) <- gsub("(beta|tau|gamma)", "hat{\\\\\\1}",
                            pnames[-1])

caption <- paste("Empirical correlations between $\\hat{\\phi}$ and",
                 "$\\hat{\\bm{\\beta}}$ for the two parametrizations",
                 "of COM-Poisson model fit to overdispersed data.")
print(xtable(corr.ngra,
             align = c("lccccccccccc"),
             caption = caption,
             digits = 3,
             label = "tab:corr-soy"),
      size = "small",
      sanitize.rownames.function = identity,
      sanitize.colnames.function = function(x) sprintf("$%s$", x))

@

The observed counts and fitted curves for each humidity level with
confidence bands are show in \autoref{fig:pred-soy}. The fitted values
are identical both models, leading the same conclusions. On the other
hand, confidence bands for the Poisson model are smaller due to
equidispersion assumption. Considering COM-Poisson models and
quasi-Poisson approach, the confidence bands are quite similar, which
shows their flexibility.

<<pred-soy, fig.pos="H", fig.height=3.8, fig.width=8, fig.cap="Dispersion diagrams of grain counts as function of potassium doses and humidity levels with fitted curves and confidence intervals (95\\%).">>=

#-----------------------------------------------------------------------
# Prediction

# Data for prediction
pred <- with(soyBeans,
             expand.grid(
                 bloc = factor(levels(bloc)[1], levels = levels(bloc)),
                 umid = levels(umid),
                 K = seq(min(K), max(K), l = 20)
             ))
qn <- qnorm(0.975) * c(fit = 0, lwr = -1, upr = 1)

# Design matrix for prediction
X <- model.matrix(update(form2, NULL ~ .), pred)
bl <- attr(X, "assign") == 1
X[, bl] <- X[, bl] + 1/(sum(bl) + 1)

# Considering Poisson
aux <- exp(confint(
    glht(m2PO, linfct = X), calpha = univariate_calpha())$confint)
colnames(aux) <- c("fit", "lwr", "upr")
aux <- data.frame(modelo = "Poisson", aux)
predPO.ngra <- cbind(pred, aux)

# Considering COM-Poisson
aux <- predictcm(m2C1, newdata = X)
aux <- data.frame(modelo = "COM-Poisson", aux)
predC1.ngra <- cbind(pred, aux)

# Considering COM-Poisson (mean parametrization)
aux <- predictcm(m2C2, newdata = X)
aux <- data.frame(modelo = "COM-Poisson2", aux)
predC2.ngra <- cbind(pred, aux)

# Considering Quasi-Poisson
aux <- exp(confint(
    glht(m2QP, linfct = X), calpha = univariate_calpha())$confint)
colnames(aux) <- c("fit", "lwr", "upr")
aux <- data.frame(modelo = "Quasi-Poisson", aux)
predQP.ngra <- cbind(pred, aux)

# Representing the confidence intervals
pred.ngra <- rbind(predPO.ngra, predC1.ngra, predC2.ngra, predQP.ngra)

# Legend
key <- list(columns = 4,
            cex = 0.9,
            lines = list(col = 1, lty = rev(1:4)),
            text = list(parse(
                text = c("'Poisson'", "'COM-Poisson'",
                         "'COM-Poisson'[mu]", "'Quasi-Poisson'"))
                )
            )

# Graph
update(xy1, layout = c(NA, 1), type = c("p", "g"),
       alpha = 0.6, key = key) +
    as.layer(
        xyplot(fit ~ K | umid,
               data = pred.ngra,
               groups = modelo,
               type = "l",
               col = 1,
               ly = pred.ngra$lwr,
               uy = pred.ngra$upr,
               cty = "bands",
               fill = "gray80",
               alpha = 0.1,
               panel = panel.superpose,
               panel.groups = panel.cbH,
               prepanel = cmpreg::prepanel.cbH,
               lty = rev(1:4))
    )

@

\subsection{Assessing toxicity of pollutants in aquatic systems}

The data set come from an experiment to measure the reproductive
toxicity of a herbicide, nitrofen, on a species of zooplankton
(\textit{Ceriodaphnia dubia}). 50 animals were randomized into batches
of 10 and each batch was put in a solution with a measured concentration
of nitrofen (0, 0.8, 1.6, 2.35 e 3.10 mug$/10^2$litre)
(\texttt{dose}). Then the number of total live offspring was recorded
\cite{Bailer1994}.

<<fit-ovos, cahe=TRUE, include=FALSE>>=

#-----------------------------------------------------------------------
# Load data
# data(nitrofen, package = "boot")
# data(Paula, package = "labestData")
# nitrofen <- PaulaEx4.6.20
nitrofen <- read.table("./data/nitrofen.txt",
                       header = TRUE, sep = "\t")
nitrofen <- transform(nitrofen, dose = dose / 100)

#-----------------------------------------------------------------------
# Fit models
mnames <- c("PO", "C1", "C2", "QP")

# Predictors
form31 <-  novos ~ dose
form32 <-  novos ~ dose + I(dose^2)
form33 <-  novos ~ dose + I(dose^2) + I(dose^3)

predictors <- list("pred1" = form31, "pred2" = form32, "pred3" = form33)
fmodels.ovos <- lapply(predictors, function(form) {
    PO <- glm(form, data = nitrofen, family = poisson)
    C1 <- fitcm(form, data = nitrofen, model = "CP", sumto = 100)
    C2 <- fitcm(form, data = nitrofen, model = "CP2", sumto = 100)
    QP <- glm(form, data = nitrofen, family = quasipoisson)
    list("PO" = PO, "C1" = C1, "C2" = C2, "QP" = QP)
})

#-----------------------------------------------------------------------
# LRT for nested models

# Poisson
auxPO <- lapply(fmodels.ovos, function(x) x$PO)
do.call("getAnova", auxPO)

# COM-Poisson standard
auxC1 <- lapply(fmodels.ovos, function(x) x$C1)
do.call("getAnova", auxC1)

# COM-Poisson mean-parameterized
auxC2 <- lapply(fmodels.ovos, function(x) x$C2)
do.call("getAnova", auxC2)

# Quasi-Poisson
auxQP <- lapply(fmodels.ovos, function(x) x$QP)
do.call("getAnova", auxQP)

#--------------------------------------------
# Separe the choose models
form3 <- form33
m3PO <- fmodels.ovos$pred3$PO
m3C1 <- fmodels.ovos$pred3$C1
m3C2 <- fmodels.ovos$pred3$C2
m3QP <- fmodels.ovos$pred3$QP

models.ovos <- list(m3PO, m3C1, m3C2, m3QP)
names(models.ovos) <- mnames

# Numbers of calls to loglik and numerical gradient
models.ovos$C1@details$counts
models.ovos$C2@details$counts

@

For this data set we proposed three linear predictors. We compare the
performance of models using the likelihood ratio tests to select the
systematic part of model. The linear predictors are
\begin{center}
\begin{minipage}{12cm}
Predictor 1: $\log(\mu_i) = \beta_0 + \beta_1 \texttt{dose}_i$\\
Predictor 2: $\log(\mu_i) = \beta_0 + \beta_1 \texttt{dose}_i +
             \beta_2 \texttt{dose}_i^2$\\
Predictor 3: $\log(\mu_i) = \beta_0 + \beta_1 \texttt{dose}_i +
             \beta_2 \texttt{dose}_i^2 + \beta_3 \texttt{dose}_i^3$.
\end{minipage}
\end{center}

\begin{table}[ht]
\centering
\caption{Model fit measures and comparisons between predictors and models.}
\label{tab:anova-ovos}
\begin{tabularx}{\textwidth}{lCCCCCrC}
  \toprule
 Poisson & np & $\ell$ & AIC & 2(diff $\ell$) & diff np & P($>\rchi^2$) & \\
 \midrule
<<anova-ovos1, results="asis">>=

auxPO <- lapply(fmodels.ovos, function(x) x$PO)
tab <- do.call("getAnova", c(print = FALSE, auxPO))
tab <- cbind(tab, NA)
rownames(tab) <- paste("Preditor", rownames(tab))
digits <- c(1, 0, 3, 3, 3, 0, -2, 3)
print(xtable(tab, digits = digits),
      include.colnames = FALSE,
      hline.after = NULL,
      only.contents = TRUE)

@
\specialrule{0em}{0.5em}{0em} %% Apenas para espaçamento
  COM-Poisson & np & $\ell$ & AIC & 2(diff $\ell$) & diff np &
  P($>\rchi^2$) & $\hat{\phi}$ \\
  \midrule
<<anova-ovos2, results="asis">>=

auxC1 <- lapply(fmodels.ovos, function(x) x$C1)
tab <- do.call("getAnova", c(print = FALSE, auxC1))
tab <- cbind(tab, sapply(auxC1, function(x) coef(x)[1]))
rownames(tab) <- paste("Preditor", rownames(tab))
print(xtable(tab, digits = digits),
      include.colnames = FALSE,
      hline.after = NULL,
      only.contents = TRUE)

@
\specialrule{0em}{0.5em}{0em} %% Apenas para espaçamento
  COM-Poisson$_\mu$ & np & $\ell$ & AIC & 2(diff $\ell$) & diff np &
  P($>\rchi^2$) & $\hat{\phi}$ \\
  \midrule
<<anova-ovos3, results="asis">>=

auxC2 <- lapply(fmodels.ovos, function(x) x$C2)
tab <- do.call("getAnova", c(print = FALSE, auxC2))
tab <- cbind(tab, sapply(auxC2, function(x) coef(x)[1]))
rownames(tab) <- paste("Preditor", rownames(tab))
print(xtable(tab, digits = digits),
      include.colnames = FALSE,
      hline.after = NULL,
      only.contents = TRUE)

@
\specialrule{0em}{0.5em}{0em} %% Apenas para espaçamento
  Quasi-Poisson & np & QDev & AIC & F & diff np & P($>F$) & $\hat{\sigma}$ \\
  \midrule
<<anova-ovos4, results="asis">>=

auxQP <- lapply(fmodels.ovos, function(x) x$QP)
tab <- do.call("getAnova", c(print = FALSE, auxQP))
tab <- cbind(tab, sapply(auxQP, function(x) summary(x)$dispersion))
rownames(tab) <- paste("Preditor", rownames(tab))
print(xtable(tab, digits = digits),
      include.colnames = FALSE,
      hline.after = NULL,
      only.contents = TRUE)

@
 \bottomrule
\end{tabularx}
\vspace{-1mm}

\footnotesize \raggedright np, number of parameters; diff $\ell$,
difference in log-likelihoods; QDev, quasi-deviance, F, F statistics
based on quasi-deviances; diff np, difference in np.
\end{table}

\autoref{tab:anova-ovos} summarizes the results of fitted models and
likelihood ratio tests comparing the sequence of predictors. All models
indicate the cubic effect of nitrofen dosing. Considering this
predictor, there is an evidence of equidispersed counts, $\phi$
estimates of COM-Poisson close to zero and $\sigma$ of quasi-Poisson, to
one. It is interesting to note that if we omit the effects of a greater
order on the predictor, the modes show evidence of overdispersion
counts. This exemplifies the discussion about causes of overdispersion
made in \autoref{introduction}. We can also note that the quasi-Poisson
approach, although robust to equisdispersion assumption, shows a higher
descriptive levels ($p$-values) than parametric models, that is, the
tests under parametric models are most powerful than quasi-Poisson
models in the equidispersed case.

In \autoref{tab:coef-ovos} we presented the estimates of the regression
parameters considering the cubic dose effect (third predictor). The
interpretations are similar to the discussed in the others cases
studies, but in this the Poisson model is also suitable for indicating
the significance of the effects. In addition, note that the parameters
of the standard COM-Poisson model are comparable to the others models,
this occurs because it is the particular case $\phi = 0$ which implies
$\lambda = \mu$. \autoref{fig:pred-ovos} shows the number of live
off-springs observed in experiment with fitted curves and confidence
bands for all model strategies adopted.

\begin{table}[H]
\centering
\caption{Parameter estimates (Est) and ratio between estimate and
  standard error (SE) for the four model strategies for the nitrofen
  experiment.}
\label{tab:coef-ovos}
\begin{tabularx}{\textwidth}{lCCCCCCCC}
  \toprule
  & \multicolumn{2}{c}{Poisson} &
    \multicolumn{2}{c}{COM-Poisson} &
    \multicolumn{2}{c}{COM-Poisson$_\mu$} &
    \multicolumn{2}{c}{Quasi-Poisson} \\
\cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7} \cmidrule(lr){8-9}
<<coef-ovos, results="asis">>=

#-----------------------------------------------------------------------
# Goodness of fit measures and estimate parameters

# GoF measures
measures.ovos <- lapply(models.ovos, function(x)
    c("LogLik" = logLik(x), "AIC" = AIC(x), "BIC" = BIC(x)))

# Get the estimates
co2 <- coef(m2C2)
est <- lapply(models.ovos, FUN = function(x) getCoefs(x))
est.ovos <- do.call(cbind, est)[-1, ]

# Organize in table
pnames <- paste0("\\beta_{", 0:3, "}")
rownames(est.ovos) <- paste0("$", pnames, "$")
print.xtable(xtable(est.ovos, digits = 4),
             hline.after = c(0, nrow(est.ovos)),
             only.contents = TRUE)

@
\end{tabularx}
\end{table}

The fitted values and confidence bands in \autoref{fig:pred-ovos} come
to have a full overlay. This shows that even in case where is not
necessary to estimate the precision parameter $\phi$, but estimating it
does not lead to incorrect analysis.

<<pred-ovos, fig.height=3.5, fig.width=5.5, out.width="0.7\\textwidth", fig.cap="Number of live offsprings observed for each nitrofen concentration level with fitted curves and 95\\% confidence intervals.">>=

#-----------------------------------------------------------------------
# Prediction

# Data for prediction
pred <- with(nitrofen,
             data.frame("dose" = seq(min(dose), max(dose),
                                     length.out = 100)))
qn <- qnorm(0.975) * c(fit = 0, lwr = -1, upr = 1)

# Design matrix for prediction
X <- model.matrix(update(form3, NULL ~ .), pred)

# Considering Poisson
aux <- exp(confint(
    glht(m3PO, linfct = X), calpha = univariate_calpha())$confint)
colnames(aux) <- c("fit", "lwr", "upr")
aux <- data.frame(modelo = "Poisson", aux)
predPO.novos <- cbind(pred, aux)

# Considering COM-Poisson
aux <- predictcm(m3C1, newdata = X)
aux <- data.frame(modelo = "COM-Poisson", aux)
predC1.novos <- cbind(pred, aux)

# Considering COM-Poisson (mean parametrization)
aux <- predictcm(m3C2, newdata = X)
aux <- data.frame(modelo = "COM-Poisson2", aux)
predC2.novos <- cbind(pred, aux)

# Considering Quasi-Poisson
aux <- exp(confint(
    glht(m3QP, linfct = X), calpha = univariate_calpha())$confint)
colnames(aux) <- c("fit", "lwr", "upr")
aux <- data.frame(modelo = "Quasi-Poisson", aux)
predQP.novos <- cbind(pred, aux)

# Representing the confidence intervals
pred.novos <- rbind(predPO.novos, predC1.novos, predC2.novos, predQP.novos)
ord <- order(pred.novos$dose, pred.novos$modelo)
pred.novos <- pred.novos[ord, ]

# Legend
key <- list(columns = 2,
            lines = list(col = 1, lty = rev(1:4)),
            text = list(parse(
                text = c("'Poisson'", "'COM-Poisson'",
                         "'COM-Poisson'[mu]", "'Quasi-Poisson'"))
                )
            )

# Graph
xyplot(novos ~ dose,
       data = nitrofen,
       xlab = "Nitrofen concentration level",
       ylab = "Number of live offspring",
       grid = TRUE,
       alpha = 0.6,
       key = key,
       spread = 0.05,
       panel = panel.beeswarm) +
    as.layer(
        xyplot(fit ~ dose,
               auto.key = TRUE,
               data = pred.novos,
               groups = modelo,
               type = "l",
               col = 1,
               ly = pred.novos$lwr,
               uy = pred.novos$upr,
               cty = "bands",
               fill = "gray80",
               alpha = 0.1,
               panel = panel.superpose,
               panel.groups = panel.cbH,
               prepanel = prepanel.cbH,
               lty = rev(1:4))
    )

@

Finally, in \autoref{tab:corr-ovos} we presented the empirical
correlations between regression and precision parameters. The results
show that even in the particular case ($\phi=0$), the empirical
correlations for the standard models are not zero. For the
reparametrized model, as discussed in the previous sections, the
correlations are practically null.

<<corr-ovos, results="asis">>=

#-----------------------------------------------------------------------
# Correlation between estimates
corr.ovos <- do.call("rbind",
                     lapply(models.ovos[c("C1", "C2")],
                            function(x) cov2cor(vcov(x))[1, -1]))

# Organize on table
rownames(corr.ovos) <- paste0("COM-Poisson", c("", "$_\\mu$"))
colnames(corr.ovos) <- gsub("(beta)", "hat{\\\\\\1}", pnames)

caption <- paste("Empirical correlations between $\\hat{\\phi}$ and",
                 "$\\hat{\\bm{\\beta}}$ for the two parametrizations",
                 "of COM-Poisson model fit to equidispersed data.")
print(xtable(corr.ovos,
             align = c("lCCCC"),
             caption = caption,
             digits = 3,
             label = "tab:corr-ovos"),
      # size = "small",
      sanitize.rownames.function = identity,
      sanitize.colnames.function = function(x) sprintf("$%s$", x),
      tabular.environment = "tabularx",
      width = "\\textwidth")

@

\section{Concluding remarks}
\label{conclusion}

In this paper, we argued the parametrization of COM-Poisson distribution
to analyse dispersed count data. We proposed a novel reparametrization
based only mean approximation gives by \cite{Shmueli2005}. The
likelihood-based method to estimation and inference is presented. We
assess the properties of these estimators by a simulation
study. Applications of the proposed model are showed by three case
studies with over, under and equidispersed counts. In the data analysis
we compared standard Poisson, original COM-Poisson, repametrized
COM-Poisson and quasi-Poisson models.

We computed the quadratic errors for the mean approximation. The results
showed that it is well accurate supports our proposed
reparametrization. The results of a simulation study showed that the
estimators of the regression and precision parameters are consistent and
unbiased and asymptotically unbiased respectively. The empirical
coverage rates of the confidence intervals for the parameters, compute
by the asymptotical distrition of the estimators, are close to the
nominal level for sizes greather than 100. The worst scenario is when we
have small sample size and strong overdispersed counts. In general, we
recommend the use of Wald intervals for computational simplicity and
good results.

The data analysis in the case studies has shown that COM-Poisson is a
suitable distribution for over, under and equidispersion situations. The
observed empirical correlation between regression and precision
parameters suggest the orthogonality between $\mu$ and $\phi$ in
COM-Poisson$\mu$ distribution against the original parametrization. Due
to this feature the computational procedure is faster under the
reparameterized model. Another advantage, perhaps the main, of the
reparametrization is that the estimated regression parameters have
interpretation of rate ratios, same of the Poisson model. In comparison
of the quasi-Poisson approach the results was similar in both
simulations, nevertheless the COM-Poisson is a full parametric model,
allowing the calculation of probabilities for example.

In general, the results presented by the reparameterized COM-Poisson
models were satisfactory and superior to the conventional
approaches. Therefore, its use in the analysis of count data is
encouraged. The computational routines for fitting of original and
reparametrized COM-Poisson models are available in the supplementary
material\textsuperscript{\ref{papercompanion}}.

There are many possible extensions and possible topics for futher
investigation to model discussed in the present paper, including
simulation study to assess model robustness against distribution
miss-specification, asses theoretical approximations for
$Z(\lambda, \nu)$ (or $Z(\mu,\phi))$, in order to avoid the selection of
sum's upper bound, propose a double GLM based on the reparametrized
COM-Poisson model and a accomodate random effects since we have a model
parametrized on mean.

%-----------------------------------------------------------------------
\newpage

\small
\bibliography{references.bib}
\end{document}

### 2017-09-18 codes/01-simulation.R
Explain the choices in the simulation study

@@ -25,24 +25,42 @@ sizes <- c(50, 100, 300, 1000)
 names(sizes) <- sprintf("n=%s", sizes)
 
 #-------------------------------------------
-# Visualize the averages counts used in simulations
-x1 <- seq(0, 1, length.out = 1000)
-x2 <- rep(letters[1:3], length.out = 1000)
+# Justify the choices of the parameters simulation
+x1 <- seq(0, 1, length.out = 100)
+x2 <- rep(letters[1:3], length.out = 100)
 X <- model.matrix(~ x1 + x2)
 mu <- exp(X %*% beta)
 
-xyplot(mu ~ x1, groups = x2,
-       type = c("g", "l"),
-       lwd = 2,
-       xlab = "Continous variable",
-       ylab = "Counts average",
-       auto.key = list(
-           column = 3,
-           points = FALSE,
-           lines = TRUE,
-           title = "Categorical variable",
-           cex.title = 1.1
-       ))
+daexplain <- ldply(lapply(phis, function(phi) {
+    va <- compute_variance(mu, phi, sumto = 1000)
+    data.frame(mu = mu, va = va, di = va / mu, x1 = x1, x2 = x2)
+}), .id = "phi")
+
+fl <- parse(text = gsub("=", "==", levels(daexplain$phi)))
+xy1 <- xyplot(mu ~ x1, groups = x2,
+              type = c("g", "l"),
+              lwd = 2,
+              xlab = "Continous variable",
+              ylab = "Counts average",
+              auto.key = list(
+                  column = 3,
+                  points = FALSE,
+                  lines = TRUE,
+                  title = "Categorical variable",
+                  cex.title = 1.1
+              ))
+xy2 <- xyplot(di ~ x1 | phi, groups = x2,
+              type = c("g", "l"),
+              lwd = 2,
+              # scales = "free",
+              layout = c(2, 2),
+              xlab = "Continous variable",
+              ylab = "Dispersion index",
+              data = daexplain,
+              strip = strip.custom(factor.levels = fl))
+
+print(xy1, split = c(1, 1, 2, 1), more = TRUE)
+print(xy2, split = c(2, 1, 2, 1), more = FALSE)
 
 #-------------------------------------------
 # Random generation for the COM-Poisson distribution

### 2017-09-18 article.Rnw
Improve simulation study section
 - Remove algorithm enviroment to show steps;
 - Include a figure to explain the choice of the parameters.

@@
 In this section we conducted a simulation study to assess the properties
-of the proposed model and estimators0. We considered average counts varying
-from 3 to 27 according to a regression model structure with a continuous
-variable and a categorical variable with three levels. With respect to
-the precision parameter $\phi$, we fixed it to four values $-1.6$,
+of the proposed model and estimators. We considered average counts
+varying from 3 to 27 according to a regression model structure with a
+continuous variable $\bm{x}_1$ between 0 and 1 and a categorical
+variable $\bm{x}_2$ with three levels. Then the systematic componente is
+given by $\bm{\mu} = \exp(\beta_0 + \beta_1 \bm{x}_1 + \beta_{21}
+\bm{x}_{21} + \beta_{22} \bm{x}_{22})$, where $\bm{x}_{21}$ and
+$\bm{x}_{22}$ are \textit{dummy} variable for $\bm{x}_2$. With respect
+to the precision parameter $\phi$, we fixed it to four values $-1.6$,
 $-1.0$, $0.0$ and $1.8$, in order to have strong overdispersed, moderate
-overdispersed, equidispersed and underdispersed counts respectively. For
-each value of precision parameter, we considered four different sample
-sizes, 50, 100, 300 and 1000, generating 1000 datasets in each case. The
-\autoref{alg:simulation} shows the details of the steps in simulation
-study. In \autoref{fig:bias-plot} we show the bias of the estimators for
-each scenario (combination between values of the dispersion parameter
-and samples sizes) together with a confidence interval calculated as
-average bias plus and minus $\Phi(0.975)$ times the standard error. The
-scales are standardized for each parameter by dividing the bias by the
-standard error obtained for the sample of size 50.
-
-\begin{algorithm}
-\caption{Steps in simulation study.}
-\label{alg:simulation}
-  \Begin{
-  $\bm{\beta} = \begin{bmatrix} \beta_0 & \beta_1 & \beta_{21} &
-    \beta_{22} \end{bmatrix}^\top = \begin{bmatrix} 2.0 &
-    0.5 & 0.8 & -0.8 \end{bmatrix}^\top$\;
-  \For{$n \in \{50, 100, 300, 1000\}$}{
-    set $\bm{x}_1$ as a sequence, with $n$ elements, between $0$ and
-    $1$\;
-    set $\bm{x}_2$ as a repetition, with $n$ elements, of three
-    categories\;
-    compute $\bm{\mu}$ using $\bm{\mu} = \exp(\beta_0 + \beta_1 \bm{x}_1
-    +  \beta_{21} \bm{x}_{21} + \beta_{22} \bm{x}_{22})$, where
-    $\bm{x}_{21}$ and $\bm{x}_{22}$ are \textit{dummy} variable for
-    $\bm{x}_2$\;
-    \For{$\phi \in \{-1.6, -1.0, 0.0, 1.8\}$}{
-      \Repeat{$1000$ times}{
-        simulate $\bm{y}$ from COM-Poisson distribution with $\bm{\mu}$
-        and $\phi$ parameters\;
-        fit COM-Poisson$_\mu$ regression model to $\bm{y}$ data with
-        $\bm{X} = \begin{bmatrix} \bm{1} & \bm{x}_1 & \bm{x}_{21} &
-          \bm{x}_{22} \end{bmatrix}$ design matrix\;
-        get $\hat{\bm{\theta}} = \begin{bmatrix} \hat{\phi} & \hat{\beta}_0 &
-          \hat{\beta}_1 & \hat{\beta}_{21} & \hat{\beta}_{22}
-        \end{bmatrix}^\top$\;
-        get confidence intervals for $\hat{\bm{\theta}}$ by quadratic
-        approximation at the maximum likelihood estimate
-        (assumes $\hat{\bm{\theta}} \sim \mathcal{N}(\hat{\bm{\theta}},
-        \sqrt{-\bm{v}})$, where $\bm{v}$ is a diagonal of the inverse of
-        the Hessian matrix)\;
-      }
-    }
-  }
-}
-\end{algorithm}
+overdispersed, equidispersed and underdispersed counts
+respectively. \autoref{fig:justpars} shows the variation of the average
+counts (left) and variation of dispersion index (E$(Y)$ / var$(Y)$) for
+each fixed value of the precision parameter $\phi$. This configuration
+allows us to assess the proprierties of the estimators in extreme
+situation such as This configuration allows us to evaluate the
+properties of the estimators in extreme situations such as high counts
+and low dispersion, and low counts and high dispersion.
+
+In order to assess the consistency of the estimators we considered four
+different sample sizes, 50, 100, 300 and 1000, generating 1000 datasets
+in each case. In \autoref{fig:bias-plot} we show the bias of the
+estimators for each scenario (combination between values of the
+precision parameter and samples sizes) together with a confidence
+interval calculated as average bias plus and minus $\Phi(0.975)$ times
+the standard error. The scales are standardized for each parameter by
+dividing the bias by the standard error obtained for the sample of size
+50.
+
+%-----------------------------------------------------------------------
+% Explain the simulation
+% \begin{algorithm}
+% \caption{Steps in simulation study.}
+% \label{alg:simulation}
+%   \Begin{
+%   $\bm{\beta} = \begin{bmatrix} \beta_0 & \beta_1 & \beta_{21} &
+%     \beta_{22} \end{bmatrix}^\top = \begin{bmatrix} 2.0 &
+%     0.5 & 0.8 & -0.8 \end{bmatrix}^\top$\;
+%   \For{$n \in \{50, 100, 300, 1000\}$}{
+%     set $\bm{x}_1$ as a sequence, with $n$ elements, between $0$ and
+%     $1$\;
+%     set $\bm{x}_2$ as a repetition, with $n$ elements, of three
+%     categories\;
+%     compute $\bm{\mu}$ using $\bm{\mu} = \exp(\beta_0 + \beta_1 \bm{x}_1
+%     +  \beta_{21} \bm{x}_{21} + \beta_{22} \bm{x}_{22})$, where
+%     $\bm{x}_{21}$ and $\bm{x}_{22}$ are \textit{dummy} variable for
+%     $\bm{x}_2$\;
+%     \For{$\phi \in \{-1.6, -1.0, 0.0, 1.8\}$}{
+%       \Repeat{$1000$ times}{
+%         simulate $\bm{y}$ from COM-Poisson distribution with $\bm{\mu}$
+%         and $\phi$ parameters\;
+%         fit COM-Poisson$_\mu$ regression model to $\bm{y}$ data with
+%         $\bm{X} = \begin{bmatrix} \bm{1} & \bm{x}_1 & \bm{x}_{21} &
+%           \bm{x}_{22} \end{bmatrix}$ design matrix\;
+%         get $\hat{\bm{\theta}} = \begin{bmatrix} \hat{\phi} & \hat{\beta}_0 &
+%           \hat{\beta}_1 & \hat{\beta}_{21} & \hat{\beta}_{22}
+%         \end{bmatrix}^\top$\;
+%         get confidence intervals for $\hat{\bm{\theta}}$ by quadratic
+%         approximation at the maximum likelihood estimate
+%         (assumes $\hat{\bm{\theta}} \sim \mathcal{N}(\hat{\bm{\theta}},
+%         \sqrt{-\bm{v}})$, where $\bm{v}$ is a diagonal of the inverse of
+%         the Hessian matrix)\;
+%       }
+%     }
+%   }
+% }
+% \end{algorithm}
 
@@
 results <- readRDS("./codes/simulation.rds")
+
+@
+
+<<justpars, fig.height=4, fig.width=9, fig.cap="Average counts (left) and dispersion indexes (right) for each scenario considered in the simulation study.">>=
+
+#-------------------------------------------
+# Justify the choices of the parameters simulation
+x1 <- seq(0, 1, length.out = 100)
+x2 <- rep(letters[1:3], length.out = 100)
+X <- model.matrix(~ x1 + x2)
+mu <- exp(X %*% beta)
+
+daexplain <- ldply(lapply(phis, function(phi) {
+    va <- compute_variance(mu, phi, sumto = 300)
+    data.frame(mu = mu, va = va, di = va / mu, x1 = x1, x2 = x2)
+}), .id = "phi")
+
+fl <- parse(text = gsub("=", "==", levels(daexplain$phi)))
+xy1 <- xyplot(mu ~ x1, groups = x2,
+              type = c("g", "l"),
+              lwd = 2,
+              xlab = "Continous variable",
+              ylab = "Average count",
+              auto.key = list(
+                  column = 3,
+                  points = FALSE,
+                  lines = TRUE,
+                  title = "Categorical variable",
+                  cex.title = 1.1
+              ))
+xy2 <- xyplot(di ~ x1 | phi, groups = x2,
+              type = c("g", "l"),
+              lwd = 2,
+              # scales = "free",
+              layout = c(2, 2),
+              xlab = "Continous variable",
+              ylab = "Dispersion index",
+              data = daexplain,
+              strip = strip.custom(factor.levels = fl))
+
+print(xy1, split = c(1, 1, 2, 1), more = TRUE)
+print(xy2, split = c(2, 1, 2, 1), more = FALSE)
 
@@
 unbiased, in small samples the parameter is overestimated and the
-empirical distribution is right-skewed. \autoref{fig:coverage-plot}
-presents the confidence interval coverage rate based on quadratic
-approximation at the maximum likelihood estimate by sample size and
-dispersed levels.
+empirical distribution is slightly right-skewed.
+\autoref{fig:coverage-plot} presents the confidence interval coverage
+rate based on quadratic approximation at the maximum likelihood estimate
+(assumes $\hat{\bm{\theta}} \sim \mathcal{N}(\hat{\bm{\theta}},
+\sqrt{-\bm{v}})$, where $\bm{v}$ is a diagonal of the inverse of the
+Hessian matrix) by sample size and dispersed levels.
 

### 2017-09-18 article.Rnw
Spell check

@@
     \hspace*{.8mm}\thanks{Department of Exact Sciences, ESALQ-USP,
-    Piracicaba, São Paulo, Brazil} \and
+    Piracicaba, S\~{a}o Paulo, Brazil} \and
   Walmes M. Zeviani\thanks{Department of Statistics, UFPR, Curitiba,
-    Paraná, Brazil} \and
+    Paran\'{a}, Brazil} \and
   Wagner H. Bonat\footnotemark[3] \and
@@
   Bernoulli distribution as a limiting case. It can deal with both under
-  and overdispersed count data. The main disadvantage of COM-Poisson
-  distribution is that the location parameter does not represent the
-  average making difficult interpretation in a regression model. In this
-  paper we present a very simple reparametrization based on mean
-  approximation. The estimation and inference is based on likelihood and
-  simulation study showed that the maximum likelihood estimators are
-  unbiased and consistent for both regression and precision
-  parameters. Furthermore we observe empirical correlation between
-  regression and precision parameters close to 0 suggest the
-  orthogonality for the proposed reparametrization. We presents
-  applications of the proposed model by three case studies with over,
-  under and equidispersed counts. Results show improvements over the
-  conventional approaches. Therefore, its use in the analysis of count
-  data is encouraged. The computational routines for fitting of original
-  and reparametrized COM-Poisson models are available in the
+  and overdispersed count data. Therefore it is a suitable distribution
+  to analysis of dispersed count data. In spite of the nice properties
+  of COM-Poisson, the location parameter of the distribution does not
+  represent the average making difficult interpretation in a regression
+  structure. In this paper we present a very simple reparametrization
+  based on mean approximation. The estimation and inference is based on
+  likelihood and simulation study showed that the maximum likelihood
+  estimators are unbiased and consistent for both regression and
+  precision parameters. Furthermore, we observe empirical correlation
+  between regression and precision parameters close to zero, suggest the
+  orthogonality feature in the proposed reparametrization model. We
+  presents applications of the proposed model by three case studies with
+  over, under and equidispersed counts. Results show improvements over
+  the conventional approaches. Therefore, its use in the analysis of
+  count data is encouraged. The computational routines for fitting of
+  original and reparametrized COM-Poisson models are available in the
   supplementary material.
@@
 underdispersed, this motivated the class of duration dependence models
-\citep{Winkelmann1995}. Other possibity to explain underdispersion are
+\citep{Winkelmann1995}. Other possibility to explain underdispersion are
 order statistics. Suppose $Y_i$ is Poisson distributed then maximum of
@@
 performed using the software R \citep{Rcore2017}. The source code and
-data sets are avaliable in supplementary material\footnote{Available on
+data sets are available in supplementary material\footnote{Available on
   \texttt{\url{http://www.leg.ufpr.br/~eduardojr/papercompanions}}
@@
 continuous variable $\bm{x}_1$ between 0 and 1 and a categorical
-variable $\bm{x}_2$ with three levels. Then the systematic componente is
+variable $\bm{x}_2$ with three levels. Then the systematic component is
 given by $\bm{\mu} = \exp(\beta_0 + \beta_1 \bm{x}_1 + \beta_{21}
@@
 each fixed value of the precision parameter $\phi$. This configuration
-allows us to assess the proprierties of the estimators in extreme
+allows us to assess the properties of the estimators in extreme
 situation such as This configuration allows us to evaluate the
@@
 studies with over, under and equidispersed counts. In the data analysis
-we compared standard Poisson, original COM-Poisson, repametrized
+we compared standard Poisson, original COM-Poisson, reparametrized
 COM-Poisson and quasi-Poisson models.
@@
 coverage rates of the confidence intervals for the parameters, compute
-by the asymptotical distrition of the estimators, are close to the
-nominal level for sizes greather than 100. The worst scenario is when we
+by the asymptotic distribution of the estimators, are close to the
+nominal level for sizes greater than 100. The worst scenario is when we
 have small sample size and strong overdispersed counts. In general, we
@@
 to this feature the computational procedure is faster under the
-reparameterized model. Another advantage, perhaps the main, of the
+reparametrized model. Another advantage, perhaps the main, of the
 reparametrization is that the estimated regression parameters have
@@
 
-In general, the results presented by the reparameterized COM-Poisson
+In general, the results presented by the reparametrized COM-Poisson
 models were satisfactory and superior to the conventional
@@
 
-There are many possible extensions and possible topics for futher
+There are many possible extensions and possible topics for further
 investigation to model discussed in the present paper, including
@@
 sum's upper bound, propose a double GLM based on the reparametrized
-COM-Poisson model and a accomodate random effects since we have a model
+COM-Poisson model and a accommodate random effects since we have a model
 parametrized on mean.

### 2017-10-11 article.Rnw
Corrections and suggestions indicated by Walmes

@@
   and overdispersed count data. Therefore it is a suitable distribution
-  to analysis of dispersed count data. In spite of the nice properties
-  of COM-Poisson, the location parameter of the distribution does not
-  represent the average making difficult interpretation in a regression
-  structure. In this paper we present a very simple reparametrization
-  based on mean approximation. The estimation and inference is based on
-  likelihood and simulation study showed that the maximum likelihood
-  estimators are unbiased and consistent for both regression and
-  precision parameters. Furthermore, we observe empirical correlation
-  between regression and precision parameters close to zero, suggest the
-  orthogonality feature in the proposed reparametrization model. We
-  presents applications of the proposed model by three case studies with
-  over, under and equidispersed counts. Results show improvements over
-  the conventional approaches. Therefore, its use in the analysis of
-  count data is encouraged. The computational routines for fitting of
-  original and reparametrized COM-Poisson models are available in the
-  supplementary material.
+  to the analysis of dispersed count data. In spite of the nice
+  properties of COM-Poisson, the location parameter of the distribution
+  does not represent the conditional mean making difficult
+  interpretation in a regression structure. In this paper we present a
+  very simple reparametrization based on mean approximation. The
+  estimation and inference is based on likelihood and simulation study
+  showed that the maximum likelihood estimators are unbiased and
+  consistent for both regression and precision parameters. Furthermore,
+  we observed empirical correlation between regression and precision
+  parameters close to zero suggesting the orthogonality feature in the
+  proposed mean reparametrized model. We present applications of the
+  proposed model through three case studies with over, under and
+  equidispersed counts. Results show improvements over the conventional
+  approaches. Therefore, its use in the analysis of count data is
+  encouraged. The computational routines for fitting of original and
+  reparametrized COM-Poisson models are available in the supplementary
+  material.
 
@@
 Count data are variables that assume non-negative integer values and
-representing the number of times an event occurs in the observation
-unit. This kind of data are common in experimental research such as
-number of grains produced by a plant, number of fruits produced by a
-tree, number of insects on a particular cell, among others. Since the
+represent the number of times an event occurs in the observation
+unit. This kind of data is common in experimental research such as
+the number of grains produced by a plant, number of fruits produced by a
+tree, number of insects captured by a trap, among others. Since the
 seminal paper by \citet{Nelder1972} introduced the class of generalized
 linear models (GLM), analysis of count data often consider the Poisson
-regression model. These models provide a suitable distribution for a
-counting random variable and an efficient Newton scoring algorithm are
-used to fit them. Furthermore, the theoretical background for GLMs is
+regression model. This model provides a suitable distribution for a
+counting random variable and an efficient Newton scoring algorithm is
+used to fit it. Furthermore, the theoretical background for GLMs is
 well established in the class of dispersion models \citep{Jorgensen1997}
@@
 distribution has only one parameter, which represents the mean, and
-also, the variance. This relationship is referred as equidispersion.
-However, this assumption can be unsuitable to many situations and when
-the Poisson model is applied in these cases the standard errors for the
-parameters estimates are inconsistent \citep{Winkelmann1995}.
+also, the variance of the count random variable. This relationship is
+referred as equidispersion.  However, this assumption can be unsuitable
+to many situations and when the Poisson model is applied in these cases
+the standard errors for the parameters estimates are inconsistent
+\citep{Winkelmann1995}.
 
@@
 relevant covariates, heterogeneity of sampling units, different domain
-not considered and excess of zero \citep{Hinde1998}. Lesser reported,
-but that has received attention from the statistical community, is a
-underdispersion case, when the variance is less than the mean. The
-process that reduce the variability are not as well known that lead to
-extra variability. For the same reason, there a few approaches to deal
-underdispersed cases. The explanatory mechanisms to underdispersion may
-be related to Poisson process. When the time between events is not
-exponential distributed, the number of events can be over or
-underdispersed, this motivated the class of duration dependence models
-\citep{Winkelmann1995}. Other possibility to explain underdispersion are
-order statistics. Suppose $Y_i$ is Poisson distributed then maximum of
-$Y_i$ is underdispersed.
+size not considered and excess of zero \citep{Hinde1998}. Lesser
+reported, but that has received attention from the statistical
+community, is a underdispersion case, when the variance is less than the
+mean. The processes that reduce the variability are not as well known as
+those lead to extra variability. For the same reason, there are a few
+approaches to deal underdispersed cases. The explanatory mechanisms to
+underdispersion may be related to Poisson process. When the time between
+events is not exponential distributed, the number of events can be over
+or underdispersed, this motivated the class of duration dependence
+models \citep{Winkelmann1995}. Other possibility to explain
+underdispersion are order statistics. Suppose $Y_i$ is Poisson
+distributed then maximum of $Y_i$ is underdispersed.
 
@@
 overdispersion case Poisson mixed models are widely applied. One popular
-example of this approach is the negative-binomial model, where the
-natural is assumed to follow a gamma distribution. However, other
-distributions for the random effects may be considered, for example the
-model Poisson-Tweedie \citep{Bonat2017} and
+example of this approach is the negative-binomial model, where the mean
+is added a random noise that is assumed to follow a gamma
+distribution. However, other distributions for the random effects may be
+considered, for example the model Poisson-Tweedie \citep{Bonat2017} and
 Poisson-Inverse-Gaussian. For underdispersion, the Gamma-Count
-distribution assumes a gamma distribution for the time between events in
-Poisson process, thus it can deal with dispersed counts
-\citep{Zeviani2014}. The COM-Poisson distribution is given by a
-generalization the Poisson distribution allowed for a non-linear
-decrease in ratios of successive probabilities \citep{Shmueli2005}.
+distribution assumes a gamma distribution for the time between events,
+thus it can deal with dispersed counts \citep{Zeviani2014}. The
+COM-Poisson distribution is given by a generalization the Poisson
+distribution allowed for a non-linear decrease in ratios of successive
+probabilities \citep{Shmueli2005}.
 
@@
 COM-Poisson distribution is that the location parameter does not
-represent the average making difficult interpretation in a regression
-structure. \citep{Huang2017} proposed a mean-parametrization COM-Poisson
-distribution to avoid this issue, in this approach the mean parameter is
-obtained by solving the root of an equation defined as a infinite sum.
+represent the regression conditional mean making difficult
+interpretation in a regression structure. \citet{Huang2017} proposed a
+mean-parametrization COM-Poisson distribution to avoid this issue, in
+this approach the mean parameter is obtained by solving the root of an
+equation defined as a infinite sum.
 
@@
 \citet{Shmueli2005}. In this parametrization the probability mass
-function is wrote with $\mu$ and $\phi$ as function of original
+function is written with $\mu$ and $\phi$ as function of original
 parameters, where $\mu$ is the mean and $\phi$ is the precision or
@@
 by \cite{Huang2017} because in this case it is not necessary find $\mu$
-as the root of an equation involving an infinite sum. In simulation
-study we show the properties about maximum likelihood estimators in
-proposed and
+as the root of an equation involving an infinite sum. We conduce a
+simulation study to performe the properties of the maximum likelihood
+estimators for the proposed mean-parametrized COM-Poisson regression
+model.
 
@@
 the COM-Poisson distribution and \cite{Huang2017} parametrization. The
-proposed reparametrization and assessment of moments approximations is
+proposed reparametrization and assessement of moments approximations is
 considered in the \autoref{reparametrization}. In the
@@
 terms of the ratio between the probabilities of two consecutive values,
-with addition of a precision parameter \cite{Sellers2010}. Let $Y$ a
+with addition of a precision parameter \cite{Sellers2010}. Let $Y$ be a
 COM-Poisson random variable, then
@@
 Poisson distribution this ratio is $\frac{y}{\lambda}$. This allows the
-distribution for overdispersion or underdispersion. The probability mass
+distribution deals with overdispersion or underdispersion cases. The probability mass
 function for the COM-Poisson random variables takes the form
@@
   \Pr(Y=y \mid \lambda, \nu) = \frac{\lambda^y}{(y!)^\nu Z(\lambda, \nu)},
-  \qquad y = 0, 1, 2, \ldots
-\end{equation},
+  \qquad y = 0, 1, 2, \ldots,
+\end{equation}
 where $\lambda > 0$, $\nu \geq 0$ and $Z(\lambda, \nu) =
 \sum_{j=0}^\infty \frac{\lambda^j}{(j!)^\nu}$ is a normalizing
-constant.
+constant that depends on both parameters.
 
@@
 normalizing constants using a thousand increments, that is,
-$\sum_{j=0}^{1000}\lambda^j/(j!)^\nu$ for different of $\lambda$ and
-$\phi$ \footnote{For $\lambda=1$ e $\nu=0$ we show Inf, although the
+$\sum_{j=0}^{1000}\lambda^j/(j!)^\nu$ for different values of $\lambda$
+and $\phi$ \footnote{For $\lambda=1$ e $\nu=0$ we show Inf, although the
   numerically value is $1000$, since the series clearly diverges.}.
@@
               " \\multicolumn{7}{c}{$\\bm{\\lambda}$} \\\\\n",
-              "$\\bm{\\nu}$ & 0,5 & 1 & 5 & 10 & 30 & 50 \\\\\n")
+              "$\\bm{\\nu}$ & 0.5 & 1 & 5 & 10 & 30 & 50 \\\\\n")
       ))
@@
 \citet{Sellers2010} using an asymptotic approximation for
-$Z(\lambda,\nu)$, shows that an approximate forms can be closed by
+$Z(\lambda,\nu)$, showed that an approximate form can be
 \begin{equation}
@@
 relationship can be approximate by $\frac{1}{\nu}E(Y)$. In
-\autoref{reparametrization} we assess the approximations.
+\autoref{reparametrization} we assess the accucary of these
+approximations.
 
@@
 indirectly. \citet{Huang2017} shows how the means of COM-Poisson
-distributions can be modelled directly. in the \autoref{eqn:pmf-cmp},
-\Citeauthor{Huang2017} proposes that the parameter $\lambda$ is a
+distributions can be modelled directly. In the \autoref{eqn:pmf-cmp},
+\Citeauthor{Huang2017} proposes that the parameter $\lambda$ as a
 function of $\mu$ and $\nu$, given by the solution to
@@
 Thus the mean-parametrized COM-Poisson regression is $\eta(E(Y_i)) =
-\log(\mu_i)$. In this article, we presented a alternative
+\log(\mu_i)$. In this article, we presented an alternative
 mean-parametrization of COM-Poisson models to avoid the limitations of
@@
 \end{equation}
-Precision parameter is taken on the log scale to avoid restrictions on the
-parameter space, $\phi = \log(\nu), \phi \in \mathbb{R}$. The
+The precision parameter is taken on the log scale to avoid restrictions
+on the parameter space, $\phi = \log(\nu)$, $\phi \in \mathbb{R}$. The
 interpretation of $\phi$ are the same to $\nu$, but in other scale. For
-$\phi<0$ we have overdispersed case and $\phi>0$ overdispersed, relatives
-to Poisson distribution. $\phi=0$ represents the special case that the
-COM-Poisson comes down to Poisson.
+$\phi<0$ we have the overdispersed case and $\phi>0$ underdispersed,
+relatives to Poisson distribution. $\phi=0$ represents the special case
+that the COM-Poisson recovers to Poisson.
 
@@
 aux <- expand.grid(
-    mu = seq(2, 30, length.out = 50),
-    phi = seq(log(0.3), log(2.5), length.out = 50))
+    mu = seq(3, 30, length.out = 50),
+    phi = seq(-1.5, 1.8, length.out = 50))
 
@@
 
-In order to assess the accuracy of the moments approximations (
-\autoref{eqn:mean-approx}) we presented the quadratic errors of moments
+In order to assess the accuracy of the moments approximations
+(\autoref{eqn:mean-approx}) we presented the quadratic errors of moments
 approximations in \autoref{fig:approx-plot}(a) and
-\ref{fig:approx-plot}(b). Quadratic errors are obtained by $[\mu -
-E(Y)]^2$ for expectation and by $\left[ \frac{\mu}{\exp(\phi)} V(Y)
-  \right ]^2$, where
-$E(Y)$ and $V(Y)$ are numerically calculated using the moments
-definition and dotted lines represents the regions $\nu \leq 1$ and
-$\lambda > 10^\nu$ in the $(\mu\,,\phi)$ scale.
+\ref{fig:approx-plot}(b). Quadratic errors are obtained by
+$[\mu - E(Y)]^2$ for expectation and by
+$[ \mu \exp(-\phi) - \text{var}(Y)]^2$, where $E(Y)$ and var$(Y)$ are
+numerically calculated using the moments definitio. The dotted lines
+represent the border between the regions $\nu \leq 1$ and
+$\lambda > 10^\nu$, in the $(\mu\,,\phi)$ scale.
 
 <<approx-plot, fig.height=3, fig.width=8, fig.cap="Quadratic errors for expectation (a) and variance (b) approximation, dotted lines representing the restriction for good approximations by \\cite{Shmueli2005}.
-Mean-variance relationship (c), dotted line represents $E(Y)=V(Y)$.">>=
+Mean-variance relationship (c), dotted line represents $E(Y)=\text{var}(Y)$.">>=
 
@@
               axis = axis.grid,
-              xlab = expression(E(X)),
-              ylab = expression(V(X)),
+              xlab = expression(E(Y)),
+              ylab = expression(var(Y)),
               sub = "(c)",
@@
 
-The results in \autoref{fig:approx-plot} shows that the mean
-approximation is accurate, see maximum quadratic errors is
-\Sexpr{max(grid[["emu"]])}. For the variance approximation the maximum
-is \Sexpr{max(grid[["eva"]])} and this occurs to negative values of
-$\phi$.  Interestingly, the errors are bigger to negative values of
-$\phi$ and appear to have no relation with $\mu$ against the regions
-gives by \citet{Shmueli2005} ($\phi \leq 0$ and
-$\mu > 10 - \frac{\exp(\phi) - 1}{2\exp(\phi)}$). The relationship
+The results in \autoref{fig:approx-plot} show that the mean
+approximation is accurate, the maximum of the quadratic errors is
+\Sexpr{max(grid[["emu"]])} in the range evaluated. For the variance
+approximation, the maximum is \Sexpr{max(grid[["eva"]])} and this occurs
+to negative values of $\phi$. Interestingly, the errors are bigger to
+negative values of $\phi$ and appear to have no relation with $\mu$
+against the regions gives by \citet{Shmueli2005} ($\phi \leq 0$ and
+$\mu > 10 - \frac{\exp(\phi) - 1}{2\exp(\phi)}$).
+\autoref{fig:approx-plot}(c) shows that the relationship
 between mean and variance is linear with slope controlled by precision
-parameter $\phi$ (\autoref{fig:approx-plot}(c)).
+parameter $\phi$, but not in the form $\mu \exp(-\phi)$. In terms of
+moments, this leads to a specification indistinguishable from the
+quasi-Poisson.
 
@@
 proposed reparametrization. Replacing $\lambda$ and $\nu$ as function of
-$\mu$ and $\phi$ in \autoref{eqn:pmf-cmp} the reparametrized
+$\mu$ and $\phi$ in \autoref{eqn:pmf-cmp}, the reparametrized
 distribution takes the form
@@
   \frac{(y!)^{-e^\phi}}{Z(\mu, \phi)},
-  \qquad y = 0, 1, 2, \ldots\,
+  \qquad y = 0, 1, 2, \ldots\,,
 \end{equation} where $\mu > 0$. We denotes this distribution as
@@
     \sum_{i=1}^n \log(y_i!) \right ] -
-  \sum_{i=1}^n \log(Z_i(\mu_i, \phi)),
+  \sum_{i=1}^n \log(Z(\mu_i, \phi)),
 \end{equation}
@@
 known covariates for the $i$-th observation, and $(\bm{\beta},\, \phi)
-\in \mathbb{R}^{p+1}$. The normalizing constant $Z_i(\mu_i, \phi)$ is
+\in \mathbb{R}^{p+1}$. The normalizing constant $Z(\mu_i, \phi)$ is
 given by
 \begin{equation*}
-  Z_i(\mu_i, \phi) = \sum_{j=0}^\infty \left [ \left (
+  Z(\mu_i, \phi) = \sum_{j=0}^\infty \left [ \left (
     \mu_i + \frac{e^\phi - 1}{2e^\phi} \right )^{je^\phi}
@@
 \end{equation*}
-To compute the log-likelihood function we need the calculation of an
-infinite series for each observation, which makes the computation
-expensive for parametric space regions that the convergence of sum is
-slow.
+To compute the log-likelihood function we need the computation of an
+infinite serie for each observation, which makes the computation
+expensive for regions in the parameter space where the convergence of
+sum $Z$ is slow.
 
@@
 standard COM-Poisson distribution, the precision parameter is taken on
-the log scale, to avoid restriction of the parametric space.
+the log scale, to avoid restriction of the parameter space.
 
@@
 $\bm{x}_{22}$ are \textit{dummy} variable for $\bm{x}_2$. With respect
-to the precision parameter $\phi$, we fixed it to four values $-1.6$,
-$-1.0$, $0.0$ and $1.8$, in order to have strong overdispersed, moderate
+to the precision parameter $\phi$, we fixed it to four values: $-1.6$,
+$-1.0$, $0.0$ and $1.8$; in orderto have strong overdispersed, moderate
 overdispersed, equidispersed and underdispersed counts
 respectively. \autoref{fig:justpars} shows the variation of the average
-counts (left) and variation of dispersion index (E$(Y)$ / var$(Y)$) for
+counts (left) and variation of dispersion index (E$(Y)$/var$(Y)$) for
 each fixed value of the precision parameter $\phi$. This configuration
-allows us to assess the properties of the estimators in extreme
-situation such as This configuration allows us to evaluate the
-properties of the estimators in extreme situations such as high counts
-and low dispersion, and low counts and high dispersion.
+allows us to evaluate the properties of the estimators in extreme
+situations such as high counts and low dispersion, and low counts and
+high dispersion.
 
 In order to assess the consistency of the estimators we considered four
-different sample sizes, 50, 100, 300 and 1000, generating 1000 datasets
+different sample sizes: 50, 100, 300 and 1000; generating 1000 datasets
 in each case. In \autoref{fig:bias-plot} we show the bias of the
@@
 
+labx2 <- expression("Level of categorical variable"~(x[2]))
+labx1 <- expression("Values of continous variable"~(x[1]))
+
 fl <- parse(text = gsub("=", "==", levels(daexplain$phi)))
@@
               lwd = 2,
-              xlab = "Continous variable",
+              xlab = labx1,
               ylab = "Average count",
@@
                   lines = TRUE,
-                  title = "Categorical variable",
+                  title = labx2,
                   cex.title = 1.1
@@
               layout = c(2, 2),
-              xlab = "Continous variable",
+              xlab = labx1,
               ylab = "Dispersion index",
@@
 (assumes $\hat{\bm{\theta}} \sim \mathcal{N}(\hat{\bm{\theta}},
-\sqrt{-\bm{v}})$, where $\bm{v}$ is a diagonal of the inverse of the
-Hessian matrix) by sample size and dispersed levels.
+\sqrt{\bm{v}})$, where $\bm{v}$ is a diagonal of the inverse of the
+observed information matrix) by sample size and dispersed levels.
 
@@
        par.settings = list(
-           superpose.symbol = list(pch = 19),
            layout.heights = list(strip = 1.5)
        ),
-       auto.key = list(
+       key = list(
            column = 4,
-           lines = TRUE,
-           text = fl
+           type = "o",
+           divide = 1,
+           text = list(fl),
+           lines = list(pch = 19, col = cols[1:4])
        ),
@@
 In this section, we report three illustrative examples of real count
-data analysis obtained of the experimental research. We considered as
+data analysis obtained from experimental research. We considered as
 alternative models for the analysis the standard Poisson model, the
@@
 submitted to five levels of artificial defoliation (\texttt{des}) and
-observed on five growth stages (\texttt{est}). The purpose of this
+crossed with five growth stages (\texttt{est}). The purpose of this
 research was to study the effect of defoliation levels at different
-growth stages of cotton plants on the observed number of bolls
-produced. The study was conducted in a greenhouse and the experimental
-design was completely randomized with five replicates. This data set was
-analysed in \citet{Zeviani2014} using the \textit{Gamma-Count}
-distribution.
+growth stages of cotton plants on the cotton production, expressed by
+the number of bools produced. The study was conducted in a greenhouse
+and the experimental design was completely randomized with five
+replicates. This data set was analysed by \citet{Zeviani2014} using the
+\textit{Gamma-Count} distribution.
 
@@
 goodness-of-fit measures for the Poisson, COM-Poisson, COM-Poisson$_\mu$
-e quasi-Poisson models are presented in \autoref{tab:coef-cotton}.
+and quasi-Poisson models are presented in \autoref{tab:coef-cotton}.
 
@@
 \autoref{tab:coef-cotton} also shows the advantage of COM-Poisson$_\mu$
-(reparametrized), since the estimates are quite similar to obtained by
+(reparametrized), since the estimates are quite similar those obtained by
 the Poisson model, whereas estimates obtained by the COM-Poisson model
 in original parametrization are on a non-interpretable scale. The ratios
-between estimates and their respective standard deviations for the
+between estimates and their respective standard errors for the
 COM-Poisson models are very close to ratios obtained by quasi-Poisson
@@
 presents the empirical correlations between precision and regression
-parameters, obtained by variance and covariance matrix. The correlations
-are practically null considering the COM-Poisson$_\mu$, whereas to the
-original parametrization they are quite high. This reflects a better
-performance of the maximization algorithm in mean-parametrization model,
-in this case COM-Poisson$_\mu$ fit was \Sexpr{time12[3] * 100 /
-  time11[3]}\% slower then COM-Poisson model.
+parameters, obtained from the estimated variance and covariance
+matrix. The correlations are practically null considering the
+COM-Poisson$_\mu$, whereas to the original parametrization they are
+quite high. This reflects a better performance of the maximization
+algorithm in mean-parametrization model, in this case COM-Poisson$_\mu$
+fit was \Sexpr{time12[3] * 100 / time11[3]}\% slower then COM-Poisson
+model. It is important to note that the initial values for the BFGS
+algorithm are provided by Poisson fitted model, then in the COM-Poisson
+mean-parametrized model the initial values are practically the maximum
+likelihood estimates and the effort of maximization is on the precision
+parameter $\phi$.
 
@@
 
-The second example is a experiment completely randomized block with
-treatments in a $5\times 3$ factorial arrangement about soy bean
-culture. The aim of this study was to evaluate the effects of potassium
-doses (\texttt{K}) applied to soil (0, 0.3, 0.6, 1.2 e 1.8 100mg
-dm$^{-3}$) and soil humidity (\texttt{umid}) levels (37.5, 50, 62.5\%,
-which represent few, ideal and abundance water respectively) on the
-soy bean production. The experiment was carried out in a greenhouse, in
-pots with two plants, and the count variable measured was the number of
-grains per pot \citep{Serafim2012}. \autoref{fig:desc-soy} (left) shows
-the number of grains recorded for each combination of potassium dose and
-humidity level, it is important to note the indication of a quadratic
-level of the potassium levels. Most points in the sample means and
-variances dispersion diagram (right) are above the identity line,
-suggesting overdispersion.
+The second example is an experiment in a completely randomized block
+with treatments in a $5\times 3$ factorial arrangement about soybean
+(\emph{Glicine Max}). The aim of this study was to evaluate the effects
+of potassium doses (\texttt{K}) applied to soil (0, 0.3, 0.6, 1.2 and
+1.8 $\times$ 100mg dm$^{-3}$) and soil moisture (\texttt{umid}) levels
+(37.5, 50, 62.5\%) on the soybean production. The experiment was
+carried out in a greenhouse, in pots with two plants, and the count
+variable measured was the number of grains per pot
+\citep{Serafim2012}. \autoref{fig:desc-soy} (left) shows the number of
+grains recorded for each combination of potassium dose and moisture
+level, it is important to note the indication of a quadratic level of
+the potassium levels. Most points in the sample variance vs. sample
+means dispersion diagram (right) are above the identity line, suggesting
+overdispersion (block effect not yet removed).
 
@@
 
-<<desc-soy, fig.pos="H", fig.height=4, fig.width=8, fig.cap="Number of grains per pot for each potassium dose and humidity level (left) and sample mean agains sample variance of the five replicates for each experimental treatment (right).">>=
+<<desc-soy, fig.pos="H", fig.height=4, fig.width=8, fig.cap="Number of grains per pot for each potassium dose and moisture level (left) and sample mean agains sample variance of the five replicates for each experimental treatment (right).">>=
 
@@
               strip = strip.custom(
-                  strip.names = TRUE, var.name = "humidity",
+                  strip.names = TRUE, var.name = "moisture",
                   factor.levels = paste0(levels(soyBeans$umid), "%")))
@@
 dm$^{-3}$, where $\gamma_i$ is the effect of $i$-th block, $\tau_j$ is
-the effect of $j$-th humidity level and $\beta_{3j}$ is the first order
-potassium effect (\texttt{K}) for the $j$-th humidity level
+the effect of $j$-th moisture level and $\beta_{3j}$ is the first order
+potassium effect (\texttt{K}) for the $j$-th moisture level
 (\texttt{umid}). \autoref{tab:coef-soy} presents the estimates, ratio
-between estimate and standard deviation and goodness-of-fit measures for
+between estimate and standard error and goodness-of-fit measures for
 the alternative models fitted.
@@
 regression parameters, the similarities between models are analogous to
-the previous section. Both models agree on the indication of effects,
-however the Poisson model indicates the effects with greater
-significance, because it does not fit the extra variability.
+the previous section. Both models indicate effects of block, potassium
+dose and moisture level, however the Poisson model indicates the effects
+with greater significance, because it does not fit the extra
+variability.
 
@@
 \caption{Parameter estimates (Est) and ratio between estimate and
-  standard error (SE) for the four model strategies for the soy bean
+  standard error (SE) for the four model strategies for the soybean
   experiment.}
@@
 \Sexpr{time22[3]} seconds respectively. This may be due the
-well-behaviour of likelihood function for orthogonal parameters. In
-\autoref{tab:corr-soy} we present the empirical correlations between
-regression and precision parameter. The correlations are close to zero
-for the COM-Poisson$_\mu$ parametrization, indicating the empirical
-orthogonality between $\mu$ and $\phi$ on COM-Poisson$_\mu$
-distribution.
+well-behaviour of likelihood function for orthogonal parameters also,
+best start values from Poisson model. In \autoref{tab:corr-soy} we
+present the empirical correlations between regression and precision
+parameter. The correlations are close to zero for the COM-Poisson$_\mu$
+parametrization, indicating the empirical orthogonality between $\mu$
+and $\phi$ on COM-Poisson$_\mu$ distribution.
 
@@
 confidence bands are show in \autoref{fig:pred-soy}. The fitted values
-are identical both models, leading the same conclusions. On the other
-hand, confidence bands for the Poisson model are smaller due to
+are identical for both models, leading to the same conclusions. On the
+other hand, confidence bands for the Poisson model are smaller due to
 equidispersion assumption. Considering COM-Poisson models and
@@
 
-The data set come from an experiment to measure the reproductive
+The data set comes from an experiment to measure the reproductive
 toxicity of a herbicide, nitrofen, on a species of zooplankton
@@
 likelihood ratio tests comparing the sequence of predictors. All models
-indicate the cubic effect of nitrofen dosing. Considering this
+indicate the cubic effect of nitrofen concentration. Considering this
 predictor, there is an evidence of equidispersed counts, $\phi$
@@
 to analyse dispersed count data. We proposed a novel reparametrization
-based only mean approximation gives by \cite{Shmueli2005}. The
+based only mean approximation indicated by \cite{Shmueli2005}. The
 likelihood-based method to estimation and inference is presented. We
@@
 parameters suggest the orthogonality between $\mu$ and $\phi$ in
-COM-Poisson$\mu$ distribution against the original parametrization. Due
+COM-Poisson$_\mu$ distribution against the original parametrization. Due
 to this feature the computational procedure is faster under the

### 2017-10-31 article.Rnw
Modify preamble according to SMJ class

@@
-\documentclass[a4paper]{article}
-
-\usepackage{amssymb}
+%-------------------------------------------
+% Document class
+\documentclass[submit]{smj}
+
+%-------------------------------------------
+% Personal packages and definitions
 \usepackage{amsfonts}
-\usepackage{amsmath}
 \usepackage{bm}
@@
 
-\usepackage{hyperref}
-\hypersetup{colorlinks=true, linkcolor=blue, citecolor=blue,
-  urlcolor=blue}
-
-\usepackage[bottom]{footmisc}
-\usepackage{geometry}
-\geometry{lmargin=3cm, rmargin=3cm, tmargin=3cm, bmargin=3cm}
-
 % Algorithm
@@
 \SetAlCapSty{}
-% \SetAlCapFnt{\footnotesize}
-% \SetAlCapNameFnt{\footnotesize}
-
-\usepackage{natbib}
-% \bibliographystyle{humannat}
-\bibliographystyle{agsm}
-
-\usepackage{lipsum}
+
 \usepackage{float}
-\usepackage{mathpazo}
 
@@
 
-%-----------------------------------------------------------------------
-\title{Reparametrization of COM-Poisson Regression Models with
+% -------------------------------------------
+% Authors and affiliations
+\Author{Eduardo E. Ribeiro Jr\Affil{1},
+  Walmes M. Zeviani\Affil{2},
+  Wagner H. Bonat\Affil{2},
+  Clarice G. B. Dem\'{e}trio\Affil{1}
+  and John Hinde\Affil{3}
+}
+\AuthorRunning{Ribeiro Jr \textrm{et al.}}
+
+\Affiliations{
+\item Department of Exact Sciences,
+  University of S\~{a}o Paulo - ESALQ,
+  Piracicaba,
+  SP, Brazil
+\item Department of Statistics,
+  Paran\'{a} Federal University,
+  Curitiba,
+  PR, Brazil
+\item School of Mathematics, Statistics and Applied Mathematics,
+  National University of Ireland,
+  Galway,
+  Galway, Ireland
+}
+
+%-------------------------------------------
+% Corresponding author
+\CorrAddress{Eduardo E. Ribeiro Jr,
+             Department of Exact Sciences,
+             Luiz de Queiroz College of Agriculture - ESALQ,
+             University of S\~{a}o Paulo - USP,
+             Piracicaba, S\~{a}o Paulo,
+             P\'{a}dua Dias, 11, Avenue, CEP--13.418-900
+             Brazil}
+\CorrEmail{jreduardo@usp.br}
+\CorrPhone{(+55)\;41\;9\;8711\;9034}
+\CorrFax{(+55)\;19\;3447\;6021}
+
+%-------------------------------------------
+% Title and shor title
+\Title{Reparametrization of COM-Poisson Regression Models with
   Applications in the Analysis of Experimental Data}
-
-\author{
-  Eduardo E. R. Junior\thanks{Corresponding author:
-    \href{mailto:jreduardo@usp.br}{\nolinkurl{jreduardo@usp.br}}}
-    \hspace*{.8mm}\thanks{Department of Exact Sciences, ESALQ-USP,
-    Piracicaba, S\~{a}o Paulo, Brazil} \and
-  Walmes M. Zeviani\thanks{Department of Statistics, UFPR, Curitiba,
-    Paran\'{a}, Brazil} \and
-  Wagner H. Bonat\footnotemark[3] \and
-  Clarice G. B. Dem\'{e}trio\footnotemark[2]
-}
-
-\date{\today}
-
-<<setup, include=FALSE>>=
-
-library(knitr)
-library(xtable)
-options(digits = 3, OutDec = ".",
-        xtable.caption.placement = "top",
-        xtable.booktabs = TRUE,
-        xtable.sanitize.text.function = identity)
-opts_chunk$set(
-    warning = FALSE,
-    message = FALSE,
-    echo = FALSE,
-    results = "hide",
-    fig.width = 7,
-    fig.height = 5,
-    out.width = "1\\textwidth",
-    fig.align = "center",
-    fig.pos = "H",
-    # dev = "tikz"
-    dev.args = list(family = "Palatino")
-    )
-
-source("./codes/lattice-panels.R")
-source("./codes/functions.R")
-
-# Colors for legends
-cols <- trellis.par.get("superpose.line")$col
-
-# Useful packages
-library(bbmle)
-library(multcomp)
-library(plyr)
-library(tidyr)
-library(dplyr)
-
-@
-
-%-----------------------------------------------------------------------
-
-\begin{document}
-\maketitle
-
-\begin{abstract}
-
+\TitleRunning{Reparametrization of COM-Poisson Regression Models}
+
+%-------------------------------------------
+% Abstract
+\Abstract{
   In the analysis of count data seldom the equidispersion assumption is
@@
   material.
-
-\vspace{0.1cm}
-\noindent\textbf{keywords:} COM-Poisson, Over/Under-dispersion, Count
-data, Likelihood inference.
-\end{abstract}
-
-\pagebreak
+}
+
+%-------------------------------------------
+% Key-words
+\Keywords{
+COM-Poisson; Count data; Over/Under-dispersion; Likelihood inference
+}
+
+<<setup, include=FALSE>>=
+
+library(knitr)
+library(xtable)
+options(digits = 3, OutDec = ".",
+        xtable.caption.placement = "top",
+        xtable.booktabs = TRUE,
+        xtable.sanitize.text.function = identity)
+opts_chunk$set(
+    warning = FALSE,
+    message = FALSE,
+    echo = FALSE,
+    results = "hide",
+    fig.width = 7,
+    fig.height = 5,
+    out.width = "1\\textwidth",
+    fig.align = "center",
+    fig.pos = "H",
+    # dev = "tikz"
+    dev.args = list(family = "Palatino")
+    )
+
+source("./codes/lattice-panels.R")
+source("./codes/functions.R")
+
+# Colors for legends
+cols <- trellis.par.get("superpose.line")$col
+
+# Useful packages
+library(bbmle)
+library(multcomp)
+library(plyr)
+library(tidyr)
+library(dplyr)
+
+@
+
+%-----------------------------------------------------------------------
+
+\begin{document}
+\maketitle
 

### 2017-10-31 article.Rnw
Modify the size, align and position of tables

@@
         xtable.booktabs = TRUE,
-        xtable.sanitize.text.function = identity)
+        xtable.sanitize.text.function = identity,
+        xtable.size = "small",
+        xtable.math.style.negative = TRUE)
 opts_chunk$set(
@@
     fig.align = "center",
-    fig.pos = "H",
+    fig.pos = "!ht",
     # dev = "tikz"
@@
 \begin{document}
+
 \maketitle
@@
 
-\begin{table}[h]
-\centering
+\begin{table}[!ht]
+\centering \small
 \caption{Parameter estimates (Est) and ratio between estimate and
@@
 \label{tab:coef-cotton}
-\begin{tabularx}{\textwidth}{lCCCCCCCC}
+\begin{tabularx}{\textwidth}{lrrrrrrrr}
   \toprule
@@
     x <- formatC(x, getOption("digits"), format = "f")
-    x <- gsub("NA", "---", x)
-    paste(paste0("\\multicolumn{2}{c}{", x, "}"),
+    x <- gsub("NA", "-", x)
+    paste(paste0("\\multicolumn{2}{c}{$", x, "$}"),
           collapse = " & ")
@@
 
-<<desc-soy, fig.pos="H", fig.height=4, fig.width=8, fig.cap="Number of grains per pot for each potassium dose and moisture level (left) and sample mean agains sample variance of the five replicates for each experimental treatment (right).">>=
+<<desc-soy, fig.pos="!ht", fig.height=4, fig.width=8, fig.cap="Number of grains per pot for each potassium dose and moisture level (left) and sample mean agains sample variance of the five replicates for each experimental treatment (right).">>=
 
@@
 
-\begin{table}[ht]
-\centering
+\begin{table}[!ht]
+\centering \small
 \caption{Parameter estimates (Est) and ratio between estimate and
@@
 \label{tab:coef-soy}
-\begin{tabularx}{\textwidth}{lCCCCCCCC}
+\begin{tabularx}{\textwidth}{lrrrrrrrr}
   \toprule
@@
 
-<<pred-soy, fig.pos="H", fig.height=3.8, fig.width=8, fig.cap="Dispersion diagrams of grain counts as function of potassium doses and humidity levels with fitted curves and confidence intervals (95\\%).">>=
+<<pred-soy, fig.pos="!ht", fig.height=3.8, fig.width=8, fig.cap="Dispersion diagrams of grain counts as function of potassium doses and humidity levels with fitted curves and confidence intervals (95\\%).">>=
 
@@
 
-\begin{table}[ht]
-\centering
+\begin{table}[!ht]
+\centering \small
 \caption{Model fit measures and comparisons between predictors and models.}
 \label{tab:anova-ovos}
-\begin{tabularx}{\textwidth}{lCCCCCrC}
+\begin{tabularx}{\textwidth}{lcccccrc}
   \toprule
@@
 
-\begin{table}[H]
+\begin{table}[!ht]
 \centering
@@
 %-----------------------------------------------------------------------
-\newpage
-
-\small
 \bibliography{references.bib}
+
 \end{document}

### 2017-11-06 article.Rnw
Divide tables to fit on page and modify to right alignment

@@
 % Document class
-\documentclass[submit]{smj}
+\documentclass[publish]{smj}
 
@@
 \usepackage{amsfonts}
+\usepackage{float}
 \usepackage{bm}
@@
 
-\usepackage{float}
-
 \usepackage{booktabs}
@@
 \newcolumntype{C}{>{\centering\arraybackslash}X}
-
-\makeatletter
-\def\and{%
-  \end{tabular}%
-  \hskip 0em \@plus.17fil \protect$\bullet$
-  \begin{tabular}[t]{c}}%
-\makeatother
 
@@
 
+divide_matrix <- function(mat, divide) {
+    mat1 <- mat[, 1:divide]
+    mat2 <- mat[, (divide + 1):ncol(mat)]
+    nc1 <- ncol(mat1)
+    nc2 <- ncol(mat2)
+    cdiff <- abs(ncol(mat1) - ncol(mat2))
+    if (cdiff) {
+        complete <- matrix(NA, nrow = nrow(mat), ncol = cdiff)
+        colnames(complete) <- rep(" ", cdiff)
+        if (nc1 < nc2) {
+            mat1 <- cbind(mat1, complete)
+        } else {
+            mat2 <- cbind(mat2, complete)
+        }
+    }
+    return(list(mat1, mat2))
+}
+
 source("./codes/lattice-panels.R")
@@
              align = "C|CCCCCC",
-             # align = "ccccccc",
+             # align = "c|cccccc",
              label = "tab:convergenceZ"),
@@
 \label{tab:coef-cotton}
-\begin{tabularx}{\textwidth}{lrrrrrrrr}
+\begin{tabular}{lrrrrrrrr}
   \toprule
@@
 @
-\end{tabularx}
+\end{tabular}
 \end{table}
@@
 colnames(corr.ncap) <- gsub("beta", "hat{\\\\beta}", pnames[-1])
-
-caption <- paste("Empirical correlations between $\\hat{\\phi}$ and",
-                 "$\\hat{\\bm{\\beta}}$ for the two parametrizations",
-                 "of COM-Poisson model fit to underdispersed data.")
-print(xtable(corr.ncap,
-             align = c("lccccccccccc"),
-             caption = caption,
-             digits = 3,
-             label = "tab:corr-cotton"),
-      size = "small",
-      sanitize.rownames.function = identity,
+corrncap_aux <- divide_matrix(corr.ncap, divide = 6)
+
+# caption <- paste("Empirical correlations between $\\hat{\\phi}$ and",
+#                  "$\\hat{\\bm{\\beta}}$ for the two parametrizations",
+#                  "of COM-Poisson model fit to underdispersed data.")
+# print(xtable(corr.ncap,
+#              align = c("lrrrrrrrrrrr"),
+#              caption = caption,
+#              digits = 3,
+#              label = "tab:corr-cotton"),
+#       size = "small",
+#       sanitize.rownames.function = identity,
+#       sanitize.colnames.function = function(x) sprintf("$%s$", x))
+
+@
+
+\begin{table}[ht]
+\centering
+\caption{Empirical correlations between $\hat{\phi}$ and $\hat{\bm{\beta}}$ for the two parametrizations of COM-Poisson model fit to equidispersed data.}
+\label{tab:corr-cotton}
+\begingroup\small
+\begin{tabular}{lrrrrrr}
+  \toprule
+<<results="asis">>=
+
+print(xtable(corrncap_aux[[1]], digits = 4),
+      only.contents = TRUE,
+      hline.after = FALSE,
       sanitize.colnames.function = function(x) sprintf("$%s$", x))
-
-@
+cat("\\midrule", sep = "\n")
+
+print(xtable(corrncap_aux[[2]], digits = 4),
+      only.contents = TRUE,
+      sanitize.colnames.function = function(x) sprintf("$%s$", x))
+
+@
+\end{tabular}
+\endgroup
+\end{table}
 
@@
 \label{tab:coef-soy}
-\begin{tabularx}{\textwidth}{lrrrrrrrr}
+\begin{tabular}{lrrrrrrrr}
   \toprule
@@
                             pnames[-1])
-
-caption <- paste("Empirical correlations between $\\hat{\\phi}$ and",
-                 "$\\hat{\\bm{\\beta}}$ for the two parametrizations",
-                 "of COM-Poisson model fit to overdispersed data.")
-print(xtable(corr.ngra,
-             align = c("lccccccccccc"),
-             caption = caption,
-             digits = 3,
-             label = "tab:corr-soy"),
-      size = "small",
-      sanitize.rownames.function = identity,
+corrngra_aux <- divide_matrix(corr.ncap, divide = 7)
+
+# caption <- paste("Empirical correlations between $\\hat{\\phi}$ and",
+#                  "$\\hat{\\bm{\\beta}}$ for the two parametrizations",
+#                  "of COM-Poisson model fit to overdispersed data.")
+# print(xtable(corr.ngra,
+#              align = c("lccccccccccc"),
+#              caption = caption,
+#              digits = 3,
+#              label = "tab:corr-soy"),
+#       size = "small",
+#       sanitize.rownames.function = identity,
+#       sanitize.colnames.function = function(x) sprintf("$%s$", x))
+
+@
+
+\begin{table}[ht]
+\centering
+\caption{Empirical correlations between $\hat{\phi}$ and $\hat{\bm{\beta}}$ for the two parametrizations of COM-Poisson model fit to overdispersed data.}
+\label{tab:corr-soy}
+\begingroup\small
+\begin{tabular}{lrrrrrrr}
+  \toprule
+<<results="asis">>=
+
+print(xtable(corrngra_aux[[1]], digits = 4),
+      only.contents = TRUE,
+      hline.after = FALSE,
       sanitize.colnames.function = function(x) sprintf("$%s$", x))
-
-@
+cat("\\midrule", sep = "\n")
+
+print(xtable(corrngra_aux[[2]], digits = 4),
+      only.contents = TRUE,
+      sanitize.colnames.function = function(x) sprintf("$%s$", x))
+
+@
+\end{tabular}
+\endgroup
+\end{table}
 
@@
 \label{tab:anova-ovos}
-\begin{tabularx}{\textwidth}{lcccccrc}
+\begin{tabularx}{\textwidth}{lCCCCCrC}
   \toprule
@@
 \end{tabularx}
-\vspace{-1mm}
+%\vspace{-1mm}
 
@@
 \label{tab:coef-ovos}
-\begin{tabularx}{\textwidth}{lCCCCCCCC}
+\begin{tabular}{lrrrrrrrr}
   \toprule
@@
 print(xtable(corr.ovos,
-             align = c("lCCCC"),
+             align = c("lrrrr"),
              caption = caption,
-             digits = 3,
+             digits = 4,
              label = "tab:corr-ovos"),
@@
       sanitize.colnames.function = function(x) sprintf("$%s$", x),
-      tabular.environment = "tabularx",
-      width = "\\textwidth")
+      # width = "\\textwidth,"
+      tabular.environment = "tabular")
 

### 2017-11-13 article.Rnw
Improve abstract and introduction

@@
 
-%-------------------------------------------
+% -------------------------------------------
 % Abstract
-\Abstract{
-  In the analysis of count data seldom the equidispersion assumption is
-  suitable, hence the Poisson regression model is inappropriate in many
-  situations. The COM-Poisson is a member of the exponential family, has
-  the Poisson and geometric distributions as special cases and the
-  Bernoulli distribution as a limiting case. It can deal with both under
-  and overdispersed count data. Therefore it is a suitable distribution
-  to the analysis of dispersed count data. In spite of the nice
-  properties of COM-Poisson, the location parameter of the distribution
-  does not represent the conditional mean making difficult
-  interpretation in a regression structure. In this paper we present a
-  very simple reparametrization based on mean approximation. The
-  estimation and inference is based on likelihood and simulation study
-  showed that the maximum likelihood estimators are unbiased and
-  consistent for both regression and precision parameters. Furthermore,
-  we observed empirical correlation between regression and precision
-  parameters close to zero suggesting the orthogonality feature in the
-  proposed mean reparametrized model. We present applications of the
-  proposed model through three case studies with over, under and
-  equidispersed counts. Results show improvements over the conventional
-  approaches. Therefore, its use in the analysis of count data is
-  encouraged. The computational routines for fitting of original and
-  reparametrized COM-Poisson models are available in the supplementary
-  material.
+\Abstract{ In the analysis of count data often the equidispersion
+  assumption is not suitable, hence the Poisson regression model is
+  inappropriate.  As a generalization of the Poisson distribution the
+  COM-Poisson distribution can deal with under-, equi- and overdispersed
+  count data.  It is a member of the exponential family of distributions
+  and has the Poisson and geometric distributions as special cases as
+  well as the Bernoulli distribution as a limiting case.  In spite of
+  the nice properties of the COM-Poisson distribution, its location
+  parameter does not correspond to the expectation, which in turn
+  becomes challenging the interpretation of the regression models
+  specified using such a distribution. In this paper, we propose a
+  straightforward reparametrization of the COM-Poisson distribution
+  based on an approximation to the expectation of this distribution. The
+  main advantage of our new parametrization is the straightforward
+  interpretation of the regression coefficients in terms of the
+  expectation of the count response variable, as usual in the context of
+  generalized linear models. Furthermore, the estimation and inference
+  for the new COM-Poisson regression model can be done based on the
+  likelihood paradigm. We carried out simulation studies to verify the
+  finite sample properties of the maximum likelihood estimators. The
+  results from our simulation study show that the maximum likelihood
+  estimators are unbiased and consistent for both regression and
+  dispersion parameters. We observed that the empirical correlation
+  between the regression and dispersion parameter estimators is close to
+  zero, which suggests that these parameters are orthogonal. We
+  illustrate the application of the proposed model through the analysis
+  of three data sets with over-, under- and equidispersed count data.
+  The study of distribution properties through a consideration of
+  dispersion, zero-inflated and heavy tail indices, and the results of
+  data analysis show the flexibility over standard
+  approaches. Therefore, we encourage the application of the new
+  parametrization for the analysis of count data in the context of
+  COM-Poisson regression models.  The computational routines for fitting
+  the original and new version of the COM-Poisson regression model and
+  the analyzed data sets are available in the supplementary material.
 }
@@
 % Key-words
-\Keywords{
-COM-Poisson; Count data; Over/Under-dispersion; Likelihood inference
-}
+\Keywords{ COM-Poisson, Overdispersion, Underdispersion, Count data,
+  Likelihood inference.  }
 
@@
 
-Count data are variables that assume non-negative integer values and
-represent the number of times an event occurs in the observation
-unit. This kind of data is common in experimental research such as
-the number of grains produced by a plant, number of fruits produced by a
-tree, number of insects captured by a trap, among others. Since the
-seminal paper by \citet{Nelder1972} introduced the class of generalized
-linear models (GLM), analysis of count data often consider the Poisson
-regression model. This model provides a suitable distribution for a
-counting random variable and an efficient Newton scoring algorithm is
-used to fit it. Furthermore, the theoretical background for GLMs is
-well established in the class of dispersion models \citep{Jorgensen1997}
-as a generalization of the exponential family of distributions.
-
-In spite of the advantages of the GLM framework, the Poisson
-distribution has only one parameter, which represents the mean, and
-also, the variance of the count random variable. This relationship is
-referred as equidispersion.  However, this assumption can be unsuitable
-to many situations and when the Poisson model is applied in these cases
-the standard errors for the parameters estimates are inconsistent
-\citep{Winkelmann1995}.
-
-In practice, overdispersion, when the variance greater than the mean, is
-largely reported in the literature and may occur due to the absence of
-relevant covariates, heterogeneity of sampling units, different domain
-size not considered and excess of zero \citep{Hinde1998}. Lesser
-reported, but that has received attention from the statistical
-community, is a underdispersion case, when the variance is less than the
-mean. The processes that reduce the variability are not as well known as
-those lead to extra variability. For the same reason, there are a few
-approaches to deal underdispersed cases. The explanatory mechanisms to
-underdispersion may be related to Poisson process. When the time between
-events is not exponential distributed, the number of events can be over
-or underdispersed, this motivated the class of duration dependence
-models \citep{Winkelmann1995}. Other possibility to explain
-underdispersion are order statistics. Suppose $Y_i$ is Poisson
-distributed then maximum of $Y_i$ is underdispersed.
+Count data are random variables that assume non-negative integer values
+and represent the number of times an event occurs in the observation
+unit. This kind of data is common in crop sciences such as the number of
+grains produced by a plant, number of fruits produced by a tree, number
+of insects captured by a trap, to cite but a few.  Since the seminal
+paper of \citet{Nelder1972} where the class of the generalized linear
+models (GLMs) was introduced, the analysis of count data often employs
+the Poisson regression model. This model provides a suitable strategy
+for the analysis of count data and an efficient Newton scoring algorithm
+can be used for fitting the model.
+
+In spite of the advantages of the Poisson regression model, the Poisson
+distribution has only one parameter, which represents both the
+expectation and variance of the count random variable.  This restriction
+in the relationship between the expectation and variance induced by the
+Poisson distribution is referred as equidispersion.  However, in
+practical data analysis such a restriction can be unsuitable, since the
+observed data can present both the variance smaller or larger than the
+expectation, leading to the cases of under and overdispersion,
+respectively.  The main problem of the application of the Poisson
+regression model to non equidispersed count data is that the standard
+errors associated with the regression coefficients are inconsistently
+estimated, which in turn can lead to misleading
+inferences~\citep{Winkelmann1995, Bonat2017}.
+
+In practice, overdispersion is largely reported in the literature and
+may occur due to the absence of relevant covariates, heterogeneity of
+sampling units, different domain size not considered in the analysis and
+excess of zeros~\citep{Hinde1998}.  The case of underdispersion is less
+report in the literature, however, it has been of increasing interest in
+the statistical community. The processes that reduce the variability are
+not as well-known as those leading to extra variability.  For the same
+reason, there are a few approaches to deal with underdispersed count
+data. The explanatory mechanisms leading to underdispersion may be
+related to the underline Poisson process generating the count data.
+% Here, you gave the idea that you are going to explain all mechanisms.
+% Perhaps, you can improve on it!! Have a look at my material for
+% regression models school.
+When the time between events is not exponential distributed, the number
+of events can be over or underdispersed, this process motivated the
+class of duration dependence models~\citep{Winkelmann1995}.  Another
+possibility to explain the underdispersion phenomenon is order
+statistics.  Suppose $Y_i$ is Poisson distributed then the maximum of
+$Y_i$ is underdispersed. (NEED REF).
 
 The strategies for constructing alternative count distributions are
-related with the causes of the non-equidispersion. Particular to
-overdispersion case Poisson mixed models are widely applied. One popular
-example of this approach is the negative-binomial model, where the mean
-is added a random noise that is assumed to follow a gamma
-distribution. However, other distributions for the random effects may be
-considered, for example the model Poisson-Tweedie \citep{Bonat2017} and
-Poisson-Inverse-Gaussian. For underdispersion, the Gamma-Count
-distribution assumes a gamma distribution for the time between events,
-thus it can deal with dispersed counts \citep{Zeviani2014}. The
-COM-Poisson distribution is given by a generalization the Poisson
-distribution allowed for a non-linear decrease in ratios of successive
-probabilities \citep{Shmueli2005}.
-
-The COM-Poisson is a member of the exponential family, has the Poisson
-and geometric distributions as special cases and the Bernoulli
-distribution as a limiting case. It can deal with both under and
-overdispersed count data. Some recently applications of the COM-Poisson
-distribution are related by \citet{Lord2010} for analysis of traffic
-crash data, \citet{Sellers2010} to model airfreight breakage and book
-purchases and \citet{Huang2017} with analysis of attendance data,
-takeover birds and cotton bolls count. The main disadvantage of
-COM-Poisson distribution is that the location parameter does not
-represent the regression conditional mean making difficult
-interpretation in a regression structure. \citet{Huang2017} proposed a
-mean-parametrization COM-Poisson distribution to avoid this issue, in
-this approach the mean parameter is obtained by solving the root of an
-equation defined as a infinite sum.
-
-The main goal of this article is to proposed a novel COM-Poisson
-parametrization based on mean approximation given by
-\citet{Shmueli2005}. In this parametrization the probability mass
-function is written with $\mu$ and $\phi$ as function of original
-parameters, where $\mu$ is the mean and $\phi$ is the precision or
-dispersion parameter. Opposite the original parametrization, the
-proposed parametrization leads to interpretability of the parameters in
-a regression model and it is more simple than mean-parametrization gives
-by \cite{Huang2017} because in this case it is not necessary find $\mu$
-as the root of an equation involving an infinite sum. We conduce a
-simulation study to performe the properties of the maximum likelihood
-estimators for the proposed mean-parametrized COM-Poisson regression
-model.
+related with the causes of the non-equidispersion. In special to the
+overdispersion case the Poisson mixture models are widely applied.  One
+popular example of this approach is the negative-binomial model, where a
+gamma random effect is added to the expectation of the Poisson
+distribution.  However, other distributions can be used to represent the
+random effects.  For example the Poisson-Tweedie model~\citep{Bonat2017}
+and its special cases as the Poisson inverse-Gaussian and Neyman-Type A
+consider that the random effects are Tweedie, inverse Gaussian and
+Poisson distributed, respectively.  The Gamma-Count distribution assumes
+a gamma distribution for the time between events, thus it can deal with
+underdispersed as well as overdispersed count
+data~\citep{Zeviani2014}. The COM-Poisson distribution is obtained by a
+generalization of the Poisson distribution allowing for a non-linear
+decrease in the ratios of successive probabilities~\citep{Shmueli2005}.
+
+The COM-Poisson distribution is a member of the exponential family and
+it has the Poisson and geometric distributions as special cases as well
+as the Bernoulli distribution as a limiting case. It can deal with both
+under and overdispersed count data.  Some recently applications of the
+COM-Poisson distribution include \citet{Lord2010} for the analysis of
+traffic crash data, \citet{Sellers2010} for the modelling of airfreight
+breakage and book purchases and \citet{Huang2017} to the analysis of
+attendance data, takeover birds and cotton bolls count.  The main
+disadvantage of the COM-Poisson regression model as presented
+in~\citet{Sellers2010} is that its location parameter does not
+correspond to the expectation of the distribution, which in turn becomes
+the interpretation of the regression coefficients challenging and no
+comparable with standard approaches as the Poisson and negative binomial
+regression models. \citet{Huang2017} proposed a mean-parametrization of
+the COM-Poisson distribution in order to avoid such an issue.  In this
+approach the mean parameter is obtained by solving an non-linear
+equation defined as an infinite sum.  Consequently, it is
+computationally demanding and liable to numerical problems.
+
+The main goal of this article is to propose a novel COM-Poisson
+parametrization based on the mean approximation presented by
+\citet{Shmueli2005}. In this parametrization, the probability mass
+function is written in terms of $\mu$ and $\phi$, where $\mu$ is the
+expectation and $\phi$ is the dispersion parameter. In contrast to the
+original parametrization, the proposed parametrization leads to
+regression coefficients directly associated with the expectation of the
+response variable as usual in the context of generalized linear models.
+Consequently, the obtained regression coefficients are comparable with
+the ones obtained by standard approaches as the Poisson and negative
+binomial regression models.  Furthermore, our novel COM-Poisson
+parametrization is simpler than the strategy proposed
+by~\cite{Huang2017}, since it does not involve any numerical method for
+solving non-linear equations.
 
 This paper is organized as follows. In \autoref{background} we present
-the COM-Poisson distribution and \cite{Huang2017} parametrization. The
-proposed reparametrization and assessement of moments approximations is
-considered in the \autoref{reparametrization}. In the
-\autoref{estimation-and-inference} we show the estimation and inference
-methods based on likelihood function whose properties are assess in
-\autoref{simulation-study} by simulations. Applications of
-reparametrized COM-Poisson regression model are illustrated in
-\autoref{case-studies} with three case studies. The analysis are
-performed using the software R \citep{Rcore2017}. The source code and
-data sets are available in supplementary material\footnote{Available on
+the COM-Poisson distribution and the strategy proposed
+by~\cite{Huang2017}.  The proposed reparametrization, assessment of
+moment approximations, and study of distribution properties are
+considered in the~\autoref{reparametrization}. In
+the~\autoref{estimation-and-inference} we present the estimation and
+inference for the novel COM-Poisson regression model based on the
+likelihood paradigm. The properties of the maximum likelihood estimators
+are assess in~\autoref{simulation-study} through simulation studies.  We
+illustrate the application of the new COM-Poisson regression model
+through the analysis of three data sets. We provide an \texttt{R}
+implementation of the COM-Poisson and reparameterized COM-Poisson
+regression models as well as the analyzed data sets in the supplementary
+material.\footnote{Available on
   \texttt{\url{http://www.leg.ufpr.br/~eduardojr/papercompanions}}
@@
 print(xy2, split = c(2, 1, 3, 1), more = TRUE)
-print(xy3, split = c(3, 1, 3, 1), more = TRUE)
+print(xy3, split = c(3, 1, 3, 1), more = FALSE)
 

### 2017-11-13 article.Rnw
Improve background section

@@
 The COM-Poisson distribution generalizes the Poisson distribution in
-terms of the ratio between the probabilities of two consecutive values,
-with addition of a precision parameter \cite{Sellers2010}. Let $Y$ be a
-COM-Poisson random variable, then
+terms of the ratio between the probabilities of two consecutive events
+by adding an extra dispersion parameter~\cite{Sellers2010}.  Let $Y$ be
+a COM-Poisson random variable, then
 $$\frac{\Pr(Y=y-1)}{\Pr(Y=y)} = \frac{y^\nu}{\lambda}$$ while for the
-Poisson distribution this ratio is $\frac{y}{\lambda}$. This allows the
-distribution deals with overdispersion or underdispersion cases. The probability mass
-function for the COM-Poisson random variables takes the form
+Poisson distribution this ratio is $\frac{y}{\lambda}$. It allows the
+COM-Poisson distribution deals with non equidispersed count data.  The
+probability mass function of the COM-Poisson distribution is given by
 \begin{equation}
@@
 with large of $\lambda$, the sum is so huge it causes
-\textit{overflow}. \autoref{tab:convergenceZ} shows the calculated
-normalizing constants using a thousand increments, that is,
+\textit{overflow}. \autoref{tab:convergenceZ} shows the values of the
+normalizing constants using one thousand increments, that is,
 $\sum_{j=0}^{1000}\lambda^j/(j!)^\nu$ for different values of $\lambda$
-and $\phi$ \footnote{For $\lambda=1$ e $\nu=0$ we show Inf, although the
-  numerically value is $1000$, since the series clearly diverges.}.
-
-<<convergenceZ, echo=FALSE, results="asis">>=
+and $\phi$.
+
+<<convergenceZ, echo=FALSE>>=
 
@@
 xt <- xtabs(z ~ nu + lambda, data = grid)
+
+@
+
+\begin{table}[ht]
+\centering
+\caption{Values for $Z(\lambda, \nu)$ constant (numerically computed) for values of $\lambda$ (0.5 to 50) and $\phi$ (0 to 1)}
+\label{tab:convergenceZ}
+\begingroup\small
+\begin{tabularx}{\textwidth}{C|CCCCCC}
+  \toprule
+
+<<results="asis">>=
 caption <- paste(
-    "Values for $Z(\\lambda, \\nu)$ constant (calculated numerically)",
+    "Values for $Z(\\lambda, \\nu)$ constant (numerically computed)",
     "for values of $\\lambda$ (0.5 to 50) and $\\phi$ (0 to 1)")
@@
 xt2 <- gsub("e\\+00", "    ", format(xt))
-xt2 <- gsub("\\|([0-9].{2}) \\|", "\\|**\\1** \\|", xt2)
+xt2 <- gsub("Inf", "divergent$^{**}$", format(xt2))
+xt2[1, -1] <- "divergent$^{*\\,\\,\\,}$"
+
 print(xtable(xt2, digits = -2,
@@
       tabular.environment = "tabularx",
+      only.contents = TRUE,
       width = "\\textwidth",
@@
 
-An undesirable feature of COM-Poisson distribution is that the moments
-cannot be solved in closed form. \citet{Shmueli2005} and
+\end{tabularx}
+\endgroup
+
+\footnotesize \raggedright
+divergent$^{*}$ is a mathematically divergent serie; and
+divergent$^{**}$ a numerically divergent serie.
+\end{table}
+
+In the first line of \autoref{tab:convergenceZ} we have mathematically
+divergent series, because $\sum_{j=0}\lambda^y$, when $\lambda \geq 1$,
+is divergent.  In other cases the series diverges numerically, due to
+the computational storage limitation.  For both divergences it is
+impossible to compute probabilities, therefore, this acts as a
+restriction in the parameter space.
+
+An undesirable feature of the COM-Poisson distribution is that the
+moments cannot be obtained in closed form. \citet{Shmueli2005} and
 \citet{Sellers2010} using an asymptotic approximation for
-$Z(\lambda,\nu)$, showed that an approximate form can be
+$Z(\lambda,\nu)$, showed that the expectation and variance of the
+COM-Poisson distribution can be approximated by
 \begin{equation}
@@
 \end{equation} which is particularly accurate for  $\nu \leq 1$ or
-$\lambda > 10$. The authors also shows that the mean-variance
-relationship can be approximate by $\frac{1}{\nu}E(Y)$. In
-\autoref{reparametrization} we assess the accucary of these
-approximations.
-
-A regression formulation for COM-Poisson models was proposed by
-\citet{Sellers2010}, using the original parametrization. In this case,
-the COM-Poisson regression is $\eta(E(Y_i)) = \log(\lambda_i)$ and the
-relationship between $E(Y_i)$ and $\bm{x}_i$ is modelled
-indirectly. \citet{Huang2017} shows how the means of COM-Poisson
-distributions can be modelled directly. In the \autoref{eqn:pmf-cmp},
-\Citeauthor{Huang2017} proposes that the parameter $\lambda$ as a
-function of $\mu$ and $\nu$, given by the solution to
+$\lambda > 10$. The authors also argue that the mean-variance
+relationship can be approximate by
+$\frac{1}{\nu}E(Y)$. In~\autoref{reparametrization} we assess the
+accucary of these approximations.
+
+The COM-Poisson regression model was proposed by~\citet{Sellers2010},
+using the original parametrization. In this case, the COM-Poisson
+regression model is $\log(\lambda_i) = \bm{x}_i^\top \bm{\beta}$ and the
+relationship between $E(Y_i)$ and $\bm{x}_i$ is modelled indirectly.
+\citet{Huang2017} shows how to model directly the expectation of the
+COM-Poisson distribution in a suitable parametrization. In the
+\autoref{eqn:pmf-cmp}, \Citeauthor{Huang2017} proposes that the
+parameter $\lambda$ as a function of $\mu$ and $\nu$, given by the
+solution to
 \begin{equation*}
@@
 \end{equation*}
-Thus the mean-parametrized COM-Poisson regression is $\eta(E(Y_i)) =
-\log(\mu_i)$. In this article, we presented an alternative
-mean-parametrization of COM-Poisson models to avoid the limitations of
-original parametrization and complexity of \Citeauthor{Huang2017}
-parametrization.
+Thus the mean-parametrized COM-Poisson regression model is $\log(\mu_i)
+= \bm{x}_i^\top \bm{\beta}$. In this article, we propose an alternative
+mean-parametrization of the COM-Poisson distribution in order to avoid the
+limitations of the original parametrization and the numerical complexity
+of the \Citeauthor{Huang2017}'s approach.
 

### 2017-11-13 article.Rnw
Improve background section

@@
 
-\section{Reparametrization}
+\section{Reparametrized COM-Poisson regression model}
 \label{reparametrization}
@@
 mean approximation (\autoref{eqn:mean-approx}). We introduced new
-parameter $\mu$, using the mean approximation
+parameter $\mu$, using the this approximation
 \begin{equation}
@@
 \end{equation}
-The precision parameter is taken on the log scale to avoid restrictions
-on the parameter space, $\phi = \log(\nu)$, $\phi \in \mathbb{R}$. The
-interpretation of $\phi$ are the same to $\nu$, but in other scale. For
-$\phi<0$ we have the overdispersed case and $\phi>0$ underdispersed,
-relatives to Poisson distribution. $\phi=0$ represents the special case
-that the COM-Poisson recovers to Poisson.
+The dispersion parameter is taken on the log scale for computational
+convenience, thus $\phi = \log(\nu)$, $\phi \in \mathbb{R}$. The
+interpretation of $\phi$ is the same as the $\nu$, but in another
+scale. For $\phi < 0$ and $\phi > 0$ we have the overdispersed and
+underdispersed cases, respectively.  When $\phi=0$ we have Poisson
+distribution as the special case.
 
@@
 aux <- expand.grid(
-    mu = seq(3, 30, length.out = 50),
+    mu = seq(3, 30, length.out = 55),
     phi = seq(-1.5, 1.8, length.out = 50))
@@
 
-In order to assess the accuracy of the moments approximations
-(\autoref{eqn:mean-approx}) we presented the quadratic errors of moments
-approximations in \autoref{fig:approx-plot}(a) and
-\ref{fig:approx-plot}(b). Quadratic errors are obtained by
-$[\mu - E(Y)]^2$ for expectation and by
-$[ \mu \exp(-\phi) - \text{var}(Y)]^2$, where $E(Y)$ and var$(Y)$ are
-numerically calculated using the moments definitio. The dotted lines
+In order to assess the accuracy of the approximations presented in
+(\autoref{eqn:mean-approx}), \autoref{fig:approx-plot} presents the
+quadratic errors for the expectation (a) and variance (b).  The
+quadratic errors were obtained by $[\mu - E(Y)]^2$ for the expectation
+and by $[ \mu \exp(-\phi) - \text{var}(Y)]^2$ for the variance. In both
+cases $E(Y)$ and var$(Y)$ were computed numerically. The dotted lines
 represent the border between the regions $\nu \leq 1$ and
-$\lambda > 10^\nu$, in the $(\mu\,,\phi)$ scale.
-
-<<approx-plot, fig.height=3, fig.width=8, fig.cap="Quadratic errors for expectation (a) and variance (b) approximation, dotted lines representing the restriction for good approximations by \\cite{Shmueli2005}.
-Mean-variance relationship (c), dotted line represents $E(Y)=\text{var}(Y)$.">>=
+$\lambda > 10^\nu$, in the $\mu$ and $\phi$ scale.
+
+<<approx-plot, fig.height=3, fig.width=6, out.width=".8\\textwidth", fig.cap="Quadratic errors for the approximation of the expectation (a) and variance (b).
+Dotted lines represent the restriction for suitable approximations given by \\cite{Shmueli2005}.">>=
 
@@
                  aspect = "fill",
-                 col.regions = myreg(100),
+                 col.regions = myreg,
                  xlab = expression(phi),
@@
                      panel.levelplot(x, y, z, ...)
-                     panel.curve(10 - (exp(x) - 1)/(2 * exp(x)),
+                     panel.curve(10 - ( exp(x) - 1)/(2 * exp(x)),
                                  lty = 2)
@@
                      panel.levelplot(x, y, z, ...)
-                     panel.curve(10 - (exp(x) - 1)/(2 * exp(x)),
-                                 lty = 2)
+                     panel.curve((10 - ( exp(x) - 1)/
+                                  (2 * exp(x)))/exp(x), lty = 2)
                      panel.abline(v = 0, lty = 2)
@@
 
-#-------------------------------------------
-# Mean-variance relantionship
-xy3 <- xyplot(var ~ mean,
-              groups = phi,
-              data = grid,
-              type = "l",
-              lwd = 2,
-              axis = axis.grid,
-              xlab = expression(E(Y)),
-              ylab = expression(var(Y)),
-              sub = "(c)",
-              legend = list(
-                  top = list(
-                      fun = draw.colorkey,
-                      args = list(
-                          key = list(
-                              space = "top",
-                              col = myreg(length(unique(grid$phi))),
-                              at = unique(grid$phi),
-                              draw = FALSE)))),
-              par.settings = modifyList(
-                  ps2, list(superpose.line = list(
-                                col = myreg(length(unique(grid$phi))))
-                            )
-              ),
-              panel = function(x, y, ...) {
-                  panel.xyplot(x, y, ...)
-                  panel.curve(1*x, min(x), max(x), lty = 2)
-              },
-              page = function(...) {
-                  grid.text(expression(phi),
-                            just = "bottom",
-                            x = unit(0.15, "npc"),
-                            y = unit(0.83, "npc"))
-              })
-
-print(xy1, split = c(1, 1, 3, 1), more = TRUE)
-print(xy2, split = c(2, 1, 3, 1), more = TRUE)
-print(xy3, split = c(3, 1, 3, 1), more = FALSE)
+print(xy1, split = c(1, 1, 2, 1), more = TRUE)
+print(xy2, split = c(2, 1, 2, 1), more = FALSE)
 
@@
 The results in \autoref{fig:approx-plot} show that the mean
-approximation is accurate, the maximum of the quadratic errors is
-\Sexpr{max(grid[["emu"]])} in the range evaluated. For the variance
-approximation, the maximum is \Sexpr{max(grid[["eva"]])} and this occurs
-to negative values of $\phi$. Interestingly, the errors are bigger to
-negative values of $\phi$ and appear to have no relation with $\mu$
-against the regions gives by \citet{Shmueli2005} ($\phi \leq 0$ and
+approximation is accurate, the largest quadratic error is
+\Sexpr{max(grid[["emu"]])} for the parameter values evaluated. For the
+variance approximation, the largest quadratic error was
+\Sexpr{max(grid[["eva"]])} and it occurs for negative values of
+$\phi$. Interestingly, the errors are larger to negative values of
+$\phi$ and present no clear relation with $\mu$ opposite to the regions
+gives by \citet{Shmueli2005} ($\phi \leq 0$ and
 $\mu > 10 - \frac{\exp(\phi) - 1}{2\exp(\phi)}$).
-\autoref{fig:approx-plot}(c) shows that the relationship
-between mean and variance is linear with slope controlled by precision
-parameter $\phi$, but not in the form $\mu \exp(-\phi)$. In terms of
-moments, this leads to a specification indistinguishable from the
-quasi-Poisson.
 
@@
          border = "transparent",
-         scales = list(x = list(at = pretty(y))),
+         # scales = list(x = list(at = pretty(y))),
          ylab = expression(P(Y==y)),
@@
 @
+
+In order to explore the flexibility of the COM-Poisson model to deal
+with real count data, we compute indexes for dispersion (DI),
+zero-inflation (ZI) and heavy-tail (HI), which are respectively given by
+\begin{equation*}
+\text{DI} = \frac{\text{var}(Y)}{\text{E}(Y)}, \quad
+\text{ZI} = 1 + \frac{\log \Pr(Y = 0)}{\text{E}(Y)}
+  \quad \text{and} \quad
+\text{HT} = \frac{\Pr(Y=y+1)}{\Pr(Y=y)}\quad \text{for} \quad y \to
+\infty.
+\end{equation*}
+These indexes are defined in relation to the Poisson distribution. Thus,
+the dispersion index indicates overdispersion for $\text{DI} > 1$,
+underdispersion for $\text{DI} < 1$ and equidispersion for
+$\text{DI} = 1$. The zero-inflation index indicates zero-inflation for
+$\text{ZI} > 0$, zero-deflation for $\text{ZI} < 0$ and no excess of
+zeros for $\text{ZI} = 0$. Finally, heavy-tail index indicates a
+heavy-tail distribution for $\text{HT} \to 1$ when $y \to \infty$.
+
+These indexes are discussed by \citet{Bonat2017} to study the
+flexibility of Poisson-Tweedie distribution, and \citet{Puig2006} to
+describe count distributions. Regarding the COM-Poisson$_\mu$
+distribution, we presents in \autoref{fig:indexes-plot} the relationship
+between mean and variance (a), the dispersion and zero-inflation indexes
+(b--c) for differents values of $\mu$ and $\phi$ and heavy-tail index
+(d) for $\mu=25$ and differents values of $y$ and $\phi$.
+
+<<compute-indexes, include=FALSE, cache=TRUE>>=
+
+# Dispersion index
+grid <- transform(grid, dispersion_index = var / mean)
+
+# Zero-inflation index
+prob0 <- mapply(FUN = dcmp,
+                mu = grid$mu,
+                phi = grid$phi,
+                MoreArgs = list(y = 0, sumto = 300),
+                SIMPLIFY = FALSE)
+prob0 <- unlist(prob0)
+grid <- transform(grid, zero_index = 1 + log(prob0) / mean)
+
+# Heavy tail-index
+hi_fun <- Vectorize(FUN = function(x, mu, phi, sumto) {
+    probs <- dcmp(y = c(x, x + 1), mu = mu, phi = phi, sumto = sumto)
+    probs[2] / probs[1]
+}, vectorize.args = "x")
+
+x <- 40:150
+his <- mapply(FUN = hi_fun,
+              mu = grid$mu,
+              phi = grid$phi,
+              MoreArgs = list(x = x, sumto = 300),
+              SIMPLIFY = FALSE)
+
+aux1 <- lapply(his, function(k) data.frame("x" = x, "heavy_index" = k))
+aux1 <- do.call(rbind, aux1)
+aux2 <- grid[c("mu", "phi")][rep(1:nrow(grid), each = length(x)), ]
+heavy_data <- cbind(aux1, aux2)
+
+@
+
+<<indexes-plot, fig.width=9.5, fig.height=4, fig.cap="Indexes for COM-Poisson distribution.
+Mean and variance relation ship (a), dispersion, zero-inflation and heavy-tail indexes (b--d) for different parameter values.
+Dotted lines represents the Poisson special case.">>=
+
+#-----------------------------------------------------------------------
+# Study indexes on the reparametrized COM-Poisson distribution
+mygpar <- gpar(cex = 1.2)
+psaux <- modifyList(
+    ps0, list(superpose.line = list(
+                  col = myreg(length(unique(grid$phi)))),
+              layout.widths = list(
+                  left.padding = -0.5)
+              )
+)
+
+#-------------------------------------------
+# Mean and variance relationship
+xy1 <- xyplot(var ~ mu | "Mean-Variance",
+              groups = phi,
+              data = grid,
+              type = "l",
+              lwd = 2,
+              axis = axis.grid,
+              xlab = expression(mu),
+              ylab = "",
+              # ylab = expression(var(Y)),
+              scales = list(y = list(rot = 90)),
+              sub = "(a)",
+              legend = list(
+                  top = list(
+                      fun = draw.colorkey,
+                      args = list(
+                          key = list(
+                              space = "top",
+                              col = myreg(length(unique(grid$phi))),
+                              at = unique(grid$phi),
+                              draw = FALSE)))),
+              par.settings = psaux,
+              panel = function(x, y, ...) {
+                  panel.xyplot(x, y, ...)
+                  panel.curve(1*x, min(x), max(x), lty = 2)
+              },
+              page = function(...) {
+                  grid.text(expression(phi),
+                            just = "bottom",
+                            gp = mygpar,
+                            x = unit(0.020, "npc"),
+                            y = unit(0.875, "npc"))
+              })
+
+#-------------------------------------------
+# Dispersion index (DI)
+xy2 <- xyplot(var / mean ~ mu | "Dispersion index",
+              groups = phi,
+              data = grid,
+              type = "l",
+              lwd = 2,
+              axis = axis.grid,
+              xlab = expression(mu),
+              ylab = "",
+              scales = list(y = list(rot = 90)),
+              sub = "(b)",
+              legend = list(
+                  top = list(
+                      fun = draw.colorkey,
+                      args = list(
+                          key = list(
+                              space = "top",
+                              col = myreg(length(unique(grid$phi))),
+                              at = unique(grid$phi),
+                              draw = FALSE)))),
+              par.settings = psaux,
+              panel = function(x, y, ...) {
+                  panel.xyplot(x, y, ...)
+                  panel.curve(1 + 0*x, min(x), max(x), lty = 2)
+              },
+              page = function(...) {
+                  grid.text(expression(phi),
+                            just = "bottom",
+                            gp = mygpar,
+                            x = unit(0.020, "npc"),
+                            y = unit(0.875, "npc"))
+              })
+
+#-------------------------------------------
+# Zero-inflation index
+xy3 <- xyplot(zero_index ~ mu | "Zero-inflation index",
+              groups = phi,
+              data = grid,
+              type = "l",
+              lwd = 2,
+              axis = axis.grid,
+              xlab = expression(mu),
+              ylab = "",
+              scales = list(y = list(rot = 90)),
+              sub = "(c)",
+              legend = list(
+                  top = list(
+                      fun = draw.colorkey,
+                      args = list(
+                          key = list(
+                              space = "top",
+                              col = myreg(length(unique(grid$phi))),
+                              at = unique(grid$phi),
+                              draw = FALSE)))),
+              par.settings = psaux,
+              panel = function(x, y, ...) {
+                  panel.xyplot(x, y, ...)
+                  panel.curve(0*x, min(x), max(x), lty = 2)
+              },
+              page = function(...) {
+                  grid.text(expression(phi),
+                            just = "bottom",
+                            gp = mygpar,
+                            x = unit(0.020, "npc"),
+                            y = unit(0.875, "npc"))
+              })
+
+#-------------------------------------------
+# Heavy tail-index
+# Choose one specific mu of unique(grid$mu)
+choosemu <- 25
+xy4 <- xyplot(heavy_index ~ x | "Heavy-tail index",
+              groups = phi,
+              # data = heavy_data,
+              data = subset(heavy_data, mu == choosemu),
+              type = "l",
+              lwd = 2,
+              axis = axis.grid,
+              xlab = expression(y),
+              ylab = "",
+              scales = list(y = list(rot = 90)),
+              sub = "(d)",
+              legend = list(
+                  top = list(
+                      fun = draw.colorkey,
+                      args = list(
+                          key = list(
+                              space = "top",
+                              col = myreg(length(unique(grid$phi))),
+                              at = unique(grid$phi),
+                              draw = FALSE)))),
+              par.settings = psaux,
+              panel = function(x, y, ...) {
+                  panel.xyplot(x, y, ...)
+                  panel.curve(choosemu / (x + 1), from = min(x),
+                              to = max(x), lty = 2)
+              },
+              page = function(...) {
+                  grid.text(expression(phi),
+                            just = "bottom",
+                            gp = mygpar,
+                            x = unit(0.020, "npc"),
+                            y = unit(0.875, "npc"))
+              })
+
+print(xy1, split = c(1, 1, 4, 1), more = TRUE)
+print(xy2, split = c(2, 1, 4, 1), more = TRUE)
+print(xy3, split = c(3, 1, 4, 1), more = TRUE)
+print(xy4, split = c(4, 1, 4, 1), more = FALSE)
+
+midi <- min(grid$dispersion_index)
+madi <- max(grid$dispersion_index)
+
+@
+
+\autoref{fig:indexes-plot} shows that the indexes are slightly depend on
+the expected values and tend to stabilize for large values of
+$\mu$. Consequently, the mean and variance relationship
+\autoref{fig:indexes-plot}(a) is proportional to the dispersion
+parameter $\phi$. In terms of moments, this leads to a specification
+indistinguishable from the quasi-Poisson regression model. The
+dispersion indexes in \autoref{fig:indexes-plot}(b) show that the
+distribuion is suitable to deal to dispersed counts, of course. For the
+parameter values evaluated the largest DI was \Sexpr{madi} and smallest
+was \Sexpr{midi}. \autoref{fig:indexes-plot}(c) shows the COM-Poisson
+can handle with a limited amount of zero-inflation, in cases of
+overdispersion ($\phi < 0$). On the other hand, for $\phi > 0$
+(underdispersion) this distribution is suitable to deal with
+zero-deflated counts. Heavy-tail indexes in
+\autoref{fig:indexes-plot}(d) indicates the distribution is in general
+light tailed distribution, i.e. $HT \to 0$ for $y \to \infty$.
 

### 2017-11-13 article.Rnw
Improve reparametrization section

@@
 
-In this section we shall describe the estimation and inference methods
-adopted in this paper. The two forms of COM-Poisson models are fitted by
-maximization of the log-likelihood function. The log-likelihood function
-based on a set of independent observation $y_i$, $i=1,2,\ldots,n$ for
-COM-Poisson$_\mu$ has the following form,
+In this section we shall describe the estimation and inference to the
+two forms of the COM-Poisson regression model based on the maximum
+likelihood method. The log-likelihood function for a set of
+independent observations $y_i$, $i=1,2,\ldots,n$ for COM-Poisson$_\mu$
+from the COM-Poisson$_\mu$ distribution has the following form,
 \begin{equation}
@@
 where $\mu_i = \exp(\bm{x}_i^\top\bm{\beta})$, with
-$\bm{x}_i^\top = (x_{i1},\, x_{i2},\, \ldots,\, x_{ip})$ the vector of
+$\bm{x}_i^\top = (x_{i1},\, x_{i2},\, \ldots,\, x_{ip})$ is a vector of
 known covariates for the $i$-th observation, and $(\bm{\beta},\, \phi)
@@
 given by
-\begin{equation*}
+\begin{equation}
+\label{eqn:infseries}
   Z(\mu_i, \phi) = \sum_{j=0}^\infty \left [ \left (
@@
   \frac{1}{(j!)^{e^\phi}} \right ]
-\end{equation*}
-To compute the log-likelihood function we need the computation of an
-infinite serie for each observation, which makes the computation
-expensive for regions in the parameter space where the convergence of
-sum $Z$ is slow.
+\end{equation}
+The evaluation of the log-likelihood function for each observation
+involves the computation of the infinite series in
+\autoref{eqn:infseries}.  Thus, the fitting procedure is
+computational expansive for regions of the parameter space where the
+convergence of the infinite sum is slow.
 
 Parameter estimation requires numerical maximization of
-\autoref{eqn:ll-rcmp}. Since $\ell$ derivatives cannot be solved in
-closed forms, we adopted the BFGS algorithm \cite{Nocedal1995} that uses
-numerical estimate for the Hessian matrix
-$\mathcal{H}(\bm{\theta})$. Standard errors for regression coefficients
-are obtained based on the observed information matrix
+\autoref{eqn:ll-rcmp}. Since derivatives of $\ell$ cannot be obtained in
+closed forms, we adopted the BFGS algorithm~\cite{Nocedal1995} as
+implemented in the function \texttt{optim()} for the statistical
+software \texttt{R}. Standard errors for regression coefficients are
+obtained based on the observed information matrix
 $\mathcal{I}(\bm{\theta})$, where
-$\mathcal{I}(\bm{\theta}) = -\mathcal{H}(\bm{\theta})$ numerically
-calculated by BFGS algorithm. Confidence intervals for $\hat{\mu}_i$ are
-obtained by using the delta method \citep[p. 89]{Pawitan2001}.
-
-The parameter estimation for COM-Poisson model in the original
-parametrization is analogous to presented for COM-Poisson$_\mu$, however
-considering \autoref{eqn:ll-rcmp} in terms of $\lambda$. Even for the
-standard COM-Poisson distribution, the precision parameter is taken on
-the log scale, to avoid restriction of the parameter space.
-
-Values of the maximized log-likelihood, the Akaike criterion (AIC) and
-Bayesian criterion (BIC) are computed to compare the parametric
-models. As a baseline model, in the applications we fitted the
-quasi-Poisson model \citep{Wedderburn1974}. This approach is based on
+$\mathcal{I}(\bm{\theta}) = -\mathcal{H}(\bm{\theta})$ (Hessian matrix)
+is computed numerically. Confidence intervals for $\hat{\mu}_i$ are
+obtained by using the delta method~\citep[p. 89]{Pawitan2001}.
+
+The parameter estimation for the COM-Poisson regression model in the
+original parametrization is analogous to the one presented for the
+COM-Poisson$_\mu$ distribution, however, it considers
+\autoref{eqn:ll-rcmp} in terms of $\lambda$. Even for the standard
+COM-Poisson distribution, the dispersion parameter is taken on the log
+scale to avoid numerical issues.
+
+In the applications we fitted the quasi-Poisson model
+\citep{Wedderburn1974} as a baseline model. This approach is based on
 second-moment assumption that allows more flexibility to model. In this

### 2017-11-13 article.Rnw
Improve simulation study section

@@
 In this section we conducted a simulation study to assess the properties
-of the proposed model and estimators. We considered average counts
-varying from 3 to 27 according to a regression model structure with a
-continuous variable $\bm{x}_1$ between 0 and 1 and a categorical
-variable $\bm{x}_2$ with three levels. Then the systematic component is
-given by $\bm{\mu} = \exp(\beta_0 + \beta_1 \bm{x}_1 + \beta_{21}
-\bm{x}_{21} + \beta_{22} \bm{x}_{22})$, where $\bm{x}_{21}$ and
-$\bm{x}_{22}$ are \textit{dummy} variable for $\bm{x}_2$. With respect
-to the precision parameter $\phi$, we fixed it to four values: $-1.6$,
-$-1.0$, $0.0$ and $1.8$; in orderto have strong overdispersed, moderate
-overdispersed, equidispersed and underdispersed counts
-respectively. \autoref{fig:justpars} shows the variation of the average
-counts (left) and variation of dispersion index (E$(Y)$/var$(Y)$) for
-each fixed value of the precision parameter $\phi$. This configuration
-allows us to evaluate the properties of the estimators in extreme
-situations such as high counts and low dispersion, and low counts and
-high dispersion.
-
-In order to assess the consistency of the estimators we considered four
-different sample sizes: 50, 100, 300 and 1000; generating 1000 datasets
-in each case. In \autoref{fig:bias-plot} we show the bias of the
-estimators for each scenario (combination between values of the
-precision parameter and samples sizes) together with a confidence
-interval calculated as average bias plus and minus $\Phi(0.975)$ times
-the standard error. The scales are standardized for each parameter by
-dividing the bias by the standard error obtained for the sample of size
-50.
+of the maximum likelihood estimators as well as the flexibility of the
+COM-Poisson regression model to deal with non-equidispersed count data.
+
+We considered average counts varying from $3$ to $27$ according to a
+regression model with a continuous and a categorical covariate. The
+continuous covariate~($\bm{x}_1$) was generated as a sequence from $0$
+to $1$ and length equals to the sample size.  Similarly, the categorical
+covariate~($\bm{x}_2$) was generated as a sequence of three values each
+one repeated $3/n$ times (rounding up when required), where $n$ denotes
+the sample size. Thus, the expectation of the COM-Poisson random
+variable is given by
+$\bm{\mu} = \exp(\beta_0 + \beta_1 \bm{x}_1 + \beta_{21} \bm{x}_{21} +
+\beta_{22} \bm{x}_{22})$,
+where $\bm{x}_{21}$ and $\bm{x}_{22}$ are dummy representing the levels
+of $\bm{x}_2$.  The regression coefficients were fixed at the values,
+$\beta_0 = 2$, $\beta_1 = 05$, $\beta_{21} = 0.8$ and
+$\beta_{22} = -0.8$.
+
+We designed four simulation scenarios by considering four different
+values of the dispersion parameter $\phi = -1.6, -1.0, 0.0$ and $1.8$.
+Thus, we have strong and moderate overdispersion, equidispersion and
+underdispersion, respectively.  \autoref{fig:justpars} shows the
+variation of the average counts (left) and dispersion index (right) for
+each value of the dispersion parameter considered in the simulation
+study.  These configurations allow us to assess the properties of the
+maximum likelihood estimators in extreme situations such as high counts
+and low dispersion, and low counts and high dispersion, but also in the
+standard case of equidispersion.
 
@@
 
+In order to check the consistency of the estimators we considered four
+different sample sizes: $50$, $100$, $300$ and $1000$; generating $1000$
+data sets in each case.  In \autoref{fig:bias-plot} we show the bias of
+the estimators for each simulation scenario (combination between values
+of the dispersion parameter and samples sizes) along with the confidence
+intervals calculated as average bias plus and minus $\Phi(0.975)$ times
+the average standard error. The scales are standardized for each
+parameter by dividing the average bias by the average standard error
+obtained for the sample of size $50$.
+
 <<bias-data, cache=TRUE>>=
@@
 
-The results in \autoref{fig:bias-plot} show that for all dispersed
-levels, both the expected bias and standard error tend to 0 as the
+The results in \autoref{fig:bias-plot} show that for all dispersion
+levels, both the average bias and standard errors tend to $0$ as the
 sample size is increased. The estimators for the regression parameters
 are unbiased, consistency and their empirical distributions are
-symmetric. For the precision parameter, the estimator is asymptotically
+symmetric. For the dispersion parameter, the estimator is asymptotically
 unbiased, in small samples the parameter is overestimated and the
 empirical distribution is slightly right-skewed.
-\autoref{fig:coverage-plot} presents the confidence interval coverage
-rate based on quadratic approximation at the maximum likelihood estimate
-(assumes $\hat{\bm{\theta}} \sim \mathcal{N}(\hat{\bm{\theta}},
-\sqrt{\bm{v}})$, where $\bm{v}$ is a diagonal of the inverse of the
-observed information matrix) by sample size and dispersed levels.
+\autoref{fig:coverage-plot} presents the empirical coverage rate of the
+asymptotic confidence intervals.
 
@@
 regression parameters the empirical coverage rates are close to the
-nominal level of 95\% for sample sizes greater than 100 and all
-dispersed levels. For the precision parameter the empirical coverage
-rates are slightly lower than the nominal level, however, they become
-closer for large samples. The worst scenario is when we have small
-sample size and strong overdispersed counts.
+nominal level of 95\% for sample sizes greater than $100$ and all
+simulation scenarios. For the dispersion parameter the empirical
+coverage rates are slightly lower than the nominal level, however, they
+become closer to the nominal level for large samples.  The worst
+scenario is when we have small sample size and strong overdispersed
+counts.
 

### 2017-11-14 article.Rnw
Improve case studies section

@@
 
-In this section, we report three illustrative examples of real count
-data analysis obtained from experimental research. We considered as
-alternative models for the analysis the standard Poisson model, the
-COM-Poisson model in two forms (original and mean parametrization) and
-Quasi-Poisson model, based on second-moment assumptions. The data sets
-and \texttt{R} codes for their analysis are available on on-line
-supplements.
+In this section, we report three illustrative examples of count data
+analysis. We considered as alternative models for the analysis the
+standard Poisson regression model, the COM-Poisson model in the two
+forms (original and new parametrization) and the quasi-Poisson
+regression model.  The data sets and \texttt{R} codes for their analysis
+are available as supplementary material.
 
@@
 submitted to five levels of artificial defoliation (\texttt{des}) and
-crossed with five growth stages (\texttt{est}). The purpose of this
-research was to study the effect of defoliation levels at different
-growth stages of cotton plants on the cotton production, expressed by
-the number of bools produced. The study was conducted in a greenhouse
-and the experimental design was completely randomized with five
-replicates. This data set was analysed by \citet{Zeviani2014} using the
+crossed with five growth stages (\texttt{est}). The main goal of this
+study was to assess the effect of defoliation levels at different growth
+stages of cotton plants on the cotton production, expressed by the
+number of bolls produced. The study was conducted in a greenhouse and
+the experimental design was completely randomized with five
+replicates. This data set was analyzed by~\citet{Zeviani2014} using the
 \textit{Gamma-Count} distribution.
 
-Following the results of \citet{Zeviani2014}, we considered the linear
-predictor specified by
+Following~\citet{Zeviani2014}, the linear predictor is given by
 $$\log(\mu_{ij}) = \beta_0 + \beta_{1j} \textrm{def}_i + \beta_{2j}
@@
 and $j$-th growth stage ($j$ = 1: vegetative, 2: flower bud, 3: blossom,
-4: boll, 5: boll open), that is, we have a second order effect of
+4: boll, 5: boll open), i.e. we have a second order effect of
 defoliation in each growth stage. The parameters estimates and
 goodness-of-fit measures for the Poisson, COM-Poisson, COM-Poisson$_\mu$
-and quasi-Poisson models are presented in \autoref{tab:coef-cotton}.
+and quasi-Poisson regression models are presented in
+\autoref{tab:coef-cotton}.
 
@@
 The results presented in \autoref{tab:coef-cotton} show that the
-goodness-of-fit measures (maximum log-likelihood, AIC and BIC) are quite
-similar between COM-Poisson and COM-Poisson$_\mu$. This shows that the
-reparametrization does not change the model fit, as expected. The Poisson
-model is clearly unsuitable, being overly conservative. The difference
-between the log-likelihoods of the Poisson and COM-Poisson$_\mu$ models
-is \Sexpr{lrt.ncap[2, 4]}. Comparing this value with the chi-square
-distribution, with one degree of freedom, we have these models are
-significantly different ($\phi \neq 0$). The estimated value of
-precision parameter, $\hat{\phi}$ is \Sexpr{co1[1]}, indicating
-underdispersion.
-
-\autoref{tab:coef-cotton} also shows the advantage of COM-Poisson$_\mu$
-(reparametrized), since the estimates are quite similar those obtained by
-the Poisson model, whereas estimates obtained by the COM-Poisson model
-in original parametrization are on a non-interpretable scale. The ratios
-between estimates and their respective standard errors for the
-COM-Poisson models are very close to ratios obtained by quasi-Poisson
-model. However, it is important to note that COM-Poisson model is a
-full parametric approach, that is, there is a probability distribution
-associate to counts.
-
-In the \autoref{fig:pred-cotton} are presented the observed values and
-curves of fitted values with confidence intervals (95\%) as functions of
-the defoliation level for each growth stage. The fitted values are same
-for both models, but confidence intervals are bigger for Poisson model
-because the equidispersion assumption. The results from COM-Poisson
-models are consistent with those from the Gamma-Count model, fitted by
-\cite{Zeviani2014}, Poisson-Tweedie model, fitted by \cite{Bonat2017}
-and alternative COM-Poisson parametrization, fitted by \cite{Huang2017},
-in that both methods indicate underdispersion and significant effects of
-defoliation for the vegetative, blossom and boll growth stages.
+goodness-of-fit measures (log-likelihood, AIC and BIC) are quite similar
+for the COM-Poisson and COM-Poisson$_\mu$ models.  It suggests that the
+reparametrization does not change the model fit, as expected.  The
+Poisson model is clearly unsuitable, being overly conservative.  The
+difference in terms of log-likelihood value from the Poisson to the
+COM-Poisson$_\mu$ model was $\Sexpr{lrt.ncap[2, 4]}$, which in turn
+suggests the better fit of the COM-Poisson$_{\mu}$ model. A chi-square
+test also supports this statement.  Finally, the estimated value of the
+dispersion parameter $\hat{\phi} = \Sexpr{co1[1]}$ suggests underdispersion.
+
+Furthermore, results in \autoref{tab:coef-cotton} also shows the
+advantage of the COM-Poisson$_\mu$ model, since the estimates are quite
+similar to the ones obtained by the Poisson model, whereas estimates
+obtained from the COM-Poisson model in the original parametrization are
+on a non interpretable scale.  The ratios between estimates and their
+respective standard errors for the COM-Poisson models are very close to
+ratios obtained by quasi-Poisson model. However, it is important to note
+that COM-Poisson model is a full parametric approach, i.e. there is a
+probability distribution associate to counts. On the other hand, the
+quasi-Poisson model is a specification based only on second-moments
+assumptions.
+
+\autoref{fig:pred-cotton} presents the observed and fitted values with
+confidence intervals (95\%) as a function of the defoliation level for
+each growth stage. The fitted values are the same for the Poisson and
+COM-Poisson$_{\mu}$ models, however, the confidence intervals are larger
+for the Poisson model because the equidispersion assumption.  The
+results from the COM-Poisson$_{\mu}$ model are consistent with those
+from the Gamma-Count model~\cite{Zeviani2014},
+Poisson-Tweedie~\cite{Bonat2017} and the alternative parametrization of
+the COM-Poisson distribution proposed by~\cite{Huang2017}.  In all
+strategies the models indicated underdispersion and significant effects
+of defoliation for the vegetative, blossom and boll growth stages.
 
@@
 
-In order to assess the orthogonality between $\bm{\mu}$ and $\phi$ in
-the COM-Poisson$_\mu$ parametrization, \autoref{tab:corr-cotton}
-presents the empirical correlations between precision and regression
-parameters, obtained from the estimated variance and covariance
-matrix. The correlations are practically null considering the
-COM-Poisson$_\mu$, whereas to the original parametrization they are
-quite high. This reflects a better performance of the maximization
-algorithm in mean-parametrization model, in this case COM-Poisson$_\mu$
-fit was \Sexpr{time12[3] * 100 / time11[3]}\% slower then COM-Poisson
-model. It is important to note that the initial values for the BFGS
-algorithm are provided by Poisson fitted model, then in the COM-Poisson
-mean-parametrized model the initial values are practically the maximum
-likelihood estimates and the effort of maximization is on the precision
-parameter $\phi$.
+In order to assess the relation between $\bm{\mu}$ and $\phi$ in the
+COM-Poisson$_\mu$ parametrization, \autoref{tab:corr-cotton} presents
+the empirical correlations between the regression and dispersion
+parameters, as computed by the asymptotic covariance matrix of the
+estimators, i.e. the inverse of the observed information.  The
+correlations are practically null considering the COM-Poisson$_\mu$. On
+the other hand, for the original parametrization such correlations are
+quite larger, in special for the parameter $\beta_0$ (due to effects
+parametrization in the linear predictor).  This result explain the
+better performance of the maximization algorithm in the new
+parametrization.  It is important to note that the initial values for
+the \texttt{BFGS} algorithm are provided by the Poisson model, then in
+the COM-Poisson$_{\mu}$ model the initial values are practically the
+maximum likelihood estimates and the effort of maximization is on the
+dispersion parameter $\phi$ only. To compare the computational times on
+the two parametrizations we repeat $50$ times the routine to fit. In
+this case COM-Poisson$_\mu$ fit was, on average, $38$\% faster than the
+original one.
 
@@
 1.8 $\times$ 100mg dm$^{-3}$) and soil moisture (\texttt{umid}) levels
-(37.5, 50, 62.5\%) on the soybean production. The experiment was
-carried out in a greenhouse, in pots with two plants, and the count
-variable measured was the number of grains per pot
-\citep{Serafim2012}. \autoref{fig:desc-soy} (left) shows the number of
-grains recorded for each combination of potassium dose and moisture
-level, it is important to note the indication of a quadratic level of
-the potassium levels. Most points in the sample variance vs. sample
-means dispersion diagram (right) are above the identity line, suggesting
+(37.5, 50, 62.5\%) on the soybean production. The experiment was carried
+out in a greenhouse, in pots with two plants, and the count variable
+measured was the number of grains per pot~\citep{Serafim2012}.
+\autoref{fig:desc-soy} (left) shows the number of grains recorded for
+each combination of potassium dose and moisture level, it is important
+to note the indication of a quadratic effect of the potassium
+levels. Most points in the sample variance \textit{versus} sample means
+dispersion diagram (right) are above the identity line, suggesting
 overdispersion (block effect not yet removed).
@@
 
-For the analysis of this data set we proposed, based on descriptive
+For the analysis of this data set we proposed based on descriptive
 analysis \autoref{fig:desc-soy}, the linear predictor specified as a
@@
 between estimate and standard error and goodness-of-fit measures for
-the alternative models fitted.
-
-The results in \autoref{tab:coef-soy} has similar interpretations to
-results presented in \autoref{case-cotton}. The two parametrization of
-COM-Poisson presented very similar goodness-of-fit measures and better
-than Poisson model. The difference between the log-likelihoods of the
-Poisson and COM-Poisson models is \Sexpr{lrt.ngra[2, 4]}, indicating
-that $\phi$ is significantly different of zero. With respect to the
-regression parameters, the similarities between models are analogous to
-the previous section. Both models indicate effects of block, potassium
-dose and moisture level, however the Poisson model indicates the effects
-with greater significance, because it does not fit the extra
-variability.
+the alternative models.
+
+The results in \autoref{tab:coef-soy} have similar interpretations to
+the results presented in \autoref{case-cotton}. The two parametrization
+of COM-Poisson model presented very similar goodness-of-fit measures and
+better fit than the Poisson model. The difference between the
+log-likelihoods of the Poisson and COM-Poisson models was
+$\Sexpr{lrt.ngra[2, 4]}$, indicating that $\phi$ is significantly
+different from zero. Concerning to the regression parameters, the
+similarities between the models are analogous to the previous
+Section. Both models indicate effects of block, potassium dose and
+moisture level, however the Poisson model indicates the effects with
+greater significance, because it does not fit the extra variability.
 
@@
 
-The infinite sum $Z(\mu, \phi)$ in cases of oversidepersed counts needs
-a greater upper bound for the convergence, making the computation of the
-likelihood more expansive. In this case, the upper bound was 700. The
-number of likelihood computations, on the BFGS algorithm, was
-\Sexpr{c21[1]} and \Sexpr{c22[2]} for the original and
-mean-parametrization and the time to fit was \Sexpr{time21[3]} and
-\Sexpr{time22[3]} seconds respectively. This may be due the
-well-behaviour of likelihood function for orthogonal parameters also,
-best start values from Poisson model. In \autoref{tab:corr-soy} we
-present the empirical correlations between regression and precision
-parameter. The correlations are close to zero for the COM-Poisson$_\mu$
-parametrization, indicating the empirical orthogonality between $\mu$
-and $\phi$ on COM-Poisson$_\mu$ distribution.
+The infinite sum $Z(\mu, \phi)$ in the cases of overdispersed count data
+requires a larger upper bound to reach convergence. Thus, in these cases
+the computation of the log-likelihood function is computationally
+expensive.  For the data set considered, the upper bound was fixed at
+$700$.  The \texttt{BFGS} algorithm evaluated the log-likelihood
+function $\Sexpr{c21[1]}$ and $\Sexpr{c22[2]}$ times to reach
+convergence, when using the original and new parametrization of the
+COM-Poisson distribution, respectively.  In terms of computational time
+and for 50 repetitions of fit, the proposed reparametrization was on
+average $110\%$ faster than the original one.  Probably, it is due to
+the better behaviour of the log-likelihood function as well as better
+initial values obtained from the Poisson fit. In \autoref{tab:corr-soy}
+we present the empirical correlation between the regression and
+dispersion parameter estimates. The correlations are close to zero for
+the COM-Poisson$_\mu$ model, indicating the empirical orthogonality
+between $\mu$ and $\phi$.
 
@@
 
-The observed counts and fitted curves for each humidity level with
-confidence bands are show in \autoref{fig:pred-soy}. The fitted values
-are identical for both models, leading to the same conclusions. On the
-other hand, confidence bands for the Poisson model are smaller due to
-equidispersion assumption. Considering COM-Poisson models and
-quasi-Poisson approach, the confidence bands are quite similar, which
-shows their flexibility.
+The observed and fitted counts for each humidity level with confidence
+intervals are shown in \autoref{fig:pred-soy}. The fitted values are
+identical for the Poisson and COM-Poisson$_{\mu}$ models, leading to the
+same conclusions. On the other hand, confidence intervals for the
+Poisson model are smaller than the ones from the
+COM-Poisson$_{\mu}$. Probably, it is due to the equidispersion
+assumption underline the Poisson model. The confidence intervals from
+the COM-Poisson$_{\mu}$ and quasi-Poisson models are really similar,
+which in turn show the already highlighted similarity between these
+approches.
 
@@
 toxicity of a herbicide, nitrofen, on a species of zooplankton
-(\textit{Ceriodaphnia dubia}). 50 animals were randomized into batches
-of 10 and each batch was put in a solution with a measured concentration
-of nitrofen (0, 0.8, 1.6, 2.35 e 3.10 mug$/10^2$litre)
+(\textit{Ceriodaphnia dubia}). Fifty animals were randomized into
+batches of ten and each batch was put in a solution with a measured
+concentration of nitrofen (0, 0.8, 1.6, 2.35 e 3.10 mug$/10^2$litre)
 (\texttt{dose}). Then the number of total live offspring was recorded
@@
 For this data set we proposed three linear predictors. We compare the
-performance of models using the likelihood ratio tests to select the
-systematic part of model. The linear predictors are
+performance of models using the likelihood ratio tests. The linear
+predictors are
 \begin{center}
@@
 
-\autoref{tab:anova-ovos} summarizes the results of fitted models and
+\autoref{tab:anova-ovos} summarizes the results of the fitted models and
 likelihood ratio tests comparing the sequence of predictors. All models
-indicate the cubic effect of nitrofen concentration. Considering this
-predictor, there is an evidence of equidispersed counts, $\phi$
-estimates of COM-Poisson close to zero and $\sigma$ of quasi-Poisson, to
-one. It is interesting to note that if we omit the effects of a greater
-order on the predictor, the modes show evidence of overdispersion
-counts. This exemplifies the discussion about causes of overdispersion
-made in \autoref{introduction}. We can also note that the quasi-Poisson
-approach, although robust to equisdispersion assumption, shows a higher
-descriptive levels ($p$-values) than parametric models, that is, the
-tests under parametric models are most powerful than quasi-Poisson
-models in the equidispersed case.
-
-In \autoref{tab:coef-ovos} we presented the estimates of the regression
+indicate the significance of the cubic effect of the nitrofen
+concentration.  Considering this predictor, there is an evidence of
+equidispersed counts, $\phi$ estimates of the COM-Poisson close to zero
+and $\sigma$ of quasi-Poisson close to one.  It is interesting to note
+that if we omit the high order effects the models show evidence of
+overdispersion. This exemplifies the discussion about causes of
+overdispersion made in \autoref{introduction}.  We can also note that
+the quasi-Poisson approach, although robust to equidispersion
+assumption, shows a higher descriptive levels ($p$-values) than
+parametric models, that is, the tests under parametric models are more
+powerful than the ones under the quasi-Poisson model in the
+equidispersed case.
+
+In \autoref{tab:coef-ovos} we present the estimates of the regression
 parameters considering the cubic dose effect (third predictor). The
-interpretations are similar to the discussed in the others cases
-studies, but in this the Poisson model is also suitable for indicating
-the significance of the effects. In addition, note that the parameters
-of the standard COM-Poisson model are comparable to the others models,
-this occurs because it is the particular case $\phi = 0$ which implies
-$\lambda = \mu$. \autoref{fig:pred-ovos} shows the number of live
-off-springs observed in experiment with fitted curves and confidence
-bands for all model strategies adopted.
+interpretations are similar to others cases studies, however, in this
+case the Poisson model is also suitable for indicating the significance
+of the covariate effects.  In addition, note that the parameter
+estimates of the original COM-Poisson model are comparable to the others
+models. It occurs because it is the special case $\phi = 0$ which
+implies $\lambda = \mu$. \autoref{fig:pred-ovos} shows the number of
+live off-springs observed and fitted along with confidence intervals for
+all model strategies adopted.
 
@@
 
-The fitted values and confidence bands in \autoref{fig:pred-ovos} come
-to have a full overlay. This shows that even in case where is not
-necessary to estimate the precision parameter $\phi$, but estimating it
-does not lead to incorrect analysis.
+The fitted values and confidence intervals in \autoref{fig:pred-ovos}
+come to have a full overlay. It shows that the estimation of the extra
+dispersion parameter does not affect the estimation of the regression
+coefficients in the case of equidispersed counts.
 
@@
 
-Finally, in \autoref{tab:corr-ovos} we presented the empirical
-correlations between regression and precision parameters. The results
-show that even in the particular case ($\phi=0$), the empirical
-correlations for the standard models are not zero. For the
+Finally, in \autoref{tab:corr-ovos} we present the empirical
+correlations between the regression and dispersion parameter estimates.
+The results show that even in the special case ($\phi=0$), the empirical
+correlations for the original COM-Poisson model are not zero.  For the
 reparametrized model, as discussed in the previous sections, the
-correlations are practically null.
+correlations are practically null. The computational times for fifty
+repetitions of fit are similar. The average time to fit for the
+COM-Poisson$_\mu$ and COM-Poisson models is 1.19 and 1.09 seconds,
+respectively.
 

### 2017-11-14 article.Rnw
Improve conclusion section

@@
 
-In this paper, we argued the parametrization of COM-Poisson distribution
-to analyse dispersed count data. We proposed a novel reparametrization
-based only mean approximation indicated by \cite{Shmueli2005}. The
-likelihood-based method to estimation and inference is presented. We
-assess the properties of these estimators by a simulation
-study. Applications of the proposed model are showed by three case
-studies with over, under and equidispersed counts. In the data analysis
-we compared standard Poisson, original COM-Poisson, reparametrized
-COM-Poisson and quasi-Poisson models.
-
-We computed the quadratic errors for the mean approximation. The results
-showed that it is well accurate supports our proposed
-reparametrization. The results of a simulation study showed that the
-estimators of the regression and precision parameters are consistent and
-unbiased and asymptotically unbiased respectively. The empirical
-coverage rates of the confidence intervals for the parameters, compute
-by the asymptotic distribution of the estimators, are close to the
-nominal level for sizes greater than 100. The worst scenario is when we
-have small sample size and strong overdispersed counts. In general, we
-recommend the use of Wald intervals for computational simplicity and
-good results.
-
-The data analysis in the case studies has shown that COM-Poisson is a
-suitable distribution for over, under and equidispersion situations. The
-observed empirical correlation between regression and precision
-parameters suggest the orthogonality between $\mu$ and $\phi$ in
-COM-Poisson$_\mu$ distribution against the original parametrization. Due
-to this feature the computational procedure is faster under the
-reparametrized model. Another advantage, perhaps the main, of the
-reparametrization is that the estimated regression parameters have
-interpretation of rate ratios, same of the Poisson model. In comparison
-of the quasi-Poisson approach the results was similar in both
-simulations, nevertheless the COM-Poisson is a full parametric model,
-allowing the calculation of probabilities for example.
+In this paper, we presented a novel parametrization of the COM-Poisson
+distribution and the associated regression model. The novel
+parametrization was based on a simple asymptotic approximation for the
+expectation and variance of the COM-Poisson distribution. The main
+advantage of the proposed reparametrization is the simple interpretation
+of the regression coefficients in terms of the expectation of the
+response variable as usual in the generalize linear models
+context. Thus, it is possible to compare the results of the COM-Poisson
+model with the ones from standard approaches as the Poisson and
+quasi-Poisson regression models. Furthermore, in the novel
+parametrization the COM-Poisson distribution is indexed by the
+expectation $\mu$ and an extra dispersion parameter $\phi$ whose our
+data analysis suggest to be orthogonal.
+
+We evaluated the accuracy of the asymptotic approximations for the
+expectation and variance of the COM-Poisson distribution by considering
+quadratic approximation errors.  The results showed that the
+approximations are accurate for a large part of the parameter space of
+the parameters $\mu$ and $\phi$, which in turn support our
+reparametrization.  We carried out a simulation study to assess the
+properties of the reparametrized COM-Poisson model to deal with
+different levels of dispersion as well as the properties of the maximum
+likelihood estimators. The results of our simulation study suggested
+that the maximum likelihood estimators of the regression and dispersion
+parameters are unbiased and consistent.  The empirical coverage rates of
+the confidence intervals computed based on the asymptotic distribution
+of the maximum likelihood estimators are close to the nominal level for
+sample size greater than $100$.  The worst scenario is when we have
+small sample sizes and strong overdispersed counts.  In general, we
+recommend the use of the asymptotic confidence intervals for
+computational simplicity.
+
+The data analyses have shown that the COM-Poisson regression model is a
+suitable choice to deal with non equidispersed count data.  The observed
+empirical correlation between the regression and dispersion parameter
+estimators suggest orthogonality between $\mu$ and $\phi$ in
+COM-Poisson$_\mu$ distribution. Thus, the computational procedure based
+on the proposed reparametrization is faster than in the original
+parametrization.
 
 In general, the results presented by the reparametrized COM-Poisson
-models were satisfactory and superior to the conventional
-approaches. Therefore, its use in the analysis of count data is
-encouraged. The computational routines for fitting of original and
-reparametrized COM-Poisson models are available in the supplementary
+models were satisfactory and comparable to the convential approaches.
+Therefore, its use in the analysis of count data is encouraged. The
+computational routines for fitting the original and reparametrized
+COM-Poisson regression models are available in the supplementary
 material\textsuperscript{\ref{papercompanion}}.
 
-There are many possible extensions and possible topics for further
-investigation to model discussed in the present paper, including
-simulation study to assess model robustness against distribution
-miss-specification, asses theoretical approximations for
-$Z(\lambda, \nu)$ (or $Z(\mu,\phi))$, in order to avoid the selection of
-sum's upper bound, propose a double GLM based on the reparametrized
-COM-Poisson model and a accommodate random effects since we have a model
-parametrized on mean.
+There are many possible extensions to the model discussed in this paper,
+including simulation studies to assess the model robustness against
+model misspecification and to asses the theoretical approximations for
+$Z(\lambda, \nu)$ (or $Z(\mu,\phi))$. Another simple extension of the
+proposed model is to model both $\mu$ and $\phi$ parameters as functions
+of covariate in a double generalized linear models framework.  Finally,
+the reparametrized version of the COM-Poisson model also encourages the
+specification of generalized linear mixed models using this
+distribution.
 

### 2017-11-14 article.Rnw
Text corrections

@@
 related to the underline Poisson process generating the count data.
-% Here, you gave the idea that you are going to explain all mechanisms.
-% Perhaps, you can improve on it!! Have a look at my material for
-% regression models school.
 When the time between events is not exponential distributed, the number
@@
 \footnotesize \raggedright
-divergent$^{*}$ is a mathematically divergent serie; and
-divergent$^{**}$ a numerically divergent serie.
+divergent$^{*}$ is a mathematically divergent series; and
+divergent$^{**}$ a numerically divergent aerie.
 \end{table}
@@
 In the first line of \autoref{tab:convergenceZ} we have mathematically
-divergent series, because $\sum_{j=0}\lambda^y$, when $\lambda \geq 1$,
-is divergent.  In other cases the series diverges numerically, due to
-the computational storage limitation.  For both divergences it is
-impossible to compute probabilities, therefore, this acts as a
-restriction in the parameter space.
+divergent series, because $\sum_{j=0}^\infty\lambda^y$, when
+$\lambda \geq 1$, is divergent.  In other cases the series diverges
+numerically, due to the computational storage limitation.  For both
+divergences it is impossible to compute probabilities, therefore, this
+acts as a restriction in the parameter space.
 
@@
 $\frac{1}{\nu}E(Y)$. In~\autoref{reparametrization} we assess the
-accucary of these approximations.
+accuracy of these approximations.
 
@@
 between mean and variance (a), the dispersion and zero-inflation indexes
-(b--c) for differents values of $\mu$ and $\phi$ and heavy-tail index
-(d) for $\mu=25$ and differents values of $y$ and $\phi$.
+(b--c) for different values of $\mu$ and $\phi$ and heavy-tail index
+(d) for $\mu=25$ and different values of $y$ and $\phi$.
 
@@
 dispersion indexes in \autoref{fig:indexes-plot}(b) show that the
-distribuion is suitable to deal to dispersed counts, of course. For the
+distribution is suitable to deal to dispersed counts, of course. For the
 parameter values evaluated the largest DI was \Sexpr{madi} and smallest
@@
 which in turn show the already highlighted similarity between these
-approches.
+approaches.
 
@@
 batches of ten and each batch was put in a solution with a measured
-concentration of nitrofen (0, 0.8, 1.6, 2.35 e 3.10 mug$/10^2$litre)
+concentration of nitrofen ($0, 0.8, 1.6, 2.35$ e $3.10$ mug$/10^2$litre)
 (\texttt{dose}). Then the number of total live offspring was recorded
@@
 
-For this data set we proposed three linear predictors. We compare the
+For this data set we proposed three linear predictors. We compared the
 performance of models using the likelihood ratio tests. The linear
@@
 In general, the results presented by the reparametrized COM-Poisson
-models were satisfactory and comparable to the convential approaches.
+models were satisfactory and comparable to the conventional approaches.
 Therefore, its use in the analysis of count data is encouraged. The

### 2017-11-14 codes/24-timestofit.R
Add script to evaluate computational times

@@ -1,0 +1,83 @@
+# =====================================================================
+# Comparasion of the computational times
+#                                                        Eduardo Junior
+#                                                    edujrrib@gmail.com
+#                                                            2017-08-05
+# =====================================================================
+
+#-----------------------------------------------------------------------
+# Load package and functions
+library(bbmle)
+library(microbenchmark)
+source("lattice-panels.R")
+source("functions.R")
+
+# Colors for legends
+cols <- trellis.par.get("superpose.line")$col
+
+#-----------------------------------------------------------------------
+# Load datas
+
+#-------------------------------------------
+# Section 6.1
+data1 <- read.table("../data/cottonBolls.txt",
+                    header = TRUE, sep = "\t")
+form1 <- ncap ~ est:(des + I(des^2))
+
+#-------------------------------------------
+# Section 6.2
+data2 <- read.table("../data/soyaBeans.txt",
+                        header = TRUE, sep = "\t")
+data2$umid <- as.factor(data2$umid)
+data2 <- data2[-74, ] ## Incorrect observation
+data2 <- transform(data2, K = K / 100)
+form2 <- ngra ~ bloc + umid * K + I(K^2)
+
+#-------------------------------------------
+# Section 6.3
+data3 <- read.table("../data/nitrofen.txt",
+                       header = TRUE, sep = "\t")
+data3 <- transform(data3, dose = dose / 100)
+form3 <- novos ~ dose + I(dose^2) + I(dose^3)
+
+#-----------------------------------------------------------------------
+# Compare computation times by 50 repetitions
+
+# Times to the first case study
+bench1 <- microbenchmark(
+    "CMP  " = fitcm(form1, data = data1, model = "CP" , sumto = 50),
+    "CMPmu" = fitcm(form1, data = data1, model = "CP2", sumto = 50),
+    times = 50)
+
+#-------------------------------------------
+# Unit: seconds
+#   expr      min       lq     mean   median       uq      max neval cld
+#  CMP   2.134262 2.144675 2.167662 2.157547 2.187935 2.326526    50   b
+#  CMPmu 1.533308 1.549595 1.568905 1.559458 1.581462 1.627994    50  a
+#-------------------------------------------
+
+# Times to the second case study
+bench2 <- microbenchmark(
+    "CMP  " = fitcm(form2, data = data2, model = "CP" , sumto = 700),
[…]

### 2017-11-17 article.Rnw
Add an alternative graph to show bias results and fixed little things

@@
     fig.align = "center",
-    fig.pos = "!ht",
+    fig.pos = "!htp",
     # dev = "tikz"
@@
   \qquad y = 0, 1, 2, \ldots\,,
-\end{equation} where $\mu > 0$. We denotes this distribution as
+\end{equation} where $\mu > 0$. We denote this distribution as
 COM-Poisson$_\mu$. In \autoref{fig:pmf-cmp} we show the shapes of
@@
 
-Parameter estimation requires numerical maximization of
-\autoref{eqn:ll-rcmp}. Since derivatives of $\ell$ cannot be obtained in
-closed forms, we adopted the BFGS algorithm~\cite{Nocedal1995} as
-implemented in the function \texttt{optim()} for the statistical
-software \texttt{R}. Standard errors for regression coefficients are
-obtained based on the observed information matrix
-$\mathcal{I}(\bm{\theta})$, where
+Parameter estimation requires the numerical maximization of
+\autoref{eqn:ll-rcmp}. Since the derivatives of $\ell$ cannot be
+obtained in closed forms, we adopted the \texttt{BFGS}
+algorithm~\cite{Nocedal1995} as implemented in the function
+\texttt{optim()} for the statistical software \texttt{R}. Standard
+errors for the regression coefficients are obtained based on the
+observed information matrix $\mathcal{I}(\bm{\theta})$, where
 $\mathcal{I}(\bm{\theta}) = -\mathcal{H}(\bm{\theta})$ (Hessian matrix)
@@
 
-<<bias-plot, cache=TRUE, fig.height=5, fig.width=8, fig.cap="Distributions of standardized bias (gray points) and average with confidence intervals (black segments) by differents sample sizes and dispersion levels.">>=
+<<bias-plot, cache=TRUE, fig.height=6.5, fig.width=8.5, fig.cap="Distributions of standardized bias (gray points) and average with confidence intervals (black segments) by different sample sizes and dispersion levels.">>=
 
@@
 
-xyplot(param ~ bias | phi, groups = n, data = bias,
+# Alternative graph (figure with all point is very large in size memory)
+bwplot(param ~ bias | phi, groups = n, data = bias,
+       pch = "|",
+       grid = TRUE,
+       as.table = TRUE,
+       layout = c(4, 1),
+       xlab = "Standardized Bias",
+       fill = "gray80",
+       col = "red",
+       strip = strip.custom(factor.levels = fl),
        panel = panel.superpose,
-       layout = c(4, 1),
+       scales = list(y = list(labels = yl)),
+       box.width = 0.05,
+       whisker.width = 0.4,
        key = key,
-       strip = strip.custom(factor.levels = fl),
-       scales = list(y = list(labels = yl)),
-       ylab = "",
-       xlab = "Standardized Bias",
-       jitter.y = TRUE,
-       factor = 0.1,
-       gap = 0.3,
-       alpha = 0.6,
-       col = "gray80",
-       cex = 0.7,
-       panel.groups = function(x, y, group.number,
-                               subscripts, gap, ...){
-           noise <- centfac(factor(levels(bias$n)), space = gap)
-           noise <- sort(noise)
-           panel.xyplot(x, y + noise[group.number], ...)
+       par.settings = list(
+           plot.symbol = list(pch = 20, cex = 0.5, col = "gray70"),
+           box.rectangle = list(col = "gray70"),
+           box.umbrella=list(col = "gray70", lwd = 2)
+       ),
+       panel.groups = function(x, y, group.number, ...) {
+           my.panel.bwplot(x, (y - 0.05) + (group.number - 2.5) / 5, ...)
+           panel.abline(v = 0, lty = 2)
        }) +
-    as.layer(
     segplot(param ~ bias[, "lwr"] + bias[, "upr"] | phi,
@@
             })
-    )
+
+# xyplot(param ~ bias | phi, groups = n, data = bias,
+#        panel = panel.superpose,
+#        layout = c(4, 1),
+#        key = key,
+#        strip = strip.custom(factor.levels = fl),
+#        scales = list(y = list(labels = yl)),
+#        ylab = "",
+#        xlab = "Standardized Bias",
+#        jitter.y = TRUE,
+#        factor = 0.1,
+#        gap = 0.3,
+#        alpha = 0.6,
+#        col = "gray80",
+#        cex = 0.7,
+#        panel.groups = function(x, y, group.number,
+#                                subscripts, gap, ...){
+#            noise <- centfac(factor(levels(bias$n)), space = gap)
+#            noise <- sort(noise)
+#            panel.xyplot(x, y + noise[group.number], ...)
+#        }) +
+#     as.layer(
+#         segplot(param ~ bias[, "lwr"] + bias[, "upr"] | phi,
+#                 centers = bias[, "fit"],
+#                 data = cidata,
+#                 draw = FALSE,
+#                 horizontal = TRUE,
+#                 layout = c(4, 1),
+#                 groups = n, gap = 0.3,
+#                 key = key,
+#                 pch = 21:24,
+#                 cex = 0.7,
+#                 lwd = 2,
+#                 panel = function(...) {
+#                     panel.groups.segplot(...)
+#                     panel.abline(v = 0, lty = 2)
+#                     panel.abline(h = 1:5, col = "lightgray", lty = 2)
+#                 })
+#     )
+
 @
@@
 
-<<coverage-plot, fig.height=4, fig.width=9, fig.cap="Coverage rate based on confidence intervals obtained by quadratic approximation for differents sample sizes and dispersion levels.">>=
+<<coverage-plot, fig.height=4, fig.width=9, fig.cap="Coverage rate based on confidence intervals obtained by quadratic approximation for different sample sizes and dispersion levels.">>=
 
@@
 \centering
-\caption{Empirical correlations between $\hat{\phi}$ and $\hat{\bm{\beta}}$ for the two parametrizations of COM-Poisson model fit to equidispersed data.}
+\caption{Empirical correlations between $\hat{\phi}$ and $\hat{\bm{\beta}}$ for the two parametrizations of COM-Poisson model fit to underdispersed data.}
 \label{tab:corr-cotton}

### 2017-11-17 codes/01-simulation.R

@@ -71,6 +71,7 @@ rcmp <- Vectorize(FUN = function(n, mu, phi) {
 
 #-------------------------------------------
 # Simulation
+# CAUTION LONG TIME CONSUMING 12H, LOAD THE RESULTS OF COMPRESSED FILE
 library(parallel)
 results <- mclapply(sizes, function(n) {
     x1 <- seq(0, 1, length.out = n)
@@ -101,7 +102,6 @@ results <- mclapply(sizes, function(n) {
         }) # Close replicate
     }) # Close phi
 }, mc.cores = 3) # Close sizes
-
 str(results)
 
 saveRDS(results, "simulation.rds", compress = FALSE)
@@ -134,10 +134,7 @@ aux <- ldply(lapply(results, function(x) {
     })
     ldply(out, .id = "phi")
 }), .id = "n")
-
 str(aux)
-
-
 
 # Where is NA's?
 aux$na <- apply(aux, 1, function(x) as.integer(sum(is.na(x)) != 0))
@@ -243,6 +240,47 @@ xyplot(param ~ bias | phi, groups = n, data = bias,
                     panel.abline(h = 1:5, col = "lightgray", lty = 2)
                 })
     )
+
+# Alternative graph (figure with all point is very large in size memory)
+bwplot(param ~ bias | phi, groups = n, data = bias,
+       pch = "|",
+       grid = TRUE,
+       as.table = TRUE,
+       layout = c(4, 1),
+       xlab = "Standardized Bias",
+       fill = "gray80",
+       col = "red",
+       strip = strip.custom(factor.levels = fl),
+       panel = panel.superpose,
+       scales = list(y = list(labels = yl)),
+       box.width = 0.05,
+       whisker.width = 0.4,
+       key = key,
+       par.settings = list(
+           plot.symbol = list(pch = 20, cex = 0.5, col = "gray70"),
+           box.rectangle = list(col = "gray70"),
+           box.umbrella=list(col = "gray70", lwd = 2)
+       ),
+       panel.groups = function(x, y, group.number, ...) {
+           my.panel.bwplot(x, (y - 0.05) + (group.number - 2.5) / 5, ...)
+           panel.abline(v = 0, lty = 2)
+       }) +
+    segplot(param ~ bias[, "lwr"] + bias[, "upr"] | phi,
+            centers = bias[, "fit"],
+            data = cidata,
+            draw = FALSE,
+            horizontal = TRUE,
+            layout = c(4, 1),
+            groups = n, gap = 0.3,
+            key = key,
[…]

### 2017-12-03 article.Rnw
Keep figures in their respective sections
Improve text by Clarice's suggestions and some corrections

@@
 \usepackage{float}
+\usepackage[section]{placeins}
 \usepackage{bm}
@@
 % Key-words
-\Keywords{ COM-Poisson, Overdispersion, Underdispersion, Count data,
-  Likelihood inference.  }
+\Keywords{ COM-Poisson, Count data, Likelihood inference,
+  Overdispersion, Underdispersion. }
 
@@
 related to the underline Poisson process generating the count data.
-When the time between events is not exponential distributed, the number
+When the time between events is not exponentially distributed, the number
 of events can be over or underdispersed, this process motivated the
@@
 statistics.  Suppose $Y_i$ is Poisson distributed then the maximum of
-$Y_i$ is underdispersed. (NEED REF).
+$Y_i$ is underdispersed~\citep{Steutel1989}.
 
@@
 overdispersion case the Poisson mixture models are widely applied.  One
-popular example of this approach is the negative-binomial model, where a
-gamma random effect is added to the expectation of the Poisson
-distribution.  However, other distributions can be used to represent the
+popular example of this approach is the negative-binomial model, where
+the expectation of the Poisson distribution is assumed gamma
+distributed.  However, other distributions can be used to represent the
 random effects.  For example the Poisson-Tweedie model~\citep{Bonat2017}
@@
 terms of the ratio between the probabilities of two consecutive events
-by adding an extra dispersion parameter~\cite{Sellers2010}.  Let $Y$ be
+by adding an extra dispersion parameter~\citep{Sellers2010}.  Let $Y$ be
 a COM-Poisson random variable, then
@@
 with large of $\lambda$, the sum is so huge it causes
-\textit{overflow}. \autoref{tab:convergenceZ} shows the values of the
+overflow. \autoref{tab:convergenceZ} shows the values of the
 normalizing constants using one thousand increments, that is,
@@
           command = c(
-              " \\multicolumn{7}{c}{$\\bm{\\lambda}$} \\\\\n",
+              "& \\multicolumn{6}{c}{$\\bm{\\lambda}$} \\\\\n",
               "$\\bm{\\nu}$ & 0.5 & 1 & 5 & 10 & 30 & 50 \\\\\n")
@@
 divergent$^{*}$ is a mathematically divergent series; and
-divergent$^{**}$ a numerically divergent aerie.
+divergent$^{**}$ a numerically divergent series.
 \end{table}
@@
 In the first line of \autoref{tab:convergenceZ} we have mathematically
-divergent series, because $\sum_{j=0}^\infty\lambda^y$, when
+divergent series, because $\sum_{j=0}^\infty\lambda^j$, when
 $\lambda \geq 1$, is divergent.  In other cases the series diverges
@@
   \label{eqn:mean-approx}
-  E(Y) \approx \lambda^{1/\nu} - \frac{\nu - 1}{2\nu} \qquad
+  \text{E}(Y) \approx \lambda^{1/\nu} - \frac{\nu - 1}{2\nu} \qquad
   \textrm{and} \qquad
-  \textrm{var}(Y) \approx \frac{\lambda^{1/\nu}}{\nu}\,,
+  \textrm{Var}(Y) \approx \frac{\lambda^{1/\nu}}{\nu}\,,
 \end{equation} which is particularly accurate for  $\nu \leq 1$ or
@@
 relationship can be approximate by
-$\frac{1}{\nu}E(Y)$. In~\autoref{reparametrization} we assess the
-accuracy of these approximations.
+$\frac{1}{\nu}\text{E}(Y)$. In~\autoref{reparametrization}, we assess
+the accuracy of these approximations.
 
@@
 regression model is $\log(\lambda_i) = \bm{x}_i^\top \bm{\beta}$ and the
-relationship between $E(Y_i)$ and $\bm{x}_i$ is modelled indirectly.
+relationship between E$(Y_i)$ and $\bm{x}_i$ is modelled indirectly.
 \citet{Huang2017} shows how to model directly the expectation of the
@@
 mean approximation (\autoref{eqn:mean-approx}). We introduced new
-parameter $\mu$, using the this approximation
+parameter $\mu$, using this approximation,
 \begin{equation}
@@
 
-In order to assess the accuracy of the approximations presented in
+In order to assess the accuracy of the moments approximations
 (\autoref{eqn:mean-approx}), \autoref{fig:approx-plot} presents the
 quadratic errors for the expectation (a) and variance (b).  The
-quadratic errors were obtained by $[\mu - E(Y)]^2$ for the expectation
-and by $[ \mu \exp(-\phi) - \text{var}(Y)]^2$ for the variance. In both
-cases $E(Y)$ and var$(Y)$ were computed numerically. The dotted lines
-represent the border between the regions $\nu \leq 1$ and
-$\lambda > 10^\nu$, in the $\mu$ and $\phi$ scale.
+quadratic errors were obtained by $[\mu - \text{E}(Y)]^2$ for the
+expectation and by $[ \mu \exp(-\phi) - \text{Var}(Y)]^2$ for the
+variance. In both cases $\text{E}(Y)$ and Var$(Y)$ were computed
+numerically. The dotted lines represent the border between the regions
+$\nu \leq 1$ and $\lambda > 10^\nu$, in the $\mu$ and $\phi$ scale.
 
@@
 $\phi$. Interestingly, the errors are larger to negative values of
-$\phi$ and present no clear relation with $\mu$ opposite to the regions
+$\phi$ and present no clear relation with $\mu$, opposite to the regions
 gives by \citet{Shmueli2005} ($\phi \leq 0$ and
@@
 \end{equation} where $\mu > 0$. We denote this distribution as
-COM-Poisson$_\mu$. In \autoref{fig:pmf-cmp} we show the shapes of
+COM-Poisson$_\mu$. In \autoref{fig:pmf-cmp}, we show the shapes of
 COM-Poisson$_\mu$ distribution.
@@
 \begin{equation*}
-\text{DI} = \frac{\text{var}(Y)}{\text{E}(Y)}, \quad
+\text{DI} = \frac{\text{Var}(Y)}{\text{E}(Y)}, \quad
 \text{ZI} = 1 + \frac{\log \Pr(Y = 0)}{\text{E}(Y)}
@@
 describe count distributions. Regarding the COM-Poisson$_\mu$
-distribution, we presents in \autoref{fig:indexes-plot} the relationship
+distribution, we present in \autoref{fig:indexes-plot} the relationship
 between mean and variance (a), the dispersion and zero-inflation indexes
@@
 <<indexes-plot, fig.width=9.5, fig.height=4, fig.cap="Indexes for COM-Poisson distribution.
-Mean and variance relation ship (a), dispersion, zero-inflation and heavy-tail indexes (b--d) for different parameter values.
+Mean and variance relationship (a), dispersion, zero-inflation and heavy-tail indexes (b--d) for different parameter values.
 Dotted lines represents the Poisson special case.">>=
@@
               ylab = "",
-              # ylab = expression(var(Y)),
+              # ylab = expression(Var(Y)),
               scales = list(y = list(rot = 90)),
@@
 two forms of the COM-Poisson regression model based on the maximum
-likelihood method. The log-likelihood function for a set of
-independent observations $y_i$, $i=1,2,\ldots,n$ for COM-Poisson$_\mu$
-from the COM-Poisson$_\mu$ distribution has the following form,
+likelihood method. The log-likelihood function for a set of independent
+observations $y_i$, $i=1,2,\ldots,n$ from the COM-Poisson$_\mu$
+distribution has the following form,
 \begin{equation}
@@
     \mu_i + \frac{e^\phi - 1}{2e^\phi} \right )^{je^\phi}
-  \frac{1}{(j!)^{e^\phi}} \right ]
+  \frac{1}{(j!)^{e^\phi}} \right ].
 \end{equation}
 The evaluation of the log-likelihood function for each observation
-involves the computation of the infinite series in
-\autoref{eqn:infseries}.  Thus, the fitting procedure is
-computational expansive for regions of the parameter space where the
-convergence of the infinite sum is slow.
+involves the computation of the infinite series
+(\autoref{eqn:infseries}).  Thus, the fitting procedure is computational
+expensive for regions of the parameter space where the convergence of
+the infinite sum is slow.
 
@@
 obtained in closed forms, we adopted the \texttt{BFGS}
-algorithm~\cite{Nocedal1995} as implemented in the function
-\texttt{optim()} for the statistical software \texttt{R}. Standard
-errors for the regression coefficients are obtained based on the
-observed information matrix $\mathcal{I}(\bm{\theta})$, where
-$\mathcal{I}(\bm{\theta}) = -\mathcal{H}(\bm{\theta})$ (Hessian matrix)
+algorithm~\citep{Nocedal1995} as implemented in the function
+\texttt{optim()} for the statistical software \texttt{R}
+\citep{Rcore2017}. Standard errors for the regression coefficients are
+obtained based on the observed information matrix
+$\mathcal{I}(\bm{\theta})$, where
+$\mathcal{I}(\bm{\theta}) = -\mathcal{H}(\bm{\theta})$ (hessian matrix)
 is computed numerically. Confidence intervals for $\hat{\mu}_i$ are
@@
 second-moment assumption that allows more flexibility to model. In this
-case the variance of the response variable is fixed by a additional
-parameter $\sigma$, $\textrm{var}(Y_i)=\sigma \mu_i$. These models are
-fitted in software R \citep{Rcore2017} by function \texttt{glm(...,
-  family = quasipoisson)}.
+case the variance of the response variable is fixed by an additional
+parameter $\sigma$, $\textrm{Var}(Y_i)=\sigma \mu_i$. These models are
+fitted in software \texttt{R} using the function \texttt{glm(..., family
+  = quasipoisson)}.
 
@@
 
-In this section we conducted a simulation study to assess the properties
+In this section we performed a simulation study to assess the properties
 of the maximum likelihood estimators as well as the flexibility of the
@@
 covariate~($\bm{x}_2$) was generated as a sequence of three values each
-one repeated $3/n$ times (rounding up when required), where $n$ denotes
+one repeated $n/3$ times (rounding up when required), where $n$ denotes
 the sample size. Thus, the expectation of the COM-Poisson random
@@
 different sample sizes: $50$, $100$, $300$ and $1000$; generating $1000$
-data sets in each case.  In \autoref{fig:bias-plot} we show the bias of
+data sets in each case.  In \autoref{fig:bias-plot}, we show the bias of
 the estimators for each simulation scenario (combination between values
@@
 levels, both the average bias and standard errors tend to $0$ as the
-sample size is increased. The estimators for the regression parameters
-are unbiased, consistency and their empirical distributions are
+sample size increases. The estimators for the regression parameters
+are unbiased, consistent and their empirical distributions are
 symmetric. For the dispersion parameter, the estimator is asymptotically
-unbiased, in small samples the parameter is overestimated and the
+unbiased; in small samples the parameter is overestimated and the
 empirical distribution is slightly right-skewed.
-\autoref{fig:coverage-plot} presents the empirical coverage rate of the
-asymptotic confidence intervals.
-
-<<bias-plot, cache=TRUE, fig.height=6.5, fig.width=8.5, fig.cap="Distributions of standardized bias (gray points) and average with confidence intervals (black segments) by different sample sizes and dispersion levels.">>=
+
+<<bias-plot, cache=TRUE, fig.height=6.5, fig.width=8.5, fig.cap="Distributions of standardized bias (gray box-plots) and average with confidence intervals (black segments) by different sample sizes and dispersion levels.">>=
 
@@
 
-The coverage rates in \autoref{fig:coverage-plot} show that for the
+\autoref{fig:coverage-plot} presents the empirical coverage rate of the
+asymptotic confidence intervals. The results show that for the
 regression parameters the empirical coverage rates are close to the
@@
 simulation scenarios. For the dispersion parameter the empirical
-coverage rates are slightly lower than the nominal level, however, they
+coverage rates are slightly lower than the nominal level; however, they
 become closer to the nominal level for large samples.  The worst
@@
 replicates. This data set was analyzed by~\citet{Zeviani2014} using the
-\textit{Gamma-Count} distribution.
+Gamma-Count distribution.
 
 Following~\citet{Zeviani2014}, the linear predictor is given by
-$$\log(\mu_{ij}) = \beta_0 + \beta_{1j} \textrm{def}_i + \beta_{2j}
-\textrm{def}_i^2$$
+$$\log(\mu_{ij}) = \beta_0 + \beta_{1j} \texttt{def}_i + \beta_{2j}
+\texttt{def}_i^2$$
 where $\mu_{ij}$ is the expected number of cotton bolls for the $i$-th
@@
 \caption{Parameter estimates (Est) and ratio between estimate and
-  standard error (SE) for the four model strategies for the cotton
-  experiment.}
+  standard error (SE) for the four model strategies for the analysis
+  of the cotton experiment.}
 \label{tab:coef-cotton}
@@
 that COM-Poisson model is a full parametric approach, i.e. there is a
-probability distribution associate to counts. On the other hand, the
+probability distribution associated to the counts. On the other hand, the
 quasi-Poisson model is a specification based only on second-moments
@@
 results from the COM-Poisson$_{\mu}$ model are consistent with those
-from the Gamma-Count model~\cite{Zeviani2014},
-Poisson-Tweedie~\cite{Bonat2017} and the alternative parametrization of
-the COM-Poisson distribution proposed by~\cite{Huang2017}.  In all
+from the Gamma-Count model~\citep{Zeviani2014},
+Poisson-Tweedie~\citep{Bonat2017} and the alternative parametrization of
+the COM-Poisson distribution proposed by~\citet{Huang2017}.  In all
 strategies the models indicated underdispersion and significant effects
@@
 \centering
-\caption{Empirical correlations between $\hat{\phi}$ and $\hat{\bm{\beta}}$ for the two parametrizations of COM-Poisson model fit to underdispersed data.}
+\caption{Empirical correlations between $\hat{\phi}$ and
+  $\hat{\bm{\beta}}$ for the two parametrizations of COM-Poisson model
+  fit to underdispersed data.}
 \label{tab:corr-cotton}
@@
 
-\subsection{Soil moisture and potassium doses on soy bean culture}
+\subsection{Soil moisture and potassium doses on soybean culture}
 \label{case-soybean}
 
-The second example is an experiment in a completely randomized block
-with treatments in a $5\times 3$ factorial arrangement about soybean
-(\emph{Glicine Max}). The aim of this study was to evaluate the effects
+The second example is a $5\times 3$ factorial experiment in a randomized
+complete block design. The aim of this study was to evaluate the effects
 of potassium doses (\texttt{K}) applied to soil (0, 0.3, 0.6, 1.2 and
 1.8 $\times$ 100mg dm$^{-3}$) and soil moisture (\texttt{umid}) levels
-(37.5, 50, 62.5\%) on the soybean production. The experiment was carried
-out in a greenhouse, in pots with two plants, and the count variable
-measured was the number of grains per pot~\citep{Serafim2012}.
-\autoref{fig:desc-soy} (left) shows the number of grains recorded for
-each combination of potassium dose and moisture level, it is important
-to note the indication of a quadratic effect of the potassium
-levels. Most points in the sample variance \textit{versus} sample means
-dispersion diagram (right) are above the identity line, suggesting
-overdispersion (block effect not yet removed).
+(37.5, 50, 62.5\%) on the soybean (\emph{Glicine Max}) production. The
+experiment was carried out in a greenhouse, in pots with two plants, and
+the count variable measured was the number of grains per
+pot~\citep{Serafim2012}.  \autoref{fig:desc-soy} (left) shows the number
+of grains recorded for each combination of potassium dose and moisture
+level, it is important to note the indication of a quadratic effect of
+the potassium levels as shown by smoothing curves. Most points in the
+sample variance \textit{versus} sample means dispersion diagram (right)
+are above the identity line, suggesting overdispersion (block effect not
+yet removed).
 
@@
 
+co2 <- coef(m2C2)
+
 # Numbers of calls to loglik and numerical gradient
@@
 
-<<desc-soy, fig.pos="!ht", fig.height=4, fig.width=8, fig.cap="Number of grains per pot for each potassium dose and moisture level (left) and sample mean agains sample variance of the five replicates for each experimental treatment (right).">>=
+<<desc-soy, fig.pos="!ht", fig.height=4, fig.width=8, fig.cap="Number of grains per pot for each potassium dose and moisture level (left) and sample mean agains sample variance of the five replicates for each experimental treatment (right).
+Solid lines are the smoothing curves on the left and the least of squares curve on the right.">>=
 
@@
 
-For the analysis of this data set we proposed based on descriptive
-analysis \autoref{fig:desc-soy}, the linear predictor specified as a
-linear doses for each humidity level and a global quadratic dose effect,
-i.e
+For the analysis of this data set based on descriptive analysis
+(\autoref{fig:desc-soy}), we proposed the following linear predictor
 $$
@@
 $$
-with $i=$1: block II, 2: block III, 3: block IV e 4: block V; $j=$1:
+where $i=$1: block II, 2: block III, 3: block IV e 4: block V; $j=$1:
 50\% e 2: 62.5\%; and $k=$1: 0.0, 2: 0.3, 3: 0.6, 4: 1.2, 5: 1.8 100mg
-dm$^{-3}$, where $\gamma_i$ is the effect of $i$-th block, $\tau_j$ is
-the effect of $j$-th moisture level and $\beta_{3j}$ is the first order
-potassium effect (\texttt{K}) for the $j$-th moisture level
+dm$^{-3}$,
+
+where $\gamma_i$ is the effect of $i$-th block ($i=$1: block II, 2:
+block III, 3: block IV and 4: block V), $\tau_j$ is the effect of $j$-th
+moisture level ($j=$1: 50\% and 2: 62.5\%) and $\beta_{3j}$ is the first
+order potassium effect (\texttt{K}) for the $j$-th moisture level
 (\texttt{umid}). \autoref{tab:coef-soy} presents the estimates, ratio
-between estimate and standard error and goodness-of-fit measures for
-the alternative models.
-
-The results in \autoref{tab:coef-soy} have similar interpretations to
-the results presented in \autoref{case-cotton}. The two parametrization
+between estimate and standard error and goodness-of-fit measures for the
+alternative models.
+
+The results in \autoref{tab:coef-soy} show that the two parametrization
 of COM-Poisson model presented very similar goodness-of-fit measures and
@@
 $\Sexpr{lrt.ngra[2, 4]}$, indicating that $\phi$ is significantly
-different from zero. Concerning to the regression parameters, the
-similarities between the models are analogous to the previous
-Section. Both models indicate effects of block, potassium dose and
-moisture level, however the Poisson model indicates the effects with
-greater significance, because it does not fit the extra variability.
+different from zero. The estimate of $\phi$ ($\Sexpr{co2[1]}$) indicates
+overdispersion, corroborating to the descriptive analysis. Concerning to
+the regression parameters, the similarities between the models are
+analogous to the previous section. Both models indicate effects of
+block, potassium dose and moisture level, however the Poisson model
+indicates the effects with greater significance, because it does not fit
+the extra variability.
 
@@
 \caption{Parameter estimates (Est) and ratio between estimate and
-  standard error (SE) for the four model strategies for the soybean
-  experiment.}
+  standard error (SE) for the four model strategies for the analysis of
+  the soybean experiment.}
 \label{tab:coef-soy}
@@
 the better behaviour of the log-likelihood function as well as better
-initial values obtained from the Poisson fit. In \autoref{tab:corr-soy}
+initial values obtained from the Poisson fit. In \autoref{tab:corr-soy},
 we present the empirical correlation between the regression and
@@
 \centering
-\caption{Empirical correlations between $\hat{\phi}$ and $\hat{\bm{\beta}}$ for the two parametrizations of COM-Poisson model fit to overdispersed data.}
+\caption{Empirical correlations between $\hat{\phi}$ and
+  $\hat{\bm{\beta}}$ for the two parametrizations of COM-Poisson model
+  fit to overdispersed data.}
 \label{tab:corr-soy}
@@
 
-\subsection{Assessing toxicity of pollutants in aquatic systems}
+\subsection{Assessing toxicity of nitrofen in aquatic systems}
+
+Nitrofen is a herbicide that was used extensively for the control of
+broad-leaved and grass weeds in cereals and rice. Although it is
+relatively non-toxic to adult mammals, nitrofen is a significant
+tetragen and mutagen. It is also acutely toxic and reproductively toxic
+to cladoceran zooplankton. Nitrofen is no longer in commercial use in
+the U.S., having been the first pesticide to be withdrawn due to
+tetragenic effects~\citep{Bailer1994}.
 
@@
 concentration of nitrofen ($0, 0.8, 1.6, 2.35$ e $3.10$ mug$/10^2$litre)
-(\texttt{dose}). Then the number of total live offspring was recorded
-\cite{Bailer1994}.
+(\texttt{dose}). Then the number of live offspring was recorded.
 
@@
 
-For this data set we proposed three linear predictors. We compared the
-performance of models using the likelihood ratio tests. The linear
-predictors are
+For this data set we proposed three linear predictors,
 \begin{center}
@@
 \centering \small
-\caption{Model fit measures and comparisons between predictors and models.}
+\caption{Model fit measures and comparisons between predictors and
+  models fitted to the nitrofen data.}
 \label{tab:anova-ovos}
@@
 concentration.  Considering this predictor, there is an evidence of
-equidispersed counts, $\phi$ estimates of the COM-Poisson close to zero
-and $\sigma$ of quasi-Poisson close to one.  It is interesting to note
+equidispersed counts, $\phi$ estimate of the COM-Poisson is close to zero
+and $\sigma$ of quasi-Poisson is close to one.  It is interesting to note
 that if we omit the high order effects the models show evidence of
-overdispersion. This exemplifies the discussion about causes of
+overdispersion. This exemplifies the discussion on the causes of
 overdispersion made in \autoref{introduction}.  We can also note that
 the quasi-Poisson approach, although robust to equidispersion
-assumption, shows a higher descriptive levels ($p$-values) than
+assumption, shows higher descriptive levels ($p$-values) than
 parametric models, that is, the tests under parametric models are more
@@
 
-In \autoref{tab:coef-ovos} we present the estimates of the regression
+In \autoref{tab:coef-ovos}, we present the estimates of the regression
 parameters considering the cubic dose effect (third predictor). The
@@
 estimates of the original COM-Poisson model are comparable to the others
-models. It occurs because it is the special case $\phi = 0$ which
-implies $\lambda = \mu$. \autoref{fig:pred-ovos} shows the number of
-live off-springs observed and fitted along with confidence intervals for
-all model strategies adopted.
+models. It occurs because it is the special case, where $\phi = 0$, which
+implies $\lambda = \mu$.
 
@@
 \caption{Parameter estimates (Est) and ratio between estimate and
-  standard error (SE) for the four model strategies for the nitrofen
-  experiment.}
+  standard error (SE) for the four model strategies for the analysis of
+  the nitrofen experiment.}
 \label{tab:coef-ovos}
@@
 
-The fitted values and confidence intervals in \autoref{fig:pred-ovos}
-come to have a full overlay. It shows that the estimation of the extra
-dispersion parameter does not affect the estimation of the regression
-coefficients in the case of equidispersed counts.
+\autoref{fig:pred-ovos} shows the number of live off-springs observed
+and fitted along with confidence intervals for all model strategies
+adopted.  The fitted values and confidence intervals come to have a full
+overlay. It shows that the estimation of the extra dispersion parameter
+does not affect the estimation of the regression coefficients in the
+case of equidispersed counts.
 
@@
 of the regression coefficients in terms of the expectation of the
-response variable as usual in the generalize linear models
+response variable as usual in the generalized linear models
 context. Thus, it is possible to compare the results of the COM-Poisson
@@
 quadratic approximation errors.  The results showed that the
-approximations are accurate for a large part of the parameter space of
-the parameters $\mu$ and $\phi$, which in turn support our
-reparametrization.  We carried out a simulation study to assess the
-properties of the reparametrized COM-Poisson model to deal with
-different levels of dispersion as well as the properties of the maximum
-likelihood estimators. The results of our simulation study suggested
-that the maximum likelihood estimators of the regression and dispersion
-parameters are unbiased and consistent.  The empirical coverage rates of
-the confidence intervals computed based on the asymptotic distribution
-of the maximum likelihood estimators are close to the nominal level for
-sample size greater than $100$.  The worst scenario is when we have
-small sample sizes and strong overdispersed counts.  In general, we
-recommend the use of the asymptotic confidence intervals for
-computational simplicity.
+approximations are accurate for a large part of the parameter space,
+which in turn support our reparametrization.  We carried out a
+simulation study to assess the properties of the reparametrized
+COM-Poisson model to deal with different levels of dispersion as well as
+the properties of the maximum likelihood estimators. The results of our
+simulation study suggested that the maximum likelihood estimators of the
+regression and dispersion parameters are unbiased and consistent.  The
+empirical coverage rates of the confidence intervals computed based on
+the asymptotic distribution of the maximum likelihood estimators are
+close to the nominal level for sample size greater than $100$.  The
+worst scenario is when we have small sample sizes and strong
+overdispersed counts.  In general, we recommend the use of the
+asymptotic confidence intervals for computational simplicity.
 
 The data analyses have shown that the COM-Poisson regression model is a
-suitable choice to deal with non equidispersed count data.  The observed
+suitable choice to deal with dispersed count data.  The observed
 empirical correlation between the regression and dispersion parameter
@@
 including simulation studies to assess the model robustness against
-model misspecification and to asses the theoretical approximations for
+model misspecification and to assess the theoretical approximations for
 $Z(\lambda, \nu)$ (or $Z(\mu,\phi))$. Another simple extension of the

### 2018-01-27 codes/01-simulation.R
Add summary of the covariance matrix obtained with the simulations

@@ -13,6 +13,13 @@ library(plyr)
 library(tidyr)
 library(bbmle)
 
+#-------------------------------------------
+# Random generation for the COM-Poisson distribution
+rcmp <- Vectorize(FUN = function(n, mu, phi) {
+    par <- mu2lambda(mu, phi)
+    with(par, compoisson::rcom(n = n, lambda, nu))
+}, vectorize.args = "mu")
+
 #-----------------------------------------------------------------------
 # Simulation study
 B <- 1000
@@ -61,13 +68,6 @@ xy2 <- xyplot(di ~ x1 | phi, groups = x2,
 
 print(xy1, split = c(1, 1, 2, 1), more = TRUE)
 print(xy2, split = c(2, 1, 2, 1), more = FALSE)
-
-#-------------------------------------------
-# Random generation for the COM-Poisson distribution
-rcmp <- Vectorize(FUN = function(n, mu, phi) {
-    par <- mu2lambda(mu, phi)
-    with(par, compoisson::rcom(n = n, lambda, nu))
-}, vectorize.args = "mu")
 
 #-------------------------------------------
 # Simulation
@@ -398,3 +398,67 @@ useOuterStrips(
        factor.levels =  yl
     )
 )
+
+#-----------------------------------------------------------------------
+# Obtain the covariance between dispersion and regression parameters
+aux <- ldply(
+    lapply(results, function(x) {
+    ind <- names(x); names(ind) <- ind
+    out <- lapply(ind, function(y) {
+        vcovs <- lapply(x[[y]], "[[", "vcov")
+        index <- vapply(lapply(vcovs, is.na), sum, integer(1))
+        vcovs <- vcovs[!index]
+        corrs <- lapply(vcovs, cov2cor)
+        covs <- do.call(rbind, lapply(vcovs, "[", 1, 2:5))
+        cors <- do.call(rbind, lapply(corrs, "[", 1, 2:5))
+        ldply(list("cov" = covs, "cor" = cors), .id = "scale")
+    })
+    ldply(out, .id = "phi")
+}), .id = "n")
+str(aux)
+
+# Organize the results
+dacov <- gather(aux, param, value, b0:b22, factor_key = TRUE)
+dacov$phi <- ordered(dacov$phi,
+                     c("phi=-1.6", "phi=-1", "phi=0", "phi=1.8"))
+str(dacov)
+
+# Distributions of covariance/correlations
[…]

### 2018-01-27 article.Rnw
Improve text by Hinde's suggestions and some corrections

@@
   count data.  It is a member of the exponential family of distributions
-  and has the Poisson and geometric distributions as special cases as
+  and has the Poisson and geometric distributions as special cases, as
   well as the Bernoulli distribution as a limiting case.  In spite of
   the nice properties of the COM-Poisson distribution, its location
-  parameter does not correspond to the expectation, which in turn
-  becomes challenging the interpretation of the regression models
-  specified using such a distribution. In this paper, we propose a
-  straightforward reparametrization of the COM-Poisson distribution
-  based on an approximation to the expectation of this distribution. The
-  main advantage of our new parametrization is the straightforward
+  parameter does not correspond to the expectation, which complicates
+  the interpretation of regression models specified using this
+  distribution. In this paper, we propose a straightforward
+  reparametrization of the COM-Poisson distribution based on an
+  approximation to the expectation of this distribution. The main
+  advantage of our new parametrization is the straightforward
   interpretation of the regression coefficients in terms of the
@@
   finite sample properties of the maximum likelihood estimators. The
-  results from our simulation study show that the maximum likelihood
+  results from our simulation study show that the maximum likeli-hood
   estimators are unbiased and consistent for both regression and
@@
   The study of distribution properties through a consideration of
-  dispersion, zero-inflated and heavy tail indices, and the results of
-  data analysis show the flexibility over standard
+  dispersion, zero-inflated and heavy tail indices, together with the
+  results of data analysis show the flexibility over standard
   approaches. Therefore, we encourage the application of the new
@@
 and represent the number of times an event occurs in the observation
-unit. This kind of data is common in crop sciences such as the number of
-grains produced by a plant, number of fruits produced by a tree, number
-of insects captured by a trap, to cite but a few.  Since the seminal
-paper of \citet{Nelder1972} where the class of the generalized linear
-models (GLMs) was introduced, the analysis of count data often employs
-the Poisson regression model. This model provides a suitable strategy
-for the analysis of count data and an efficient Newton scoring algorithm
-can be used for fitting the model.
+period. This kind of data is common in crop sciences, such as the number
+of grains produced by a plant, number of fruits produced by a tree,
+number of insects captured by a trap, to cite but a few.  Since the
+seminal paper of \citet{Nelder1972} where the class of the generalized
+linear models (GLMs) was introduced, the analysis of count data often
+employs the Poisson regression model. This model provides a suitable
+strategy for the analysis of count data and an efficient Newton scoring
+algorithm can be used for fitting the model.
 
@@
 expectation and variance of the count random variable.  This restriction
-in the relationship between the expectation and variance induced by the
+on the relationship between the expectation and variance induced by the
 Poisson distribution is referred as equidispersion.  However, in
 practical data analysis such a restriction can be unsuitable, since the
-observed data can present both the variance smaller or larger than the
+observed data can present  variance both smaller or larger than the
 expectation, leading to the cases of under and overdispersion,
 respectively.  The main problem of the application of the Poisson
-regression model to non equidispersed count data is that the standard
+regression model to non-equidispersed count data is that the standard
 errors associated with the regression coefficients are inconsistently
@@
 may occur due to the absence of relevant covariates, heterogeneity of
-sampling units, different domain size not considered in the analysis and
-excess of zeros~\citep{Hinde1998}.  The case of underdispersion is less
-report in the literature, however, it has been of increasing interest in
-the statistical community. The processes that reduce the variability are
-not as well-known as those leading to extra variability.  For the same
-reason, there are a few approaches to deal with underdispersed count
-data. The explanatory mechanisms leading to underdispersion may be
-related to the underline Poisson process generating the count data.
-When the time between events is not exponentially distributed, the number
-of events can be over or underdispersed, this process motivated the
-class of duration dependence models~\citep{Winkelmann1995}.  Another
-possibility to explain the underdispersion phenomenon is order
-statistics.  Suppose $Y_i$ is Poisson distributed then the maximum of
-$Y_i$ is underdispersed~\citep{Steutel1989}.
+sampling units, different observational periods/regions not considered
+in the analysis, and excess of zeros~\citep{Hinde1998}.  The case of
+underdispersion is less report in the literature, however, it has been
+of increasing interest in the statistical community. The processes that
+reduce the variability are not as well-known as those leading to extra
+variability.  For this reason, there are few approaches to deal with
+underdispersed count data. The explanatory mechanisms leading to
+underdispersion may be related to the underlying stochastic process
+generating the count data.  When the time between events is not
+exponentially distributed, the number of events can be over or
+underdispersed, this process motivated the class of duration dependence
+models~\citep{Winkelmann1995}.  Another possible explanation of
+underdispersion is when the responses correspond to order statistics of
+component observations, such as maxima of Poisson distributed
+counts~\citep{Steutel1989}.
 
 The strategies for constructing alternative count distributions are
-related with the causes of the non-equidispersion. In special to the
-overdispersion case the Poisson mixture models are widely applied.  One
+related with the causes of the non-equidispersion. Specfically for the
+overdispersion case Poisson mixture models are widely applied.  One
 popular example of this approach is the negative-binomial model, where
-the expectation of the Poisson distribution is assumed gamma
+the expectation of the Poisson distribution is assumed to be gamma
 distributed.  However, other distributions can be used to represent the
-random effects.  For example the Poisson-Tweedie model~\citep{Bonat2017}
-and its special cases as the Poisson inverse-Gaussian and Neyman-Type A
-consider that the random effects are Tweedie, inverse Gaussian and
-Poisson distributed, respectively.  The Gamma-Count distribution assumes
-a gamma distribution for the time between events, thus it can deal with
-underdispersed as well as overdispersed count
-data~\citep{Zeviani2014}. The COM-Poisson distribution is obtained by a
-generalization of the Poisson distribution allowing for a non-linear
-decrease in the ratios of successive probabilities~\citep{Shmueli2005}.
+random variation.  For example the Poisson-Tweedie
+model~\citep{Bonat2017} and its special cases as the Poisson
+inverse-Gaussian and Neyman-Type A assume that the random effects are
+Tweedie, inverse Gaussian and Poisson distributed, respectively.  The
+Gamma-Count distribution assumes a gamma distribution for the time
+between events, thus it can deal with underdispersed as well as
+overdispersed count data~\citep{Zeviani2014}. The COM-Poisson
+distribution is obtained by a generalization of the Poisson distribution
+allowing for a non-linear decrease in the ratios of successive
+probabilities~\citep{Shmueli2005}.
 
 The COM-Poisson distribution is a member of the exponential family and
-it has the Poisson and geometric distributions as special cases as well
+it has the Poisson and geometric distributions as special cases, as well
 as the Bernoulli distribution as a limiting case. It can deal with both
@@
 traffic crash data, \citet{Sellers2010} for the modelling of airfreight
-breakage and book purchases and \citet{Huang2017} to the analysis of
-attendance data, takeover birds and cotton bolls count.  The main
+breakage and book purchases, and \citet{Huang2017} to the analysis of
+attendance data, takeover birds and cotton boll counts.  The main
 disadvantage of the COM-Poisson regression model as presented
 in~\citet{Sellers2010} is that its location parameter does not
-correspond to the expectation of the distribution, which in turn becomes
-the interpretation of the regression coefficients challenging and no
-comparable with standard approaches as the Poisson and negative binomial
-regression models. \citet{Huang2017} proposed a mean-parametrization of
-the COM-Poisson distribution in order to avoid such an issue.  In this
-approach the mean parameter is obtained by solving an non-linear
-equation defined as an infinite sum.  Consequently, it is
-computationally demanding and liable to numerical problems.
+correspond to the expectation of the distribution, which complicates the
+interpretation of regression models and means that they are not
+comparable with standard approaches such as the Poisson and negative
+binomial regression models. \citet{Huang2017} proposed a
+mean-parametrization of the COM-Poisson distribution in order to avoid
+such an issue.  In this approach the mean parameter is obtained by
+solving an non-linear equation defined as an infinite sum.
+Consequently, it is computationally demanding and liable to numerical
+problems.
 
@@
 function is written in terms of $\mu$ and $\phi$, where $\mu$ is the
-expectation and $\phi$ is the dispersion parameter. In contrast to the
+expectation and $\phi$ is a dispersion parameter. In contrast to the
 original parametrization, the proposed parametrization leads to
 regression coefficients directly associated with the expectation of the
-response variable as usual in the context of generalized linear models.
+response variable, as usual in the context of generalized linear models.
 Consequently, the obtained regression coefficients are comparable with
-the ones obtained by standard approaches as the Poisson and negative
-binomial regression models.  Furthermore, our novel COM-Poisson
+the ones obtained by standard approaches, such as the Poisson and
+negative binomial regression models.  Furthermore, our novel COM-Poisson
 parametrization is simpler than the strategy proposed
-by~\cite{Huang2017}, since it does not involve any numerical method for
-solving non-linear equations.
+by~\cite{Huang2017}, since it does not require any numerical method for
+solving non-linear equations, and we show the attractive properties like
+the orthogonality between dispersion and regression parameters and
+consistency and asymptotic normality of the maximum likelihood
+estimators are retained.
 
@@
 considered in the~\autoref{reparametrization}. In
-the~\autoref{estimation-and-inference} we present the estimation and
+the~\autoref{estimation-and-inference} we present estimation and
 inference for the novel COM-Poisson regression model based on the
 likelihood paradigm. The properties of the maximum likelihood estimators
-are assess in~\autoref{simulation-study} through simulation studies.  We
-illustrate the application of the new COM-Poisson regression model
-through the analysis of three data sets. We provide an \texttt{R}
-implementation of the COM-Poisson and reparameterized COM-Poisson
-regression models as well as the analyzed data sets in the supplementary
+and the orthogonality property are assessed
+in~\autoref{simulation-study} through simulation studies.  We illustrate
+the application of the new COM-Poisson regression model through the
+analysis of three data sets. We provide an \texttt{R} implementation of
+the COM-Poisson and reparameterized COM-Poisson regression models as
+well as the analyzed data sets in the supplementary
 material.\footnote{Available on
@@
 $$\frac{\Pr(Y=y-1)}{\Pr(Y=y)} = \frac{y^\nu}{\lambda}$$ while for the
-Poisson distribution this ratio is $\frac{y}{\lambda}$. It allows the
-COM-Poisson distribution deals with non equidispersed count data.  The
-probability mass function of the COM-Poisson distribution is given by
+Poisson distribution this ratio is $\frac{y}{\lambda}$ corresponding to
+$\nu=1$.  It allows the COM-Poisson distribution deals with
+non-equidispersed count data.  The probability mass function of the
+COM-Poisson distribution is given by
 \begin{equation}
@@
 and $\lambda \geq 1$, but numerically for small values of $\nu$ combined
-with large of $\lambda$, the sum is so huge it causes
-overflow. \autoref{tab:convergenceZ} shows the values of the
-normalizing constants using one thousand increments, that is,
+with large values of $\lambda$, the sum is so huge it causes
+overflow. \autoref{tab:convergenceZ} shows the values of the normalizing
+constants using one thousand increments, that is,
 $\sum_{j=0}^{1000}\lambda^j/(j!)^\nu$ for different values of $\lambda$
@@
 In the first line of \autoref{tab:convergenceZ} we have mathematically
-divergent series, because $\sum_{j=0}^\infty\lambda^j$, when
-$\lambda \geq 1$, is divergent.  In other cases the series diverges
-numerically, due to the computational storage limitation.  For both
-divergences it is impossible to compute probabilities, therefore, this
-acts as a restriction in the parameter space.
+divergent series, because $\sum_{j=0}^\infty\lambda^j$ is divergent when
+$\lambda \geq 1$.  In other cases the series diverges numerically, due
+to the computational storage limitation.  For both forms of divergence
+it is impossible to compute probabilities, therefore, this acts as a
+restriction on the parameter space.
 
@@
 \autoref{eqn:pmf-cmp}, \Citeauthor{Huang2017} proposes that the
-parameter $\lambda$ as a function of $\mu$ and $\nu$, given by the
+parameter $\lambda$ as a function of $\mu$ and $\nu$, is given by the
 solution to
@@
 The proposed reparametrization of COM-Poisson models is based on
-mean approximation (\autoref{eqn:mean-approx}). We introduced new
+the mean approximation (\autoref{eqn:mean-approx}). We introduced a new
 parameter $\mu$, using this approximation,
@@
 convenience, thus $\phi = \log(\nu)$, $\phi \in \mathbb{R}$. The
-interpretation of $\phi$ is the same as the $\nu$, but in another
+interpretation of $\phi$ is the same as the $\nu$, but on another
 scale. For $\phi < 0$ and $\phi > 0$ we have the overdispersed and
 underdispersed cases, respectively.  When $\phi=0$ we have Poisson
-distribution as the special case.
+distribution as a special case.
 
@@
 
-In order to assess the accuracy of the moments approximations
+In order to assess the accuracy of the moment approximations
 (\autoref{eqn:mean-approx}), \autoref{fig:approx-plot} presents the
-quadratic errors for the expectation (a) and variance (b).  The
+quadratic errors for (a) expectation  and (b) variance.  The
 quadratic errors were obtained by $[\mu - \text{E}(Y)]^2$ for the
@@
 
-<<approx-plot, fig.height=3, fig.width=6, out.width=".8\\textwidth", fig.cap="Quadratic errors for the approximation of the expectation (a) and variance (b).
+<<approx-plot, fig.height=3, fig.width=6, out.width=".8\\textwidth", fig.cap="Quadratic errors for the approximation of the (a) expectation and (b) variance.
 Dotted lines represent the restriction for suitable approximations given by \\cite{Shmueli2005}.">>=
@@
 \Sexpr{max(grid[["eva"]])} and it occurs for negative values of
-$\phi$. Interestingly, the errors are larger to negative values of
-$\phi$ and present no clear relation with $\mu$, opposite to the regions
-gives by \citet{Shmueli2005} ($\phi \leq 0$ and
+$\phi$. Interestingly, the errors are larger for negative values of
+$\phi$ and present no clear relation with $\mu$, as opposed to the
+regions gives by \citet{Shmueli2005} ($\phi \leq 0$ and
 $\mu > 10 - \frac{\exp(\phi) - 1}{2\exp(\phi)}$).
@@
 describe count distributions. Regarding the COM-Poisson$_\mu$
-distribution, we present in \autoref{fig:indexes-plot} the relationship
-between mean and variance (a), the dispersion and zero-inflation indexes
-(b--c) for different values of $\mu$ and $\phi$ and heavy-tail index
-(d) for $\mu=25$ and different values of $y$ and $\phi$.
+distribution, in \autoref{fig:indexes-plot} we present the relationship
+between (a) mean and variance, (b--c) the dispersion and zero-inflation
+indexes for different values of $\mu$ and $\phi$, and (d) heavy-tail
+index for $\mu=25$ and different values of $y$ and $\phi$.
 
@@
 <<indexes-plot, fig.width=9.5, fig.height=4, fig.cap="Indexes for COM-Poisson distribution.
-Mean and variance relationship (a), dispersion, zero-inflation and heavy-tail indexes (b--d) for different parameter values.
+(a) Mean and variance relationship, (b--d) dispersion, zero-inflation and heavy-tail indexes for different parameter values.
 Dotted lines represents the Poisson special case.">>=
@@
 
-\autoref{fig:indexes-plot} shows that the indexes are slightly depend on
-the expected values and tend to stabilize for large values of
+\autoref{fig:indexes-plot} shows that the indexes are slightly dependent
+on the expected values and tend to stabilize for large values of
 $\mu$. Consequently, the mean and variance relationship
@@
 was \Sexpr{midi}. \autoref{fig:indexes-plot}(c) shows the COM-Poisson
-can handle with a limited amount of zero-inflation, in cases of
+can handle a limited amount of zero-inflation, in cases of
 overdispersion ($\phi < 0$). On the other hand, for $\phi > 0$
@@
 zero-deflated counts. Heavy-tail indexes in
-\autoref{fig:indexes-plot}(d) indicates the distribution is in general
-light tailed distribution, i.e. $HT \to 0$ for $y \to \infty$.
+\autoref{fig:indexes-plot}(d) indicate the distribution is in general a
+light-tailed distribution, i.e. $HT \to 0$ for $y \to \infty$.
 
@@
 
-In this section we shall describe the estimation and inference to the
+In this section we describe the estimation and inference for the
 two forms of the COM-Poisson regression model based on the maximum
@@
 involves the computation of the infinite series
-(\autoref{eqn:infseries}).  Thus, the fitting procedure is computational
+(\autoref{eqn:infseries}).  Thus, the fitting procedure is computationally
 expensive for regions of the parameter space where the convergence of
@@
 In the applications we fitted the quasi-Poisson model
-\citep{Wedderburn1974} as a baseline model. This approach is based on
-second-moment assumption that allows more flexibility to model. In this
-case the variance of the response variable is fixed by an additional
-parameter $\sigma$, $\textrm{Var}(Y_i)=\sigma \mu_i$. These models are
-fitted in software \texttt{R} using the function \texttt{glm(..., family
-  = quasipoisson)}.
+\citep{Wedderburn1974} as a baseline model. This approach is based on a
+second-moment assumption that allows more flexibility to the model. In
+this case the variance of the response variable is fixed by an
+additional parameter $\sigma$, $\textrm{Var}(Y_i)=\sigma \mu_i$. These
+models are fitted in the \texttt{R} software using the function
+\texttt{glm(..., family = quasipoisson)}.
 
@@
 continuous covariate~($\bm{x}_1$) was generated as a sequence from $0$
-to $1$ and length equals to the sample size.  Similarly, the categorical
+to $1$ and of length equal to the sample size.  Similarly, the categorical
 covariate~($\bm{x}_2$) was generated as a sequence of three values each
@@
 of $\bm{x}_2$.  The regression coefficients were fixed at the values,
-$\beta_0 = 2$, $\beta_1 = 05$, $\beta_{21} = 0.8$ and
+$\beta_0 = 2$, $\beta_1 = 0.5$, $\beta_{21} = 0.8$ and
 $\beta_{22} = -0.8$.
 
-We designed four simulation scenarios by considering four different
+We designed four simulation scenarios by considering different
 values of the dispersion parameter $\phi = -1.6, -1.0, 0.0$ and $1.8$.
-Thus, we have strong and moderate overdispersion, equidispersion and
+Thus, we have strong and moderate overdispersion, equidispersion, and
 underdispersion, respectively.  \autoref{fig:justpars} shows the
@@
 study.  These configurations allow us to assess the properties of the
-maximum likelihood estimators in extreme situations such as high counts
+maximum likelihood estimators in extreme situations, such as high counts
 and low dispersion, and low counts and high dispersion, but also in the
@@
 levels, both the average bias and standard errors tend to $0$ as the
-sample size increases. The estimators for the regression parameters
-are unbiased, consistent and their empirical distributions are
+sample size increases. The estimators for the regression parameters are
+unbiased, consistent and their empirical distributions are
 symmetric. For the dispersion parameter, the estimator is asymptotically
@@
 forms (original and new parametrization) and the quasi-Poisson
-regression model.  The data sets and \texttt{R} codes for their analysis
+regression model.  The data sets and \texttt{R} code for their analysis
 are available as supplementary material.
@@
 difference in terms of log-likelihood value from the Poisson to the
-COM-Poisson$_\mu$ model was $\Sexpr{lrt.ncap[2, 4]}$, which in turn
+COM-Poisson$_\mu$ model was $$\Sexpr{lrt.ncap[2, 4]}$$, which in turn
 suggests the better fit of the COM-Poisson$_{\mu}$ model. A chi-square
 test also supports this statement.  Finally, the estimated value of the
-dispersion parameter $\hat{\phi} = \Sexpr{co1[1]}$ suggests underdispersion.
-
-Furthermore, results in \autoref{tab:coef-cotton} also shows the
+dispersion parameter $\\hat{\phi} = \Sexpr{co1[1]}$ suggests
+underdispersion.
+
+Furthermore, results in \autoref{tab:coef-cotton} also show the
 advantage of the COM-Poisson$_\mu$ model, since the estimates are quite
@@
 that COM-Poisson model is a full parametric approach, i.e. there is a
-probability distribution associated to the counts. On the other hand, the
-quasi-Poisson model is a specification based only on second-moments
+probability distribution associated to the counts. On the other hand,
+the quasi-Poisson model is a specification based only on second-moment
 assumptions.
@@
 the other hand, for the original parametrization such correlations are
-quite larger, in special for the parameter $\beta_0$ (due to effects
+quite large, in particular for the parameter $\beta_0$ (due to effects
 parametrization in the linear predictor).  This result explain the
@@
 dispersion parameter $\phi$ only. To compare the computational times on
-the two parametrizations we repeat $50$ times the routine to fit. In
+the two parametrizations we repeat the fitting $50$ times. In
 this case COM-Poisson$_\mu$ fit was, on average, $38$\% faster than the
@@
 1.8 $\times$ 100mg dm$^{-3}$) and soil moisture (\texttt{umid}) levels
-(37.5, 50, 62.5\%) on the soybean (\emph{Glicine Max}) production. The
+(37.5, 50, 62.5\%) on  soybean (\emph{Glicine Max}) production. The
 experiment was carried out in a greenhouse, in pots with two plants, and
-the count variable measured was the number of grains per
+the count variable measured was the number of bean seeds per
 pot~\citep{Serafim2012}.  \autoref{fig:desc-soy} (left) shows the number
-of grains recorded for each combination of potassium dose and moisture
+of bean seeds recorded for each combination of potassium dose and moisture
 level, it is important to note the indication of a quadratic effect of
@@
 
-<<desc-soy, fig.pos="!ht", fig.height=4, fig.width=8, fig.cap="Number of grains per pot for each potassium dose and moisture level (left) and sample mean agains sample variance of the five replicates for each experimental treatment (right).
+<<desc-soy, fig.pos="!ht", fig.height=4, fig.width=8, fig.cap="Number of bean seeds per pot for each potassium dose and moisture level (left) and sample mean against sample variance of the five replicates for each experimental treatment (right).
 Solid lines are the smoothing curves on the left and the least of squares curve on the right.">>=
@@
 
-For the analysis of this data set based on descriptive analysis
+For the analysis of this data set based on the descriptive analysis
 (\autoref{fig:desc-soy}), we proposed the following linear predictor
@@
 50\% e 2: 62.5\%; and $k=$1: 0.0, 2: 0.3, 3: 0.6, 4: 1.2, 5: 1.8 100mg
-dm$^{-3}$,
-
-where $\gamma_i$ is the effect of $i$-th block ($i=$1: block II, 2:
-block III, 3: block IV and 4: block V), $\tau_j$ is the effect of $j$-th
-moisture level ($j=$1: 50\% and 2: 62.5\%) and $\beta_{3j}$ is the first
-order potassium effect (\texttt{K}) for the $j$-th moisture level
-(\texttt{umid}). \autoref{tab:coef-soy} presents the estimates, ratio
-between estimate and standard error and goodness-of-fit measures for the
-alternative models.
+dm$^{-3}$, where $\gamma_i$ is the effect of $i$-th block ($i=$1: block
+II, 2: block III, 3: block IV and 4: block V), $\tau_j$ is the effect of
+$j$-th moisture level ($j=$1: 50\% and 2: 62.5\%) and $\beta_{3j}$ is
+interaction of the first order potassium effect (\texttt{K}) for the
+$j$-th moisture level (\texttt{umid}). \autoref{tab:coef-soy} presents
+the estimates, ratio between estimate and standard error and
+goodness-of-fit measures for the alternative models.
 
@@
 different from zero. The estimate of $\phi$ ($\Sexpr{co2[1]}$) indicates
-overdispersion, corroborating to the descriptive analysis. Concerning to
+overdispersion, corroborating the descriptive analysis. Concerning to
 the regression parameters, the similarities between the models are
@@
 block, potassium dose and moisture level, however the Poisson model
-indicates the effects with greater significance, because it does not fit
-the extra variability.
+indicates the effects with greater significance, because it does not
+take account of the extra variability.
 
@@
 convergence, when using the original and new parametrization of the
-COM-Poisson distribution, respectively.  In terms of computational time
-and for 50 repetitions of fit, the proposed reparametrization was on
-average $110\%$ faster than the original one.  Probably, it is due to
-the better behaviour of the log-likelihood function as well as better
-initial values obtained from the Poisson fit. In \autoref{tab:corr-soy},
-we present the empirical correlation between the regression and
-dispersion parameter estimates. The correlations are close to zero for
-the COM-Poisson$_\mu$ model, indicating the empirical orthogonality
-between $\mu$ and $\phi$.
+COM-Poisson distribution, respectively.  In terms of computational time,
+for 50 repetitions of fit, the proposed reparametrization was on average
+$110\%$ faster than the original one.  Probably, it is due to the better
+behaviour of the log-likelihood function as well as better initial
+values obtained from the Poisson fit. In \autoref{tab:corr-soy}, we
+present the empirical correlation between the regression and dispersion
+parameter estimates. The correlations are close to zero for the
+COM-Poisson$_\mu$ model, indicating the empirical orthogonality between
+$\mu$ and $\phi$.
 
@@
 same conclusions. On the other hand, confidence intervals for the
-Poisson model are smaller than the ones from the
-COM-Poisson$_{\mu}$. Probably, it is due to the equidispersion
-assumption underline the Poisson model. The confidence intervals from
-the COM-Poisson$_{\mu}$ and quasi-Poisson models are really similar,
-which in turn show the already highlighted similarity between these
-approaches.
-
-<<pred-soy, fig.pos="!ht", fig.height=3.8, fig.width=8, fig.cap="Dispersion diagrams of grain counts as function of potassium doses and humidity levels with fitted curves and confidence intervals (95\\%).">>=
+Poisson model are smaller than the ones from the COM-Poisson$_{\mu}$,
+due to the equidispersion assumption underlying the Poisson model. The
+confidence intervals from the COM-Poisson$_{\mu}$ and quasi-Poisson
+models are really similar, which in turn shows the already highlighted
+similarity between these approaches, however only the COM-Poisson
+model$_{\mu}$ corresponds to a fully specified probability model.
+
+<<pred-soy, fig.pos="!ht", fig.height=3.8, fig.width=8, fig.cap="Dispersion diagrams of been seeds counts as function of potassium doses and humidity levels with fitted curves and confidence intervals (95\\%).">>=
 
@@
 The data set comes from an experiment to measure the reproductive
-toxicity of a herbicide, nitrofen, on a species of zooplankton
+toxicity of the herbicide, nitrofen, on a species of zooplankton
 (\textit{Ceriodaphnia dubia}). Fifty animals were randomized into
 batches of ten and each batch was put in a solution with a measured
-concentration of nitrofen ($0, 0.8, 1.6, 2.35$ e $3.10$ mug$/10^2$litre)
-(\texttt{dose}). Then the number of live offspring was recorded.
+concentration of nitrofen ($0, 0.8, 1.6, 2.35$ and $3.10$
+$\mu$g$/10^2$litre) (\texttt{dose}). Then the number of live offspring
+was recorded.
 
@@
 
-For this data set we proposed three linear predictors,
+For this data set we consider three models with linear predictors,
 \begin{center}
 \begin{minipage}{12cm}
-Predictor 1: $\log(\mu_i) = \beta_0 + \beta_1 \texttt{dose}_i$\\
-Predictor 2: $\log(\mu_i) = \beta_0 + \beta_1 \texttt{dose}_i +
+Linear: $\log(\mu_i) = \beta_0 + \beta_1 \texttt{dose}_i$\\
+Quadratic: $\log(\mu_i) = \beta_0 + \beta_1 \texttt{dose}_i +
              \beta_2 \texttt{dose}_i^2$\\
-Predictor 3: $\log(\mu_i) = \beta_0 + \beta_1 \texttt{dose}_i +
+Cubic: $\log(\mu_i) = \beta_0 + \beta_1 \texttt{dose}_i +
              \beta_2 \texttt{dose}_i^2 + \beta_3 \texttt{dose}_i^3$.
@@
 concentration.  Considering this predictor, there is an evidence of
-equidispersed counts, $\phi$ estimate of the COM-Poisson is close to zero
-and $\sigma$ of quasi-Poisson is close to one.  It is interesting to note
-that if we omit the high order effects the models show evidence of
-overdispersion. This exemplifies the discussion on the causes of
+equidispersed counts, the $\phi$ estimate of the COM-Poisson is close to
+zero and $\sigma$ of quasi-Poisson is close to one.  It is interesting
+to note that if we omit the high order effects the models show evidence
+of overdispersion. This exemplifies the discussion on the causes of
 overdispersion made in \autoref{introduction}.  We can also note that
 the quasi-Poisson approach, although robust to equidispersion
-assumption, shows higher descriptive levels ($p$-values) than
-parametric models, that is, the tests under parametric models are more
-powerful than the ones under the quasi-Poisson model in the
-equidispersed case.
+assumption, shows higher descriptive levels ($p$-values) than parametric
+models, that is, the tests under parametric models are more powerful
+than the ones under the quasi-Poisson model in the equidispersed case.
 
 In \autoref{tab:coef-ovos}, we present the estimates of the regression
-parameters considering the cubic dose effect (third predictor). The
-interpretations are similar to others cases studies, however, in this
-case the Poisson model is also suitable for indicating the significance
-of the covariate effects.  In addition, note that the parameter
-estimates of the original COM-Poisson model are comparable to the others
-models. It occurs because it is the special case, where $\phi = 0$, which
-implies $\lambda = \mu$.
+parameters considering the cubic dose model. The interpretations are
+similar to others cases studies, however, in this case the Poisson model
+is also suitable for indicating the significance of the covariate
+effects.  In addition, note that the parameter estimates of the original
+COM-Poisson model are comparable to the others models. This occurs
+because we are in the particular case, where $\phi = 0$, which implies
+$\lambda = \mu$.
 
@@
 
-\autoref{fig:pred-ovos} shows the number of live off-springs observed
-and fitted along with confidence intervals for all model strategies
-adopted.  The fitted values and confidence intervals come to have a full
-overlay. It shows that the estimation of the extra dispersion parameter
-does not affect the estimation of the regression coefficients in the
-case of equidispersed counts.
+\autoref{fig:pred-ovos} shows the number of live off-spring observed and
+fitted curves along with confidence intervals for all model strategies
+adopted.  The fitted values and confidence intervals are identical and
+have a complete overlap. It shows that the estimation of the extra
+dispersion parameter does not affect the estimation of the regression
+coefficients in the case of equidispersed counts.
 
@@
 parametrization the COM-Poisson distribution is indexed by the
-expectation $\mu$ and an extra dispersion parameter $\phi$ whose our
+expectation $\mu$ and an extra dispersion parameter $\phi$ which our
 data analysis suggest to be orthogonal.

### 2018-01-27 codes/01-simulation.R
Add source code and compress file to orthogonality simulation study

@@ -401,6 +401,7 @@ useOuterStrips(
 
 #-----------------------------------------------------------------------
 # Obtain the covariance between dispersion and regression parameters
+
 aux <- ldply(
     lapply(results, function(x) {
     ind <- names(x); names(ind) <- ind
@@ -462,3 +463,183 @@ useOuterStrips(
 )
 
 #-----------------------------------------------------------------------
+# Likelihood surfaces (simulation with simple no covariate case)
+
+# Settings of the simulation
+fixed_mu <- 10
+fixed_size <- 300L
+
+# Simulate 1000 observations of the DW distribution
+set.seed(97380)
+das <- lapply(phis, function(phi) {
+    rcmp(fixed_size, mu = fixed_mu, phi = phi)
+})
+
+# Compute approximately expected dispersion indexes
+dis <- vapply(phis, compute_variance, double(1),
+              mu = fixed_mu, sumto = 500L) / 10
+dis
+
+# Verify mean, variance and dispersion indexes in the simulations
+ldply(lapply(das, function(x) {
+    c("Mean" = mean(x), "Var" = var(x), "DI" = var(x)/mean(x))
+}), .id = "phi") %>%
+    mutate(out, "Expected DI" = dis)
+
+# Fit models
+index <- seq_along(phis)
+names(index) <- phis
+
+# Fit on original parametrization
+models_orpar <- lapply(index, function(i) {
+    lambda <- mu2lambda(fixed_mu, phis[i])[[1]]
+    start <- c("phi" = phis[i], "(Intercept)" = log(lambda))
+    y <- das[[i]]
+    fitcm(y ~ 1, sumto = 300L, model = "CP", start = start)
+})
+
+# Fit on propose reparametrization
+models_repar <- lapply(index, function(i) {
+    start <- c("phi2" = phis[i], "(Intercept)" = log(fixed_mu))
+    y <- das[[i]]
+    fitcm(y ~ 1, sumto = 300L, model = "CP2", start = start)
+})
+
+# Same maximum loglikelihood
+cbind("Original" = vapply(models_orpar, logLik, numeric(1)),
+      "Propose"  = vapply(models_repar, logLik, numeric(1)))
+
+#-------------------------------------------
+# Deviance contours
+
+# Original parametrization
+devs_orpar <- purrr::map_df(index, function(i) {
+    m0 <- models_orpar[[i]]
+    co <- unname(coef(m0))
+    # Wald intervals
+    interval <- confint(m0, level = 0.999, method = "quad")
[…]

### 2018-01-27 codes/00-approx.R
Remove code to simulate data and plot deviance surfaces

@@ -303,211 +303,3 @@ print(xy1, split = c(1, 1, 4, 1), more = TRUE)
 print(xy2, split = c(2, 1, 4, 1), more = TRUE)
 print(xy3, split = c(3, 1, 4, 1), more = TRUE)
 print(xy4, split = c(4, 1, 4, 1), more = FALSE)
-
-#-----------------------------------------------------------------------
-# Likelihood surfaces
-
-# Random generation
-rcmp <- function(n, mu, phi, upper = 500L, sumto = 300L) {
-    x <- 0:upper
-    fx <- dcmp(x, mu, phi, sumto = sumto)
-    sample(x, size = n, replace = TRUE, prob = fx)
-}
-
-
-# Moments function
-moments_compois <- Vectorize(
-    function(mu, phi, upper = 500L, sumto = 300L) {
-        x <- 0:upper
-        fx <- dcmp(x, mu, phi, sumto = sumto)
-        me <- sum(x * fx)
-        va <- sum(x^2 * fx) - me^2
-        return(c("Expectation" = me, "Variance" = va))
-    }, vectorize.args = c("mu", "phi"))
-
-# System equation to find parameters
-library(rootSolve)
-system_equation <- function(param, Expec, DI, trace = FALSE) {
-    par1 <- param[1]
-    par2 <- param[2]
-    if (trace) print(param)
-    moments <- moments_compois(par1, par2)
-    eq1 <- moments[1] -  Expec
-    eq2 <- moments[2] / moments[1] - DI
-    return(c(eq1, eq2))
-}
-
-# Dispersion indexes
-dis <- c(0.5, 1, 3)
-names(dis) <- dis
-
-# Find parameters for fixed expectation and different DI's
-pars <- lapply(dis, function(di) {
-    multiroot(system_equation, start = c(5, 0),
-              Expec = 5, DI = di)$root
-})
-
-# Simulate 1000 observations of the DW distribution
-set.seed(97380)
-das <- lapply(pars, function(obj) {
-    rcmp(1000L, mu = obj[1], phi = obj[2])
-})
-
-# Verify simulations
-do.call("cbind", lapply(das, function(x) {
-    c("Mean" = mean(x), "Var" = var(x), "DI" = var(x)/mean(x))
-}))
-
-# Fit models
-index <- seq_along(dis)
-names(index) <- dis
-
-# Fit on original parametrization
-models_orpar <- lapply(index, function(i) {
[…]

### 2018-01-28 codes/01-simulation.R
Change parameter for deviance surfaces and save compressed

@@ -466,8 +466,8 @@ useOuterStrips(
 # Likelihood surfaces (simulation with simple no covariate case)
 
 # Settings of the simulation
-fixed_mu <- 10
-fixed_size <- 300L
+fixed_mu <- 5
+fixed_size <- 1000L
 
 # Simulate 1000 observations of the DW distribution
 set.seed(97380)
@@ -557,15 +557,15 @@ devs_repar <- purrr::map_df(index, function(i) {
               "fitphi" = co[1])
 })
 
-#-------------------------------------------
-# Plots deviance surfaces
-output <- readRDS("orthogonality.rds")
-devs_orpar <- output$devs_orpar
-devs_repar <- output$devs_repar
+# #-------------------------------------------
+# # Plots deviance surfaces
+# output <- readRDS("orthogonality.rds")
+# devs_orpar <- output$devs_orpar
+# devs_repar <- output$devs_repar
 
 niveis <- c(0.9, 0.95, 0.99)
 cortes <- qchisq(niveis, df = 2)
-fl <- parse(text = gsub("=", "==", names(phis)))
+fl <- parse(text = paste("phi==", sort(phis)))
 
 # Original parametrization
 devs_orpar$fphi <- ordered(devs_orpar$realphi)

### 2018-01-28 article.Rnw
Add orthogonality study to the article

@@
 In this section we performed a simulation study to assess the properties
-of the maximum likelihood estimators as well as the flexibility of the
-COM-Poisson regression model to deal with non-equidispersed count data.
+of the maximum likelihood estimators and orthogonality of the
+reparametrized model as well as the flexibility of the COM-Poisson
+regression model to deal with non-equidispersed count data.
 
@@
 
-We designed four simulation scenarios by considering different
-values of the dispersion parameter $\phi = -1.6, -1.0, 0.0$ and $1.8$.
-Thus, we have strong and moderate overdispersion, equidispersion, and
+We designed four simulation scenarios by considering different values of
+the dispersion parameter $\phi = -1.6, -1.0, 0.0$ and $1.8$.  Thus, we
+have strong and moderate overdispersion, equidispersion, and
 underdispersion, respectively.  \autoref{fig:justpars} shows the
@@
 
+To check the orthogonality property we compute the covariance matrix
+between maximum likelihood estimators
+$\hat{\bm{\theta}} = (\hat{\bm{\beta}}, \phi)$, obtained by observed
+information matrix,
+Cov$(\hat{\bm{\theta}}) = \mathcal{I}^{-1}(\bm{\theta})$.
+\autoref{fig:ortho-plot} shows the covariance between regression and
+dispersion parameter estimators for each simulation scenario, on
+the correlation scale. The correlations are close to zero in all cases
+suggesting the orthogonality property for the reparametrized
+model. Interesting results in the first panel shows that
+cov($\hat{\beta}_{22}, \hat{\phi}$) is not very close to zero (values
+between $-0.4$ and $0.2$) for strong overdispersion ($\phi = -1.6$).
+
+<<ortho-plot, fig.height=4.5, fig.width=9, fig.cap="Empirical correlations between regression and dispersion parameters by different sample sizes and dispersion levels.">>=
+
+#-----------------------------------------------------------------------
+# Obtain the covariance between dispersion and regression parameters
+
+aux <- ldply(
+    lapply(results, function(x) {
+    ind <- names(x); names(ind) <- ind
+    out <- lapply(ind, function(y) {
+        vcovs <- lapply(x[[y]], "[[", "vcov")
+        index <- vapply(lapply(vcovs, is.na), sum, integer(1))
+        vcovs <- vcovs[!index]
+        corrs <- lapply(vcovs, cov2cor)
+        covs <- do.call(rbind, lapply(vcovs, "[", 1, 2:5))
+        cors <- do.call(rbind, lapply(corrs, "[", 1, 2:5))
+        ldply(list("cov" = covs, "cor" = cors), .id = "scale")
+    })
+    ldply(out, .id = "phi")
+}), .id = "n")
+str(aux)
+
+# Organize the results
+dacov <- gather(aux, param, value, b0:b22, factor_key = TRUE)
+dacov$phi <- ordered(dacov$phi,
+                     c("phi=-1.6", "phi=-1", "phi=0", "phi=1.8"))
+str(dacov)
+
+# Distributions of covariance/correlations
+fl <- parse(text = gsub("=", "==", levels(dacov$phi)))
+yl <- parse(text = paste0("hat(beta)[", c(0, 1, 21, 22), "]"))
+
+# Boxplots
+bwplot(param ~ value | phi,
+       groups = n,
+       data = subset(dacov, scale == "cor"),
+       box.width = 0.12,
+       pch = "|",
+       grid = TRUE,
+       as.table = TRUE,
+       layout = c(4, 1),
+       xlab = "Empirical correlations with dispersion parameter estimator",
+       fill = colorRampPalette(c("gray80",  "gray20"))(4),
+       panel = panel.superpose,
+       scales = list(y = list(labels = yl), x = "free"),
+       whisker.width = 0.4,
+       strip = strip.custom(factor.levels = fl),
+       auto.key = list(
+           column = 4,
+           rectangles = TRUE,
+           points = FALSE,
+           title = "Sample size",
+           cex.title = 1
+       ),
+       par.settings = list(
+           plot.symbol = list(pch = 19, cex = 0.1)
+       ),
+       panel.groups = function(x, y, group.number, ...) {
+           my.panel.bwplot(x, y + (group.number - 2.5) / 5, ...)
+           panel.abline(v = 0, lty = 2)
+       })
+
+@
+
+To ilustrate the orthogonality, \autoref{fig:ortho-surf} the deviance
+surfaces for four simulated data set of size 1000, $\mu=5$ and different
+values of the dispersion parameters. The shapes of the deviance function
+show that the proposed parametrization is better for both computation
+and asymptotic (normal-based) inference. Furthermore, it is interesting
+to note that the deviance function shape on strong overdispersion
+($\phi=-1.8$) not as well behaved as the others. This is due to the
+difficulty of the distribution in dealing with strong overdispersion in
+low counts (see dispersion index plot in the
+\autoref{fig:indexes-plot}). This also explains the results of
+Cov$(\hat{\beta}_{22}, \phi)$ in the first panel of
+\autoref{fig:ortho-plot}, since $\beta_{22}$ is negative and associate
+with low counts.
+
+<<ortho-surf, fig.height=5.1, fig.width=9, fig.cap="Deviance surfaces under original and proposed parametrization for simulated data sets of size 1000 an different dispersion parameters.
+The ellipses are confidence regions (90, 95 and 99\\%), dotted lines are the maximum likelihood estimates, and points are the real parameters used in the simulation.">>=
+
+#-------------------------------------------
+# Plots deviance surfaces
+output <- readRDS("codes/orthogonality.rds")
+devs_orpar <- output$devs_orpar
+devs_repar <- output$devs_repar
+
+niveis <- c(0.9, 0.95, 0.99)
+cortes <- qchisq(niveis, df = 2)
+fl <- parse(text = paste("phi==", sort(phis)))
+
+# Original parametrization
+devs_orpar$fphi <- ordered(devs_orpar$realphi)
+xy1 <- useOuterStrips(
+    levelplot(
+        deviance ~ lambda + phi | fphi + "Original parametrization",
+        data = devs_orpar,
+        scales = list(y = list(rot = 90, relation = "free"),
+                      x = "free"),
+        cuts = 30,
+        xlab = expression(lambda),
+        ylab = expression(phi),
+        colorkey = FALSE,
+        panel = function(x, y, z, at, region, ...,
+                         subscripts = subscripts) {
+            ##
+            flambda <- devs_orpar$fitlambda[subscripts]
+            fphi <- devs_orpar$fitphi[subscripts]
+            ##
+            tlambda <- devs_orpar$reallambda[subscripts]
+            tphi <- devs_orpar$realphi[subscripts]
+            ##
+            panel.levelplot(x, y, z, at = at, region = TRUE,
+                            ..., subscripts = subscripts)
+            panel.contourplot(x, y, z, ..., at = cortes,
+                              contour = TRUE, region = FALSE,
+                              subscripts = subscripts)
+            panel.abline(v = flambda, h = fphi, lty = 2)
+            panel.points(x = tlambda, y = tphi,
+                         lty = 2, pch = 19)
+        }),
+    strip = strip.custom(factor.levels = fl),
+    strip.left = strip.custom(par.strip.text=list(cex = 0.9))
+)
+
+# Propose parametrization
+devs_repar$fphi <- ordered(devs_repar$realphi)
+xy2 <- useOuterStrips(
+    levelplot(
+        deviance ~ mu + phi2 | fphi + "Proposed parametrization",
+        data = devs_repar,
+        scales = list(y = list(rot = 90, relation = "free"),
+                      x = "free"),
+        cuts = 30,
+        xlab = expression(mu),
+        ylab = expression(phi),
+        colorkey = FALSE,
+        panel = function(x, y, z, at, region, ...,
+                         subscripts = subscripts) {
+            ##
+            fmu <- devs_repar$fitmu[subscripts]
+            fphi <- devs_repar$fitphi[subscripts]
+            ##
+            tmu <- devs_repar$realmu[subscripts]
+            tphi <- devs_repar$realphi[subscripts]
+            ##
+            panel.levelplot(x, y, z, at = at, region = TRUE,
+                            ..., subscripts = subscripts)
+            panel.contourplot(x, y, z, ..., at = cortes,
+                              contour = TRUE, region = FALSE,
+                              subscripts = subscripts)
+            panel.abline(v = fmu, h = fphi, lty = 2)
+            panel.points(x = tmu, y = tphi,
+                         lty = 2, pch = 19)
+        }),
+    strip = strip.custom(factor.levels = fl),
+    strip.left = strip.custom(par.strip.text=list(cex = 0.9))
+)
+
+# Organize plots
+print(xy1, position = c(0.0, 0.48, 1.0, 1.0), more = TRUE)
+print(xy2, position = c(0.0, 0.0, 1.0, 0.52), more = FALSE)
+
+@
+
 \section{Case studies}
@@
 difference in terms of log-likelihood value from the Poisson to the
-COM-Poisson$_\mu$ model was $$\Sexpr{lrt.ncap[2, 4]}$$, which in turn
+COM-Poisson$_\mu$ model was $\Sexpr{lrt.ncap[2, 4]}$, which in turn
 suggests the better fit of the COM-Poisson$_{\mu}$ model. A chi-square

### 2018-01-28 article.Rnw
Improve the concluding remarks section

@@
 
-In this paper, we presented a novel parametrization of the COM-Poisson
-distribution and the associated regression model. The novel
-parametrization was based on a simple asymptotic approximation for the
-expectation and variance of the COM-Poisson distribution. The main
+In this paper, we presented and characterized a novel parametrization of
+the COM-Poisson distribution and the associated regression model. The
+novel parametrization was based on a simple asymptotic approximation for
+the expectation and variance of the COM-Poisson distribution. The main
 advantage of the proposed reparametrization is the simple interpretation
@@
 expectation $\mu$ and an extra dispersion parameter $\phi$ which our
-data analysis suggest to be orthogonal.
+data analysis suggest being orthogonal. This is similar to
+\Citeauthor{Huang2017}'s (\citeyear{Huang2017}) parametrization but is
+simpler, because the $\mu$ is obtained from simple algebra.
 
@@
 approximations are accurate for a large part of the parameter space,
-which in turn support our reparametrization.  We carried out a
-simulation study to assess the properties of the reparametrized
+which in turn support our reparametrization. Moreover, we discuss the
+properties and flexibility of the distribution to deal with count data
+although dispersion, zero-inflation and heavy-tail indexes. We carried
+out a simulation study to assess the properties of the reparametrized
 COM-Poisson model to deal with different levels of dispersion as well as

### 2018-01-28 article.Rnw
Make minor editions

@@
 
-To ilustrate the orthogonality, \autoref{fig:ortho-surf} the deviance
-surfaces for four simulated data set of size 1000, $\mu=5$ and different
-values of the dispersion parameters. The shapes of the deviance function
-show that the proposed parametrization is better for both computation
-and asymptotic (normal-based) inference. Furthermore, it is interesting
-to note that the deviance function shape on strong overdispersion
-($\phi=-1.8$) not as well behaved as the others. This is due to the
-difficulty of the distribution in dealing with strong overdispersion in
-low counts (see dispersion index plot in the
-\autoref{fig:indexes-plot}). This also explains the results of
-Cov$(\hat{\beta}_{22}, \phi)$ in the first panel of
-\autoref{fig:ortho-plot}, since $\beta_{22}$ is negative and associate
-with low counts.
-
-<<ortho-surf, fig.height=5.1, fig.width=9, fig.cap="Deviance surfaces under original and proposed parametrization for simulated data sets of size 1000 an different dispersion parameters.
+<<ortho-surf, fig.height=5, fig.width=9, fig.cap="Deviance surfaces under original and proposed parametrization for simulated data sets of size 1000 an different dispersion parameters.
 The ellipses are confidence regions (90, 95 and 99\\%), dotted lines are the maximum likelihood estimates, and points are the real parameters used in the simulation.">>=
@@
     strip = strip.custom(factor.levels = fl),
-    strip.left = strip.custom(par.strip.text=list(cex = 0.9))
+    strip.left = strip.custom(par.strip.text=list(cex = 0.85))
 )
@@
 @
+
+To ilustrate the orthogonality, \autoref{fig:ortho-surf} the deviance
+surfaces for four simulated data set of size 1000, $\mu=5$ and different
+values of the dispersion parameters. The shapes of the deviance function
+show that the proposed parametrization is better for both computation
+and asymptotic (normal-based) inference. Furthermore, it is interesting
+to note that the deviance function shape on strong overdispersion
+($\phi=-1.6$) not as well behaved as the others. This is due to the
+difficulty of the distribution in dealing with strong overdispersion in
+low counts (see dispersion index plot in the
+\autoref{fig:indexes-plot}). This also explains the results of
+Cov$(\hat{\beta}_{22}, \phi)$ in the first panel of
+\autoref{fig:ortho-plot}, since $\beta_{22}$ is negative and associate
+with low counts.
 

### 2018-01-29 article.Rnw
Correct spelling

@@
 between maximum likelihood estimators
-$\hat{\bm{\theta}} = (\hat{\bm{\beta}}, \phi)$, obtained by observed
-information matrix,
+$\hat{\bm{\theta}} = (\hat{\bm{\beta}}, \phi)$, obtained from the
+observed information matrix,
 Cov$(\hat{\bm{\theta}}) = \mathcal{I}^{-1}(\bm{\theta})$.
 \autoref{fig:ortho-plot} shows the covariance between regression and
-dispersion parameter estimators for each simulation scenario, on
-the correlation scale. The correlations are close to zero in all cases
+dispersion parameter estimators for each simulation scenario, on the
+correlation scale. The correlations are close to zero in all cases
 suggesting the orthogonality property for the reparametrized
-model. Interesting results in the first panel shows that
+model. Interestingly, results in the first panel show that
 cov($\hat{\beta}_{22}, \hat{\phi}$) is not very close to zero (values
@@
 
-<<ortho-surf, fig.height=5, fig.width=9, fig.cap="Deviance surfaces under original and proposed parametrization for simulated data sets of size 1000 an different dispersion parameters.
+<<ortho-surf, fig.height=5, fig.width=9, fig.cap="Deviance surfaces contour plots under original and proposed parametrization for four simulated data sets of size 1000 with different dispersion parameters.
 The ellipses are confidence regions (90, 95 and 99\\%), dotted lines are the maximum likelihood estimates, and points are the real parameters used in the simulation.">>=
@@
 
-To ilustrate the orthogonality, \autoref{fig:ortho-surf} the deviance
-surfaces for four simulated data set of size 1000, $\mu=5$ and different
-values of the dispersion parameters. The shapes of the deviance function
-show that the proposed parametrization is better for both computation
-and asymptotic (normal-based) inference. Furthermore, it is interesting
-to note that the deviance function shape on strong overdispersion
-($\phi=-1.6$) not as well behaved as the others. This is due to the
-difficulty of the distribution in dealing with strong overdispersion in
-low counts (see dispersion index plot in the
-\autoref{fig:indexes-plot}). This also explains the results of
-Cov$(\hat{\beta}_{22}, \phi)$ in the first panel of
-\autoref{fig:ortho-plot}, since $\beta_{22}$ is negative and associate
-with low counts.
+To illustrate the orthogonality, \autoref{fig:ortho-surf} displays
+contour plots of the deviance surfaces for four simulated data set of
+size 1000, $\mu=5$ and different values of the dispersion
+parameters. The shapes of the deviance function show that the proposed
+parametrization is better for both computation and asymptotic
+(normal-based) inference. Furthermore, it is interesting to note that
+the deviance function shape under strong overdispersion ($\phi=-1.6$) is
+not as well behaved as the others; this is due to the difficulty of the
+distribution in dealing with strong overdispersion in low counts (see
+dispersion index plot in the \autoref{fig:indexes-plot}). This also
+explains the results of Cov$(\hat{\beta}_{22}, \phi)$ in the first panel
+of \autoref{fig:ortho-plot}, since $\beta_{22}$ is negative and
+associated with low counts.
 
@@
 
-<<desc-soy, fig.pos="!ht", fig.height=4, fig.width=8, fig.cap="Number of bean seeds per pot for each potassium dose and moisture level (left) and sample mean against sample variance of the five replicates for each experimental treatment (right).
+<<desc-soy, fig.height=4, fig.width=8, fig.cap="Number of bean seeds per pot for each potassium dose and moisture level (left) and sample mean against sample variance of the five replicates for each experimental treatment (right).
 Solid lines are the smoothing curves on the left and the least of squares curve on the right.">>=
@@
 
-<<pred-soy, fig.pos="!ht", fig.height=3.8, fig.width=8, fig.cap="Dispersion diagrams of been seeds counts as function of potassium doses and humidity levels with fitted curves and confidence intervals (95\\%).">>=
+<<pred-soy, fig.height=3.8, fig.width=8, fig.cap="Dispersion diagrams of been seeds counts as function of potassium doses and humidity levels with fitted curves and confidence intervals (95\\%).">>=
 

### 2018-01-31 article.Rnw
Do minor corrections and add submission version of the paper

@@
 }
-\AuthorRunning{Ribeiro Jr \textrm{et al.}}
+\AuthorRunning{Eduardo E. Ribeiro Jr \textrm{et al.}}
 
@@
   The study of distribution properties through a consideration of
-  dispersion, zero-inflated and heavy tail indices, together with the
+  dispersion, zero-inflated and heavy tail indexes, together with the
   results of data analysis show the flexibility over standard
@@
 contour plots of the deviance surfaces for four simulated data set of
-size 1000, $\mu=5$ and different values of the dispersion
+size 1000 with $\mu=5$ and different values of the dispersion
 parameters. The shapes of the deviance function show that the proposed

### 2018-08-16 article.Rnw
Correct the curves of the restriction

@@
                      panel.levelplot(x, y, z, ...)
-                     panel.curve(10 - ( exp(x) - 1)/(2 * exp(x)),
+                     panel.curve(10 - (exp(x) - 1)/(2 * exp(x)),
                                  lty = 2)
@@
                      panel.levelplot(x, y, z, ...)
-                     panel.curve((10 - ( exp(x) - 1)/
-                                  (2 * exp(x)))/exp(x), lty = 2)
+                     panel.curve(10 - (exp(x) - 1)/(2 * exp(x)),
+                                 lty = 2)
                      panel.abline(v = 0, lty = 2)

### 2018-08-20 review/review.Rnw
Add first draft of the point-by-point responses

\documentclass[12pt, a4paper]{article}
\usepackage[margin=2cm]{geometry}

% Language
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}

% Math packages
\usepackage{amsmath, amsthm, amsfonts, amssymb, amsxtra}
\usepackage{bm}

% Tables
\usepackage{booktabs}

% Fonts
\usepackage{mathpazo}
\setlength\parindent{0pt}

% Utils
\usepackage{float}
\usepackage{lipsum}
\usepackage{hyperref}
\hypersetup{
  pdftitle={Reparametrization of the COM-Poisson Regression Models},
  colorlinks=true,
  linkcolor=red,
  citecolor=blue,
  filecolor=magenta,
  urlcolor=blue,
  bookmarksdepth=4
}

% Biliography
\usepackage{natbib}
\bibliographystyle{agsm}

% Metadata
\title{\vspace{-0.5cm}
  Reparametrization of COM-Poisson Regression Models with
  Applications in the Analysis of Experimental Data\\[0.5cm]
  \Large Replies to Reviewer's Comments
}
\author{
  Eduardo E. Ribeiro Jr\and
  Walmes M. Zeviani\and
  Wagner H. Bonat\and
  Clarice G. B. Dem\'{e}trio\and
  John Hinde}

%-----------------------------------------------------------------------
% Init document

<<include=FALSE>>=
knitr::read_chunk("review.R")
@

<<setup, echo=FALSE>>=
@

<<load-packages>>=
@

\begin{document}

\maketitle

\section{Replies to Referee 1}

\textbf{\underline{Overall reviewer's comment}:} \textit{Overall, the
  authors did a good job while presenting theirs results. However, I
  believe that the manuscript at the current state requires a
  considerable revision. Mainly, it requires a rigorous discussion with
  numerical results that when the proposed approach is useful and when
  it is not.}
\vspace{0.3cm}

\textbf{\underline{Response}:} We appreciate the referee for his
revision. All suggestions have been taken into account. Regarding the
referees' main comment, we propose a reparametrization of the
COM-Poisson model that is always valid. Only the interpretation of the
$\mu$ parameter depends on the approximation accuracy. We make it clear
in the updated version of the paper. The applicability of the model is
discussed at the end of Section 3 through the study of the dispersion
(DI), zero-inflation (ZI) and heavy-tail (HI) indexes (Figure 3).
\vspace{0.5cm}

\textbf{\underline{Specific reviewer's comments}:}
\begin{enumerate}
\item \textit{There has been some active research to find a better
    approximation for normalizing constant $Z(\lambda, \nu)$ from which
    the expressions for both mean and variance will be derived. Please
    refer to the Gaunt et al. (2016) for the latest results.}\\[0.1cm]
  \textbf{\underline{Response:}}
\item \textit{It seems that the paper missing some latest literature on
    CMP distribution. For example, the distributional properties of CMP
    were found in Daly and Gaunt (2015).}\\[0.1cm]
  \textbf{\underline{Response:}}
\item \textit{The likelihood equation (4.1) needs more details. Can we
    derive score equations? It seems like the likelihood has become even
    more complicated. Some of the nice properties of the CMP likelihood
    such as the canonical link are lost.}\\[0.1cm]
  \textbf{\underline{Response:}} The log-likelihood function cannot be
  derived in any parametrization. We obtain the Hessian matrix by
  central finite differences (as said in ...). Regarding the expression
  of reparametrized log-likelihood (Equation 3.1), although it seems
  more complicated, this function is better behaved than the function
  expressed in the original form (see Figure 8). Consider the
  exponential family of distributions, with density given by
  $$f(y) = \exp\{\sigma[\theta y - b(\theta)] - c(y, \phi)\},$$
  the COM-Poisson belongs to this family with $\sigma=1$,
  $\theta = \log(\lambda)$, $b(\theta) = \log(Z(\lambda, \nu))$ and
  $c(y, \phi)=-\nu\log(y!)$, only if $\nu$ is known. Since we don't have
  $b'(\theta) = \partial b(\theta)/\partial \theta$ in closed form, the
  canonical link function has no an important rule. In this case, the
  link function just maps $\theta$ in the parameter space of $\lambda$.
  In the reparametrized version of the COM-Poisson, we adopted log link
  function to map the parameter space and to have good interpretations
  (in terms of rate ratios).
\item \textit{For model fitting, the authors used general purpose
    optimization function in R. Did you observe any non convergence
    situations during your numerical studies.}\\[0.1cm]
  \textbf{\underline{Response:}} Yes, we have 37 situations where the
  algorithm did not converge in the simulation study (see table
  below). This occurred only for strong overdispersion and small sample
  size. We report these situations in the updated version of the paper.
  \begin{table}[H]
    \centering
    \begin{tabular}{lcccc}
      \hline
      & $\phi=-1.6$ & $\phi=-1$ & $\phi=0$ & $\phi=1.8$ \\
      \hline
      $n=50$ &    29 &  0 &   0 &   0 \\
      $n=100$ &   8  &  0 &   0 &   0 \\
      $n=300$ &   0  &  0 &   0 &   0 \\
      $n=1000$ &  0  &  0 &   0 &   0 \\
      \hline
    \end{tabular}
  \end{table}
\item \textit{Numerical derivatives with CMP tend to be very slow and
    prone to errors. It would be good to have a computation times from
    the method for different sample sizes and dispersion
    levels.}\\[0.1cm]
  \textbf{\underline{Response:}}

\item \textit{In Section 5, the simulated counts are ranging from 3 to
    27. This is too limited. Please increase the range.}\\[0.1cm]
  \textbf{\underline{Response}:} The simulation study was performed with
  averages varying from 3 to 27. Therefore, considering the different
  scenarios ($\phi=-1.6, \phi=-1, \phi=0$ and $\phi=1.8$), we cover a
  wide range of counts. The Figure below shows the variation of the
  counts for $\mu=27$ in the different scenarios.

  \hspace*{-1cm}
\begin{minipage}{\textwidth}
<<counts-ranging, fig.width=9, fig.height=4, cache=TRUE>>=
@
\end{minipage}

\item \textit{I believe that the authors simulated data according to the
    reparametrized distribution. Please generate the data according to
    the original distribution. Otherwise, the results not useful. For
    details on simulating from original CMP distribution, please refer
    to Chatla and Shmueli (2018).}\\[0.1cm]
  \textbf{\underline{Response:}} We simulated data according to the
  original parametrization. The steps in the simulation study were i)
  obtain the $\mu_i = \exp(\bm{x}_i^\top\bm{\beta})$; ii) Back to the
  original parametrization $\lambda = (\mu + (\nu - 1)/(2\nu))^\nu$
  (Equation 3.1); and iii) simulate data according COM-Poisson
  distribution using the probability integral transform theorem as
  implemented in \texttt{compoisson} package \citep{compoisson2012}.

\item \textit{For real world data, the over dispersion case is too mild
    ($\phi=-0.77$). In practice over dispersion can be really high. For
    example, even lesser than $-1.4$. It would be nice to have such
    case, at least in the simulations.}\\[0.1cm]
  \textbf{\underline{Response:}} In the simulation study we consider the
  $\phi=-1.6$ and $\phi=-1$ that leads to dispersion indexes greater
  than four (see Figure 4).

\item \textit{From the presentation point of view, focusing more on
    applications would be good as there is not much novelty in the
    methodology.}\\[0.1cm]
  \textbf{\underline{Response:}} In this paper, we propose a
  mean-parametrization of the COM-Poisson model. This parametrization
  introduced $\mu = \lambda^{1/\nu} -(\nu-1)/(2\nu)$, a simple function
  of the original parameters $\lambda$ and $\nu$. We showed that the new
  parameter space has nice properties like i) the orthogonality between
  $\mu$ and $\phi$ (and consequently between $\beta_j$ and $\phi$ in a
  regression setting), leading to better both computation and asymptotic
  (normal-based) inference \citep{Ross1970}; and ii) interpretability of
  the $\beta_j$ in terms of the rate ratios since
  $\mu_i\approx \text{E}(Y_i)$. Moreover, we characterized the
  COM-Poisson distribution in terms of the dispersion, zero-inflation
  and heavy-tail indexes that illustrate its applicability in the real
  count data. These are the methodological novelties of the article
  besides the exploration of the results. However, in the updated
  version of the paper, we provided a better discussion of the case
  studies.
\end{enumerate}

\section{Replies to Referee 2}

\textbf{\underline{Second paragraph}:} \textit{The authors suggest the
  bijective reparameterization, $(\mu_i, \phi) = k(\lambda_i, \nu)$,
  where $\mu_i = \lambda_i^{1/\nu} - (\nu-1)/(2\nu)$. (Equation 3.1 plus
  the sentence after it gives this bijection. Note: The notation
  $h(\lambda, \nu)$, should probably be changed to $h_\nu(\lambda)$ to
  better hint that the symbol $h^{-1}$ refers to the inverse with
  respect to $\lambda$, with $\nu$ fixed.) The symbol $\mu_i$ is
  presumably used to remind the analyst that this parameter is, by
  Shmueli et al., approximately equal to $E(Y_i)$; we could write
  $m(\lambda_i, \nu) = E(Y_i) \approx \mu_i \equiv aE(Y_i)$. Shmueli et
  al. also showed that $var(Y_i)\approx \lambda_i^{1/\nu}\equiv a
  var(Y_i)$.}\\[0.1cm]
\textbf{\underline{Response:}} Thank you for this comment. We have
changed the notation of the reparameterization function in the updated
version.
\vspace{0.4cm}

\textbf{\underline{Third paragraph}:} \textit{The authors also hint that
  $var(Y_i)\approx \mu_i \exp(-\phi)$ (see the quadratic error analysis
  leading to Fig. 1). This is curious given that
  $avar(Y_i)=\mu\exp(-\phi) + (\exp(\phi)-1)/(2\exp(2\phi))$ and the
  latter summand converges to $-\infty$ as $\phi\to-\infty$. This latter
  summand is 0 or close to 0 for $\phi=0$ or $\phi>0$. Taken together,
  these last two sentences imply that the $\mu_i\exp(-\phi)$ may be a
  reasonable approximation to the variance when there is equi- or under-
  dispersion, but it may be unreasonable when there is lots of
  overdispersion. It seems that some of the results, simulations, and
  sample analyses reflect this (e.g. Fig 3, Fig 6-8, ``deviance function
  shape under strong overdispersion $\phi=-1.6$ is not as well
  behaved...'', p. 19, etc.). This needs to be addressed.}\\[0.3cm]
\textbf{\underline{Response:}} We thanks the referee for exploring this
point. However, we didn't use the approximation of variance
($a\text{Var}(Y_i)$) to repametrize or fit the COM-Poisson model. To
compute de variances, that is showed in the Figure 3(a) and (b) we used
$Var(Y)=\sum_{y=0}^{500} y^2p(y) - [\sum_{y=0}^{500} y p(y)]^2$. As
highlighted in the Figure 1(b), the approximation is not accurate,
mainly for overdispersion ($\phi<0$). The Figure (a) and (b) below shows
the behaviour of $a\text{Var}(Y_i)$ function by fixing $\mu$ and $\phi$,
repectivaly, illustrating the referees' statement. The $a\text{Var}(Y)$
leads to negative values for variance for small $\mu$ and $\phi<0$ that
shows that the approximation is not reasonable for this region of the
parameter space.

\hspace*{-0.0cm}
\begin{minipage}{\textwidth}
<<approx-var, fig.width=9, fig.height=3.5, cache=TRUE>>=
@
\end{minipage}
\vspace{0.3cm}

The results shown in Figures 7 and 8 reflect that orthogonality property
is slightly lost for strong overdispersion and small $\mu$. This is
related to accuracy of the $aE(Y_i)$. When $aE(Y_i)$ is accurate, $\mu$
represents the expectation of the $Y_i$ and, consequently, $\mu$ and
$\phi$ (or $\beta_j$ and $\phi$ in a regression setting) are
orthogonal. We improved the discussion about this in the current version
of the article.
\vspace{0.4cm}

\textbf{\underline{Fifth paragraph}:} \textit{This last point also hints
  that a quasi-Poisson model based on the assumption that
  $var(Y_i)\propto\mu_i$ (see p. 14) may not be reasonable for analyzing
  overdispersed data if the $COMPo$ model truly holds. Is this the
  case?}\\[0.1cm]
\textbf{\underline{Response:}} The quasi-Poisson model is specified by
second-moment assumptions (expectation and variance). As highlighted in
Figure 3(a) (and in referee's previous comment), the mean-variance
relationship for COM-Poisson is linear. Therefore, is suitable to
analyse data generated according to the COM-Poisson distribution, using
the assumption $Var(Y_i)\propto \mu_i$. The advantage of the COM-Poisson
approach is it corresponds to a fully specified probability model
allowing to compute the likelihood and its resulting results (deviance,
AIC, BIC).
\vspace{0.4cm}

\textbf{\underline{Sixth paragraph}:} \textit{Consider the example of
  Section 6.1. If I understand matters correctly, it is surprising that
  the $\beta$ estimates under the
  $[COMPo(\lambda_i, \nu), g(m(\lambda_i, \nu)) = \beta^\top x_i]$ model
  are so different from those under the
  $[COMPo(\mu_i, \phi), g(\mu_i) = \beta^\top x_i]$. See the results in
  Table 2, for example. I would think that if the approximation
  $m(\lambda_i, \nu) \approx \mu_i$ is reasonable, which your quadratic
  error results seem to indicate (see Fig 1), then the ML estimates of
  $\beta$ would be very similar. Exactly what models are being fitted
  here? This needs clarification.}\\[0.1cm]
\textbf{\underline{Response:}} Note the $\lambda_i$ does not represente
the expectation, nor approximately. The quadratic erros in Figure 1(a)
is obtained by $[a\text{E(Y)} - \sum_{y=0}^{500}y p(y)]^2$.  The models
fitted in the case studies are
$[\text{COMPo}(\lambda_i, \nu), \log(\lambda_i) = \beta^\top x_i]$ and
$[\text{COMPo}(\mu_i, \phi), \log(\mu_i) = \beta^\top x_i]$. Therefore,
the $\beta$ estimates are only comparable if $\phi=0$ (Poisson special
case) otherwise the $\beta$ estimates are on different scales. We make
it clear which parameters are being placed the linear predictor in
the current version of the paper. To understand the relationship of the
coefficients, consider $\log(\lambda_i) = \beta^o$ and $\log(\mu_i) =
\beta^*$, so $\beta^*$ expressed in terms of $\beta^o$ is
$\beta^* = \exp(\beta^o) \exp{-\phi} - (\exp(\phi) - 1)/(2\exp(\phi))$.
\vspace{0.4cm}

\textbf{\underline{Specific reviewer's comments}:}

\begin{enumerate}
\item \textit{Consider the constraint ``$\mu > 0$'', after equation
    (3.2). What constraints does this impose on the $\lambda$ and
    $\nu$. Is this reasonable?}\\[0.1cm]
  \textbf{\underline{Response:}} The constraint $\mu>0$ implies that
  $\lambda > [(\nu-1)/(2\nu)]^\nu$. The Figure below show (a) the
  constraint parameter space; and (b) the expectated values
  (obtained by $\sum y p(y)$). The constraint lies on the very small
  $\lambda$ and $\nu>1$ (underdispersion). This parameter region is
  related to small expected values (smaller than 0.1) and
  underdispersion. Therefore, although this constraint is undesirable,
  it does not prejudice the application of the model.

\hspace*{-1.0cm}
\begin{minipage}{\textwidth}
<<explore-constrain, fig.width=8, fig.height=4, cache=TRUE>>=
@
\end{minipage}
\vspace{0.3cm}

We thank the referee for highlighting this point. We have discussed it
in the updated version.

\item \textit{The Shmueli approximations to the mean and variance hold
    under certain conditions. Are those conditions met in practice? For
    your examples?}\\[0.1cm]
  \textbf{\underline{Response:}} We used only the mean approximation
  given by Shmueli to reparametrize the COM-Poisson model. Moreover, as
  discussed in Figure 1(a), the the errors are close to 0 for the
  parameter grid evaluated and present no clear relation with regions
  gives by \cite{Shmueli2005} ($\phi \leq 0$ and
  $\mu > 10 - (\exp(\phi) - 1)/(2\exp(\phi))$. Except for
  counts with small averages (close to 0), the reparametrized model fits
  well. For examples presented in the article, the parameter estimates
  are summarizes below.

<<contraints-case-studies, results="asis", cache=TRUE>>=
@

\item \textit{The symbol $\mu$ is used to represent both $E(Y)$ and the
    parameter $\lambda^{1/\nu}-(\nu-1)/(2\nu) = aE(Y)$. This causes
    confusion. Is there a better alternative? As an example, in
    Simulation study of Section 5, at the bottom of p 14, does the
    symbol $\mu$ represent $E(Y)$ or $aE(Y)$? I assume the
    former.}\\[0.1cm]
  \textbf{\underline{Response:}} Thank you for advertising this
  misunderstanding. At the bottom of page 14 the $\bm{\mu}$ is
  $a\text{E}(\bm{Y})$. We keep symbol $\mu$ to the introduced parameter
  in order to highlight that it is related to the expectation. However,
  we make clear that $\mu_i = a\text{E}(Y_i)$ to avoid misunderstanding.

\item \textit{The abstract could be improved by making it more concise
    (shorter).  It should be written from the third-person perspective
    as well.}\\[0.1cm]
\textbf{\underline{Response:}}

\item \textit{I assume that the “38\% faster” algorithm means that,
    e.g., instead of 10 seconds it takes just 6.2 seconds. As this is
    not an order of magnitude, this may not be enough to warrant a
    change to the “simpler” parameterization given that this latter
    parameterization models $aE(Y_i)$ rather than $E(Y_i)$.}\\[0.1cm]
  \textbf{\underline{Response:}} The time to fit the
  COM-Poisson$(\mu_i, \phi)$ model was compared to the time to fit the
  COM-Poisson$(\lambda_i, \phi)$ model. The computational times for 50
  repetitions of fit of the models in the three case studies are
  presented in Figure below.

\hspace*{-1.0cm}
\begin{minipage}{\textwidth}
<<computation-times, fig.width=9, fig.height=3.5, cache=TRUE>>=
@
\end{minipage}
\vspace{0.3cm}

The times to adjust under the proposed parameterization are 110\% faster
than the original, in the overdispersed case. In addition, the nice
properties induced by the new parameterization like the orthogonality
between $\mu$ and $\phi$ and interpretation of $\mu$ are the main
advantages to warrant the use of the new parameterization besides
computational times.

\end{enumerate}

\bibliography{../references.bib}

\end{document}

### 2018-08-20 review/review.R

@@ -1,0 +1,231 @@
+#-----------------------------------------------------------------------
+# Load package and functions
+
+## ---- setup
+library(knitr)
+library(xtable)
+options(digits = 3, OutDec = ".",
+        xtable.caption.placement = "top",
+        xtable.booktabs = TRUE,
+        xtable.sanitize.text.function = identity,
+        xtable.size = "small",
+        xtable.math.style.negative = TRUE)
+opts_chunk$set(
+    warning = FALSE,
+    message = FALSE,
+    echo = FALSE,
+    results = "hide",
+    fig.width = 7,
+    fig.height = 5,
+    out.width = "1\\textwidth",
+    fig.align = "center",
+    fig.pos = "H",
+    # dev = "tikz"
+    dev.args = list(family = "Palatino")
+    )
+
+## ---- load-packages
+source("../codes/lattice-panels.R")
+source("../codes/functions.R")
+library(purrr)
+library(dplyr)
+library(tidyr)
+cols <- trellis.par.get("superpose.line")$col
+
+## ---- counts-ranging
+rcmp <- Vectorize(FUN = function(n, mu, phi) {
+    par <- mu2lambda(mu, phi)
+    with(par, compoisson::rcom(n = n, lambda, nu))
+}, vectorize.args = "mu")
+
+
+phis <- c(-1.6, -1, 0, 1.8)
+names(phis) <- sprintf("phi=%s", phis)
+fl <- parse(text = gsub("=", "==", names(phis)))
+da <- map_dfr(phis, function(p) {
+    tibble(phi = factor(p, levels = phis),
+           y   = c(rcmp(1000, 27, p)))
+})
+
+histogram(~y | phi,
+          type="density",
+          layout = c(4, 1),
+          axis = axis.grid,
+          breaks = seq(0, max(da$y) + 1, 1),
+          strip = strip.custom(factor.levels = fl),
+          panel = function(x, ...) {
+              panel.histogram(x, ...)
+          },
+          data = da)
+
+
+## ----approx-var
+compute_apvar <- function(mu, phi) {
+    mu * exp(-phi) + (exp(phi) - 1) / (2 * exp(2 * phi))
+}
+
+mu <- seq(1, 30, length.out = 500)
+mus <- c(2, 8, 20)
+names(mus) <- sprintf("mu=%s", mus)
+
+phi <- seq(-3.5, 5, length.out = 500)
[…]

### 2018-08-20 article.Rnw
Externalize codes of the article.Rnw file

@@
 
+<<include=FALSE>>=
+knitr::read_chunk("article.R")
+@
+
 <<setup, include=FALSE>>=
-
-library(knitr)
-library(xtable)
-options(digits = 3, OutDec = ".",
-        xtable.caption.placement = "top",
-        xtable.booktabs = TRUE,
-        xtable.sanitize.text.function = identity,
-        xtable.size = "small",
-        xtable.math.style.negative = TRUE)
-opts_chunk$set(
-    warning = FALSE,
-    message = FALSE,
-    echo = FALSE,
-    results = "hide",
-    fig.width = 7,
-    fig.height = 5,
-    out.width = "1\\textwidth",
-    fig.align = "center",
-    fig.pos = "!htp",
-    # dev = "tikz"
-    dev.args = list(family = "Palatino")
-    )
-
-divide_matrix <- function(mat, divide) {
-    mat1 <- mat[, 1:divide]
-    mat2 <- mat[, (divide + 1):ncol(mat)]
-    nc1 <- ncol(mat1)
-    nc2 <- ncol(mat2)
-    cdiff <- abs(ncol(mat1) - ncol(mat2))
-    if (cdiff) {
-        complete <- matrix(NA, nrow = nrow(mat), ncol = cdiff)
-        colnames(complete) <- rep(" ", cdiff)
-        if (nc1 < nc2) {
-            mat1 <- cbind(mat1, complete)
-        } else {
-            mat2 <- cbind(mat2, complete)
-        }
-    }
-    return(list(mat1, mat2))
-}
-
-source("./codes/lattice-panels.R")
-source("./codes/functions.R")
-
-# Colors for legends
-cols <- trellis.par.get("superpose.line")$col
-
-# Useful packages
-library(bbmle)
-library(multcomp)
-library(plyr)
-library(tidyr)
-library(dplyr)
-
 @
@@
 <<convergenceZ, echo=FALSE>>=
-
-#-----------------------------------------------------------------------
-# Convergence of Z(lambda, nu) constant
-computeZ <- function(lambda, nu, maxit = 1e4, tol = 1e-5) {
-    z <- vector("numeric", maxit)
-    j = 1
-    z[j] <- exp(j * log(lambda) - nu * lfactorial(j))
-    while (abs(z[j] - 0) > tol && j <= maxit) {
-        j = j + 1
-        z[j] <- exp(j * log(lambda) - nu * lfactorial(j))
-    }
-    if (lambda == 1 & nu == 0) z <- Inf
-    return(sum(z, na.rm = TRUE) + 1)
-}
-
-grid <- expand.grid(
-    lambda = c(0.5, 1, 5, 10, 30, 50),
-    nu = seq(0, 1, length.out = 11))
-grid$z <- apply(grid, 1, function(par) {
-    computeZ(lambda = par[1], nu = par[2], maxit = 1e4)
-})
-
-xt <- xtabs(z ~ nu + lambda, data = grid)
-
 @
@@
 
-<<results="asis">>=
-caption <- paste(
-    "Values for $Z(\\lambda, \\nu)$ constant (numerically computed)",
-    "for values of $\\lambda$ (0.5 to 50) and $\\phi$ (0 to 1)")
-
-# Remove excessive scientific notation
-xt2 <- gsub("e\\+00", "    ", format(xt))
-xt2 <- gsub("Inf", "divergent$^{**}$", format(xt2))
-xt2[1, -1] <- "divergent$^{*\\,\\,\\,}$"
-
-print(xtable(xt2, digits = -2,
-             caption = caption,
-             align = "C|CCCCCC",
-             # align = "c|cccccc",
-             label = "tab:convergenceZ"),
-      include.colnames = FALSE,
-      tabular.environment = "tabularx",
-      only.contents = TRUE,
-      width = "\\textwidth",
-      add.to.row = list(
-          pos = list(0, 0),
-          command = c(
-              "& \\multicolumn{6}{c}{$\\bm{\\lambda}$} \\\\\n",
-              "$\\bm{\\nu}$ & 0.5 & 1 & 5 & 10 & 30 & 50 \\\\\n")
-      ))
-
+<<results-Z, results="asis">>=
 @
@@
 <<data-approx, include=FALSE, cache=TRUE>>=
-
-#-----------------------------------------------------------------------
-# Study the approximation
-
-#-------------------------------------------
-# Mean and variance relationship
-aux <- expand.grid(
-    mu = seq(3, 30, length.out = 55),
-    phi = seq(-1.5, 1.8, length.out = 50))
-
-moments <- mapply(FUN = calc_moments,
-                  mu = aux$mu,
-                  phi = aux$phi,
-                  MoreArgs = list(sumto = 300),
-                  SIMPLIFY = FALSE)
-grid <- cbind(aux, t(do.call(cbind, moments)))
-grid <- transform(grid, va = mu / exp(phi))
-
 @
@@
 Dotted lines represent the restriction for suitable approximations given by \\cite{Shmueli2005}.">>=
-
-#-------------------------------------------
-# Errors in approximations for E[Y] and V[Y]
-grid <- transform(grid,
-                  emu = (mu - mean)^2,
-                  eva = (va - var)^2)
-
-myreg <- colorRampPalette(c("gray90",  "gray20"))
-xy1 <- levelplot(emu ~ phi + mu, data = grid,
-                 aspect = "fill",
-                 col.regions = myreg,
-                 xlab = expression(phi),
-                 ylab = expression(mu),
-                 sub = "(a)",
-                 colorkey = list(space = "top"),
-                 par.settings = ps2,
-                 panel = function(x, y, z, ...) {
-                     panel.levelplot(x, y, z, ...)
-                     panel.curve(10 - (exp(x) - 1)/(2 * exp(x)),
-                                 lty = 2)
-                     panel.abline(v = 0, lty = 2)
-                 })
-
-xy2 <- levelplot(eva ~ phi + mu, data = grid,
-                 aspect = "fill",
-                 col.regions = myreg,
-                 xlab = expression(phi),
-                 ylab = expression(mu),
-                 sub = "(b)",
-                 colorkey = list(space = "top"),
-                 par.settings = ps2,
-                 panel = function(x, y, z, ...) {
-                     panel.levelplot(x, y, z, ...)
-                     panel.curve(10 - (exp(x) - 1)/(2 * exp(x)),
-                                 lty = 2)
-                     panel.abline(v = 0, lty = 2)
-                 })
-
-print(xy1, split = c(1, 1, 2, 1), more = TRUE)
-print(xy2, split = c(2, 1, 2, 1), more = FALSE)
-
 @
@@
 <<pmf-cmp, fig.height=3, fig.width=7, fig.cap="Shapes of the COM-Poisson distribution for different parameter values.">>=
-
-#-------------------------------------------
-# COM-Poisson probabilities
-parg <- expand.grid(mu = c(2, 8, 15), phi = c(-0.7, 0, 0.7))
-y <- 0:30
-py <- mapply(FUN = dcmp,
-             mu = parg$mu,
-             phi = parg$phi,
-             MoreArgs = list(y = y, sumto = 100),
-             SIMPLIFY = FALSE)
-parg <- cbind(parg[rep(1:nrow(parg), each = length(y)), ],
-              y = y, py = unlist(py))
-
-leg_phi <- parse(
-    text = paste("phi == \"",
-                 formatC(unique(parg$phi), 1, format = "f"),
-                 "\""))
-barchart(py ~ y | factor(mu),
-         groups = factor(phi),
-         data = parg,
-         horizontal = FALSE,
-         layout = c(NA, 1),
-         as.table = TRUE,
-         axis = axis.grid,
-         origin = 0,
-         xlim = extendrange(y, f = 0.01),
-         border = "transparent",
-         # scales = list(x = list(at = pretty(y))),
-         ylab = expression(P(Y==y)),
-         xlab = expression(y),
-         par.settings = list(
-             superpose.polygon = list(
-                 col = c("gray40", "gray60", "gray10")),
-             superpose.line = list(
-                 col = c("gray40", "gray60", "gray10"),
-                 lwd = 2)
-         ),
-         auto.key = list(
-             columns = 3,
-             rectangles = FALSE,
-             lines = TRUE,
-             text = leg_phi
-         ),
-         strip = strip.custom(
-             strip.names = TRUE,
-             var.name = expression(mu == ""),
-             sep = ""))
-
 @
@@
 <<compute-indexes, include=FALSE, cache=TRUE>>=
-
-# Dispersion index
-grid <- transform(grid, dispersion_index = var / mean)
-
-# Zero-inflation index
-prob0 <- mapply(FUN = dcmp,
-                mu = grid$mu,
-                phi = grid$phi,
-                MoreArgs = list(y = 0, sumto = 300),
-                SIMPLIFY = FALSE)
-prob0 <- unlist(prob0)
-grid <- transform(grid, zero_index = 1 + log(prob0) / mean)
-
-# Heavy tail-index
-hi_fun <- Vectorize(FUN = function(x, mu, phi, sumto) {
-    probs <- dcmp(y = c(x, x + 1), mu = mu, phi = phi, sumto = sumto)
-    probs[2] / probs[1]
-}, vectorize.args = "x")
-
-x <- 40:150
-his <- mapply(FUN = hi_fun,
-              mu = grid$mu,
-              phi = grid$phi,
-              MoreArgs = list(x = x, sumto = 300),
-              SIMPLIFY = FALSE)
-
-aux1 <- lapply(his, function(k) data.frame("x" = x, "heavy_index" = k))
-aux1 <- do.call(rbind, aux1)
-aux2 <- grid[c("mu", "phi")][rep(1:nrow(grid), each = length(x)), ]
-heavy_data <- cbind(aux1, aux2)
-
 @
@@
 Dotted lines represents the Poisson special case.">>=
-
-#-----------------------------------------------------------------------
-# Study indexes on the reparametrized COM-Poisson distribution
-mygpar <- gpar(cex = 1.2)
-psaux <- modifyList(
-    ps0, list(superpose.line = list(
-                  col = myreg(length(unique(grid$phi)))),
-              layout.widths = list(
-                  left.padding = -0.5)
-              )
-)
-
-#-------------------------------------------
-# Mean and variance relationship
-xy1 <- xyplot(var ~ mu | "Mean-Variance",
-              groups = phi,
-              data = grid,
-              type = "l",
-              lwd = 2,
-              axis = axis.grid,
-              xlab = expression(mu),
-              ylab = "",
-              # ylab = expression(Var(Y)),
-              scales = list(y = list(rot = 90)),
-              sub = "(a)",
-              legend = list(
-                  top = list(
-                      fun = draw.colorkey,
-                      args = list(
-                          key = list(
-                              space = "top",
-                              col = myreg(length(unique(grid$phi))),
-                              at = unique(grid$phi),
-                              draw = FALSE)))),
-              par.settings = psaux,
-              panel = function(x, y, ...) {
-                  panel.xyplot(x, y, ...)
-                  panel.curve(1*x, min(x), max(x), lty = 2)
-              },
-              page = function(...) {
-                  grid.text(expression(phi),
-                            just = "bottom",
-                            gp = mygpar,
-                            x = unit(0.020, "npc"),
-                            y = unit(0.875, "npc"))
-              })
-
-#-------------------------------------------
-# Dispersion index (DI)
-xy2 <- xyplot(var / mean ~ mu | "Dispersion index",
-              groups = phi,
-              data = grid,
-              type = "l",
-              lwd = 2,
-              axis = axis.grid,
-              xlab = expression(mu),
-              ylab = "",
-              scales = list(y = list(rot = 90)),
-              sub = "(b)",
-              legend = list(
-                  top = list(
-                      fun = draw.colorkey,
-                      args = list(
-                          key = list(
-                              space = "top",
-                              col = myreg(length(unique(grid$phi))),
-                              at = unique(grid$phi),
-                              draw = FALSE)))),
-              par.settings = psaux,
-              panel = function(x, y, ...) {
-                  panel.xyplot(x, y, ...)
-                  panel.curve(1 + 0*x, min(x), max(x), lty = 2)
-              },
-              page = function(...) {
-                  grid.text(expression(phi),
-                            just = "bottom",
-                            gp = mygpar,
-                            x = unit(0.020, "npc"),
-                            y = unit(0.875, "npc"))
-              })
-
-#-------------------------------------------
-# Zero-inflation index
-xy3 <- xyplot(zero_index ~ mu | "Zero-inflation index",
-              groups = phi,
-              data = grid,
-              type = "l",
-              lwd = 2,
-              axis = axis.grid,
-              xlab = expression(mu),
-              ylab = "",
-              scales = list(y = list(rot = 90)),
-              sub = "(c)",
-              legend = list(
-                  top = list(
-                      fun = draw.colorkey,
-                      args = list(
-                          key = list(
-                              space = "top",
-                              col = myreg(length(unique(grid$phi))),
-                              at = unique(grid$phi),
-                              draw = FALSE)))),
-              par.settings = psaux,
-              panel = function(x, y, ...) {
-                  panel.xyplot(x, y, ...)
-                  panel.curve(0*x, min(x), max(x), lty = 2)
-              },
-              page = function(...) {
-                  grid.text(expression(phi),
-                            just = "bottom",
-                            gp = mygpar,
-                            x = unit(0.020, "npc"),
-                            y = unit(0.875, "npc"))
-              })
-
-#-------------------------------------------
-# Heavy tail-index
-# Choose one specific mu of unique(grid$mu)
-choosemu <- 25
-xy4 <- xyplot(heavy_index ~ x | "Heavy-tail index",
-              groups = phi,
-              # data = heavy_data,
-              data = subset(heavy_data, mu == choosemu),
-              type = "l",
-              lwd = 2,
-              axis = axis.grid,
-              xlab = expression(y),
-              ylab = "",
-              scales = list(y = list(rot = 90)),
-              sub = "(d)",
-              legend = list(
-                  top = list(
-                      fun = draw.colorkey,
-                      args = list(
-                          key = list(
-                              space = "top",
-                              col = myreg(length(unique(grid$phi))),
-                              at = unique(grid$phi),
-                              draw = FALSE)))),
-              par.settings = psaux,
-              panel = function(x, y, ...) {
-                  panel.xyplot(x, y, ...)
-                  panel.curve(choosemu / (x + 1), from = min(x),
-                              to = max(x), lty = 2)
-              },
-              page = function(...) {
-                  grid.text(expression(phi),
-                            just = "bottom",
-                            gp = mygpar,
-                            x = unit(0.020, "npc"),
-                            y = unit(0.875, "npc"))
-              })
-
-print(xy1, split = c(1, 1, 4, 1), more = TRUE)
-print(xy2, split = c(2, 1, 4, 1), more = TRUE)
-print(xy3, split = c(3, 1, 4, 1), more = TRUE)
-print(xy4, split = c(4, 1, 4, 1), more = FALSE)
-
-midi <- min(grid$dispersion_index)
-madi <- max(grid$dispersion_index)
-
 @
@@
 <<load-simulation, cache=TRUE>>=
-
-# Configuration
-B <- 1000
-beta <- c("b0" = 2, "b1" = 0.5, "b21" = 0.8, "b22" = -0.8)
-phis <- c(0, -1.6, -1, 1.8)
-names(phis) <- sprintf("phi=%s", phis)
-
-sizes <- c(50, 100, 300, 1000)
-names(sizes) <- sprintf("n=%s", sizes)
-
-# Load results
-results <- readRDS("./codes/simulation.rds")
-
 @
@@
 <<justpars, fig.height=4, fig.width=9, fig.cap="Average counts (left) and dispersion indexes (right) for each scenario considered in the simulation study.">>=
-
-#-------------------------------------------
-# Justify the choices of the parameters simulation
-x1 <- seq(0, 1, length.out = 100)
-x2 <- rep(letters[1:3], length.out = 100)
-X <- model.matrix(~ x1 + x2)
-mu <- exp(X %*% beta)
-
-daexplain <- ldply(lapply(phis, function(phi) {
-    va <- compute_variance(mu, phi, sumto = 300)
-    data.frame(mu = mu, va = va, di = va / mu, x1 = x1, x2 = x2)
-}), .id = "phi")
-
-labx2 <- expression("Level of categorical variable"~(x[2]))
-labx1 <- expression("Values of continous variable"~(x[1]))
-
-fl <- parse(text = gsub("=", "==", levels(daexplain$phi)))
-xy1 <- xyplot(mu ~ x1, groups = x2,
-              type = c("g", "l"),
-              lwd = 2,
-              xlab = labx1,
-              ylab = "Average count",
-              auto.key = list(
-                  column = 3,
-                  points = FALSE,
-                  lines = TRUE,
-                  title = labx2,
-                  cex.title = 1.1
-              ))
-xy2 <- xyplot(di ~ x1 | phi, groups = x2,
-              type = c("g", "l"),
-              lwd = 2,
-              # scales = "free",
-              layout = c(2, 2),
-              xlab = labx1,
-              ylab = "Dispersion index",
-              data = daexplain,
-              strip = strip.custom(factor.levels = fl))
-
-print(xy1, split = c(1, 1, 2, 1), more = TRUE)
-print(xy2, split = c(2, 1, 2, 1), more = FALSE)
-
 @
@@
 <<bias-data, cache=TRUE>>=
-
-#-----------------------------------------------------------------------
-# Compute standardized bias
-std <- lapply(results["n=50"], function(x) {
-    ind <- names(x); names(ind) <- ind
-    lapply(ind, function(y) {
-        bhat <- t(vapply(x[[y]], "[[", double(5), "coef"))
-        real <- matrix(c(phis[y], beta), byrow = TRUE,
-                       nrow = B, ncol = 5)
-        matrix(apply(bhat - real, 2, sd, na.rm = TRUE),
-               byrow = TRUE, nrow = B, ncol = 5)
-    })
-})
-
-aux <- ldply(lapply(results, function(x) {
-    ind <- names(x); names(ind) <- ind
-    out <- lapply(ind, function(y) {
-        bhat <- t(vapply(x[[y]], "[[", double(5), "coef"))
-        real <- matrix(c(phis[y], beta), byrow = TRUE,
-                       nrow = B, ncol = 5)
-        # (bhat - real)                 # raw bias
-        (bhat - real) / std[[1]][[y]] # standardized bias
-    })
-    ldply(out, .id = "phi")
-}), .id = "n")
-
-# Organize the results
-aux <- na.omit(aux)
-bias <- gather(aux, param, bias, phi2:b22, factor_key = TRUE)
-bias$phi <- ordered(bias$phi, c("phi=-1.6", "phi=-1", "phi=0",
-                                "phi=1.8"))
-
 @
@@
 <<bias-plot, cache=TRUE, fig.height=6.5, fig.width=8.5, fig.cap="Distributions of standardized bias (gray box-plots) and average with confidence intervals (black segments) by different sample sizes and dispersion levels.">>=
-
-# Distributions of bias and average with confidence intervals
-ci <- function(x) {
-    ci <- mean(x) + c(-1, 0, 1) * qnorm(0.975) * sd(x)
-    names(ci) <- c("lwr", "fit", "upr")
-    ci
-}
-cidata <- aggregate(bias ~ n + phi + param, data = bias, ci)
-
-key <- list(
-    type = "o",
-    divide = 1,
-    columns = 4,
-    title = "Sample size",
-    cex.title = 1.1,
-    lines = list(pch = c(21:24), cex = 0.8),
-            text = list(names(sizes))
-)
-
-fl <- parse(text = gsub("=", "==", levels(bias$phi)))
-yl <- parse(text = c("hat(phi)",
-                     paste0("hat(beta)[", c(0, 1, 21, 22), "]")))
-
-# Alternative graph (figure with all point is very large in size memory)
-bwplot(param ~ bias | phi, groups = n, data = bias,
-       pch = "|",
-       grid = TRUE,
-       as.table = TRUE,
-       layout = c(4, 1),
-       xlab = "Standardized Bias",
-       fill = "gray80",
-       col = "red",
-       strip = strip.custom(factor.levels = fl),
-       panel = panel.superpose,
-       scales = list(y = list(labels = yl)),
-       box.width = 0.05,
-       whisker.width = 0.4,
-       key = key,
-       par.settings = list(
-           plot.symbol = list(pch = 20, cex = 0.5, col = "gray70"),
-           box.rectangle = list(col = "gray70"),
-           box.umbrella=list(col = "gray70", lwd = 2)
-       ),
-       panel.groups = function(x, y, group.number, ...) {
-           my.panel.bwplot(x, (y - 0.05) + (group.number - 2.5) / 5, ...)
-           panel.abline(v = 0, lty = 2)
-       }) +
-    segplot(param ~ bias[, "lwr"] + bias[, "upr"] | phi,
-            centers = bias[, "fit"],
-            data = cidata,
-            draw = FALSE,
-            horizontal = TRUE,
-            layout = c(4, 1),
-            groups = n, gap = 0.3,
-            key = key,
-            pch = 21:24,
-            cex = 0.7,
-            lwd = 2,
-            panel = function(...) {
-                panel.groups.segplot(...)
-                panel.abline(v = 0, lty = 2)
-                panel.abline(h = 1:5, col = "lightgray", lty = 2)
-            })
-
-# xyplot(param ~ bias | phi, groups = n, data = bias,
-#        panel = panel.superpose,
-#        layout = c(4, 1),
-#        key = key,
-#        strip = strip.custom(factor.levels = fl),
-#        scales = list(y = list(labels = yl)),
-#        ylab = "",
-#        xlab = "Standardized Bias",
-#        jitter.y = TRUE,
-#        factor = 0.1,
-#        gap = 0.3,
-#        alpha = 0.6,
-#        col = "gray80",
-#        cex = 0.7,
-#        panel.groups = function(x, y, group.number,
-#                                subscripts, gap, ...){
-#            noise <- centfac(factor(levels(bias$n)), space = gap)
-#            noise <- sort(noise)
-#            panel.xyplot(x, y + noise[group.number], ...)
-#        }) +
-#     as.layer(
-#         segplot(param ~ bias[, "lwr"] + bias[, "upr"] | phi,
-#                 centers = bias[, "fit"],
-#                 data = cidata,
-#                 draw = FALSE,
-#                 horizontal = TRUE,
-#                 layout = c(4, 1),
-#                 groups = n, gap = 0.3,
-#                 key = key,
-#                 pch = 21:24,
-#                 cex = 0.7,
-#                 lwd = 2,
-#                 panel = function(...) {
-#                     panel.groups.segplot(...)
-#                     panel.abline(v = 0, lty = 2)
-#                     panel.abline(h = 1:5, col = "lightgray", lty = 2)
-#                 })
-#     )
-
 @
@@
 <<coverage-data>>=
-
-#-----------------------------------------------------------------------
-# Compute coverage rate
-aux <- ldply(lapply(results, function(x) {
-    ind <- names(x); names(ind) <- ind
-    out <- lapply(ind, function(y) {
-        ind <- lapply(x[[y]], function(z) {
-            real <- c(phis[[y]], beta)
-            cint <- z[["cint"]]
-            ind <- as.integer(cint[, 1] < real & cint[, 2] > real)
-            names(ind) <- c("phi2", names(beta))
-            ind
-        })
-        do.call("rbind", ind)
-    })
-    ldply(out, .id = "phi")
-}), .id = "n")
-
-# Organize the results
-aux <- na.omit(aux)
-coverage <- gather(aux, param, coverage, phi2:b22, factor_key = TRUE)
-coverage$phi <- ordered(coverage$phi, c("phi=-1.6", "phi=-1", "phi=0",
-                                        "phi=1.8"))
-coverage$n <- as.numeric(gsub("n=([0-9]+)", "\\1", coverage$n))
-covdata <- aggregate(coverage ~ n + phi + param, data = coverage,
-                     function(x) sum(x == 1) / length(x))
-
 @
@@
 <<coverage-plot, fig.height=4, fig.width=9, fig.cap="Coverage rate based on confidence intervals obtained by quadratic approximation for different sample sizes and dispersion levels.">>=
-
-fl <- parse(text = gsub("=", "==", levels(bias$phi)))
-yl <- parse(text = c("hat(phi)",
-                     paste0("hat(beta)[", c(0, 1, 21, 22), "]")))
-xyplot(coverage ~ n | param,
-       type = c("g", "p", "l"),
-       groups = phi,
-       data = covdata,
-       pch = 19,
-       lwd = 1.5,
-       layout = c(NA, 1),
-       xlab = "Sample size",
-       ylab = "Coverage rate",
-       par.settings = list(
-           layout.heights = list(strip = 1.5)
-       ),
-       key = list(
-           column = 4,
-           type = "o",
-           divide = 1,
-           text = list(fl),
-           lines = list(pch = 19, col = cols[1:4])
-       ),
-       strip = strip.custom(
-           factor.levels = yl
-       ),
-       panel = function(x, y, ...) {
-           panel.xyplot(x, y, ...)
-           panel.abline(h = 0.95, lty = 2)
-       })
-
 @
@@
 <<ortho-plot, fig.height=4.5, fig.width=9, fig.cap="Empirical correlations between regression and dispersion parameters by different sample sizes and dispersion levels.">>=
-
-#-----------------------------------------------------------------------
-# Obtain the covariance between dispersion and regression parameters
-
-aux <- ldply(
-    lapply(results, function(x) {
-    ind <- names(x); names(ind) <- ind
-    out <- lapply(ind, function(y) {
-        vcovs <- lapply(x[[y]], "[[", "vcov")
-        index <- vapply(lapply(vcovs, is.na), sum, integer(1))
-        vcovs <- vcovs[!index]
-        corrs <- lapply(vcovs, cov2cor)
-        covs <- do.call(rbind, lapply(vcovs, "[", 1, 2:5))
-        cors <- do.call(rbind, lapply(corrs, "[", 1, 2:5))
-        ldply(list("cov" = covs, "cor" = cors), .id = "scale")
-    })
-    ldply(out, .id = "phi")
-}), .id = "n")
-str(aux)
-
-# Organize the results
-dacov <- gather(aux, param, value, b0:b22, factor_key = TRUE)
-dacov$phi <- ordered(dacov$phi,
-                     c("phi=-1.6", "phi=-1", "phi=0", "phi=1.8"))
-str(dacov)
-
-# Distributions of covariance/correlations
-fl <- parse(text = gsub("=", "==", levels(dacov$phi)))
-yl <- parse(text = paste0("hat(beta)[", c(0, 1, 21, 22), "]"))
-
-# Boxplots
-bwplot(param ~ value | phi,
-       groups = n,
-       data = subset(dacov, scale == "cor"),
-       box.width = 0.12,
-       pch = "|",
-       grid = TRUE,
-       as.table = TRUE,
-       layout = c(4, 1),
-       xlab = "Empirical correlations with dispersion parameter estimator",
-       fill = colorRampPalette(c("gray80",  "gray20"))(4),
-       panel = panel.superpose,
-       scales = list(y = list(labels = yl), x = "free"),
-       whisker.width = 0.4,
-       strip = strip.custom(factor.levels = fl),
-       auto.key = list(
-           column = 4,
-           rectangles = TRUE,
-           points = FALSE,
-           title = "Sample size",
-           cex.title = 1
-       ),
-       par.settings = list(
-           plot.symbol = list(pch = 19, cex = 0.1)
-       ),
-       panel.groups = function(x, y, group.number, ...) {
-           my.panel.bwplot(x, y + (group.number - 2.5) / 5, ...)
-           panel.abline(v = 0, lty = 2)
-       })
-
 @
@@
 The ellipses are confidence regions (90, 95 and 99\\%), dotted lines are the maximum likelihood estimates, and points are the real parameters used in the simulation.">>=
-
-#-------------------------------------------
-# Plots deviance surfaces
-output <- readRDS("codes/orthogonality.rds")
-devs_orpar <- output$devs_orpar
-devs_repar <- output$devs_repar
-
-niveis <- c(0.9, 0.95, 0.99)
-cortes <- qchisq(niveis, df = 2)
-fl <- parse(text = paste("phi==", sort(phis)))
-
-# Original parametrization
-devs_orpar$fphi <- ordered(devs_orpar$realphi)
-xy1 <- useOuterStrips(
-    levelplot(
-        deviance ~ lambda + phi | fphi + "Original parametrization",
-        data = devs_orpar,
-        scales = list(y = list(rot = 90, relation = "free"),
-                      x = "free"),
-        cuts = 30,
-        xlab = expression(lambda),
-        ylab = expression(phi),
-        colorkey = FALSE,
-        panel = function(x, y, z, at, region, ...,
-                         subscripts = subscripts) {
-            ##
-            flambda <- devs_orpar$fitlambda[subscripts]
-            fphi <- devs_orpar$fitphi[subscripts]
-            ##
-            tlambda <- devs_orpar$reallambda[subscripts]
-            tphi <- devs_orpar$realphi[subscripts]
-            ##
-            panel.levelplot(x, y, z, at = at, region = TRUE,
-                            ..., subscripts = subscripts)
-            panel.contourplot(x, y, z, ..., at = cortes,
-                              contour = TRUE, region = FALSE,
-                              subscripts = subscripts)
-            panel.abline(v = flambda, h = fphi, lty = 2)
-            panel.points(x = tlambda, y = tphi,
-                         lty = 2, pch = 19)
-        }),
-    strip = strip.custom(factor.levels = fl),
-    strip.left = strip.custom(par.strip.text=list(cex = 0.85))
-)
-
-# Propose parametrization
-devs_repar$fphi <- ordered(devs_repar$realphi)
-xy2 <- useOuterStrips(
-    levelplot(
-        deviance ~ mu + phi2 | fphi + "Proposed parametrization",
-        data = devs_repar,
-        scales = list(y = list(rot = 90, relation = "free"),
-                      x = "free"),
-        cuts = 30,
-        xlab = expression(mu),
-        ylab = expression(phi),
-        colorkey = FALSE,
-        panel = function(x, y, z, at, region, ...,
-                         subscripts = subscripts) {
-            ##
-            fmu <- devs_repar$fitmu[subscripts]
-            fphi <- devs_repar$fitphi[subscripts]
-            ##
-            tmu <- devs_repar$realmu[subscripts]
-            tphi <- devs_repar$realphi[subscripts]
-            ##
-            panel.levelplot(x, y, z, at = at, region = TRUE,
-                            ..., subscripts = subscripts)
-            panel.contourplot(x, y, z, ..., at = cortes,
-                              contour = TRUE, region = FALSE,
-                              subscripts = subscripts)
-            panel.abline(v = fmu, h = fphi, lty = 2)
-            panel.points(x = tmu, y = tphi,
-                         lty = 2, pch = 19)
-        }),
-    strip = strip.custom(factor.levels = fl),
-    strip.left = strip.custom(par.strip.text=list(cex = 0.85))
-)
-
-# Organize plots
-print(xy1, position = c(0.0, 0.48, 1.0, 1.0), more = TRUE)
-print(xy2, position = c(0.0, 0.0, 1.0, 0.52), more = FALSE)
-
 @
@@
 <<fit-cotton, cache=TRUE, include=FALSE>>=
-
-#-----------------------------------------------------------------------
-# Load data
-# data(cottonBolls, package = "cmpreg")
-cottonBolls <- read.table("./data/cottonBolls.txt",
-                          header = TRUE, sep = "\t")
-cottonBolls$est <- ordered(
-    cottonBolls$est,
-    c("vegetative", "flower bud", "blossom", "boll", "boll open")
-)
-
-#-----------------------------------------------------------------------
-# Fit models
-mnames <- c("PO", "C1", "C2", "QP")
-
-# Predictor, following Zeviani et al. (2014)
-form1 <- ncap ~ est:(des + I(des^2))
-
-m1PO <- glm(form1, data = cottonBolls, family = poisson)
-time11 <- system.time(
-    m1C1 <- fitcm(form1, data = cottonBolls, model = "CP", sumto = 50)
-)
-time12 <- system.time(
-    m1C2 <- fitcm(form1, data = cottonBolls, model = "CP2", sumto = 50)
-)
-m1QP <- glm(form1, data = cottonBolls, family = quasipoisson)
-
-models.ncap <- list(m1PO, m1C1, m1C2, m1QP)
-names(models.ncap) <- mnames
-
-# Numbers of calls to loglik and numerical gradient
-c11 <- models.ncap$C1@details$counts
-c12 <- models.ncap$C2@details$counts
-
-# LRT between Poisson and COM-Poisson (test: phi == 0)
-lrt.ncap <- getAnova(m1PO, m1C2)
-
 @
@@
 <<results-cotton, results="asis">>=
-
-#-----------------------------------------------------------------------
-# Goodness of fit measures and estimate parameters
-
-# GoF measures
-measures.ncap <- lapply(models.ncap, function(x)
-    c("LogLik" = logLik(x), "AIC" = AIC(x), "BIC" = BIC(x)))
-
-# Get the estimates
-co1 <- coef(m1C2)
-est <- lapply(models.ncap, FUN = function(x) getCoefs(x))
-est.ncap <- do.call(cbind, est)
-
-# Organize in table
-pnames <- c("\\phi\\,,\\,\\sigma", "\\beta_0",
-            paste0("\\beta_{1", 1:5, "}"),
-            paste0("\\beta_{2", 1:5, "}"))
-rownames(est.ncap) <- paste0("$", pnames, "$")
-meds <- apply(do.call(cbind, measures.ncap), 1, function(x) {
-    x <- formatC(x, getOption("digits"), format = "f")
-    x <- gsub("NA", "-", x)
-    paste(paste0("\\multicolumn{2}{c}{$", x, "$}"),
-          collapse = " & ")
-})
-text_gof <- paste(paste(names(meds), "&", meds),
-                  collapse = "\\\\\n ")
-
-append_gof <- list(
-    pos = list(nrow(est.ncap)),
-    command = paste("\\specialrule{0.01em}{0.3em}{0.3em} \n",
-                    text_gof, "\\\\\n",
-                    "\\bottomrule"))
-print.xtable(xtable(est.ncap, digits = 4,
-                    label = "tab:coef-cotton"),
-             hline.after = 0,
-             only.contents = TRUE,
-             add.to.row = append_gof)
-
 @
@@
 <<pred-cotton, fig.height=3.5, fig.width=8.5, fig.cap="Scatterplots of the observed data and curves of fitted values with 95\\% confidence intervals as functions of the defoliation level for each growth stage.">>=
-
-#-----------------------------------------------------------------------
-# Prediction
-
-# Data for prediction
-pred <- with(cottonBolls,
-             expand.grid(
-                 est = levels(est),
-                 des = seq(min(des), max(des), l = 20)
-             ))
-qn <- qnorm(0.975) * c(fit = 0, lwr = -1, upr = 1)
-
-# Design matrix for prediction
-X <- model.matrix(update(form1, NULL~.), pred)
-
-# Considering Poisson
-aux <- exp(confint(
-    glht(m1PO, linfct = X), calpha = univariate_calpha())$confint)
-colnames(aux) <- c("fit", "lwr", "upr")
-aux <- data.frame(modelo = "Poisson", aux)
-predPO.ncap <- cbind(pred, aux)
-
-# Considering COM-Poisson
-aux <- predictcm(m1C1, newdata = X)
-aux <- data.frame(modelo = "COM-Poisson", aux)
-predC1.ncap <- cbind(pred, aux)
-
-# Considering COM-Poisson (mean parametrization)
-aux <- predictcm(m1C2, newdata = X)
-aux <- data.frame(modelo = "COM-Poisson2", aux)
-predC2.ncap <- cbind(pred, aux)
-
-# Considering Quasi-Poisson
-aux <- exp(confint(
-    glht(m1QP, linfct = X), calpha = univariate_calpha())$confint)
-colnames(aux) <- c("fit", "lwr", "upr")
-aux <- data.frame(modelo = "Quasi-Poisson", aux)
-predQP.ncap <- cbind(pred, aux)
-
-# Representing the confidence intervals
-pred.ncap <- rbind(predPO.ncap, predC1.ncap, predC2.ncap, predQP.ncap)
-
-# Legend
-key <- list(columns = 4,
-            cex = 0.9,
-            lines = list(col = 1, lty = rev(1:4)),
-            text = list(parse(
-                text = c("'Poisson'", "'COM-Poisson'",
-                         "'COM-Poisson'[mu]", "'Quasi-Poisson'"))
-                ))
-
-# Graph
-xyplot(ncap ~ des | est,
-       data = cottonBolls,
-       layout = c(NA, 1),
-       as.table = TRUE,
-       grid = TRUE,
-       type = "p",
-       xlab = "Artificial defoliation level",
-       ylab = "Number of bolls produced",
-       spread = 0.05,
-       key = key,
-       alpha = 0.6,
-       panel = panel.beeswarm) +
-    as.layer(
-        xyplot(fit ~ des | est,
-               auto.key = TRUE,
-               data = pred.ncap,
-               groups = modelo,
-               type = "l",
-               layout = c(NA, 1),
-               as.table = TRUE,
-               col = 1,
-               ly = pred.ncap$lwr,
-               uy = pred.ncap$ upr,
-               cty = "bands",
-               fill = "gray80",
-               alpha = 0.1,
-               panel = panel.superpose,
-               panel.groups = panel.cbH,
-               prepanel = prepanel.cbH,
-               lty = rev(1:4))
-    )
-
 @
@@
 <<corr-cotton, results="asis">>=
-
-#-----------------------------------------------------------------------
-# Correlation between estimates
-corr.ncap <- do.call("rbind",
-                     lapply(models.ncap[c("C1", "C2")],
-                            function(x) cov2cor(vcov(x))[1, -1]))
-
-# Organize on table
-rownames(corr.ncap) <- paste0("COM-Poisson", c("", "$_\\mu$"))
-colnames(corr.ncap) <- gsub("beta", "hat{\\\\beta}", pnames[-1])
-corrncap_aux <- divide_matrix(corr.ncap, divide = 6)
-
-# caption <- paste("Empirical correlations between $\\hat{\\phi}$ and",
-#                  "$\\hat{\\bm{\\beta}}$ for the two parametrizations",
-#                  "of COM-Poisson model fit to underdispersed data.")
-# print(xtable(corr.ncap,
-#              align = c("lrrrrrrrrrrr"),
-#              caption = caption,
-#              digits = 3,
-#              label = "tab:corr-cotton"),
-#       size = "small",
-#       sanitize.rownames.function = identity,
-#       sanitize.colnames.function = function(x) sprintf("$%s$", x))
-
 @
@@
   \toprule
-<<results="asis">>=
-
-print(xtable(corrncap_aux[[1]], digits = 4),
-      only.contents = TRUE,
-      hline.after = FALSE,
-      sanitize.colnames.function = function(x) sprintf("$%s$", x))
-cat("\\midrule", sep = "\n")
-
-print(xtable(corrncap_aux[[2]], digits = 4),
-      only.contents = TRUE,
-      sanitize.colnames.function = function(x) sprintf("$%s$", x))
-
+<<cres-coton, results="asis">>=
 @
@@
 <<fit-soy, cache=TRUE, include=FALSE>>=
-
-#-----------------------------------------------------------------------
-# Load data
-# data(soyaBeans, package = "cmpreg")
-soyBeans <- read.table("./data/soyaBeans.txt",
-                        header = TRUE, sep = "\t")
-soyBeans$umid <- as.factor(soyBeans$umid)
-soyBeans <- soyBeans[-74, ] # Incorrect observation
-soyBeans <- transform(soyBeans, K = K / 100)
-
-#-----------------------------------------------------------------------
-# Fit models
-
-# Predictor
-form2 <-  ngra ~ bloc + umid * K + I(K^2)
-
-m2PO <- glm(form2, data = soyBeans, family = poisson)
-time21 <- system.time(
-    m2C1 <- fitcm(form2, data = soyBeans, model = "CP", sumto = 700)
-)
-time22 <- system.time(
-    m2C2 <- fitcm(form2, data = soyBeans, model = "CP2", sumto = 700)
-)
-m2QP <- glm(form2, data = soyBeans, family = quasipoisson)
-
-models.ngra <- list(m2PO, m2C1, m2C2, m2QP)
-names(models.ngra) <- mnames
-
-co2 <- coef(m2C2)
-
-# Numbers of calls to loglik and numerical gradient
-c21 <- models.ngra$C1@details$counts
-c22 <- models.ngra$C2@details$counts
-
-# # Profile extra parameter
-# profs.ngra <- lapply(list(c(m2C1, "phi"), c(m2C2, "phi2")),
-#                      function(x) myprofile(x[[1]], x[[2]]))
-# profs.ngra <- do.call("rbind", profs.ngra)
-
-# LRT between Poisson and COM-Poisson (test: phi == 0)
-lrt.ngra <- getAnova(m2PO, m2C2)
-
 @
@@
 Solid lines are the smoothing curves on the left and the least of squares curve on the right.">>=
-
-#-----------------------------------------------------------------------
-# Exploratory analysis
-
-# Scatter plot
-xy1 <- xyplot(ngra ~ K | umid,
-              data = soyBeans,
-              xlab = "Potassium fertilization level",
-              ylab = "Number of grains per pot",
-              type = c("p", "g", "smooth"),
-              as.table =  TRUE,
-              layout = c(2, 2),
-              strip = strip.custom(
-                  strip.names = TRUE, var.name = "moisture",
-                  factor.levels = paste0(levels(soyBeans$umid), "%")))
-
-# Sample variance vs sample mean (evidence in favor of the
-# overdispersion).
-mv <- soyBeans %>%
-    group_by(K, umid) %>%
-    summarise(mu = mean(ngra), va = var(ngra))
-xlim <- ylim <- extendrange(c(mv$mu, mv$va), f = 0.05)
-
-xy2 <- xyplot(va ~ mu,
-              data = mv,
-              type = c("p", "r", "g"),
-              xlim = xlim,
-              ylim = ylim,
-              xlab = expression("Sample"~"mean"~(bar(y))),
-              ylab = expression("Sample"~"variance"~(s^2)),
-              panel = function(...) {
-                  panel.xyplot(...)
-                  panel.abline(a = 0, b = 1, lty = 2)
-              })
-
-print(xy1, split = c(1, 1, 2, 1), more = TRUE)
-print(xy2, split = c(2, 1, 2, 1), more = FALSE)
-
 @
@@
 <<results-soya, results="asis">>=
-
-#-----------------------------------------------------------------------
-# Goodness of fit measures and estimate parameters
-
-# GoF measures
-measures.ngra <- lapply(models.ngra, function(x)
-    c("LogLik" = logLik(x), "AIC" = AIC(x), "BIC" = BIC(x)))
-
-# Get the estimates
-co2 <- coef(m2C2)
-est <- lapply(models.ngra, FUN = function(x) getCoefs(x))
-est.ngra <- do.call(cbind, est)
-
-# Organize in table
-pnames <- c("\\phi\\,,\\,\\sigma", "\\beta_0",
-            paste0("\\gamma_{", 1:4, "}"),
-            paste0("\\tau_{", 1:2, "}"),
-            "\\beta_1", "\\beta_2",
-            paste0("\\beta_{3", 1:2, "}"))
-
-rownames(est.ngra) <- paste0("$", pnames, "$")
-meds <- apply(do.call(cbind, measures.ngra), 1, function(x) {
-    x <- formatC(x, getOption("digits"), format = "f")
-    x <- gsub("NA", "-", x)
-    paste(paste0("\\multicolumn{2}{c}{$", x, "$}"),
-          collapse = " & ")
-})
-text_gof <- paste(paste(names(meds), "&", meds),
-                  collapse = "\\\\\n ")
-
-append_gof <- list(
-    pos = list(nrow(est.ngra)),
-    command = paste("\\specialrule{0.01em}{0.3em}{0.3em} \n",
-                    text_gof, "\\\\\n",
-                    "\\bottomrule"))
-print.xtable(xtable(est.ngra, digits = 4,
-                    label = "tab:coef-soy"),
-             hline.after = 0,
-             only.contents = TRUE,
-             add.to.row = append_gof)
-
 @
@@
 <<corr-soy, results="asis">>=
-
-##----------------------------------------------------------------------
-## Correlation between estimates
-corr.ngra <- do.call("rbind",
-                     lapply(models.ngra[c("C1", "C2")],
-                            function(x) cov2cor(vcov(x))[1, -1]))
-
-## Organize on table
-rownames(corr.ngra) <- paste0("COM-Poisson", c("", "$_\\mu$"))
-colnames(corr.ngra) <- gsub("(beta|tau|gamma)", "hat{\\\\\\1}",
-                            pnames[-1])
-corrngra_aux <- divide_matrix(corr.ncap, divide = 7)
-
-# caption <- paste("Empirical correlations between $\\hat{\\phi}$ and",
-#                  "$\\hat{\\bm{\\beta}}$ for the two parametrizations",
-#                  "of COM-Poisson model fit to overdispersed data.")
-# print(xtable(corr.ngra,
-#              align = c("lccccccccccc"),
-#              caption = caption,
-#              digits = 3,
-#              label = "tab:corr-soy"),
-#       size = "small",
-#       sanitize.rownames.function = identity,
-#       sanitize.colnames.function = function(x) sprintf("$%s$", x))
-
 @
@@
   \toprule
-<<results="asis">>=
-
-print(xtable(corrngra_aux[[1]], digits = 4),
-      only.contents = TRUE,
-      hline.after = FALSE,
-      sanitize.colnames.function = function(x) sprintf("$%s$", x))
-cat("\\midrule", sep = "\n")
-
-print(xtable(corrngra_aux[[2]], digits = 4),
-      only.contents = TRUE,
-      sanitize.colnames.function = function(x) sprintf("$%s$", x))
-
+<<cres-soy, results="asis">>=
 @
@@
 <<pred-soy, fig.height=3.8, fig.width=8, fig.cap="Dispersion diagrams of been seeds counts as function of potassium doses and humidity levels with fitted curves and confidence intervals (95\\%).">>=
-
-#-----------------------------------------------------------------------
-# Prediction
-
-# Data for prediction
-pred <- with(soyBeans,
-             expand.grid(
-                 bloc = factor(levels(bloc)[1], levels = levels(bloc)),
-                 umid = levels(umid),
-                 K = seq(min(K), max(K), l = 20)
-             ))
-qn <- qnorm(0.975) * c(fit = 0, lwr = -1, upr = 1)
-
-# Design matrix for prediction
-X <- model.matrix(update(form2, NULL ~ .), pred)
-bl <- attr(X, "assign") == 1
-X[, bl] <- X[, bl] + 1/(sum(bl) + 1)
-
-# Considering Poisson
-aux <- exp(confint(
-    glht(m2PO, linfct = X), calpha = univariate_calpha())$confint)
-colnames(aux) <- c("fit", "lwr", "upr")
-aux <- data.frame(modelo = "Poisson", aux)
-predPO.ngra <- cbind(pred, aux)
-
-# Considering COM-Poisson
-aux <- predictcm(m2C1, newdata = X)
-aux <- data.frame(modelo = "COM-Poisson", aux)
-predC1.ngra <- cbind(pred, aux)
-
-# Considering COM-Poisson (mean parametrization)
-aux <- predictcm(m2C2, newdata = X)
-aux <- data.frame(modelo = "COM-Poisson2", aux)
-predC2.ngra <- cbind(pred, aux)
-
-# Considering Quasi-Poisson
-aux <- exp(confint(
-    glht(m2QP, linfct = X), calpha = univariate_calpha())$confint)
-colnames(aux) <- c("fit", "lwr", "upr")
-aux <- data.frame(modelo = "Quasi-Poisson", aux)
-predQP.ngra <- cbind(pred, aux)
-
-# Representing the confidence intervals
-pred.ngra <- rbind(predPO.ngra, predC1.ngra, predC2.ngra, predQP.ngra)
-
-# Legend
-key <- list(columns = 4,
-            cex = 0.9,
-            lines = list(col = 1, lty = rev(1:4)),
-            text = list(parse(
-                text = c("'Poisson'", "'COM-Poisson'",
-                         "'COM-Poisson'[mu]", "'Quasi-Poisson'"))
-                )
-            )
-
-# Graph
-update(xy1, layout = c(NA, 1), type = c("p", "g"),
-       alpha = 0.6, key = key) +
-    as.layer(
-        xyplot(fit ~ K | umid,
-               data = pred.ngra,
-               groups = modelo,
-               type = "l",
-               col = 1,
-               ly = pred.ngra$lwr,
-               uy = pred.ngra$upr,
-               cty = "bands",
-               fill = "gray80",
-               alpha = 0.1,
-               panel = panel.superpose,
-               panel.groups = panel.cbH,
-               prepanel = cmpreg::prepanel.cbH,
-               lty = rev(1:4))
-    )
-
 @
@@
 <<fit-ovos, cahe=TRUE, include=FALSE>>=
-
-#-----------------------------------------------------------------------
-# Load data
-# data(nitrofen, package = "boot")
-# data(Paula, package = "labestData")
-# nitrofen <- PaulaEx4.6.20
-nitrofen <- read.table("./data/nitrofen.txt",
-                       header = TRUE, sep = "\t")
-nitrofen <- transform(nitrofen, dose = dose / 100)
-
-#-----------------------------------------------------------------------
-# Fit models
-mnames <- c("PO", "C1", "C2", "QP")
-
-# Predictors
-form31 <-  novos ~ dose
-form32 <-  novos ~ dose + I(dose^2)
-form33 <-  novos ~ dose + I(dose^2) + I(dose^3)
-
-predictors <- list("pred1" = form31, "pred2" = form32, "pred3" = form33)
-fmodels.ovos <- lapply(predictors, function(form) {
-    PO <- glm(form, data = nitrofen, family = poisson)
-    C1 <- fitcm(form, data = nitrofen, model = "CP", sumto = 100)
-    C2 <- fitcm(form, data = nitrofen, model = "CP2", sumto = 100)
-    QP <- glm(form, data = nitrofen, family = quasipoisson)
-    list("PO" = PO, "C1" = C1, "C2" = C2, "QP" = QP)
-})
-
-#-----------------------------------------------------------------------
-# LRT for nested models
-
-# Poisson
-auxPO <- lapply(fmodels.ovos, function(x) x$PO)
-do.call("getAnova", auxPO)
-
-# COM-Poisson standard
-auxC1 <- lapply(fmodels.ovos, function(x) x$C1)
-do.call("getAnova", auxC1)
-
-# COM-Poisson mean-parameterized
-auxC2 <- lapply(fmodels.ovos, function(x) x$C2)
-do.call("getAnova", auxC2)
-
-# Quasi-Poisson
-auxQP <- lapply(fmodels.ovos, function(x) x$QP)
-do.call("getAnova", auxQP)
-
-#--------------------------------------------
-# Separe the choose models
-form3 <- form33
-m3PO <- fmodels.ovos$pred3$PO
-m3C1 <- fmodels.ovos$pred3$C1
-m3C2 <- fmodels.ovos$pred3$C2
-m3QP <- fmodels.ovos$pred3$QP
-
-models.ovos <- list(m3PO, m3C1, m3C2, m3QP)
-names(models.ovos) <- mnames
-
-# Numbers of calls to loglik and numerical gradient
-models.ovos$C1@details$counts
-models.ovos$C2@details$counts
-
 @
@@
 <<anova-ovos1, results="asis">>=
-
-auxPO <- lapply(fmodels.ovos, function(x) x$PO)
-tab <- do.call("getAnova", c(print = FALSE, auxPO))
-tab <- cbind(tab, NA)
-rownames(tab) <- paste("Preditor", rownames(tab))
-digits <- c(1, 0, 3, 3, 3, 0, -2, 3)
-print(xtable(tab, digits = digits),
-      include.colnames = FALSE,
-      hline.after = NULL,
-      only.contents = TRUE)
-
 @
@@
 <<anova-ovos2, results="asis">>=
-
-auxC1 <- lapply(fmodels.ovos, function(x) x$C1)
-tab <- do.call("getAnova", c(print = FALSE, auxC1))
-tab <- cbind(tab, sapply(auxC1, function(x) coef(x)[1]))
-rownames(tab) <- paste("Preditor", rownames(tab))
-print(xtable(tab, digits = digits),
-      include.colnames = FALSE,
-      hline.after = NULL,
-      only.contents = TRUE)
-
 @
@@
 <<anova-ovos3, results="asis">>=
-
-auxC2 <- lapply(fmodels.ovos, function(x) x$C2)
-tab <- do.call("getAnova", c(print = FALSE, auxC2))
-tab <- cbind(tab, sapply(auxC2, function(x) coef(x)[1]))
-rownames(tab) <- paste("Preditor", rownames(tab))
-print(xtable(tab, digits = digits),
-      include.colnames = FALSE,
-      hline.after = NULL,
-      only.contents = TRUE)
-
 @
@@
 <<anova-ovos4, results="asis">>=
-
-auxQP <- lapply(fmodels.ovos, function(x) x$QP)
-tab <- do.call("getAnova", c(print = FALSE, auxQP))
-tab <- cbind(tab, sapply(auxQP, function(x) summary(x)$dispersion))
-rownames(tab) <- paste("Preditor", rownames(tab))
-print(xtable(tab, digits = digits),
-      include.colnames = FALSE,
-      hline.after = NULL,
-      only.contents = TRUE)
-
 @
@@
 <<coef-ovos, results="asis">>=
-
-#-----------------------------------------------------------------------
-# Goodness of fit measures and estimate parameters
-
-# GoF measures
-measures.ovos <- lapply(models.ovos, function(x)
-    c("LogLik" = logLik(x), "AIC" = AIC(x), "BIC" = BIC(x)))
-
-# Get the estimates
-co2 <- coef(m2C2)
-est <- lapply(models.ovos, FUN = function(x) getCoefs(x))
-est.ovos <- do.call(cbind, est)[-1, ]
-
-# Organize in table
-pnames <- paste0("\\beta_{", 0:3, "}")
-rownames(est.ovos) <- paste0("$", pnames, "$")
-print.xtable(xtable(est.ovos, digits = 4),
-             hline.after = c(0, nrow(est.ovos)),
-             only.contents = TRUE)
-
 @
@@
 <<pred-ovos, fig.height=3.5, fig.width=5.5, out.width="0.7\\textwidth", fig.cap="Number of live offsprings observed for each nitrofen concentration level with fitted curves and 95\\% confidence intervals.">>=
-
-#-----------------------------------------------------------------------
-# Prediction
-
-# Data for prediction
-pred <- with(nitrofen,
-             data.frame("dose" = seq(min(dose), max(dose),
-                                     length.out = 100)))
-qn <- qnorm(0.975) * c(fit = 0, lwr = -1, upr = 1)
-
-# Design matrix for prediction
-X <- model.matrix(update(form3, NULL ~ .), pred)
-
-# Considering Poisson
-aux <- exp(confint(
-    glht(m3PO, linfct = X), calpha = univariate_calpha())$confint)
-colnames(aux) <- c("fit", "lwr", "upr")
-aux <- data.frame(modelo = "Poisson", aux)
-predPO.novos <- cbind(pred, aux)
-
-# Considering COM-Poisson
-aux <- predictcm(m3C1, newdata = X)
-aux <- data.frame(modelo = "COM-Poisson", aux)
-predC1.novos <- cbind(pred, aux)
-
-# Considering COM-Poisson (mean parametrization)
-aux <- predictcm(m3C2, newdata = X)
-aux <- data.frame(modelo = "COM-Poisson2", aux)
-predC2.novos <- cbind(pred, aux)
-
-# Considering Quasi-Poisson
-aux <- exp(confint(
-    glht(m3QP, linfct = X), calpha = univariate_calpha())$confint)
-colnames(aux) <- c("fit", "lwr", "upr")
-aux <- data.frame(modelo = "Quasi-Poisson", aux)
-predQP.novos <- cbind(pred, aux)
-
-# Representing the confidence intervals
-pred.novos <- rbind(predPO.novos, predC1.novos, predC2.novos, predQP.novos)
-ord <- order(pred.novos$dose, pred.novos$modelo)
-pred.novos <- pred.novos[ord, ]
-
-# Legend
-key <- list(columns = 2,
-            lines = list(col = 1, lty = rev(1:4)),
-            text = list(parse(
-                text = c("'Poisson'", "'COM-Poisson'",
-                         "'COM-Poisson'[mu]", "'Quasi-Poisson'"))
-                )
-            )
-
-# Graph
-xyplot(novos ~ dose,
-       data = nitrofen,
-       xlab = "Nitrofen concentration level",
-       ylab = "Number of live offspring",
-       grid = TRUE,
-       alpha = 0.6,
-       key = key,
-       spread = 0.05,
-       panel = panel.beeswarm) +
-    as.layer(
-        xyplot(fit ~ dose,
-               auto.key = TRUE,
-               data = pred.novos,
-               groups = modelo,
-               type = "l",
-               col = 1,
-               ly = pred.novos$lwr,
-               uy = pred.novos$upr,
-               cty = "bands",
-               fill = "gray80",
-               alpha = 0.1,
-               panel = panel.superpose,
-               panel.groups = panel.cbH,
-               prepanel = prepanel.cbH,
-               lty = rev(1:4))
-    )
-
 @
@@
 <<corr-ovos, results="asis">>=
-
-#-----------------------------------------------------------------------
-# Correlation between estimates
-corr.ovos <- do.call("rbind",
-                     lapply(models.ovos[c("C1", "C2")],
-                            function(x) cov2cor(vcov(x))[1, -1]))
-
-# Organize on table
-rownames(corr.ovos) <- paste0("COM-Poisson", c("", "$_\\mu$"))
-colnames(corr.ovos) <- gsub("(beta)", "hat{\\\\\\1}", pnames)
-
-caption <- paste("Empirical correlations between $\\hat{\\phi}$ and",
-                 "$\\hat{\\bm{\\beta}}$ for the two parametrizations",
-                 "of COM-Poisson model fit to equidispersed data.")
-print(xtable(corr.ovos,
-             align = c("lrrrr"),
-             caption = caption,
-             digits = 4,
-             label = "tab:corr-ovos"),
-      # size = "small",
-      sanitize.rownames.function = identity,
-      sanitize.colnames.function = function(x) sprintf("$%s$", x),
-      # width = "\\textwidth,"
-      tabular.environment = "tabular")
-
 @

### 2018-08-20 article.R

@@ -1,0 +1,1381 @@
+
+#=======================================================================
+# General settings
+#=======================================================================
+
+## ----setup
+library(knitr)
+library(xtable)
+options(digits = 3, OutDec = ".",
+        xtable.caption.placement = "top",
+        xtable.booktabs = TRUE,
+        xtable.sanitize.text.function = identity,
+        xtable.size = "small",
+        xtable.math.style.negative = TRUE)
+opts_chunk$set(
+    warning = FALSE,
+    message = FALSE,
+    echo = FALSE,
+    results = "hide",
+    fig.width = 7,
+    fig.height = 5,
+    out.width = "1\\textwidth",
+    fig.align = "center",
+    fig.pos = "!htp",
+    # dev = "tikz"
+    dev.args = list(family = "Palatino")
+    )
+
+divide_matrix <- function(mat, divide) {
+    mat1 <- mat[, 1:divide]
+    mat2 <- mat[, (divide + 1):ncol(mat)]
+    nc1 <- ncol(mat1)
+    nc2 <- ncol(mat2)
+    cdiff <- abs(ncol(mat1) - ncol(mat2))
+    if (cdiff) {
+        complete <- matrix(NA, nrow = nrow(mat), ncol = cdiff)
+        colnames(complete) <- rep(" ", cdiff)
+        if (nc1 < nc2) {
+            mat1 <- cbind(mat1, complete)
+        } else {
+            mat2 <- cbind(mat2, complete)
+        }
+    }
+    return(list(mat1, mat2))
+}
+
+source("./codes/lattice-panels.R")
+source("./codes/functions.R")
+
+# Colors for legends
+cols <- trellis.par.get("superpose.line")$col
+
+# Useful packages
+library(bbmle)
+library(multcomp)
+library(plyr)
+library(tidyr)
+library(dplyr)
+
+#=======================================================================
+# Convergence of Z
+#=======================================================================
+
+## ----convergenceZ
+# Convergence of Z(lambda, nu) constant
+computeZ <- function(lambda, nu, maxit = 1e4, tol = 1e-5) {
+    z <- vector("numeric", maxit)
+    j = 1
+    z[j] <- exp(j * log(lambda) - nu * lfactorial(j))
+    while (abs(z[j] - 0) > tol && j <= maxit) {
+        j = j + 1
[…]

### 2018-08-27 codes/01-simulation.R
Compute times to fit (in the simulation and case studies)

@@ -81,34 +81,38 @@ results <- mclapply(sizes, function(n) {
     lapply(phis, function(phi) {
         lapply(1:B, function(i) {
             y <- rcmp(1, mu, phi)
-            fit <- try(fitcm(y ~ x1 + x2, start = c(phi, beta),
-                             model = "CP2", sumto = 200))
-                if (! class(fit) == "try-error") {
-                    ll <- -fit@min
-                    vc <- fit@vcov
-                    co <- fit@coef
-                    ci <- confint(fit, method = "quad")
-                } else {
-                    ll <- NA
-                    vc <- matrix(NA, ncol = length(beta) + 1,
-                                 nrow = length(beta) + 1)
-                    co <- rep(NA, length(beta) + 1)
-                    ci <- matrix(NA, nrow = length(beta) + 1, ncol = 2)
-                }
+            ti <- system.time(
+                fit <- try(fitcm(y ~ x1 + x2, start = c(phi, beta),
+                                 model = "CP2", sumto = 200))
+            )
+            if (! class(fit) == "try-error") {
+                ll <- -fit@min
+                vc <- fit@vcov
+                co <- fit@coef
+                ci <- confint(fit, method = "quad")
+            } else {
+                ll <- NA
+                vc <- matrix(NA, ncol = length(beta) + 1,
+                             nrow = length(beta) + 1)
+                co <- rep(NA, length(beta) + 1)
+                ci <- matrix(NA, nrow = length(beta) + 1, ncol = 2)
+            }
             colnames(ci) <- c("lwr", "upr")
-            out <- list("coef" = co, "cint" = ci,
-                        "loglik" = ll, "vcov" = vc)
+            out <- list("coef" = co,
+                        "cint" = ci,
+                        "loglik" = ll,
+                        "vcov" = vc,
+                        "ctime" = ti)
             return(out)
         }) # Close replicate
     }) # Close phi
-}, mc.cores = 3) # Close sizes
-str(results)
-
+}, mc.cores = 4) # Close sizes
 saveRDS(results, "simulation.rds", compress = FALSE)
 
 #-------------------------------------------
 # Load the results
 results <- readRDS("simulation.rds")
+results_time <- readRDS("simulation2.rds")
 
 #-----------------------------------------------------------------------
 # Compute bias
[…]

### 2018-08-27 codes/24-timestofit.R

@@ -81,3 +81,22 @@ bench3 <- microbenchmark(
 #  CMP   1.125297 1.143321 1.189389 1.160909 1.237385 1.357786    50   b
 #  CMPmu 1.034631 1.041504 1.091136 1.071659 1.128976 1.258067    50  a
 #-------------------------------------------
+
+#-----------------------------------------------------------------------
+# Organize results
+benchs <- list("Cotton (under)" = bench1,
+               "Soybean (over)" = bench2,
+               "Nitrofen (equi)" = bench3)
+benchs <- purrr::map_dfr(benchs, identity, .id = "case")
+
+class(benchs) <- "data.frame"
+saveRDS(benchs, "comparetimes.rds")
+
+xlabs <- expression("COM-Poisson", "COM-Poisson"[mu])
+bwplot(time ~ expr | case,
+       ylab = "Time (seconds)",
+       scales = list(
+           y = list(relation = "free"),
+           x = list(at = 1:2, labels = xlabs)
+       ),
+       data = benchs)

### 2018-08-27 article.Rnw
Make changes in the paper (improves and corrects the text)

@@
 % Abstract
-\Abstract{ In the analysis of count data often the equidispersion
-  assumption is not suitable, hence the Poisson regression model is
-  inappropriate.  As a generalization of the Poisson distribution the
-  COM-Poisson distribution can deal with under-, equi- and overdispersed
-  count data.  It is a member of the exponential family of distributions
-  and has the Poisson and geometric distributions as special cases, as
-  well as the Bernoulli distribution as a limiting case.  In spite of
-  the nice properties of the COM-Poisson distribution, its location
-  parameter does not correspond to the expectation, which complicates
-  the interpretation of regression models specified using this
-  distribution. In this paper, we propose a straightforward
-  reparametrization of the COM-Poisson distribution based on an
-  approximation to the expectation of this distribution. The main
-  advantage of our new parametrization is the straightforward
-  interpretation of the regression coefficients in terms of the
-  expectation of the count response variable, as usual in the context of
-  generalized linear models. Furthermore, the estimation and inference
-  for the new COM-Poisson regression model can be done based on the
+\Abstract{The COM-Poisson distribution is a two-parameter generalization
+  of Poisson distribution can deal with under-, equi- and overdispersed
+  count data. Its location parameter does not correspond to the
+  expectation, which complicates the interpretation of regression models
+  specified using this distribution. In this paper, we propose a
+  straightforward reparametrization of the COM-Poisson distribution
+  based on an approximation to the expectation of this distribution. The
+  main advantage of our new parametrization is the interpretation of the
+  regression coefficients in terms of the expectation of the count
+  response variable, as usual in the context of generalized linear
+  models. Furthermore, the estimation and inference for the
+  reparametrized COM-Poisson regression models can be done based on the
   likelihood paradigm. We carried out simulation studies to verify the
   finite sample properties of the maximum likelihood estimators. The
-  results from our simulation study show that the maximum likeli-hood
+  results from our simulation study show that the maximum likelihood
   estimators are unbiased and consistent for both regression and
-  dispersion parameters. We observed that the empirical correlation
-  between the regression and dispersion parameter estimators is close to
-  zero, which suggests that these parameters are orthogonal. We
-  illustrate the application of the proposed model through the analysis
-  of three data sets with over-, under- and equidispersed count data.
-  The study of distribution properties through a consideration of
-  dispersion, zero-inflated and heavy tail indexes, together with the
-  results of data analysis show the flexibility over standard
-  approaches. Therefore, we encourage the application of the new
-  parametrization for the analysis of count data in the context of
-  COM-Poisson regression models.  The computational routines for fitting
-  the original and new version of the COM-Poisson regression model and
-  the analyzed data sets are available in the supplementary material.
-}
+  dispersion parameters. The empirical correlation between the
+  regression and dispersion parameter estimators and deviance surfaces
+  suggests that these parameters are orthogonal. We illustrate the
+  application of the proposed model through the analysis of three data
+  sets with over-, under- and equidispersed count data. The study of
+  distribution properties through a consideration of dispersion,
+  zero-inflated and heavy tail indexes, together with the results of
+  data analysis show the flexibility over standard approaches.  The
+  computational routines for fitting the original and new version of the
+  COM-Poisson regression model and the analyzed data sets are available
+  in the supplementary material.  }
 
@@
 Poisson distribution is referred as equidispersion.  However, in
-practical data analysis such a restriction can be unsuitable, since the
+practical data analysis such restriction can be unsuitable, since the
 observed data can present  variance both smaller or larger than the
-expectation, leading to the cases of under and overdispersion,
+expectation, leading to the cases of under- and overdispersion,
 respectively.  The main problem of the application of the Poisson
@@
 The strategies for constructing alternative count distributions are
-related with the causes of the non-equidispersion. Specfically for the
-overdispersion case Poisson mixture models are widely applied.  One
+related with the causes of the non-equidispersion. Specifically for the
+overdispersion, Poisson mixture models are widely applied.  One
 popular example of this approach is the negative-binomial model, where
@@
 
-The COM-Poisson distribution is a member of the exponential family and
-it has the Poisson and geometric distributions as special cases, as well
-as the Bernoulli distribution as a limiting case. It can deal with both
-under and overdispersed count data.  Some recently applications of the
-COM-Poisson distribution include \citet{Lord2010} for the analysis of
-traffic crash data, \citet{Sellers2010} for the modelling of airfreight
-breakage and book purchases, and \citet{Huang2017} to the analysis of
-attendance data, takeover birds and cotton boll counts.  The main
-disadvantage of the COM-Poisson regression model as presented
-in~\citet{Sellers2010} is that its location parameter does not
-correspond to the expectation of the distribution, which complicates the
-interpretation of regression models and means that they are not
-comparable with standard approaches such as the Poisson and negative
-binomial regression models. \citet{Huang2017} proposed a
-mean-parametrization of the COM-Poisson distribution in order to avoid
-such an issue.  In this approach the mean parameter is obtained by
-solving an non-linear equation defined as an infinite sum.
+The COM-Poisson distribution has the Poisson and geometric distributions
+as special cases, as well as the Bernoulli distribution as a limiting
+case. It can deal with both under- and overdispersed count data.  Some
+recently applications of the COM-Poisson distribution include
+\citet{Lord2010} for the analysis of traffic crash data,
+\citet{Sellers2010} for the modelling of airfreight breakage and book
+purchases, \citet{Huang2017} to the analysis of attendance data,
+takeover birds and cotton boll counts, and \citet{Chatla2018} to model
+counts from bike sharing systems. Theoretical results and approximations
+derived from this distribution is discussed by \citet{Shmueli2005},
+\citet{Daly2016} and \citet{Gaunt2017}. The main disadvantage of the
+COM-Poisson regression model as presented in~\citet{Sellers2010} is that
+its location parameter does not correspond to the expectation of the
+distribution, which complicates the interpretation of regression models
+and means that they are not comparable with standard approaches such as
+the Poisson and negative binomial regression models. \citet{Huang2017}
+proposed a mean-parametrization of the COM-Poisson distribution in order
+to avoid such an issue.  In this approach the mean parameter is obtained
+by solving an non-linear equation defined as an infinite sum.
 Consequently, it is computationally demanding and liable to numerical
@@
 function is written in terms of $\mu$ and $\phi$, where $\mu$ is the
-expectation and $\phi$ is a dispersion parameter. In contrast to the
-original parametrization, the proposed parametrization leads to
-regression coefficients directly associated with the expectation of the
-response variable, as usual in the context of generalized linear models.
-Consequently, the obtained regression coefficients are comparable with
-the ones obtained by standard approaches, such as the Poisson and
-negative binomial regression models.  Furthermore, our novel COM-Poisson
-parametrization is simpler than the strategy proposed
-by~\cite{Huang2017}, since it does not require any numerical method for
-solving non-linear equations, and we show the attractive properties like
-the orthogonality between dispersion and regression parameters and
-consistency and asymptotic normality of the maximum likelihood
-estimators are retained.
+approximated expectation and $\phi$ is a dispersion parameter. In
+contrast to the original parametrization, the proposed parametrization
+leads to regression coefficients directly associated with the
+expectation of the response variable, as usual in the context of
+generalized linear models.  Consequently, the obtained regression
+coefficients are comparable with the ones obtained by standard
+approaches, such as the Poisson and negative binomial regression models.
+Furthermore, our novel COM-Poisson parametrization is simpler than the
+strategy proposed by~\cite{Huang2017}, since it does not require any
+numerical method for solving non-linear equations, and we show the
+attractive properties like the orthogonality between dispersion and
+regression parameters and consistency and asymptotic normality of the
+maximum likelihood estimators are retained.
 
@@
 \begin{equation*}
-  \sum_{j=0}^{\infty} (j - \mu) \frac{\lambda^j}{(y!)^\nu} = 0\,.
+  \sum_{j=0}^{\infty} (j - \mu) \frac{\lambda^j}{(j!)^\nu} = 0\,.
 \end{equation*}
@@
   \label{eqn:repar-cmp}
-  \mu = h(\lambda, \nu) = \lambda^{1/\nu} - \frac{\nu - 1}{2\nu}
+  \mu = h_\nu(\lambda) = \lambda^{1/\nu} - \frac{\nu - 1}{2\nu}
   \quad \Rightarrow \quad
-  \lambda = h^{-1}(\mu, \nu) = \left (\mu +
+  \lambda = h_\nu^{-1}(\mu) = \left (\mu +
     \frac{(\nu - 1)}{2\nu} \right )^\nu.
@@
 (\autoref{eqn:mean-approx}), \autoref{fig:approx-plot} presents the
-quadratic errors for (a) expectation  and (b) variance.  The
-quadratic errors were obtained by $[\mu - \text{E}(Y)]^2$ for the
-expectation and by $[ \mu \exp(-\phi) - \text{Var}(Y)]^2$ for the
-variance. In both cases $\text{E}(Y)$ and Var$(Y)$ were computed
-numerically. The dotted lines represent the border between the regions
-$\nu \leq 1$ and $\lambda > 10^\nu$, in the $\mu$ and $\phi$ scale.
+quadratic errors for (a) expectation and (b) variance.  The quadratic
+errors were obtained by $[\mu - \text{E}(Y)]^2$ for the expectation and
+by $[ \mu \exp(-\phi) - \text{Var}(Y)]^2$ for the variance. In both
+cases $\text{E}(Y)$ and Var$(Y)$ were computed numerically\footnote{We
+  compute E$(Y) = \sum_{y=0}^{500}yp(y)$ and
+  E$(Y^2) = \sum_{y=0}^{500}y^2p(y)$.}. The dotted lines represent the
+border between the regions $\nu \leq 1$ and $\lambda > 10^\nu$, in the
+$\mu$ and $\phi$ scale.
 
@@
   \qquad y = 0, 1, 2, \ldots\,,
-\end{equation} where $\mu > 0$. We denote this distribution as
+\end{equation}
+where $\mu > 0$. We denote this distribution as
 COM-Poisson$_\mu$. In \autoref{fig:pmf-cmp}, we show the shapes of
@@
 @
+
+The constraint $\mu>0$ imposes an undesirable restriction in the
+original parameter space, $\lambda^\nu>(\nu-1)/(2\nu)$. However, this
+restricted region is related to small underdispersed counts (averages
+smaller than 0.1 and $\nu>1$) a parameter region unexplored in practice.
 
@@
 \autoref{fig:indexes-plot}(d) indicate the distribution is in general a
-light-tailed distribution, i.e. $HT \to 0$ for $y \to \infty$.
+light-tailed distribution, i.e. $\text{HT} \to 0$ for $y \to \infty$.
 
@@
 \autoref{eqn:ll-rcmp}. Since the derivatives of $\ell$ cannot be
-obtained in closed forms, we adopted the \texttt{BFGS}
-algorithm~\citep{Nocedal1995} as implemented in the function
-\texttt{optim()} for the statistical software \texttt{R}
-\citep{Rcore2017}. Standard errors for the regression coefficients are
-obtained based on the observed information matrix
-$\mathcal{I}(\bm{\theta})$, where
-$\mathcal{I}(\bm{\theta}) = -\mathcal{H}(\bm{\theta})$ (hessian matrix)
+obtained in closed forms, we obtain the Hessian matrix by central finite
+differences. We adopted the \texttt{BFGS} algorithm~\citep{Nocedal1995}
+as implemented in the function \texttt{optim()} for the statistical
+software \texttt{R} \citep{Rcore2017}. Standard errors for the
+regression coefficients are obtained based on the observed information
+matrix $\mathcal{I}(\bm{\theta})$, where
+$\mathcal{I}(\bm{\theta}) = -\mathcal{H}(\bm{\theta})$ (Hessian matrix)
 is computed numerically. Confidence intervals for $\hat{\mu}_i$ are
@@
 continuous covariate~($\bm{x}_1$) was generated as a sequence from $0$
-to $1$ and of length equal to the sample size.  Similarly, the categorical
-covariate~($\bm{x}_2$) was generated as a sequence of three values each
-one repeated $n/3$ times (rounding up when required), where $n$ denotes
-the sample size. Thus, the expectation of the COM-Poisson random
-variable is given by
+to $1$ and of length equal to the sample size.  Similarly, the
+categorical covariate~($\bm{x}_2$) was generated as a sequence of three
+values each one repeated $n/3$ times (rounding up when required), where
+$n$ denotes the sample size. The parameter $\mu$ of the reparametrized
+COM-Poisson random variable is given by
 $\bm{\mu} = \exp(\beta_0 + \beta_1 \bm{x}_1 + \beta_{21} \bm{x}_{21} +
 \beta_{22} \bm{x}_{22})$,
-where $\bm{x}_{21}$ and $\bm{x}_{22}$ are dummy representing the levels
-of $\bm{x}_2$.  The regression coefficients were fixed at the values,
+where $\bm{x}_{21}$ and $\bm{x}_{22}$ are dummy representing the levels of
+$\bm{x}_2$.  The regression coefficients were fixed at the values,
 $\beta_0 = 2$, $\beta_1 = 0.5$, $\beta_{21} = 0.8$ and
@@
 
-%-----------------------------------------------------------------------
-% Explain the simulation
-% \begin{algorithm}
-% \caption{Steps in simulation study.}
-% \label{alg:simulation}
-%   \Begin{
-%   $\bm{\beta} = \begin{bmatrix} \beta_0 & \beta_1 & \beta_{21} &
-%     \beta_{22} \end{bmatrix}^\top = \begin{bmatrix} 2.0 &
-%     0.5 & 0.8 & -0.8 \end{bmatrix}^\top$\;
-%   \For{$n \in \{50, 100, 300, 1000\}$}{
-%     set $\bm{x}_1$ as a sequence, with $n$ elements, between $0$ and
-%     $1$\;
-%     set $\bm{x}_2$ as a repetition, with $n$ elements, of three
-%     categories\;
-%     compute $\bm{\mu}$ using $\bm{\mu} = \exp(\beta_0 + \beta_1 \bm{x}_1
-%     +  \beta_{21} \bm{x}_{21} + \beta_{22} \bm{x}_{22})$, where
-%     $\bm{x}_{21}$ and $\bm{x}_{22}$ are \textit{dummy} variable for
-%     $\bm{x}_2$\;
-%     \For{$\phi \in \{-1.6, -1.0, 0.0, 1.8\}$}{
-%       \Repeat{$1000$ times}{
-%         simulate $\bm{y}$ from COM-Poisson distribution with $\bm{\mu}$
-%         and $\phi$ parameters\;
-%         fit COM-Poisson$_\mu$ regression model to $\bm{y}$ data with
-%         $\bm{X} = \begin{bmatrix} \bm{1} & \bm{x}_1 & \bm{x}_{21} &
-%           \bm{x}_{22} \end{bmatrix}$ design matrix\;
-%         get $\hat{\bm{\theta}} = \begin{bmatrix} \hat{\phi} & \hat{\beta}_0 &
-%           \hat{\beta}_1 & \hat{\beta}_{21} & \hat{\beta}_{22}
-%         \end{bmatrix}^\top$\;
-%         get confidence intervals for $\hat{\bm{\theta}}$ by quadratic
-%         approximation at the maximum likelihood estimate
-%         (assumes $\hat{\bm{\theta}} \sim \mathcal{N}(\hat{\bm{\theta}},
-%         \sqrt{-\bm{v}})$, where $\bm{v}$ is a diagonal of the inverse of
-%         the Hessian matrix)\;
-%       }
-%     }
-%   }
-% }
-% \end{algorithm}
-
 <<load-simulation, cache=TRUE>>=
@@
 different sample sizes: $50$, $100$, $300$ and $1000$; generating $1000$
-data sets in each case.  In \autoref{fig:bias-plot}, we show the bias of
-the estimators for each simulation scenario (combination between values
-of the dispersion parameter and samples sizes) along with the confidence
-intervals calculated as average bias plus and minus $\Phi(0.975)$ times
-the average standard error. The scales are standardized for each
-parameter by dividing the average bias by the average standard error
-obtained for the sample of size $50$.
+data sets in each case. For sample sizes 50 and 100, we have 29 and 8
+simulations of the 1000 where algorithm did not converge. These non
+convergence situations occurred for $\phi=-1.6$. In
+\autoref{fig:bias-plot}, we show the bias of the estimators for each
+simulation scenario (combination between values of the dispersion
+parameter and samples sizes) along with the confidence intervals
+calculated as average bias plus and minus $\Phi(0.975)$ times the
+average standard error. The scales are standardized for each parameter
+by dividing the average bias by the average standard error obtained for
+the sample of size $50$.
 
@@
 the deviance function shape under strong overdispersion ($\phi=-1.6$) is
-not as well behaved as the others; this is due to the difficulty of the
-distribution in dealing with strong overdispersion in low counts (see
-dispersion index plot in the \autoref{fig:indexes-plot}). This also
-explains the results of Cov$(\hat{\beta}_{22}, \phi)$ in the first panel
-of \autoref{fig:ortho-plot}, since $\beta_{22}$ is negative and
-associated with low counts.
+not as well behaved as the others; this is due to the accurary of mean
+approximation \autoref{fig:approx-plot}. Although the reparametrized
+model is always valid, $\mu$ and $\phi$ are orthogonal only when $\mu$
+is the expectation of the distribution, in other words, when
+approximation $\lambda - (\nu-1)/(2\nu)$ is accurate. This also explains
+the results of Cov$(\hat{\beta}_{22}, \phi)$ in the first panel of
+\autoref{fig:ortho-plot}, since $\beta_{22}$ is negative and associated
+with low counts.
 
@@
 Following~\citet{Zeviani2014}, the linear predictor is given by
-$$\log(\mu_{ij}) = \beta_0 + \beta_{1j} \texttt{def}_i + \beta_{2j}
+$$\eta_i = \beta_0 + \beta_{1j} \texttt{def}_i + \beta_{2j}
 \texttt{def}_i^2$$
-where $\mu_{ij}$ is the expected number of cotton bolls for the $i$-th
+where $\eta_i = \log(\lambda_i)$ for the original COM-Poisson and
+$\eta_i = \log(\mu_{ij})$ for COM-Poisson$_\mu$. The parameter $\mu_i$
+is the (approximated) expected number of cotton bolls for the $i$-th
 defoliation level ($i=$ 1: 0\%, 2: 25\%, 3: 50\%, 4: 75\% e 5: 100\%)
@@
   \toprule
-<<cres-coton, results="asis">>=
+<<cres-cotton, results="asis">>=
 @
@@
 $$
-\log(\mu_{ijk}) = \beta_0 + \gamma_i + \tau_j +
+\eta_i = \beta_0 + \gamma_i + \tau_j +
   \beta_{1}\texttt{K}_k + \beta_{2}\texttt{K}_k^2 +
@@
 interaction of the first order potassium effect (\texttt{K}) for the
-$j$-th moisture level (\texttt{umid}). \autoref{tab:coef-soy} presents
-the estimates, ratio between estimate and standard error and
-goodness-of-fit measures for the alternative models.
+$j$-th moisture level (\texttt{umid}); The predictor $\eta_i$ is
+$\log(\lambda_i)$ for the original COM-Poisson and $\log(\mu_{ij})$ for
+COM-Poisson$_\mu$.\autoref{tab:coef-soy} presents the estimates, ratio
+between estimate and standard error and goodness-of-fit measures for the
+alternative models.
 
@@
 \begin{minipage}{12cm}
-Linear: $\log(\mu_i) = \beta_0 + \beta_1 \texttt{dose}_i$\\
-Quadratic: $\log(\mu_i) = \beta_0 + \beta_1 \texttt{dose}_i +
+Linear: $\eta_i = \beta_0 + \beta_1 \texttt{dose}_i$\\
+Quadratic: $\eta_i = \beta_0 + \beta_1 \texttt{dose}_i +
              \beta_2 \texttt{dose}_i^2$\\
-Cubic: $\log(\mu_i) = \beta_0 + \beta_1 \texttt{dose}_i +
+Cubic: $\eta_i = \beta_0 + \beta_1 \texttt{dose}_i +
              \beta_2 \texttt{dose}_i^2 + \beta_3 \texttt{dose}_i^3$.
@@
 \end{center}
+where $\eta_i = \log(\lambda_i)$ for the original COM-Poisson and
+$\eta_i = \log(\mu_{ij})$ for COM-Poisson$_\mu$.
 
@@
 novel parametrization was based on a simple asymptotic approximation for
-the expectation and variance of the COM-Poisson distribution. The main
-advantage of the proposed reparametrization is the simple interpretation
-of the regression coefficients in terms of the expectation of the
-response variable as usual in the generalized linear models
-context. Thus, it is possible to compare the results of the COM-Poisson
-model with the ones from standard approaches as the Poisson and
-quasi-Poisson regression models. Furthermore, in the novel
-parametrization the COM-Poisson distribution is indexed by the
-expectation $\mu$ and an extra dispersion parameter $\phi$ which our
-data analysis suggest being orthogonal. This is similar to
-\Citeauthor{Huang2017}'s (\citeyear{Huang2017}) parametrization but is
-simpler, because the $\mu$ is obtained from simple algebra.
-
-We evaluated the accuracy of the asymptotic approximations for the
-expectation and variance of the COM-Poisson distribution by considering
-quadratic approximation errors.  The results showed that the
-approximations are accurate for a large part of the parameter space,
-which in turn support our reparametrization. Moreover, we discuss the
-properties and flexibility of the distribution to deal with count data
-although dispersion, zero-inflation and heavy-tail indexes. We carried
-out a simulation study to assess the properties of the reparametrized
-COM-Poisson model to deal with different levels of dispersion as well as
-the properties of the maximum likelihood estimators. The results of our
-simulation study suggested that the maximum likelihood estimators of the
-regression and dispersion parameters are unbiased and consistent.  The
-empirical coverage rates of the confidence intervals computed based on
-the asymptotic distribution of the maximum likelihood estimators are
-close to the nominal level for sample size greater than $100$.  The
-worst scenario is when we have small sample sizes and strong
-overdispersed counts.  In general, we recommend the use of the
-asymptotic confidence intervals for computational simplicity.
+the expectation of the COM-Poisson distribution. The main advantage of
+the proposed reparametrization is the simple interpretation of the
+regression coefficients in terms of the expectation of the response
+variable as usual in the generalized linear models context. Thus, it is
+possible to compare the results of the COM-Poisson model with the ones
+from standard approaches as the Poisson and quasi-Poisson regression
+models. Furthermore, in the novel parametrization the COM-Poisson
+distribution is indexed by $\mu$ and an extra dispersion parameter
+$\phi$ which our data analysis suggest being orthogonal. This is similar
+to \Citeauthor{Huang2017}'s (\citeyear{Huang2017}) parametrization but
+is simpler, because the $\mu$ is obtained from simple algebra.
+
+The proposed parametrization is always valid, only the interpretation of
+$\mu$ parameter, as expectation of the distribution, depends on the
+accuracy of the approximation. We evaluated the accuracy of the
+approximations for the expectation and variance of the COM-Poisson
+distribution by considering quadratic approximation errors.  The results
+showed that the approximation for the expectation is accurate for a
+large part of the parameter space, which in turn support our
+reparametrization. Moreover, we discuss the properties and flexibility
+of the distribution to deal with count data although dispersion,
+zero-inflation and heavy-tail indexes. We carried out a simulation study
+to assess the properties of the reparametrized COM-Poisson model to deal
+with different levels of dispersion as well as the properties of the
+maximum likelihood estimators. The results of our simulation study
+suggested that the maximum likelihood estimators of the regression and
+dispersion parameters are unbiased and consistent.  The empirical
+coverage rates of the confidence intervals computed based on the
+asymptotic distribution of the maximum likelihood estimators are close
+to the nominal level for sample size greater than $100$.  The worst
+scenario is when we have small sample sizes and strong overdispersed
+counts.  In general, we recommend the use of the asymptotic confidence
+intervals for computational simplicity.
 
@@
 empirical correlation between the regression and dispersion parameter
-estimators suggest orthogonality between $\mu$ and $\phi$ in
-COM-Poisson$_\mu$ distribution. Thus, the computational procedure based
-on the proposed reparametrization is faster than in the original
-parametrization.
+estimators and deviance surfaces suggest orthogonality between $\mu$ and
+$\phi$ in COM-Poisson$_\mu$ distribution. Thus, the computational
+procedure based on the proposed reparametrization is faster than in the
+original parametrization.
 
@@
 model misspecification and to assess the theoretical approximations for
-$Z(\lambda, \nu)$ (or $Z(\mu,\phi))$. Another simple extension of the
-proposed model is to model both $\mu$ and $\phi$ parameters as functions
-of covariate in a double generalized linear models framework.  Finally,
-the reparametrized version of the COM-Poisson model also encourages the
-specification of generalized linear mixed models using this
-distribution.
+$Z(\lambda, \nu)$ (or $Z(\mu,\phi))$ \citep{Gaunt2017}. Another simple
+extension of the proposed model is to model both $\mu$ and $\phi$
+parameters as functions of covariate in a double generalized linear
+models framework.  Finally, the reparametrized version of the
+COM-Poisson model also encourages the specification of generalized
+linear mixed models using this distribution.
 

### 2018-08-27 review/review.Rnw
Update the revision

@@
-\documentclass[12pt, a4paper]{article}
-\usepackage[margin=2cm]{geometry}
+\documentclass[a4paper]{article}
+\usepackage[left=2.2cm, right=2.2cm,
+            top=2.2cm, bottom=2.5cm]{geometry}
 
@@
 % Fonts
-\usepackage{mathpazo}
+%\usepackage{mathpazo}
 \setlength\parindent{0pt}
@@
 
-\textbf{\underline{Response}:} We appreciate the referee for his
+\textbf{\underline{Response}:} We appreciate the referee for her/his
 revision. All suggestions have been taken into account. Regarding the
@@
     refer to the Gaunt et al. (2016) for the latest results.}\\[0.1cm]
-  \textbf{\underline{Response:}}
+  \textbf{\underline{Response:}} Thanks for this reference. We have
+  cited the paper published in the Annals of the Institute of
+  Statistical Mathematics \cite{Gaunt2017}.
 \item \textit{It seems that the paper missing some latest literature on
@@
     were found in Daly and Gaunt (2015).}\\[0.1cm]
-  \textbf{\underline{Response:}}
+  \textbf{\underline{Response:}} Thanks for this reference. We have
+  cited the paper published in the ALEA, Latin American Journal of
+  Probability and Mathematical Statistics \cite{Daly2016}.
 \item \textit{The likelihood equation (4.1) needs more details. Can we
@@
   \textbf{\underline{Response:}} The log-likelihood function cannot be
-  derived in any parametrization. We obtain the Hessian matrix by
-  central finite differences (as said in ...). Regarding the expression
-  of reparametrized log-likelihood (Equation 3.1), although it seems
-  more complicated, this function is better behaved than the function
-  expressed in the original form (see Figure 8). Consider the
-  exponential family of distributions, with density given by
+  derived in relation ti the dispersion parameter using any
+  parametrization. We obtained the Hessian matrix by central finite
+  differences using the Richardson method implemented in
+  \texttt{R}. Although the expression of reparametrized log-likelihood
+  (Equation 3.1) seems more complicated, this function is better behaved
+  than the function expressed in the original form (see Figure
+  8). Consider the exponential family of distributions, with probability
+  function given by
   $$f(y) = \exp\{\sigma[\theta y - b(\theta)] - c(y, \phi)\},$$
@@
   $b'(\theta) = \partial b(\theta)/\partial \theta$ in closed form, the
-  canonical link function has no an important rule. In this case, the
-  link function just maps $\theta$ in the parameter space of $\lambda$.
-  In the reparametrized version of the COM-Poisson, we adopted log link
-  function to map the parameter space and to have good interpretations
-  (in terms of rate ratios).
+  canonical link function has no important rule here. In this case, the
+  link function maps $\theta$ in the parameter space of $\lambda$.  In
+  the reparametrized version of the COM-Poisson, the log link function
+  maps the parameter space and provide simple interpretations (in terms
+  of rate ratios).
 \item \textit{For model fitting, the authors used general purpose
@@
   below). This occurred only for strong overdispersion and small sample
-  size. We report these situations in the updated version of the paper.
+  sizes. We reported these situations in the updated version of the paper.
   \begin{table}[H]
@@
     levels.}\\[0.1cm]
-  \textbf{\underline{Response:}}
+  \textbf{\underline{Response:}} We have computed the times to fit the
+  models for the case studies in the paper, they have different
+  dispersion levels and sample sizes. However, the Figure below show the
+  times to fit in the simulation study. The results show that it took
+  more time to fit the models under equidispersion. This result is due
+  to choose of start values and normalizing constant upper bound. We
+  set, for all models, the upper bound 500 and we start the algorithm
+  from the real parameters.
+
+  \hspace*{-1cm}
+\begin{minipage}{\textwidth}
+<<simul-times, fig.width=9, fig.height=3.5>>=
+@
+\end{minipage}
 
@@
   \textbf{\underline{Response}:} The simulation study was performed with
-  averages varying from 3 to 27. Therefore, considering the different
-  scenarios ($\phi=-1.6, \phi=-1, \phi=0$ and $\phi=1.8$), we cover a
-  wide range of counts. The Figure below shows the variation of the
-  counts for $\mu=27$ in the different scenarios.
+  expectations varying from 3 to 27. Therefore, considering the
+  different scenarios ($\phi=-1.6, \phi=-1, \phi=0$ and $\phi=1.8$), we
+  cover a wide range of counts. The Figure below shows the variation of
+  the counts for $\mu=27$ for the different scenarios.
 
@@
   original parametrization. The steps in the simulation study were i)
-  obtain the $\mu_i = \exp(\bm{x}_i^\top\bm{\beta})$; ii) Back to the
+  obtain the $\mu_i = \exp(\bm{x}_i^\top\bm{\beta})$; ii) back to the
   original parametrization $\lambda = (\mu + (\nu - 1)/(2\nu))^\nu$
-  (Equation 3.1); and iii) simulate data according COM-Poisson
+  (Equation 3.1); and iii) simulate data according to the COM-Poisson
   distribution using the probability integral transform theorem as
@@
   $\mu$ and $\phi$ (and consequently between $\beta_j$ and $\phi$ in a
-  regression setting), leading to better both computation and asymptotic
+  regression setting), leading to better computation and asymptotic
   (normal-based) inference \citep{Ross1970}; and ii) interpretability of
@@
   and heavy-tail indexes that illustrate its applicability in the real
-  count data. These are the methodological novelties of the article
+  count data. These are the methodological novelties of the paper
   besides the exploration of the results. However, in the updated
@@
 point. However, we didn't use the approximation of variance
-($a\text{Var}(Y_i)$) to repametrize or fit the COM-Poisson model. To
-compute de variances, that is showed in the Figure 3(a) and (b) we used
-$Var(Y)=\sum_{y=0}^{500} y^2p(y) - [\sum_{y=0}^{500} y p(y)]^2$. As
-highlighted in the Figure 1(b), the approximation is not accurate,
+($a\text{Var}(Y_i)$) to reparametrize or fit the COM-Poisson model. To
+compute the variances, that is showed in the Figure 3(a) and (b) we used
+$\text{Var}(Y)=\sum_{y=0}^{500} y^2p(y) - [\sum_{y=0}^{500} y
+p(y)]^2$.
+As highlighted in the Figure 1(b), the approximation is not accurate,
 mainly for overdispersion ($\phi<0$). The Figure (a) and (b) below shows
-the behaviour of $a\text{Var}(Y_i)$ function by fixing $\mu$ and $\phi$,
-repectivaly, illustrating the referees' statement. The $a\text{Var}(Y)$
+the behavior of $a\text{Var}(Y_i)$ function by fixing $\mu$ and $\phi$,
+respectively, illustrating the referees' statement. The $a\text{Var}(Y)$
 leads to negative values for variance for small $\mu$ and $\phi<0$ that
@@
 
-The results shown in Figures 7 and 8 reflect that orthogonality property
-is slightly lost for strong overdispersion and small $\mu$. This is
-related to accuracy of the $aE(Y_i)$. When $aE(Y_i)$ is accurate, $\mu$
-represents the expectation of the $Y_i$ and, consequently, $\mu$ and
-$\phi$ (or $\beta_j$ and $\phi$ in a regression setting) are
-orthogonal. We improved the discussion about this in the current version
-of the article.
-\vspace{0.4cm}
+The results shown in Figures 7 and 8 reflect that the orthogonality
+property is slightly lost for strong overdispersion and small
+$\mu$. This is related to the accuracy of the $aE(Y_i)$. When $aE(Y_i)$
+is accurate, $\mu$ represents the expectation of the $Y_i$ and,
+consequently, $\mu$ and $\phi$ (or $\beta_j$ and $\phi$ in a regression
+setting) are orthogonal. We improved the discussion about this in the
+current version of the paper.  \vspace{0.4cm}
 
@@
 relationship for COM-Poisson is linear. Therefore, is suitable to
-analyse data generated according to the COM-Poisson distribution, using
-the assumption $Var(Y_i)\propto \mu_i$. The advantage of the COM-Poisson
-approach is it corresponds to a fully specified probability model
-allowing to compute the likelihood and its resulting results (deviance,
-AIC, BIC).
+analyze data generated according to the COM-Poisson distribution, using
+the assumption $\text{Var}(Y_i)\propto \mu_i$. The advantage of the
+COM-Poisson approach is that it corresponds to a fully specified
+probability model allowing to compute the likelihood and some useful
+measures like deviance, AIC, BIC, LRT, etc.
 \vspace{0.4cm}
@@
   here? This needs clarification.}\\[0.1cm]
-\textbf{\underline{Response:}} Note the $\lambda_i$ does not represente
-the expectation, nor approximately. The quadratic erros in Figure 1(a)
-is obtained by $[a\text{E(Y)} - \sum_{y=0}^{500}y p(y)]^2$.  The models
-fitted in the case studies are
+\textbf{\underline{Response:}} Note that the $\lambda_i$ does not
+represent the expectation, nor approximately. The quadratic errors in
+Figure 1(a) is obtained by $[a\text{E(Y)} - \sum_{y=0}^{500}y p(y)]^2$.
+The models fitted in the case studies are
 $[\text{COMPo}(\lambda_i, \nu), \log(\lambda_i) = \beta^\top x_i]$ and
 $[\text{COMPo}(\mu_i, \phi), \log(\mu_i) = \beta^\top x_i]$. Therefore,
-the $\beta$ estimates are only comparable if $\phi=0$ (Poisson special
-case) otherwise the $\beta$ estimates are on different scales. We make
-it clear which parameters are being placed the linear predictor in
-the current version of the paper. To understand the relationship of the
-coefficients, consider $\log(\lambda_i) = \beta^o$ and $\log(\mu_i) =
-\beta^*$, so $\beta^*$ expressed in terms of $\beta^o$ is
-$\beta^* = \exp(\beta^o) \exp{-\phi} - (\exp(\phi) - 1)/(2\exp(\phi))$.
+the $\beta$'s estimates are comparable only if $\phi=0$ (Poisson special
+case) otherwise they are on different scales. We make it clear in which
+parameter the linear predictor is placed in the current version of the
+paper. To understand the relationship of the coefficients, consider
+$\log(\lambda_i) = \beta^o$ and $\log(\mu_i) = \beta^*$, so $\beta^*$
+expressed in terms of $\beta^o$ is
+$\beta^* = \exp(\beta^o-\phi) - (\exp(\phi) - 1)/(2\exp(\phi))$.
 \vspace{0.4cm}
@@
   $\lambda > [(\nu-1)/(2\nu)]^\nu$. The Figure below show (a) the
-  constraint parameter space; and (b) the expectated values
+  constraint border in the parameter space; and (b) the expected values
   (obtained by $\sum y p(y)$). The constraint lies on the very small
@@
   underdispersion. Therefore, although this constraint is undesirable,
-  it does not prejudice the application of the model.
+  it does not prejudice the application of the model. We thank the
+  referee for highlighting this point. We have discussed it in the
+  updated version.
 
@@
 \vspace{0.3cm}
-
-We thank the referee for highlighting this point. We have discussed it
-in the updated version.
 
@@
   counts with small averages (close to 0), the reparametrized model fits
-  well. For examples presented in the article, the parameter estimates
+  well. For examples presented in the paper, the parameter estimates
   are summarizes below.
@@
   misunderstanding. At the bottom of page 14 the $\bm{\mu}$ is
-  $a\text{E}(\bm{Y})$. We keep symbol $\mu$ to the introduced parameter
+  $a\text{E}(\bm{Y})$. We keep symbol $\mu$ for the introduced parameter
   in order to highlight that it is related to the expectation. However,
@@
     as well.}\\[0.1cm]
-\textbf{\underline{Response:}}
+  \textbf{\underline{Response:}} We have written a more concise
+  abstract. We keep it from a first-person perspective, like other
+  papers published by the journal.
 
@@
   \textbf{\underline{Response:}} The time to fit the
-  COM-Poisson$(\mu_i, \phi)$ model was compared to the time to fit the
-  COM-Poisson$(\lambda_i, \phi)$ model. The computational times for 50
-  repetitions of fit of the models in the three case studies are
-  presented in Figure below.
+  COM-Poisson$(\mu_i, \phi)$ models were compared to the time to fit the
+  COM-Poisson$(\lambda_i, \phi)$ models in the paper. The computational
+  times for 50 repetitions of fit of the models in the three case
+  studies are presented in Figure below.
 
@@
 
-The times to adjust under the proposed parameterization are 110\% faster
-than the original, in the overdispersed case. In addition, the nice
-properties induced by the new parameterization like the orthogonality
-between $\mu$ and $\phi$ and interpretation of $\mu$ are the main
-advantages to warrant the use of the new parameterization besides
-computational times.
+The computation times under the proposed parametrization are 110\%
+faster than the original, in the overdispersed case. In addition, the
+nice properties induced by the new parametrization like the
+orthogonality between $\mu$ and $\phi$ and interpretation of $\mu$ are
+the main advantages to warrant the use of the new parametrization
+besides computational times.
 

### 2018-08-27 review/review.R

@@ -9,7 +9,8 @@ options(digits = 3, OutDec = ".",
         xtable.booktabs = TRUE,
         xtable.sanitize.text.function = identity,
         xtable.size = "small",
-        xtable.math.style.negative = TRUE)
+        xtable.math.style.negative = TRUE,
+        xtable.table.placement = "H")
 opts_chunk$set(
     warning = FALSE,
     message = FALSE,
@@ -31,6 +32,28 @@ library(purrr)
 library(dplyr)
 library(tidyr)
 cols <- trellis.par.get("superpose.line")$col
+
+## ---- simul-times
+phis <- c(0, -1.6, -1, 1.8)
+names(phis) <- sprintf("phi=%s", phis)
+sizes <- c(50, 100, 300, 1000)
+names(sizes) <- sprintf("n=%s", sizes)
+readRDS("../codes/simultimes.rds") %>%
+    mutate(size = factor(sizes[size], levels = sizes),
+           phi = factor(phi, levels = names(phis)[c(2, 3, 1, 4)])) %>%
+    na.omit() %>%
+    bwplot(time ~ size | phi,
+           # groups = phi,
+           layout = c(4, 1),
+           scales = list(y = "free"),
+           ylab = "Time (seconds)",
+           xlab = "Sample size",
+           axis = axis.grid,
+           strip = strip.custom(
+               factor.levels = parse(
+                   text = gsub("=", "==", levels(.$phi)))
+           ),
+           data = .)
 
 ## ---- counts-ranging
 rcmp <- Vectorize(FUN = function(n, mu, phi) {
@@ -222,6 +245,7 @@ xlabs <- expression("COMPo"(lambda[i], phi),
                     "COMPo"(mu[i], phi))
 bwplot(time/10e6 ~ expr | case,
        ylab = "Time (seconds)",
+       axis = axis.grid,
        scales = list(
            y = list(relation = "free"),
            x = list(at = 1:2, labels = xlabs)

### 2018-08-27 presentations/esalq2018/slides-seminar.Rnw
Add presentations

\documentclass[10pt]{beamer}

%=======================================================================
% General packages
\usepackage[british]{babel}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{multicol}
\usepackage{mathtools} %% Funcionalidades (como \dcases)
\usepackage{dsfont}    %% Para \mathds{1} Indicadora
\usepackage{bm}

%=======================================================================
% For figures
\usepackage{graphicx}
\usepackage[export]{adjustbox}

%% ======================================================================
%% Fontes
\usepackage{mathpazo}
\usepackage{inconsolata}
\usepackage{verbatim}

\usefonttheme{professionalfonts}
\usefonttheme{serif}

%=======================================================================
% For tables
\usepackage[bottom]{footmisc}
\usepackage{booktabs}
\usepackage{threeparttable}
\usepackage{tabularx}
\usepackage{colortbl}
\newcolumntype{C}{>{\centering\arraybackslash}X}
\DeclareRobustCommand{\rchi}{{\mathpalette\irchi\relax}}
\newcommand{\irchi}[2]{\raisebox{\depth}{$#1\chi$}}
\usepackage{float}
\makeatletter
\def\@xfootnote[#1]{%
  \protected@xdef\@thefnmark{#1}%
  \@footnotemark\@footnotetext}
\makeatother

%% ======================================================================
%% Cores para links
\definecolor{url}{HTML}{000080}
\definecolor{run}{HTML}{4A0082}
\definecolor{col}{HTML}{6180A6}
\hypersetup{colorlinks, allcolors=., urlcolor=url, runcolor=run}

\setbeamercolor{bibliography entry author}{fg=black}
\usepackage{natbib}
\bibliographystyle{agsm}

%=======================================================================
% Algorithms
\usepackage[english, vlined, ruled]{algorithm2e}
\SetAlCapSty{}
\SetAlCapFnt{\footnotesize\color{col}}
\SetAlCapNameFnt{\footnotesize}

% Customize captions styles
\usepackage{caption}
\captionsetup{
  font=footnotesize,
  labelfont={color=col}
}

%% ======================================================================
%% Tema e cores do documento
\usetheme{CambridgeUS}
\setbeamertemplate{itemize items}[triangle]
\setbeamertemplate{navigation symbols}{}

\setbeamertemplate{frametitle}{
  \nointerlineskip
  \begin{beamercolorbox}[sep=0.3cm, ht=1.8em,
    wd=\paperwidth]{frametitle}
    \vbox{}\vskip-2ex%
    \strut\hspace*{3ex}\Large\bfseries\insertframetitle\strut
    \vskip-0.8ex%
  \end{beamercolorbox}
}

%% Slides em geral
\setbeamercolor{frametitle}{bg=white, fg=col}
\setbeamercolor{structure}{fg=col}
\setbeamercolor{palette primary}{bg=gray!30, fg=col}
\setbeamercolor{palette tertiary}{bg=col, fg=white}
\setbeamercolor{footlinecolor}{fg=white,bg=col}
\setbeamercolor{caption name}{fg=col}

% \setbeamertemplate{frametitle continuation}{[\insertcontinuationcount]}
\setbeamertemplate{frametitle continuation}{}

%% Slide Inicial
\setbeamertemplate{title page}[default]
\setbeamercolor{title}{fg=col}
\setbeamercolor{author}{fg=black!70}
\setbeamercolor{institute}{fg=black!70}
\setbeamercolor{date}{fg=black!70}
\setbeamerfont{title}{series=\bfseries, size=\Large}

%% ======================================================================
%% Definição do cabeçalho e rodapé
\setbeamerfont{headline}{size=\fontsize{5}{5}\selectfont}
\setbeamertemplate{headline}{\bfseries
  \leavevmode%
  \hbox{%
    \begin{beamercolorbox}[wd=.5\paperwidth, ht=2.2ex, dp=1ex, right,
      rightskip=1em]{section in head/foot}
      \hspace*{2ex}\insertsectionhead
    \end{beamercolorbox}%
    \begin{beamercolorbox}[wd=.5\paperwidth, ht=2.2ex, dp=1ex, left,
      leftskip=1em]{subsection in head/foot}
      \insertsubsectionhead\hspace*{2ex}
    \end{beamercolorbox}}
  \vskip0pt
}

\setbeamerfont{footline}{size=\fontsize{5}{5}\selectfont}
\makeatletter
\setbeamertemplate{footline}{\bfseries
  \leavevmode%
  \hbox{%
    \begin{beamercolorbox}[wd=.3\paperwidth, ht=2.4ex, dp=1ex, right,
      rightskip=1em]{footlinecolor}
      \centering \insertshortauthor%
    \end{beamercolorbox}%
    \begin{beamercolorbox}[wd=.6\paperwidth, ht=2.4ex, dp=1ex, left,
      leftskip=1em]{footlinecolor}
      \hfill\insertshorttitle%
    \end{beamercolorbox}%
    \begin{beamercolorbox}[wd=.1\paperwidth, ht=2.4ex, dp=1ex, left,
      leftskip=1em]{footlinecolor}
      \centering Slide \insertframenumber
    \end{beamercolorbox}}
  \vskip0pt
}
\makeatother

%% ======================================================================
%% Layout do tableofcontents
\setbeamertemplate{section in toc}{
  {\color{col} \bfseries\inserttocsectionnumber.}~
  {\leftskip=0.5em\color{black}\inserttocsection\par}
}

\setbeamertemplate{subsection in toc}{
  {\color{col!80}
  \bfseries\inserttocsectionnumber.\inserttocsubsectionnumber}~
  \leftskip=2em{\color{black}\inserttocsubsection\par}
}

%% ======================================================================
%% Formatando slides para seções e subseções
\AtBeginSection[]{
  \begin{frame}[c, allowframebreaks, noframenumbering]
    \begin{center}
      \textcolor{col}{\thesection} \\ \vspace{0.3cm}
      \parbox{0.6\textwidth}{
        \centering \textcolor{col}{\LARGE \bf \insertsection}}\\
    \end{center}
  \end{frame}
}

\AtBeginSubsection{
  \begin{frame}[c, allowframebreaks, noframenumbering]
    \begin{center}
      \textcolor{col}{\thesection.\thesubsection} \\ \vspace{0.3cm}
      \parbox{0.6\textwidth}{
        \centering \textcolor{col!80}{\large \insertsection}\\
        \centering \textcolor{col}{\Large \bf \insertsubsection}}\\
    \end{center}
  \end{frame}
}

%% ======================================================================
%% Metadados não disponíveis no YAML com o atual default.tex template

\title[Reparametrization of COM-Poisson Regression Models]{
  Reparametrization of COM-Poisson Regression Models for Analysis of
  Count Data}

\author[Eduardo E. Ribeiro Jr]{
  Eduardo Elias Ribeiro Junior \inst{1} \inst{2} \\
  Walmes Marques Zeviani \inst{1} \\
  Wagner Hugo Bonat \inst{1} \\
  Clarice Garcia Borges Demétrio \inst{2} \\
  John Hinde \inst{3}
}

\institute{
  \inst{1} Statistics and Geoinformation Laboratory (LEG-UFPR) \\
  \inst{2} Department of Exact Sciences (ESALQ-USP) \\
  \inst{3} School of Mathematics, Statistics and Applied Mathematics (NUI-Galway)
}

\date{\small \today \\[0.2cm]
  \url{jreduardo@usp.br} | \url{edujrrib@gmail.com}}

<<setup, include=FALSE>>=

library(knitr)
library(xtable)
options(digits = 3, OutDec = ".",
        xtable.caption.placement = "top",
        xtable.booktabs = TRUE,
        xtable.sanitize.text.function = identity,
        xtable.sanitize.colnames.function = identity)

opts_chunk$set(
    warning = FALSE,
    message = FALSE,
    echo = FALSE,
    results = "asis",
    ## dev = "tikz",
    fig.width = 7,
    fig.height = 5,
    out.width = "0.99\\textwidth",
    fig.align = "center",
    fig.pos = "h",
    dev.args = list(family = "Palatino"))

thm = knit_theme$get("dusk")
knit_theme$set(thm)

@

\begin{document}

\begin{frame}[plain, noframenumbering]
  \titlepage
\end{frame}

\begin{frame}[noframenumbering]{Outline}
  \tableofcontents[hideallsubsections]
\end{frame}

\section{Background}

\begin{frame}{Poisson model and limitations}
  \vspace{-0.3cm}
  \begin{itemize}
    \item[] {\bf Count data}
    \item Number of times an event occurs in the observation unit;
    \item Mean-variance relationship.
      \vspace{0.3cm}
    \item[] {\bf GLM framework \citep{Nelder1972}}
    \item Provide suitable distribution for a counting random variables;
    \item Efficient algorithm for estimation and inference;
    \item Implemented in many software.
      \vspace{0.3cm}
    \item[] {\bf Poisson model}
    \item Relationship between mean and variance,
      $\text{E}(Y)=\text{Var}(Y)$;
      \vspace{0.3cm}
    \item[] {\bf Main limitations}
    \item Overdispersion (more common),
      $\text{E}(Y)<\text{Var}(Y)$
    \item Underdispersion (less common),
      $\text{E}(Y)>\text{Var}(Y)$
  \end{itemize}

\end{frame}

\begin{frame}{Weighted Poisson models}

  \begin{itemize}
  \item The family of weighted Poisson distributions (WPD)
    \citep{DelCastillo1998}, weights the Poisson probability function by
    a suitable function.
    \vspace{0.3cm}
  \item The probability mass function of the WPD is
    $$
    \Pr(Y = y) = \frac{\exp(-\lambda)\lambda^y}{y!}
    \frac{w(y)}{\text{E}_\lambda[w(Y)]}, \quad y \in \mathbb{N},
    $$
    where $\text{E}_\lambda(\cdot)$ denotes the mean value with respect
    to the Poisson random variable with parameter $\lambda$ and $w(y)$
    is a weight function.
    \vspace{0.3cm}
  \item The weight function may depend on extra parameter to ensure more
    flexibility to the distribution.
  \end{itemize}
\end{frame}

\begin{frame}{COM-Poisson distribution}

  \begin{itemize}
  \item The COM-Poisson \citep{Shmueli2005} belongs to the family of
    weighted Poisson distributions with the weight function
    $w(y,\nu) = (y!)^{1-\nu}$.
    \vspace{0.3cm}
  \item   The probability mass function of $Y$ a
    COM-Poisson random variable is
    \begin{equation*}
      \label{eqn:pmf-compoisson}
      \Pr(Y = y) = \frac{\lambda^y \exp(-\lambda)}{
        (y!)^\nu \text{E}_\lambda[(Y!)^{1-\nu}]} =
      \frac{\lambda^y}{(y!)^\nu \text{Z}(\lambda, \nu)}, \quad
      \text{Z}(\lambda,\nu) = \sum_{j=0}^\infty \frac{\lambda^j}{(j!)^\nu},
    \end{equation*}
    where $\nu$ is the dispersion parameter.
    \vspace{0.3cm}
  \item The moments of distribution are not obtained in closed forms.
    There is approximations proposed by \citet{Shmueli2005,
      Sellers2010}:
    $$
    \text{E}(Y) \approx \dot{\text{E}}(Y) = \lambda^{1/\nu} - \frac{\nu
      - 1}{2\nu} \qquad \textrm{and} \qquad \text{Var}(Y) \approx
    \frac{\lambda^{1/\nu}}{\nu}.
    $$
  \end{itemize}
\end{frame}

\begin{frame}{COM-Poisson regression models}

\begin{itemize}
  \item[] {\bf Model definition}
  \item Modelling the relationship between
    $\text{E}(Y_i)$ and $\bm{x}_i$ indirectly
    \citep{Sellers2010}; \\[-0.5cm]
    \begin{align*}
      Y_i \mid & \,\bm{x}_i \sim
                 \textrm{COM-Poisson}(\lambda_i,\, \nu) \\
               &\eta(\text{E}(Y_i \mid \bm{x}_i)) =
                 \log(\lambda_i) = \bm{x}_i^\top\bm{\beta} &
    \end{align*}
\end{itemize}
\vspace{0.3cm}

\begin{itemize}
  \item[] {\bf Main goals}
  \item Study distribution properties in terms of i) modelling real
    count data and ii) inference aspects.
  \item Propose a reparametrization in order to model the expectation of
    the response variable as a function of the covariate values
    directly.
\end{itemize}

\end{frame}

\section{Reparametrization}

\begin{frame}{Reparametrized COM-Poisson}

\begin{itemize}
  \item[] {\bf Reparametrization}
  \item Introduced new parameter $\mu$, using the mean
    approximation\\[-0.2cm]
    \begin{equation*}
      \label{eqn:repar-cmp}
      \mu = \dot{\text{E}}(Y) = \lambda^{1/\nu} - \frac{\nu - 1}{2\nu}
      \quad \Rightarrow \quad
      \lambda = \left (\mu +
        \frac{(\nu - 1)}{2\nu} \right )^\nu;
    \end{equation*}
  \item Precision parameter is taken on the log scale to avoid
    restrictions on the parameter space\\[-0.3cm]
    $$\phi = \log(\nu) \Rightarrow \phi \in \mathbb{R};$$\\[0.3cm]
  \item[] {\bf Probability mass function}
  \item Replacing $\lambda$ and $\nu$ as function of $\mu$ and $\phi$ in
    the pmf of COM-Poisson\\[-0.2cm]
    \begin{equation*}
      \label{eqn:pmf-cmpmu}
      \Pr(Y=y \mid \mu, \phi) =
      \left ( \mu +\frac{ e^\phi-1}{2e^\phi} \right )^{ye^\phi}
      \frac{(y!)^{-e^\phi}}{Z(\mu, \phi)}.
    \end{equation*}
\end{itemize}

\end{frame}

\begin{frame}{Study of the moments approximations}

<<approx-plot, fig.height=3, fig.width=6, out.width=".9\\textwidth", fig.cap="Quadratic errors for the approximation of the (a) expectation and (b) variance. Dotted lines represent the restriction for suitable approximations given by \\cite{Shmueli2005}.">>=
include_graphics("figure/approx-plot-1.pdf")
@

\end{frame}

\begin{frame}{COM-Poisson$_\mu$ distribution}

<<pmf-cmp, fig.height=3, fig.width=7, fig.cap="Shapes of the COM-Poisson distribution for different parameter values.">>=
include_graphics("figure/pmf-cmp-1.pdf")
@

\end{frame}

\begin{frame}{Properties of COM-Poisson distribution}
  To explore the flexibility of the COM-Poisson distribution, we
  consider the follow indexes:
  \begin{itemize}
  \item \textbf{Dispersion index:}\quad\:\;
    $\text{DI} = \text{Var}(Y)/\text{E}(Y);$\\
  \item \textbf{Zero-inflation index:}\;
    $\text{ZI} = 1 + \log \Pr(Y = 0)/\text{E}(Y);$
  \item \textbf{Heavy-tail index:}\quad\:\;\,
    $\text{HT} = \Pr(Y=y+1)/\Pr(Y=y),\:
    \text{for}\: y \to \infty.$\\
  \end{itemize}
  \vspace{0.4cm}

  These indexes are interpreted in relation to the Poisson distribution:
  \begin{itemize}
  \item
    over- ($\text{DI} > 1$), under- ($\text{DI} < 1$) and
    equidispersion ($\text{DI} = 1$);\\
  \item
    zero-inflation ($\text{ZI} > 0$) and zero-deflation ($\text{ZI} <
    0$) and
  \item
    heavy-tail distribution for $\text{HT} \to 1$ when $y \to
    \infty$.
  \end{itemize}
\end{frame}

\begin{frame}{Properties of COM-Poisson distribution}

<<indexes-plot, fig.width=9.5, fig.height=4, fig.cap="Indexes for COM-Poisson distribution. (a) Mean and variance relationship, (b--d) dispersion, zero-inflation and heavy-tail indexes for different parameter values. Dotted lines represents the Poisson special case.">>=
include_graphics("figure/indexes-plot-1")
@

\end{frame}

\begin{frame}{COM-Poisson$_\mu$ regression models}

\begin{itemize}
\item[] Let $y_i$ a set of independent observations from the COM-Poisson
  and $\bm{x}_i^\top = (x_{i1},\, x_{i2},\, \ldots,\, x_{ip})$ is a
  vector of known covariates, $i=1,2,\ldots,n$.
  \vspace{0.3cm}

\item[] {\bf Model definition}
\item Modelling relationship between $\dot{\text{E}}(Y_i)$ and
  $\bm{x}_i$
  directly\\[-0.5cm]
  \begin{align*}
    Y_i \mid & \,\bm{x}_i \sim
               \textrm{COM-Poisson}_\mu(\mu_i,\, \phi) \\
             &\log(\dot{\text{E}}(Y_i \mid \bm{x}_i)) =
               \log(\mu_i) = \bm{x}_i^\top\bm{\beta} &
  \end{align*}
  \vspace{-0.2cm}
\item[] {\bf Log-likelihood function
    ($\ell = \ell(\bm{\beta},\, \phi \mid \bm{y})$)}
  \vspace{0.05cm}
\item
  $
    \begin{aligned}[t]
    &\ell = e^\phi \left [
      \sum_{i=1}^n y_i
      \log \left( \mu_i + \frac{e^\phi-1}{2e^\phi} \right )
      -\sum_{i=1}^n \log(y_i!) \right ]
      -\sum_{i=1}^n \log(Z(\mu_i, \phi))\\
      &\text{\small where }\:
      \mu_i = \exp(\bm{x}_i^\top\bm{\beta})
      \end{aligned}
  $
\end{itemize}

\end{frame}

\begin{frame}{Estimation and inference}

  The estimation and inference is based on the method of maximum
  likelihood. Let $\bm{\theta} = (\bm{\beta}, \phi)$ the model
  parameters. \vspace{0.2cm}
  \begin{itemize}
  \item Parameter estimates are obtained by numerical maximization of
    the log-likelihood function (by BFGS algorithm);\\[0.1cm]
    $\ell(\hat{\bm{\theta}}) =
    \text{max}\,\ell(\bm{\theta}),\:
    \bm{\theta} \in \mathbb{R}^{p+1}$;
    \vspace{0.2cm}
  \item Standard errors for regression coefficients are obtained based
    on the observed information matrix;\\[0.1cm]
    $\text{Var}(\hat{\bm{\theta}}) = -\mathcal{H}^{-1}$, where
    $\mathcal{H}$ is the matrix of second partial derivatives at
    $\hat{\bm{\theta}}$; \vspace{0.2cm}
  \item Confidence intervals for $\hat{\mu}_i$ are obtained by delta
    method.\\[0.1cm]
    $\text{Var}[g(\hat{\bm{\theta}})] \doteq \bm{G}\,
    \text{Var}(\hat{\bm{\theta}}) \bm{G}^\top $,
    where
    $\bm{G}^\top = (\partial g / \partial \beta_1,
    \ldots, \partial g / \partial \beta_p)^\top$;
    \vspace{0.2cm}
  \item The Hessian matrix $\mathcal{H}$ is obtained numerically by
    finite differences.
  \end{itemize}

\end{frame}

\section{Simulation study}

\begin{frame}{Definitions on the simulation study}
  \small

  \textbf{Objective:} assess the properties of maximum likelihood
  estimators and orthogonality in the reparametrized model;
  \vspace{0.2cm}

  \textbf{Simulation:} we consider counts generated according a
  regression model with a continuous and categorical covariates and
  different dispersion scenarios.
  \vspace{-0.1cm}

  \begin{center}
  \begin{minipage}{0.9\textwidth}
  \begin{algorithm}[H]
    \footnotesize
    \caption{Steps in simulation study.}
      \For{$n \in \{50, 100, 300, 1000\}$}{ set $\bm{x}_1$ as a
        sequence, with $n$ elements, between $0$ and $1$\; set
        $\bm{x}_2$ as a repetition, with $n$ elements, of three
        categories\;
        compute $\bm{\mu}$ using
        $\bm{\mu} = \exp(\beta_0 + \beta_1 \bm{x}_1 + \beta_{21}
        \bm{x}_{21} + \beta_{22} \bm{x}_{22})$\;
        \For{$\phi \in \{-1.6, -1.0, 0.0, 1.8\}$}{
          \Repeat{$1000$ times}{
            simulate $\bm{y}$ from COM-Poisson distribution with
            $\bm{\mu}$ and $\phi$ parameters\;
            fit COM-Poisson$_\mu$ regression model to simulated
            $\bm{y}$\;
            get $\hat{\bm{\theta}} = (\hat{\phi},\;
              \hat{\beta}_0,\; \hat{\beta}_1,\; \hat{\beta}_{21},\;
              \hat{\beta}_{22})$\;
            get confidence intervals for $\hat{\bm{\theta}}$ based on
            the observed information matrix.
          }
        }
      }
  \end{algorithm}
  \end{minipage}
  \end{center}
\end{frame}

\begin{frame}{Definitions on the simulation study}

<<justpars, fig.height=4, fig.width=9, fig.cap="Average counts (left) and dispersion indexes (right) for each scenario considered in the simulation study.">>=
include_graphics("figure/justpars-1")
@

\end{frame}

\begin{frame}{Bias of the estimators}

\vspace{-0.5cm}
<<bias-plot, out.width="0.9\\textwidth", fig.height=6.5, fig.width=8.5, fig.cap="Distributions of standardized bias (gray box-plots) and average with confidence intervals (black segments) by different sample sizes and dispersion levels.">>=
include_graphics("figure/bias-plot-1")
@

\end{frame}

\begin{frame}{Coverage rate of the confidence intervals}

<<coverage-plot, fig.height=4, fig.width=9, fig.cap="Coverage rate based on confidence intervals obtained by quadratic approximation for different sample sizes and dispersion levels.">>=
include_graphics("figure/coverage-plot-1")
@

\end{frame}

\begin{frame}{Orthogonality property of the MLEs}
\vspace{-0.3cm}

<<ortho-surf, fig.height=5, fig.width=9, fig.cap="Deviance surfaces contour plots under original and proposed parametrization. The ellipses are confidence regions (90, 95 and 99\\%), dotted lines are the maximum likelihood estimates, and points are the real parameters used in the simulation.">>=
include_graphics("figure/ortho-surf-1")
@

\end{frame}

\section{Case studies}

\begin{frame}[t]{Motivating data sets and data analysis}

  \begin{itemize}
  \item Three illustrative examples of count data analysis are
    reported.
    \begin{itemize} \normalsize
    \item Assessing toxicity of nitrofen in aquatic systems, an
      equidispersed example;
    \item Soil moisture and potassium doses on soybean culture, an
      overdispersed example; and
    \item Artificial defoliation in cotton phenology, an
      underdispersed example.
    \end{itemize}
    \vspace{0.3cm}
  \item In the data analysis, we consider:
    \begin{itemize}
    \item The COM-Poisson (original parametrization) model;
    \item The COM-Poisson (new parametrization) model;
    \item Quasi-Poisson model ($\text{Var}(Y) = \sigma
      \text{E}(Y)$); and
    \item The standard Poisson regression model.
    \end{itemize}
  \end{itemize}
\end{frame}

\subsection{Artificial defoliation in cotton phenology}

\begin{frame}{Cotton bolls data}

  \begin{columns}
    \begin{column}{0.4\textwidth}
      \begin{figure}
        \includegraphics[scale=0.25]{./images/defoliation}
      \end{figure}
    \end{column}
    \begin{column}{0.6\textwidth}
      {\bf Aim:} to assess the effects of five defoliation levels on the
      bolls produced at five growth stages;\\[0.2cm]
      {\bf Design:} factorial 5 $\times$ 5, with 5 replicates;\\[0.2cm]
      {\bf Experimental unit:} a plot with 2 plants;\\[0.2cm]
      {\bf Factors:}
      \begin{itemize} \small
        \item Artificial defoliation (\texttt{des}):
        \item Growth stage (\texttt{est}):
      \end{itemize}\vspace{0.1cm}
      {\bf Response variable:} Total number of cotton bolls;\\
    \end{column}
  \end{columns}

\end{frame}

\begin{frame}{Cotton bolls data}
\vspace{-0.2cm}
<<desc-cotton, fig.width=9, fig.height=5.8, fig.cap="">>=

knitr::include_graphics("figure/desc-cotton-1")

@

\end{frame}

\begin{frame}{Model specification}

  {\bf Linear predictor:} {\small following \citet{Zeviani2014}}
  \begin{itemize}
    \item $\log(\mu_{ij}) = \beta_0 + \beta_{1j} \textrm{def}_i +
      \beta_{2j} \textrm{def}_i^2$\\[0.1cm]
    {\small
    $i$ varies in the levels of artificial defoliation;\\
    $j$ varies in the levels of growth stages.}
  \end{itemize}

  \vspace{0.5cm}

  {\bf Alternative models:}
  % {\bf Modelling strategies:}
  \begin{itemize}
    \item Poisson ($\mu_{ij}$);
    \item COM-Poisson ($\lambda_{ij}=\eta(\mu_{ij})\,,\,\phi$)
    \item COM-Poisson$_\mu$ ($\mu_{ij}\,,\,\phi$)
    \item Quasi-Poisson (Var$(Y_{ij})=\sigma\mu_{ij}$)
  \end{itemize}

\end{frame}

\begin{frame}{Parameter estimates}

  \vspace{-0.6cm}
\begin{table}[!ht]
\centering \footnotesize
\caption{Parameter estimates (Est) and ratio between estimate and
  standard error (SE).}
\vspace{-0.1cm}
\label{tab:coef-cotton}
\begin{tabular}{lrrrrrrrr}
  \toprule
  & \multicolumn{2}{c}{Poisson} &
    \multicolumn{2}{c}{COM-Poisson} &
    \multicolumn{2}{c}{COM-Poisson$_\mu$} &
    \multicolumn{2}{c}{Quasi-Poisson} \\
\cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7} \cmidrule(lr){8-9}
% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Tue Jun 26 13:34:54 2018
 & Est & Est/SE & Est & Est/SE & Est & Est/SE & Est & Est/SE \\
  \midrule
  \rowcolor{col!30}
$\phi\,,\,\sigma$ &  &  & 1.585 & 12.417 & 1.582 & 12.392 & 0.241 &  \\
  $\beta_0$ & 2.190 & 34.572 & 10.897 & 7.759 & 2.190 & 74.640 & 2.190 & 70.420 \\
  $\beta_{11}$ & 0.437 & 0.847 & 2.019 & 1.770 & 0.435 & 1.819 & 0.437 & 1.726 \\
  $\beta_{12}$ & 0.290 & 0.571 & 1.343 & 1.211 & 0.288 & 1.223 & 0.290 & 1.162 \\
  $\beta_{13}$ & $-$1.242 & $-$2.058 & $-$5.750 & $-$3.886 & $-$1.247 & $-$4.420 & $-$1.242 & $-$4.192 \\
  $\beta_{14}$ & 0.365 & 0.645 & 1.595 & 1.298 & 0.350 & 1.328 & 0.365 & 1.314 \\
  $\beta_{15}$ & 0.009 & 0.018 & 0.038 & 0.035 & 0.008 & 0.032 & 0.009 & 0.036 \\
  $\beta_{21}$ & $-$0.805 & $-$1.379 & $-$3.725 & $-$2.775 & $-$0.803 & $-$2.961 & $-$0.805 & $-$2.809 \\
  $\beta_{22}$ & $-$0.488 & $-$0.861 & $-$2.265 & $-$1.805 & $-$0.486 & $-$1.850 & $-$0.488 & $-$1.754 \\
  $\beta_{23}$ & 0.673 & 0.989 & 3.135 & 2.084 & 0.679 & 2.135 & 0.673 & 2.015 \\
  $\beta_{24}$ & $-$1.310 & $-$1.948 & $-$5.894 & $-$3.657 & $-$1.288 & $-$4.095 & $-$1.310 & $-$3.967 \\
  $\beta_{25}$ & $-$0.020 & $-$0.036 & $-$0.090 & $-$0.076 & $-$0.019 & $-$0.074 & $-$0.020 & $-$0.074 \\
   \specialrule{0.01em}{0.3em}{0.3em}
 LogLik & \multicolumn{2}{c}{$-255.803$} & \multicolumn{2}{c}{$-208.250$} & \multicolumn{2}{c}{$-208.398$} & \multicolumn{2}{c}{$  -$}\\
 AIC & \multicolumn{2}{c}{$533.606$} & \multicolumn{2}{c}{$440.500$} & \multicolumn{2}{c}{$440.795$} & \multicolumn{2}{c}{$  -$}\\
 BIC & \multicolumn{2}{c}{$564.718$} & \multicolumn{2}{c}{$474.440$} & \multicolumn{2}{c}{$474.735$} & \multicolumn{2}{c}{$  -$} \\
 \bottomrule
\end{tabular}
\end{table}
\end{frame}

\begin{frame}{Fitted curves}

<<pred-cotton, fig.height=3.5, fig.width=8.5, fig.cap="Scatterplots of the observed data and curves of fitted values with 95\\% confidence intervals as functions of the defoliation level for each growth stage.">>=
include_graphics("figure/pred-cotton-1")
@

\end{frame}

\subsection{Soil moisture and potassium doses on soybean culture}

\begin{frame}{Soybean data}

  \begin{columns}
    \begin{column}{0.4\textwidth}
      \begin{figure}
        \includegraphics[width=4cm]{./images/soybean}
      \end{figure}
    \end{column}
    \begin{column}{0.6\textwidth}
      {\bf Aim:} evaluate the effects of potassium doses applied
      to soil in different soil moisture levels;\\[0.2cm]
      {\bf Design:} factorial 5 $\times$ 3 in a randomized complete
      block design (5 blocks);\\[0.2cm]
      {\bf Experimental unit:} a pot with a plant;\\[0.2cm]
      {\bf Factors:}
      \begin{itemize} \small
        \item Potassium fertilization dose (\texttt{K}):
        \item Soil moisture level (\texttt{umid}):
      \end{itemize}\vspace{0.1cm}
      {\bf Response variable:} Total number of bean seeds per pot;\\
    \end{column}
  \end{columns}

\end{frame}

\begin{frame}{Soybean data}
\vspace{-0.2cm}
<<desc-soy, fig.width=9, fig.height=5.8, fig.cap="">>=

knitr::include_graphics("figure/desc-soy-1")

@

\end{frame}

\begin{frame}{Model specification}

  {\bf Linear predictor:} {\small based on descriptive analysis,}
  \begin{itemize}
    \item $\log(\mu_{ijk}) = \beta_0 + \gamma_i + \tau_j +
      \beta_{1}\texttt{K}_k + \beta_{2}\texttt{K}_k^2 +
      \beta_{3j}\texttt{K}_k$\\[0.1cm]
    {\small
    $i$ varies according the blocks;\\
    $j$ varies in the levels of soil moisture;\\
    $k$ varies in the levels of potassium fertilization.}
  \end{itemize}

  \vspace{0.5cm}

  {\bf Alternative models:}
  % {\bf Modelling strategies:}
  \begin{itemize}
    \item Poisson ($\mu_{ij}$);
    \item COM-Poisson ($\lambda_{ij}=\eta(\mu_{ij})\,,\,\phi$)
    \item COM-Poisson$_\mu$ ($\mu_{ij}\,,\,\phi$)
    \item Quasi-Poisson (var$(Y_{ij})=\sigma\mu_{ij}$)
  \end{itemize}

\end{frame}

\begin{frame}{Parameter estimates}

\begin{table}[!ht]
\centering \footnotesize
\caption{Parameter estimates (Est) and ratio between estimate and
  standard error (SE).}
\vspace{-0.1cm}
\label{tab:coef-soy}
\begin{tabular}{lrrrrrrrr}
  \toprule
  & \multicolumn{2}{c}{Poisson} &
    \multicolumn{2}{c}{COM-Poisson} &
    \multicolumn{2}{c}{COM-Poisson$_\mu$} &
    \multicolumn{2}{c}{Quasi-Poisson} \\
\cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7} \cmidrule(lr){8-9}
% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Wed Jun 27 14:44:34 2018
 & Est & Est/SE & Est & Est/SE & Est & Est/SE & Est & Est/SE \\
  \midrule
$\phi\,,\,\sigma$ &  &  & $-$0.779 & $-$4.721 & $-$0.782 & $-$4.737 & 2.615 &  \\
  $\beta_0$ & 4.867 & 144.289 & 2.232 & 6.042 & 4.867 & 97.781 & 4.867 & 89.225 \\
  $\gamma_{1}$ & $-$0.019 & $-$0.730 & $-$0.009 & $-$0.494 & $-$0.019 & $-$0.495 & $-$0.019 & $-$0.452 \\
  $\gamma_{2}$ & $-$0.037 & $-$1.373 & $-$0.017 & $-$0.921 & $-$0.037 & $-$0.931 & $-$0.037 & $-$0.849 \\
  $\gamma_{3}$ & $-$0.106 & $-$3.889 & $-$0.049 & $-$2.422 & $-$0.106 & $-$2.634 & $-$0.106 & $-$2.405 \\
  $\gamma_{4}$ & $-$0.092 & $-$3.300 & $-$0.042 & $-$2.102 & $-$0.092 & $-$2.237 & $-$0.092 & $-$2.040 \\
  $\tau_{1}$ & 0.132 & 3.647 & 0.061 & 2.295 & 0.132 & 2.472 & 0.132 & 2.255 \\
  $\tau_{2}$ & 0.124 & 3.432 & 0.057 & 2.177 & 0.124 & 2.326 & 0.124 & 2.122 \\
  $\beta_1$ & 0.616 & 11.014 & 0.284 & 4.729 & 0.616 & 7.464 & 0.616 & 6.811 \\
  $\beta_2$ & $-$0.276 & $-$10.250 & $-$0.127 & $-$4.589 & $-$0.276 & $-$6.946 & $-$0.276 & $-$6.338 \\
  $\beta_{31}$ & 0.146 & 4.268 & 0.067 & 2.614 & 0.146 & 2.892 & 0.146 & 2.639 \\
  $\beta_{32}$ & 0.165 & 4.829 & 0.076 & 2.884 & 0.165 & 3.272 & 0.165 & 2.986 \\
   \specialrule{0.01em}{0.3em}{0.3em}
 LogLik & \multicolumn{2}{c}{$-340.082$} & \multicolumn{2}{c}{$-325.241$} & \multicolumn{2}{c}{$-325.233$} & \multicolumn{2}{c}{$  -$}\\
 AIC & \multicolumn{2}{c}{$702.164$} & \multicolumn{2}{c}{$674.482$} & \multicolumn{2}{c}{$674.467$} & \multicolumn{2}{c}{$  -$}\\
 BIC & \multicolumn{2}{c}{$727.508$} & \multicolumn{2}{c}{$702.130$} & \multicolumn{2}{c}{$702.116$} & \multicolumn{2}{c}{$  -$} \\
  \bottomrule
\end{tabular}
\end{table}
\end{frame}

\begin{frame}{Fitted curves}

<<pred-soy, fig.height=3.8, fig.width=8, fig.cap="Dispersion diagrams of been seeds counts as function of potassium doses and humidity levels with fitted curves and confidence intervals (95\\%).">>=
include_graphics("figure/pred-soy-1")
@

\end{frame}

\subsection{Assessing toxicity of nitrofen in aquatic systems}

\begin{frame}{Nitrofen data}

  \begin{columns}
    \begin{column}{0.4\textwidth}
      \begin{figure}
        \includegraphics[width=4cm]{./images/zooplankton}
      \end{figure}
    \end{column}
    \begin{column}{0.6\textwidth}
      {\bf Aim:} measure the reproductive toxicity of the herbicide
      nitrofen on a species of zooplankton
      (\textit{Ceriodaphnia dubia});\\[0.2cm]
      {\bf Design:} completely randomized design, with 10
      replicates;\\[0.2cm]
      {\bf Experimental unit:} zooplankton animal;\\[0.2cm]
      {\bf Factors:}
      \begin{itemize} \small
        \item herbicide nitrofen dose (\texttt{dose});
      \end{itemize}\vspace{0.1cm}
      {\bf Response variable:} Total number of live offspring;\\
    \end{column}
  \end{columns}

\end{frame}

\begin{frame}{Model specification}

  \textbf{Linear predictors:}
  \begin{center}
    \begin{tabular}{ll}
      Linear:     &
          $\log(\mu_i) = \beta_0 + \beta_1 \texttt{dose}_i$,\\
      Quadratic: &
          $\log(\mu_i) = \beta_0 + \beta_1 \texttt{dose}_i +
             \beta_2 \texttt{dose}_i^2$ and\\
      Cubic:     &
          $\log(\mu_i) = \beta_0 + \beta_1 \texttt{dose}_i +
             \beta_2 \texttt{dose}_i^2 + \beta_3 \texttt{dose}_i^3.$
    \end{tabular}
  \end{center}

  \vspace{0.5cm}

  {\bf Alternative models:}
  % {\bf Modelling strategies:}
  \begin{itemize}
    \item Poisson ($\mu_{ij}$);
    \item COM-Poisson ($\lambda_{ij}=\eta(\mu_{ij})\,,\,\phi$)
    \item COM-Poisson$_\mu$ ($\mu_{ij}\,,\,\phi$)
    \item Quasi-Poisson (var$(Y_{ij})=\sigma\mu_{ij}$)
  \end{itemize}

\end{frame}

\begin{frame}{Likelihood ratio tests}

\vspace{-0.4cm}
\begin{table}[!ht]
\centering \footnotesize
\caption{Model fit measures and comparisons between linear predictors.}
\vspace{-0.2cm}
\label{tab:anova-ovos}
%\begin{tabularx}{\textwidth}{lCCCCCrC}
\begin{tabular}{lcccccrr}
\toprule
 Poisson & np & $\ell$ & AIC & 2(diff $\ell$) & diff np & P($>\rchi^2$) & \\
 \midrule
% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Wed Jun 27 09:38:32 2018
 Linear & 2 & $-$180.667 & 365.335 &  &  &  &  \\
 Quadratic & 3 & $-$147.008 & 300.016 & 67.319 & 1 & 2.31E$-$16 &  \\
 Cubic & 4 & $-$144.090 & 296.180 & 5.835 & 1 & 1.57E$-$02 &  \\

\specialrule{0em}{0.5em}{0em} %% Apenas para espaçamento
  COM-Poisson & np & $\ell$ & AIC & 2(diff $\ell$) & diff np &
  P($>\rchi^2$) & $\hat{\phi}$ \\
  \midrule
% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Wed Jun 27 09:38:32 2018
 Linear & 3 & $-$167.954 & 341.908 &  &  &  & $-$0.893 \\
  Quadratic & 4 & $-$146.964 & 301.929 & 41.980 & 1 & 9.22E$-$11 & $-$0.059 \\
  Cubic & 5 & $-$144.064 & 298.129 & 5.800 & 1 & 1.60E$-$02 & 0.048 \\

\specialrule{0em}{0.5em}{0em} %% Apenas para espaçamento
  COM-Poisson$_\mu$ & np & $\ell$ & AIC & 2(diff $\ell$) & diff np &
  P($>\rchi^2$) & $\hat{\phi}$ \\
  \midrule
% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Wed Jun 27 09:38:32 2018
 Linear & 3 & $-$167.652 & 341.305 &  &  &  & $-$0.905 \\
  Quadratic & 4 & $-$146.950 & 301.900 & 41.405 & 1 & 1.24E$-$10 & $-$0.069 \\
  Cubic & 5 & $-$144.064 & 298.127 & 5.773 & 1 & 1.63E$-$02 & 0.047 \\

\specialrule{0em}{0.5em}{0em} %% Apenas para espaçamento
  Quasi-Poisson & np & QDev & AIC & F & diff np & P($>F$) & $\hat{\sigma}$ \\
  \midrule
% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Wed Jun 27 09:38:32 2018
 Linear  & 3 & 123.929 &  &  &  &  & 2.262 \\
  Quadratic & 4 & 56.610 &  & 60.840 & 1 & 5.07E$-$10 & 1.106 \\
  Cubic & 5 & 50.774 &  & 5.659 & 1 & 2.16E$-$02 & 1.031 \\

 \bottomrule
%\end{tabularx}
\end{tabular}
\end{table}

\end{frame}

\begin{frame}{Parameter estimates and fitted values}

  \vspace{-0.4cm}
\begin{table}[!ht]
\centering \footnotesize
\caption{Parameter estimates (Est) and ratio between estimate and
  standard error (SE).}
\vspace{-0.1cm}
\label{tab:coef-ovos}
\begin{tabular}{lrrrrrrrr}
  \toprule
  & \multicolumn{2}{c}{Poisson} &
    \multicolumn{2}{c}{COM-Poisson} &
    \multicolumn{2}{c}{COM-Poisson$_\mu$} &
    \multicolumn{2}{c}{Quasi-Poisson} \\
\cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7} \cmidrule(lr){8-9}
% latex table generated in R 3.4.4 by xtable 1.8-2 package
% Wed Jun 27 09:38:32 2018
 & Est & Est/SE & Est & Est/SE & Est & Est/SE & Est & Est/SE \\
  \midrule
$\beta_{0}$ & 3.477 & 62.817 & 3.649 & 4.850 & 3.477 & 64.308 & 3.477 & 61.860 \\
  $\beta_{1}$ & $-$0.086 & $-$0.433 & $-$0.091 & $-$0.448 & $-$0.088 & $-$0.452 & $-$0.086 & $-$0.426 \\
  $\beta_{2}$ & 0.153 & 0.863 & 0.161 & 0.878 & 0.155 & 0.894 & 0.153 & 0.850 \\
  $\beta_{3}$ & $-$0.097 & $-$2.398 & $-$0.102 & $-$2.229 & $-$0.098 & $-$2.464 & $-$0.097 & $-$2.361 \\
   \bottomrule
\end{tabular}
\end{table}
  \vspace{-0.4cm}

<<pred-ovos, fig.height=3.5, fig.width=5.5, out.width="0.8\\textwidth", fig.cap="Number of live offsprings observed for each nitrofen concentration level with fitted curves and 95\\% confidence intervals.">>=
include_graphics("figure/pred-ovos-1")
@

\end{frame}

\subsection*{}

\begin{frame}{Additional results}

  To compare the computational times on the two parametrizations we
  repeat the fitting 50 times.
  \vspace{-0.2cm}

<<comp-times, fig.height=4, fig.width=8, fig.cap="Computational times to fit the models under original and reparametrized versions based on the fifty repetitions.">>=

library(latticeExtra)
source("../../codes/lattice-panels.R")

benchs <- readRDS("../../codes/comparetimes.rds")
xlabs <- expression("COM-Poisson", "COM-Poisson"[mu])
bwplot(time ~ expr | case,
       ylab = "Time (seconds)",
       scales = list(
           y = list(relation = "free"),
           x = list(at = 1:2, labels = xlabs)
       ),
       data = benchs)

@

\end{frame}

\section{Final remarks}

\begin{frame}{Concluding remarks}

  {\bf Summary}

  \begin{itemize}
    \item Over/under-dispersion needs caution;
    \item COM-Poisson is a suitable choice for these situations;
    \item The proposed reparametrization, COM-Poisson$_\mu$ has some
      advantages:
      \begin{itemize}
        \item Simple transformation of the parameter space;
        \item Leads to the orthogonality of the parameters (seen
          empirically);
        \item Full parametric approach;
        \item Empirical correlation between the estimators was
          practically null;
        \item Faster for fitting;
        \item Allows interpretation of the coefficients directly (like
          GLM-Poisson model).
      \end{itemize}
  \end{itemize}

  {\bf Future work}

  \begin{itemize}
    \item Simulation study to assess model robustness against
      distribution miss specification;
    \item Assess theoretical approximations for $Z(\lambda, \nu)$
      (or $Z(\mu, \phi)$), in order to avoid the selection of sum's
      upper bound;
    \item Propose a mixed GLM based on the COM-Poisson$_\mu$ model.
  \end{itemize}
\end{frame}

\section*{Notes}

\begin{frame}

\begin{itemize}
  \item \begin{columns}[c]
      \column{.1\linewidth}
      \begin{flushright}
        \includegraphics[height=0.9cm,valign=c]{./images/arxiv}
      \end{flushright}
      \column{.8\linewidth}
      \hspace*{-0.7cm} Full-text article is available on arXiv\\
      \hspace*{-0.7cm} \url{https://arxiv.org/abs/1801.09795}
    \end{columns}
  \item \begin{columns}[c]
      \column{.1\linewidth}
      \begin{flushright}
        \includegraphics[height=0.9cm,valign=c]{./images/github-logo}
      \end{flushright}
      \column{.8\linewidth}
      \hspace*{-0.7cm} All codes (in R) and source files are available
      on GitHub\\
      \hspace*{-0.7cm} \url{https://github.com/jreduardo/article-reparcmp}
    \end{columns}
\end{itemize}

\vspace{0.5cm}

{\bf Acknowledgements}

\begin{itemize}
    \item \begin{columns}[c]
      \column{.1\linewidth}
      \begin{flushright}
        \includegraphics[height=0.9cm,valign=c]{./images/cnpq-logo}
      \end{flushright}
      \column{.8\linewidth} \vspace{-0.3cm}
      \hspace*{-0.7cm}\begin{minipage}[c]{\textwidth}
       National Council for Scientific and Technological Development
       (CNPq), for their support.\\
      \end{minipage}
    \end{columns}
\end{itemize}

\end{frame}

\section*{Bibliography}
\begin{frame}[allowframebreaks]{References}
  \small
  \vspace*{-0.5cm}
  \bibliography{../../references.bib}
\end{frame}

\end{document}

### 2018-08-27 presentations/ibc2018/slides-ibc2018.Rnw
Add presentations

--- 2018-08-27 presentations/esalq2018/slides-seminar.Rnw
@@
 \title[Reparametrization of COM-Poisson Regression Models]{
-  Reparametrization of COM-Poisson Regression Models for Analysis of
-  Count Data}
+  Reparametrization of COM-Poisson Regression Models with Applications
+  in the Analysis of Experimental Count Data \vspace{-0.4cm}}
 
-\author[Eduardo E. Ribeiro Jr]{
+\author[Clarice Garcia Borges Demétrio]{
   Eduardo Elias Ribeiro Junior \inst{1} \inst{2} \\
@@
 
-\date{\small \today \\[0.2cm]
+\date{\small 10th July 2018 \\
   \url{jreduardo@usp.br} | \url{edujrrib@gmail.com}}
@@
 \begin{frame}[plain, noframenumbering]
+  \begin{center}
+    \bfseries
+    XXIX\textsuperscript{TH} INTERNATIONAL BIOMETRIC CONFERENCE\\
+    Barcelona, Spain, 8-13 July 2018
+  \end{center}
+  \vspace{-0.5cm}
   \titlepage
@@
 
+\begin{frame}{Count data}
+
+  Number of times an event occurs in the observation unit.\\[0.3cm]
+
+  Random variables that assume non-negative integer values.\\[0.3cm]
+
+  Let $Y$ be a counting random variable, so that $y=0,1,2,\ldots$\\[0.5cm]
+
+  Examples in experimental researches:
+
+  \begin{itemize}
+    \item number of grains produced by a plant;
+    \item number of fruits produced by a tree;
+    \item number of insects on a particular cell;
+    \item others.
+  \end{itemize}
+
+\end{frame}
+
 \begin{frame}{Poisson model and limitations}
-  \vspace{-0.3cm}
+
   \begin{itemize}
-    \item[] {\bf Count data}
-    \item Number of times an event occurs in the observation unit;
-    \item Mean-variance relationship.
-      \vspace{0.3cm}
     \item[] {\bf GLM framework \citep{Nelder1972}}
@@
     \item Implemented in many software.
-      \vspace{0.3cm}
+      \vspace{0.5cm}
     \item[] {\bf Poisson model}
@@
       $\text{E}(Y)=\text{Var}(Y)$;
-      \vspace{0.3cm}
+      \vspace{0.5cm}
     \item[] {\bf Main limitations}
@@
 
-\begin{frame}{Weighted Poisson models}
-
-  \begin{itemize}
-  \item The family of weighted Poisson distributions (WPD)
-    \citep{DelCastillo1998}, weights the Poisson probability function by
-    a suitable function.
-    \vspace{0.3cm}
-  \item The probability mass function of the WPD is
-    $$
-    \Pr(Y = y) = \frac{\exp(-\lambda)\lambda^y}{y!}
-    \frac{w(y)}{\text{E}_\lambda[w(Y)]}, \quad y \in \mathbb{N},
-    $$
-    where $\text{E}_\lambda(\cdot)$ denotes the mean value with respect
-    to the Poisson random variable with parameter $\lambda$ and $w(y)$
-    is a weight function.
-    \vspace{0.3cm}
-  \item The weight function may depend on extra parameter to ensure more
-    flexibility to the distribution.
-  \end{itemize}
-\end{frame}
-
 \begin{frame}{COM-Poisson distribution}
@@
   \begin{itemize}
-  \item The COM-Poisson \citep{Shmueli2005} belongs to the family of
-    weighted Poisson distributions with the weight function
-    $w(y,\nu) = (y!)^{1-\nu}$.
-    \vspace{0.3cm}
-  \item   The probability mass function of $Y$ a
-    COM-Poisson random variable is
+    \itemsep7.5pt\parskip0pt\parsep0pt
+  \item Probability mass function \citep{Shmueli2005} takes the form \\
+    \begin{equation*} \small
+    \label{eqn:pmf-cmp}
+    \Pr(Y = y \mid \lambda, \nu) =
+      \frac{\lambda^y}{(y!)^\nu Z(\lambda, \nu)}\,,
+    \qquad Z(\lambda, \nu) = \sum_{j=0}^\infty
+      \frac{\lambda^j}{(j!)^\nu}\,,
+  \end{equation*}
+    where $\lambda > 0$ and $\nu \geq 0$.
+  \item Moments are not available in closed form;
+  \item Expectation and variance can be closely approximated by
     \begin{equation*}
-      \label{eqn:pmf-compoisson}
-      \Pr(Y = y) = \frac{\lambda^y \exp(-\lambda)}{
-        (y!)^\nu \text{E}_\lambda[(Y!)^{1-\nu}]} =
-      \frac{\lambda^y}{(y!)^\nu \text{Z}(\lambda, \nu)}, \quad
-      \text{Z}(\lambda,\nu) = \sum_{j=0}^\infty \frac{\lambda^j}{(j!)^\nu},
+      \label{eqn:mean-aprox}
+      \text{E}(Y) \approx \lambda^{1/\nu} - \frac{\nu - 1}{2\nu} \qquad
+      \textrm{and} \qquad
+      \text{Var}(Y) \approx \frac{\lambda^{1/\nu}}{\nu}
     \end{equation*}
-    where $\nu$ is the dispersion parameter.
-    \vspace{0.3cm}
-  \item The moments of distribution are not obtained in closed forms.
-    There is approximations proposed by \citet{Shmueli2005,
-      Sellers2010}:
-    $$
-    \text{E}(Y) \approx \dot{\text{E}}(Y) = \lambda^{1/\nu} - \frac{\nu
-      - 1}{2\nu} \qquad \textrm{and} \qquad \text{Var}(Y) \approx
-    \frac{\lambda^{1/\nu}}{\nu}.
-    $$
+    with accurate approximations for $\nu \leq 1$ or $\lambda > 10^\nu$
+    \citep{Shmueli2005, Sellers2012}.
   \end{itemize}
+
 \end{frame}
@@
                  \textrm{COM-Poisson}(\lambda_i,\, \nu) \\
-               &\eta(\text{E}(Y_i \mid \bm{x}_i)) =
+               &\eta(E(Y_i \mid \bm{x}_i)) =
                  \log(\lambda_i) = \bm{x}_i^\top\bm{\beta} &
@@
       \label{eqn:repar-cmp}
-      \mu = \dot{\text{E}}(Y) = \lambda^{1/\nu} - \frac{\nu - 1}{2\nu}
+      \mu = \lambda^{1/\nu} - \frac{\nu - 1}{2\nu}
       \quad \Rightarrow \quad
@@
 \item[] {\bf Model definition}
-\item Modelling relationship between $\dot{\text{E}}(Y_i)$ and
-  $\bm{x}_i$
+\item Modelling relationship between $E(Y_i)$ and $\bm{x}_i$
   directly\\[-0.5cm]
@@
                \textrm{COM-Poisson}_\mu(\mu_i,\, \phi) \\
-             &\log(\dot{\text{E}}(Y_i \mid \bm{x}_i)) =
+             &\log(E(Y_i \mid \bm{x}_i)) =
                \log(\mu_i) = \bm{x}_i^\top\bm{\beta} &
@@
     \ldots, \partial g / \partial \beta_p)^\top$;
-    \vspace{0.2cm}
   \item The Hessian matrix $\mathcal{H}$ is obtained numerically by
@@
 \vspace{-0.3cm}
-
 <<ortho-surf, fig.height=5, fig.width=9, fig.cap="Deviance surfaces contour plots under original and proposed parametrization.
@@
       \end{itemize}
-    \vspace{0.3cm}
-  \item In the data analysis, we consider:
-    \begin{itemize}
-    \item The COM-Poisson (original parametrization) model;
-    \item The COM-Poisson (new parametrization) model;
-    \item Quasi-Poisson model ($\text{Var}(Y) = \sigma
-      \text{E}(Y)$); and
-    \item The standard Poisson regression model.
-    \end{itemize}
+    \item In the data analysis, we consider the COM-Poisson model in the
+      two forms (original and new parametrization) and the quasi-Poisson
+      regression model as alternative models for the standard Poisson
+      regression model.
   \end{itemize}
@@
 
-\begin{frame}{Cotton bolls data}
-\vspace{-0.2cm}
-<<desc-cotton, fig.width=9, fig.height=5.8, fig.cap="">>=
-
-knitr::include_graphics("figure/desc-cotton-1")
-
-@
-
-\end{frame}
-
 \begin{frame}{Model specification}
@@
     \item COM-Poisson$_\mu$ ($\mu_{ij}\,,\,\phi$)
-    \item Quasi-Poisson (Var$(Y_{ij})=\sigma\mu_{ij}$)
+    \item Quasi-Poisson (var$(Y_{ij})=\sigma\mu_{ij}$)
   \end{itemize}
@@
 
-\subsection{Soil moisture and potassium doses on soybean culture}
+\subsection*{Additional results}
 
-\begin{frame}{Soybean data}
-
-  \begin{columns}
-    \begin{column}{0.4\textwidth}
-      \begin{figure}
-        \includegraphics[width=4cm]{./images/soybean}
-      \end{figure}
-    \end{column}
-    \begin{column}{0.6\textwidth}
-      {\bf Aim:} evaluate the effects of potassium doses applied
-      to soil in different soil moisture levels;\\[0.2cm]
-      {\bf Design:} factorial 5 $\times$ 3 in a randomized complete
-      block design (5 blocks);\\[0.2cm]
-      {\bf Experimental unit:} a pot with a plant;\\[0.2cm]
-      {\bf Factors:}
-      \begin{itemize} \small
-        \item Potassium fertilization dose (\texttt{K}):
-        \item Soil moisture level (\texttt{umid}):
-      \end{itemize}\vspace{0.1cm}
-      {\bf Response variable:} Total number of bean seeds per pot;\\
-    \end{column}
-  \end{columns}
-
-\end{frame}
-
-\begin{frame}{Soybean data}
-\vspace{-0.2cm}
-<<desc-soy, fig.width=9, fig.height=5.8, fig.cap="">>=
-
-knitr::include_graphics("figure/desc-soy-1")
-
-@
-
-\end{frame}
-
-\begin{frame}{Model specification}
-
-  {\bf Linear predictor:} {\small based on descriptive analysis,}
-  \begin{itemize}
-    \item $\log(\mu_{ijk}) = \beta_0 + \gamma_i + \tau_j +
-      \beta_{1}\texttt{K}_k + \beta_{2}\texttt{K}_k^2 +
-      \beta_{3j}\texttt{K}_k$\\[0.1cm]
-    {\small
-    $i$ varies according the blocks;\\
-    $j$ varies in the levels of soil moisture;\\
-    $k$ varies in the levels of potassium fertilization.}
-  \end{itemize}
-
-  \vspace{0.5cm}
-
-  {\bf Alternative models:}
-  % {\bf Modelling strategies:}
-  \begin{itemize}
-    \item Poisson ($\mu_{ij}$);
-    \item COM-Poisson ($\lambda_{ij}=\eta(\mu_{ij})\,,\,\phi$)
-    \item COM-Poisson$_\mu$ ($\mu_{ij}\,,\,\phi$)
-    \item Quasi-Poisson (var$(Y_{ij})=\sigma\mu_{ij}$)
-  \end{itemize}
-
-\end{frame}
-
-\begin{frame}{Parameter estimates}
-
-\begin{table}[!ht]
-\centering \footnotesize
-\caption{Parameter estimates (Est) and ratio between estimate and
-  standard error (SE).}
-\vspace{-0.1cm}
-\label{tab:coef-soy}
-\begin{tabular}{lrrrrrrrr}
-  \toprule
-  & \multicolumn{2}{c}{Poisson} &
-    \multicolumn{2}{c}{COM-Poisson} &
-    \multicolumn{2}{c}{COM-Poisson$_\mu$} &
-    \multicolumn{2}{c}{Quasi-Poisson} \\
-\cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7} \cmidrule(lr){8-9}
-% latex table generated in R 3.4.4 by xtable 1.8-2 package
-% Wed Jun 27 14:44:34 2018
- & Est & Est/SE & Est & Est/SE & Est & Est/SE & Est & Est/SE \\
-  \midrule
-$\phi\,,\,\sigma$ &  &  & $-$0.779 & $-$4.721 & $-$0.782 & $-$4.737 & 2.615 &  \\
-  $\beta_0$ & 4.867 & 144.289 & 2.232 & 6.042 & 4.867 & 97.781 & 4.867 & 89.225 \\
-  $\gamma_{1}$ & $-$0.019 & $-$0.730 & $-$0.009 & $-$0.494 & $-$0.019 & $-$0.495 & $-$0.019 & $-$0.452 \\
-  $\gamma_{2}$ & $-$0.037 & $-$1.373 & $-$0.017 & $-$0.921 & $-$0.037 & $-$0.931 & $-$0.037 & $-$0.849 \\
-  $\gamma_{3}$ & $-$0.106 & $-$3.889 & $-$0.049 & $-$2.422 & $-$0.106 & $-$2.634 & $-$0.106 & $-$2.405 \\
-  $\gamma_{4}$ & $-$0.092 & $-$3.300 & $-$0.042 & $-$2.102 & $-$0.092 & $-$2.237 & $-$0.092 & $-$2.040 \\
-  $\tau_{1}$ & 0.132 & 3.647 & 0.061 & 2.295 & 0.132 & 2.472 & 0.132 & 2.255 \\
-  $\tau_{2}$ & 0.124 & 3.432 & 0.057 & 2.177 & 0.124 & 2.326 & 0.124 & 2.122 \\
-  $\beta_1$ & 0.616 & 11.014 & 0.284 & 4.729 & 0.616 & 7.464 & 0.616 & 6.811 \\
-  $\beta_2$ & $-$0.276 & $-$10.250 & $-$0.127 & $-$4.589 & $-$0.276 & $-$6.946 & $-$0.276 & $-$6.338 \\
-  $\beta_{31}$ & 0.146 & 4.268 & 0.067 & 2.614 & 0.146 & 2.892 & 0.146 & 2.639 \\
-  $\beta_{32}$ & 0.165 & 4.829 & 0.076 & 2.884 & 0.165 & 3.272 & 0.165 & 2.986 \\
-   \specialrule{0.01em}{0.3em}{0.3em}
- LogLik & \multicolumn{2}{c}{$-340.082$} & \multicolumn{2}{c}{$-325.241$} & \multicolumn{2}{c}{$-325.233$} & \multicolumn{2}{c}{$  -$}\\
- AIC & \multicolumn{2}{c}{$702.164$} & \multicolumn{2}{c}{$674.482$} & \multicolumn{2}{c}{$674.467$} & \multicolumn{2}{c}{$  -$}\\
- BIC & \multicolumn{2}{c}{$727.508$} & \multicolumn{2}{c}{$702.130$} & \multicolumn{2}{c}{$702.116$} & \multicolumn{2}{c}{$  -$} \\
-  \bottomrule
-\end{tabular}
-\end{table}
-\end{frame}
-
-\begin{frame}{Fitted curves}
-
-<<pred-soy, fig.height=3.8, fig.width=8, fig.cap="Dispersion diagrams of been seeds counts as function of potassium doses and humidity levels with fitted curves and confidence intervals (95\\%).">>=
-include_graphics("figure/pred-soy-1")
-@
-
-\end{frame}
-
-\subsection{Assessing toxicity of nitrofen in aquatic systems}
-
-\begin{frame}{Nitrofen data}
-
-  \begin{columns}
-    \begin{column}{0.4\textwidth}
-      \begin{figure}
-        \includegraphics[width=4cm]{./images/zooplankton}
-      \end{figure}
-    \end{column}
-    \begin{column}{0.6\textwidth}
-      {\bf Aim:} measure the reproductive toxicity of the herbicide
-      nitrofen on a species of zooplankton
-      (\textit{Ceriodaphnia dubia});\\[0.2cm]
-      {\bf Design:} completely randomized design, with 10
-      replicates;\\[0.2cm]
-      {\bf Experimental unit:} zooplankton animal;\\[0.2cm]
-      {\bf Factors:}
-      \begin{itemize} \small
-        \item herbicide nitrofen dose (\texttt{dose});
-      \end{itemize}\vspace{0.1cm}
-      {\bf Response variable:} Total number of live offspring;\\
-    \end{column}
-  \end{columns}
-
-\end{frame}
-
-\begin{frame}{Model specification}
-
-  \textbf{Linear predictors:}
-  \begin{center}
-    \begin{tabular}{ll}
-      Linear:     &
-          $\log(\mu_i) = \beta_0 + \beta_1 \texttt{dose}_i$,\\
-      Quadratic: &
-          $\log(\mu_i) = \beta_0 + \beta_1 \texttt{dose}_i +
-             \beta_2 \texttt{dose}_i^2$ and\\
-      Cubic:     &
-          $\log(\mu_i) = \beta_0 + \beta_1 \texttt{dose}_i +
-             \beta_2 \texttt{dose}_i^2 + \beta_3 \texttt{dose}_i^3.$
-    \end{tabular}
-  \end{center}
-
-  \vspace{0.5cm}
-
-  {\bf Alternative models:}
-  % {\bf Modelling strategies:}
-  \begin{itemize}
-    \item Poisson ($\mu_{ij}$);
-    \item COM-Poisson ($\lambda_{ij}=\eta(\mu_{ij})\,,\,\phi$)
-    \item COM-Poisson$_\mu$ ($\mu_{ij}\,,\,\phi$)
-    \item Quasi-Poisson (var$(Y_{ij})=\sigma\mu_{ij}$)
-  \end{itemize}
-
-\end{frame}
-
-\begin{frame}{Likelihood ratio tests}
-
-\vspace{-0.4cm}
-\begin{table}[!ht]
-\centering \footnotesize
-\caption{Model fit measures and comparisons between linear predictors.}
-\vspace{-0.2cm}
-\label{tab:anova-ovos}
-%\begin{tabularx}{\textwidth}{lCCCCCrC}
-\begin{tabular}{lcccccrr}
-\toprule
- Poisson & np & $\ell$ & AIC & 2(diff $\ell$) & diff np & P($>\rchi^2$) & \\
- \midrule
-% latex table generated in R 3.4.4 by xtable 1.8-2 package
-% Wed Jun 27 09:38:32 2018
- Linear & 2 & $-$180.667 & 365.335 &  &  &  &  \\
- Quadratic & 3 & $-$147.008 & 300.016 & 67.319 & 1 & 2.31E$-$16 &  \\
- Cubic & 4 & $-$144.090 & 296.180 & 5.835 & 1 & 1.57E$-$02 &  \\
-
-\specialrule{0em}{0.5em}{0em} %% Apenas para espaçamento
-  COM-Poisson & np & $\ell$ & AIC & 2(diff $\ell$) & diff np &
-  P($>\rchi^2$) & $\hat{\phi}$ \\
-  \midrule
-% latex table generated in R 3.4.4 by xtable 1.8-2 package
-% Wed Jun 27 09:38:32 2018
- Linear & 3 & $-$167.954 & 341.908 &  &  &  & $-$0.893 \\
-  Quadratic & 4 & $-$146.964 & 301.929 & 41.980 & 1 & 9.22E$-$11 & $-$0.059 \\
-  Cubic & 5 & $-$144.064 & 298.129 & 5.800 & 1 & 1.60E$-$02 & 0.048 \\
-
-\specialrule{0em}{0.5em}{0em} %% Apenas para espaçamento
-  COM-Poisson$_\mu$ & np & $\ell$ & AIC & 2(diff $\ell$) & diff np &
-  P($>\rchi^2$) & $\hat{\phi}$ \\
-  \midrule
-% latex table generated in R 3.4.4 by xtable 1.8-2 package
-% Wed Jun 27 09:38:32 2018
- Linear & 3 & $-$167.652 & 341.305 &  &  &  & $-$0.905 \\
-  Quadratic & 4 & $-$146.950 & 301.900 & 41.405 & 1 & 1.24E$-$10 & $-$0.069 \\
-  Cubic & 5 & $-$144.064 & 298.127 & 5.773 & 1 & 1.63E$-$02 & 0.047 \\
-
-\specialrule{0em}{0.5em}{0em} %% Apenas para espaçamento
-  Quasi-Poisson & np & QDev & AIC & F & diff np & P($>F$) & $\hat{\sigma}$ \\
-  \midrule
-% latex table generated in R 3.4.4 by xtable 1.8-2 package
-% Wed Jun 27 09:38:32 2018
- Linear  & 3 & 123.929 &  &  &  &  & 2.262 \\
-  Quadratic & 4 & 56.610 &  & 60.840 & 1 & 5.07E$-$10 & 1.106 \\
-  Cubic & 5 & 50.774 &  & 5.659 & 1 & 2.16E$-$02 & 1.031 \\
-
- \bottomrule
-%\end{tabularx}
-\end{tabular}
-\end{table}
-
-\end{frame}
-
-\begin{frame}{Parameter estimates and fitted values}
-
-  \vspace{-0.4cm}
-\begin{table}[!ht]
-\centering \footnotesize
-\caption{Parameter estimates (Est) and ratio between estimate and
-  standard error (SE).}
-\vspace{-0.1cm}
-\label{tab:coef-ovos}
-\begin{tabular}{lrrrrrrrr}
-  \toprule
-  & \multicolumn{2}{c}{Poisson} &
-    \multicolumn{2}{c}{COM-Poisson} &
-    \multicolumn{2}{c}{COM-Poisson$_\mu$} &
-    \multicolumn{2}{c}{Quasi-Poisson} \\
-\cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7} \cmidrule(lr){8-9}
-% latex table generated in R 3.4.4 by xtable 1.8-2 package
-% Wed Jun 27 09:38:32 2018
- & Est & Est/SE & Est & Est/SE & Est & Est/SE & Est & Est/SE \\
-  \midrule
-$\beta_{0}$ & 3.477 & 62.817 & 3.649 & 4.850 & 3.477 & 64.308 & 3.477 & 61.860 \\
-  $\beta_{1}$ & $-$0.086 & $-$0.433 & $-$0.091 & $-$0.448 & $-$0.088 & $-$0.452 & $-$0.086 & $-$0.426 \\
-  $\beta_{2}$ & 0.153 & 0.863 & 0.161 & 0.878 & 0.155 & 0.894 & 0.153 & 0.850 \\
-  $\beta_{3}$ & $-$0.097 & $-$2.398 & $-$0.102 & $-$2.229 & $-$0.098 & $-$2.464 & $-$0.097 & $-$2.361 \\
-   \bottomrule
-\end{tabular}
-\end{table}
-  \vspace{-0.4cm}
-
-<<pred-ovos, fig.height=3.5, fig.width=5.5, out.width="0.8\\textwidth", fig.cap="Number of live offsprings observed for each nitrofen concentration level with fitted curves and 95\\% confidence intervals.">>=
-include_graphics("figure/pred-ovos-1")
-@
-
-\end{frame}
-
-\subsection*{}
-
-\begin{frame}{Additional results}
+\begin{frame}{Computational times and orthogonality property}
 
@@
 <<comp-times, fig.height=4, fig.width=8, fig.cap="Computational times to fit the models under original and reparametrized versions based on the fifty repetitions.">>=
-
-library(latticeExtra)
-source("../../codes/lattice-panels.R")
-
-benchs <- readRDS("../../codes/comparetimes.rds")
-xlabs <- expression("COM-Poisson", "COM-Poisson"[mu])
-bwplot(time ~ expr | case,
-       ylab = "Time (seconds)",
-       scales = list(
-           y = list(relation = "free"),
-           x = list(at = 1:2, labels = xlabs)
-       ),
-       data = benchs)
-
+include_graphics("figure/comp-times-1")
 @

### 2018-09-13 article.Rnw
Make the changes suggested by John Hinde

@@
 \item School of Mathematics, Statistics and Applied Mathematics,
-  National University of Ireland,
-  Galway,
+  National University of Ireland Galway,
   Galway, Ireland
@@
 %-------------------------------------------
-% Title and shor title
+% Title and short title
 \Title{Reparametrization of COM-Poisson Regression Models with
@@
 \Abstract{The COM-Poisson distribution is a two-parameter generalization
-  of Poisson distribution can deal with under-, equi- and overdispersed
-  count data. Its location parameter does not correspond to the
-  expectation, which complicates the interpretation of regression models
-  specified using this distribution. In this paper, we propose a
-  straightforward reparametrization of the COM-Poisson distribution
-  based on an approximation to the expectation of this distribution. The
-  main advantage of our new parametrization is the interpretation of the
-  regression coefficients in terms of the expectation of the count
-  response variable, as usual in the context of generalized linear
-  models. Furthermore, the estimation and inference for the
-  reparametrized COM-Poisson regression models can be done based on the
-  likelihood paradigm. We carried out simulation studies to verify the
-  finite sample properties of the maximum likelihood estimators. The
-  results from our simulation study show that the maximum likelihood
-  estimators are unbiased and consistent for both regression and
-  dispersion parameters. The empirical correlation between the
-  regression and dispersion parameter estimators and deviance surfaces
-  suggests that these parameters are orthogonal. We illustrate the
-  application of the proposed model through the analysis of three data
-  sets with over-, under- and equidispersed count data. The study of
-  distribution properties through a consideration of dispersion,
-  zero-inflated and heavy tail indexes, together with the results of
-  data analysis show the flexibility over standard approaches.  The
-  computational routines for fitting the original and new version of the
-  COM-Poisson regression model and the analyzed data sets are available
-  in the supplementary material.  }
+  of the Poisson distribution that can deal with under-, equi- and
+  overdispersed count data. Unfortunately, its location parameter does
+  not correspond to the expectation, which complicates the
+  interpretation of regression models specified using this
+  distribution. In this paper, we propose a straightforward
+  reparametrization of the COM-Poisson distribution based on an
+  approximation to the expectation. The main advantage of our new
+  parametrization is the interpretation of the regression coefficients
+  in terms of the (approximate) expectation of the count response
+  variable, as is usual in the context of generalized linear
+  models. Estimation and inference for the reparametrized COM-Poisson
+  regression models can be done using the likelihood paradigm. We
+  carried out simulation studies to verify the finite sample properties
+  of the maximum likelihood estimators. The results from our simulation
+  study show that the maximum likelihood estimators are unbiased and
+  consistent for both regression and dispersion parameters. The
+  empirical correlation between the regression and dispersion parameter
+  estimators and nature of the deviance surfaces suggests that these
+  parameters are also approximately orthogonal, which is advantageous
+  for interpretation, inference, and computational efficiency. We
+  illustrate the application of the proposed model through the analysis
+  of three data sets with over-, under- and equidispersed count
+  data. The study of the distribution's properties, through a
+  consideration of dispersion, zero-inflation and heavy tail indexes,
+  together with the results of data analysis, show the flexibility over
+  standard approaches.  The computational routines for fitting the
+  original and reparameterized versions of the COM-Poisson regression
+  model and data sets are available in the supplementary material.  }
 
@@
 
-Count data are random variables that assume non-negative integer values
-and represent the number of times an event occurs in the observation
-period. This kind of data is common in crop sciences, such as the number
-of grains produced by a plant, number of fruits produced by a tree,
-number of insects captured by a trap, to cite but a few.  Since the
-seminal paper of \citet{Nelder1972} where the class of the generalized
-linear models (GLMs) was introduced, the analysis of count data often
-employs the Poisson regression model. This model provides a suitable
-strategy for the analysis of count data and an efficient Newton scoring
-algorithm can be used for fitting the model.
+Count data arise from random variables that take non-negative integer
+values and typically represent the number of times an event occurs in an
+observation period or region. This kind of data is also common in crop
+sciences, examples including the number of grains produced by a plant,
+the number of fruits on a tree, the number of insects captured by a
+trap, etc.  Since the seminal paper of \citet{Nelder1972}, where the
+class of the generalized linear models (GLMs) was introduced, the
+analysis of count data often uses a Poisson regression model. This model
+provides a suitable strategy for the analysis of count data and an
+efficient Newton scoring algorithm that can be used for fitting.
 
 In spite of the advantages of the Poisson regression model, the Poisson
-distribution has only one parameter, which represents both the
-expectation and variance of the count random variable.  This restriction
-on the relationship between the expectation and variance induced by the
-Poisson distribution is referred as equidispersion.  However, in
-practical data analysis such restriction can be unsuitable, since the
-observed data can present  variance both smaller or larger than the
-expectation, leading to the cases of under- and overdispersion,
-respectively.  The main problem of the application of the Poisson
-regression model to non-equidispersed count data is that the standard
-errors associated with the regression coefficients are inconsistently
-estimated, which in turn can lead to misleading
-inferences~\citep{Winkelmann1995, Bonat2017}.
-
-In practice, overdispersion is largely reported in the literature and
-may occur due to the absence of relevant covariates, heterogeneity of
-sampling units, different observational periods/regions not considered
-in the analysis, and excess of zeros~\citep{Hinde1998}.  The case of
-underdispersion is less report in the literature, however, it has been
-of increasing interest in the statistical community. The processes that
-reduce the variability are not as well-known as those leading to extra
-variability.  For this reason, there are few approaches to deal with
-underdispersed count data. The explanatory mechanisms leading to
-underdispersion may be related to the underlying stochastic process
-generating the count data.  When the time between events is not
-exponentially distributed, the number of events can be over or
-underdispersed, this process motivated the class of duration dependence
-models~\citep{Winkelmann1995}.  Another possible explanation of
-underdispersion is when the responses correspond to order statistics of
-component observations, such as maxima of Poisson distributed
-counts~\citep{Steutel1989}.
-
-The strategies for constructing alternative count distributions are
-related with the causes of the non-equidispersion. Specifically for the
-overdispersion, Poisson mixture models are widely applied.  One
-popular example of this approach is the negative-binomial model, where
-the expectation of the Poisson distribution is assumed to be gamma
-distributed.  However, other distributions can be used to represent the
-random variation.  For example the Poisson-Tweedie
-model~\citep{Bonat2017} and its special cases as the Poisson
-inverse-Gaussian and Neyman-Type A assume that the random effects are
-Tweedie, inverse Gaussian and Poisson distributed, respectively.  The
+distribution has only one parameter that represents both the expectation
+and variance of the count random variable.  This equality restriction on
+the relationship between the expectation and variance of the Poisson
+distribution is referred as equidispersion.  However, in practical data
+analysis such restriction can be unsuitable, since the observed data can
+present variance both smaller or larger than the mean, leading to under-
+and overdispersion, respectively.  While applying the Poisson regression
+model to non-equidispersed count data can give consistent parameter
+estimates, the associated standard errors are inconsistently estimated,
+which in turn can lead to misleading inferences~\citep{Winkelmann1995,
+  Bonat2017}.
+
+In practice, overdispersion is widely reported in the literature and may
+occur due to the absence of relevant covariates, heterogeneity of
+sampling units, different observational periods/regions not being
+considered in the analysis, and excess zeros~\citep{Hinde1998}.  The
+case of underdispersion is less often reported in the literature,
+however, it has been of increasing interest in the statistical
+community. The processes that reduce the variability are not as
+well-known as those leading to extra variability.  For this reason,
+there are few models to deal with underdispersed count data. Possible
+explanatory mechanisms leading to underdispersion may be related to the
+underlying stochastic process generating the count data.  For example,
+when the time between events is not exponentially distributed, the
+number of events can be over or underdispersed; a process that motivated
+the class of duration dependence models~\citep{Winkelmann1995}.  Another
+possible explanation for underdispersion is when the responses
+correspond to order statistics of component observations, such as maxima
+of Poisson distributed counts~\citep{Steutel1989}.
+
+Strategies for constructing alternative count distributions are related
+to the causes of the non-equidispersion. Specifically for
+overdispersion, Poisson mixture (compound) models are widely applied.
+One popular example of this approach is the negative-binomial model,
+where the expectation of the Poisson distribution is assumed to be gamma
+distributed.  However, other distributions can also be used to represent
+this additional random variation.  For example the Poisson-Tweedie
+model~\citep{Bonat2017} and its special cases the Poisson
+inverse-Gaussian and Neyman-Type A models assume that the random effects
+are Tweedie, inverse Gaussian or Poisson distributed, respectively.  The
 Gamma-Count distribution assumes a gamma distribution for the time
-between events, thus it can deal with underdispersed as well as
-overdispersed count data~\citep{Zeviani2014}. The COM-Poisson
-distribution is obtained by a generalization of the Poisson distribution
-allowing for a non-linear decrease in the ratios of successive
-probabilities~\citep{Shmueli2005}.
-
-The COM-Poisson distribution has the Poisson and geometric distributions
-as special cases, as well as the Bernoulli distribution as a limiting
-case. It can deal with both under- and overdispersed count data.  Some
-recently applications of the COM-Poisson distribution include
+between events and it can deal with underdispersed as well as
+overdispersed count data~\citep{Zeviani2014}. The subject of this paper,
+the COM-Poisson distribution, is obtained by a generalization of the
+Poisson distribution that allows for a non-linear decrease in the ratios
+of successive probabilities~\citep{Shmueli2005}.
+
+The COM-Poisson distribution includes the Poisson and geometric
+distributions as special cases, as well as the Bernoulli distribution as
+a limiting case. It can deal with both under- and overdispersed count
+data.  Some recent applications of the COM-Poisson distribution include
 \citet{Lord2010} for the analysis of traffic crash data,
 \citet{Sellers2010} for the modelling of airfreight breakage and book
-purchases, \citet{Huang2017} to the analysis of attendance data,
-takeover birds and cotton boll counts, and \citet{Chatla2018} to model
+purchases, \citet{Huang2017} on the analysis of attendance data,
+takeover bids and cotton boll counts, and \citet{Chatla2018} to model
 counts from bike sharing systems. Theoretical results and approximations
-derived from this distribution is discussed by \citet{Shmueli2005},
+derived for this distribution are discussed by \citet{Shmueli2005},
 \citet{Daly2016} and \citet{Gaunt2017}. The main disadvantage of the
@@
 its location parameter does not correspond to the expectation of the
-distribution, which complicates the interpretation of regression models
-and means that they are not comparable with standard approaches such as
-the Poisson and negative binomial regression models. \citet{Huang2017}
-proposed a mean-parametrization of the COM-Poisson distribution in order
-to avoid such an issue.  In this approach the mean parameter is obtained
-by solving an non-linear equation defined as an infinite sum.
-Consequently, it is computationally demanding and liable to numerical
+distribution. This complicates the interpretation of regression models
+and means that they are not comparable with standard approaches, such as
+the Poisson and negative binomial regression models. In order to avoid
+this issue, \citet{Huang2017} proposed a mean-parametrization of the
+COM-Poisson distribution.  In his approach the mean parameter is
+obtained as the solution of a non-linear equation defined as an infinite
+sum, which is computationally demanding and liable to numerical
 problems.
@@
 function is written in terms of $\mu$ and $\phi$, where $\mu$ is the
-approximated expectation and $\phi$ is a dispersion parameter. In
+approximate expectation and $\phi$ is a dispersion parameter. In
 contrast to the original parametrization, the proposed parametrization
-leads to regression coefficients directly associated with the
-expectation of the response variable, as usual in the context of
-generalized linear models.  Consequently, the obtained regression
-coefficients are comparable with the ones obtained by standard
+leads to regression coefficients directly associated (approximately)
+with the expectation of the response variable, as is usual in the
+context of generalized linear models.  Consequently, the resulting
+regression coefficients are comparable to those obtained by standard
 approaches, such as the Poisson and negative binomial regression models.
-Furthermore, our novel COM-Poisson parametrization is simpler than the
-strategy proposed by~\cite{Huang2017}, since it does not require any
-numerical method for solving non-linear equations, and we show the
-attractive properties like the orthogonality between dispersion and
-regression parameters and consistency and asymptotic normality of the
-maximum likelihood estimators are retained.
+Furthermore, our novel COM-Poisson parametrization is computationally
+simpler than the strategy proposed by~\citet{Huang2017}, since it does
+not require any numerical method for solving non-linear equations. Also,
+we show that attractive properties like the orthogonality between
+dispersion and regression parameters and consistency and asymptotic
+normality of the maximum likelihood estimators are retained.
 
 This paper is organized as follows. In \autoref{background} we present
-the COM-Poisson distribution and the strategy proposed
-by~\cite{Huang2017}.  The proposed reparametrization, assessment of
-moment approximations, and study of distribution properties are
-considered in the~\autoref{reparametrization}. In
-the~\autoref{estimation-and-inference} we present estimation and
-inference for the novel COM-Poisson regression model based on the
-likelihood paradigm. The properties of the maximum likelihood estimators
-and the orthogonality property are assessed
-in~\autoref{simulation-study} through simulation studies.  We illustrate
-the application of the new COM-Poisson regression model through the
-analysis of three data sets. We provide an \texttt{R} implementation of
-the COM-Poisson and reparameterized COM-Poisson regression models as
-well as the analyzed data sets in the supplementary
-material.\footnote{Available on
+the COM-Poisson distribution and the approach proposed
+by~\citet{Huang2017}.  The newly proposed reparametrization is
+considered in ~\autoref{reparametrization}, along with assessment of the
+moment approximation, and a study of distribution's
+properties. In~\autoref{estimation-and-inference} we present estimation
+and inference for the novel COM-Poisson regression model in a likelihood
+framework. The properties of the maximum likelihood estimators and their
+approximate orthogonality are assessed in~\autoref{simulation-study}
+through simulation studies.  We illustrate the application of the new
+COM-Poisson regression model with the analysis of three data sets. We
+provide an \texttt{R} implementation for the COM-Poisson and
+reparameterized COM-Poisson regression models, together with the
+analyzed data sets, in the supplementary material.\footnote{Available on
   \texttt{\url{http://www.leg.ufpr.br/~eduardojr/papercompanions}}
@@
 Poisson distribution this ratio is $\frac{y}{\lambda}$ corresponding to
-$\nu=1$.  It allows the COM-Poisson distribution deals with
+$\nu=1$.  This allows the COM-Poisson distribution to deal with
 non-equidispersed count data.  The probability mass function of the
@@
 \end{equation}
-where $\lambda > 0$, $\nu \geq 0$ and $Z(\lambda, \nu) =
-\sum_{j=0}^\infty \frac{\lambda^j}{(j!)^\nu}$ is a normalizing
-constant that depends on both parameters.
+where $\lambda > 0$, $\nu \geq 0$ and
+$Z(\lambda, \nu) = \sum_{j=0}^\infty \frac{\lambda^j}{(j!)^\nu}$ is a
+normalizing constant that depends on both parameters.
 
 The $Z(\lambda, \nu)$ series diverges theoretically only when $\nu=0$
-and $\lambda \geq 1$, but numerically for small values of $\nu$ combined
-with large values of $\lambda$, the sum is so huge it causes
-overflow. \autoref{tab:convergenceZ} shows the values of the normalizing
-constants using one thousand increments, that is,
-$\sum_{j=0}^{1000}\lambda^j/(j!)^\nu$ for different values of $\lambda$
-and $\phi$.
+and $\lambda \geq 1$, but numerically, for small values of $\nu$
+combined with large values of $\lambda$, the sum is so huge it results
+in overflow. \autoref{tab:convergenceZ} shows the values of the
+normalizing constant based on one thousand terms in the summation, that
+is $\sum_{j=0}^{1000}\lambda^j/(j!)^\nu$, for different values of
+$\lambda$ and $\phi$.
 
@@
 \centering
-\caption{Values for $Z(\lambda, \nu)$ constant (numerically computed) for values of $\lambda$ (0.5 to 50) and $\phi$ (0 to 1)}
+\caption{Values for $Z(\lambda, \nu)$ normalizing constant (computed numerically with 1000 terms in the summation) for values of $\lambda$ (0.5 to 50) and $\phi$ (0 to 1)}
 \label{tab:convergenceZ}
@@
 \footnotesize \raggedright
-divergent$^{*}$ is a mathematically divergent series; and
+divergent$^{*}$ denotes a mathematically divergent series; and
 divergent$^{**}$ a numerically divergent series.
@@
 to the computational storage limitation.  For both forms of divergence
-it is impossible to compute probabilities, therefore, this acts as a
-restriction on the parameter space.
+it is impossible to compute COM-Poisson probabilities, therefore, this
+places a restriction on the feasible parameter space.
 
@@
   \textrm{Var}(Y) \approx \frac{\lambda^{1/\nu}}{\nu}\,,
-\end{equation} which is particularly accurate for  $\nu \leq 1$ or
+\end{equation}with greatest accuracy for  $\nu \leq 1$ or
 $\lambda > 10$. The authors also argue that the mean-variance
-relationship can be approximate by
-$\frac{1}{\nu}\text{E}(Y)$. In~\autoref{reparametrization}, we assess
-the accuracy of these approximations.
+relationship can be approximated as
+$\textrm{Var}(Y) \approx\frac{1}{\nu}\text{E}(Y)$.
+In~\autoref{reparametrization}, we assess the accuracy of these
+approximations.
 
 The COM-Poisson regression model was proposed by~\citet{Sellers2010},
-using the original parametrization. In this case, the COM-Poisson
-regression model is $\log(\lambda_i) = \bm{x}_i^\top \bm{\beta}$ and the
-relationship between E$(Y_i)$ and $\bm{x}_i$ is modelled indirectly.
+using this original parametrization. Specifically, the COM-Poisson
+regression model is defined as
+$\log(\lambda_i) = \bm{x}_i^\top \bm{\beta}$ and the relationship
+between E$(Y_i)$ and $\bm{x}_i$ is modelled indirectly.
 \citet{Huang2017} shows how to model directly the expectation of the
-COM-Poisson distribution in a suitable parametrization. In the
-\autoref{eqn:pmf-cmp}, \Citeauthor{Huang2017} proposes that the
-parameter $\lambda$ as a function of $\mu$ and $\nu$, is given by the
-solution to
+COM-Poisson distribution in a suitable reparametrization. From
+\autoref{eqn:pmf-cmp}, \Citeauthor{Huang2017} notes that the parameter
+$\lambda$ is given, as a function of $\mu$ and $\nu$, by the solution to
 \begin{equation*}
@@
 \end{equation*}
-Thus the mean-parametrized COM-Poisson regression model is $\log(\mu_i)
-= \bm{x}_i^\top \bm{\beta}$. In this article, we propose an alternative
-mean-parametrization of the COM-Poisson distribution in order to avoid the
-limitations of the original parametrization and the numerical complexity
-of the \Citeauthor{Huang2017}'s approach.
+This allows \Citeauthor{Huang2017} to define a mean-parametrized
+COM-Poisson regression model with
+$\log(\mu_i) = \bm{x}_i^\top \bm{\beta}$ giving a direct relationship
+between $\mu$ and the covariates $\bm{x}$. In this article, we propose
+an alternative approximate mean-parametrization of the COM-Poisson
+distribution in order to avoid the limitations of the original
+parametrization and the numerical complexity of the
+\Citeauthor{Huang2017}'s approach.
 
@@
 
-The proposed reparametrization of COM-Poisson models is based on
-the mean approximation (\autoref{eqn:mean-approx}). We introduced a new
+The proposed reparametrization of COM-Poisson models is based on the
+mean approximation (\autoref{eqn:mean-approx}). We introduce a new
 parameter $\mu$, using this approximation,
@@
 convenience, thus $\phi = \log(\nu)$, $\phi \in \mathbb{R}$. The
-interpretation of $\phi$ is the same as the $\nu$, but on another
+interpretation of $\phi$ is the same as the $\nu$, but simply on another
 scale. For $\phi < 0$ and $\phi > 0$ we have the overdispersed and
-underdispersed cases, respectively.  When $\phi=0$ we have Poisson
+underdispersed cases, respectively.  When $\phi=0$ we have the Poisson
 distribution as a special case.
@@
 (\autoref{eqn:mean-approx}), \autoref{fig:approx-plot} presents the
-quadratic errors for (a) expectation and (b) variance.  The quadratic
-errors were obtained by $[\mu - \text{E}(Y)]^2$ for the expectation and
-by $[ \mu \exp(-\phi) - \text{Var}(Y)]^2$ for the variance. In both
-cases $\text{E}(Y)$ and Var$(Y)$ were computed numerically\footnote{We
-  compute E$(Y) = \sum_{y=0}^{500}yp(y)$ and
-  E$(Y^2) = \sum_{y=0}^{500}y^2p(y)$.}. The dotted lines represent the
-border between the regions $\nu \leq 1$ and $\lambda > 10^\nu$, in the
-$\mu$ and $\phi$ scale.
-
-<<approx-plot, fig.height=3, fig.width=6, out.width=".8\\textwidth", fig.cap="Quadratic errors for the approximation of the (a) expectation and (b) variance.
+squared errors for (a) expectation and (b) variance.  The squared errors
+are $[\mu - \text{E}(Y)]^2$ for the expectation and
+$[ \mu \exp(-\phi) - \text{Var}(Y)]^2$ for the variance. In both cases
+$\text{E}(Y)$ and Var$(Y)$ were computed numerically using 500
+terms\footnote{ E$(Y) \approx \sum_{y=0}^{500}yp(y)$ and
+  E$(Y^2) \approx \sum_{y=0}^{500}y^2p(y)$.}. The dotted lines represent
+the border between the regions $\nu \leq 1$ and $\lambda > 10^\nu$, in
+the $\mu$ and $\phi$ scale.
+
+<<approx-plot, fig.height=3, fig.width=6, out.width=".8\\textwidth", fig.cap="Squared errors for the approximation of the (a) expectation and (b) variance.
 Dotted lines represent the restriction for suitable approximations given by \\cite{Shmueli2005}.">>=
@@
 The results in \autoref{fig:approx-plot} show that the mean
-approximation is accurate, the largest quadratic error is
-\Sexpr{max(grid[["emu"]])} for the parameter values evaluated. For the
-variance approximation, the largest quadratic error was
-\Sexpr{max(grid[["eva"]])} and it occurs for negative values of
-$\phi$. Interestingly, the errors are larger for negative values of
-$\phi$ and present no clear relation with $\mu$, as opposed to the
-regions gives by \citet{Shmueli2005} ($\phi \leq 0$ and
+approximation is accurate, the largest squared error, for the parameter
+values evaluated, is $0.038$.  For the variance approximation, the
+largest squared error was $XXX$ and it occurs for negative values of
+$\phi$. Interestingly, the errors for the variance are larger for
+negative values of $\phi$ and present no clear relation with $\mu$, as
+opposed to the regions gives by \citet{Shmueli2005} ($\phi \leq 0$ and
 $\mu > 10 - \frac{\exp(\phi) - 1}{2\exp(\phi)}$).
@@
 The results presented in \autoref{fig:approx-plot}(a) support the
-proposed reparametrization. Replacing $\lambda$ and $\nu$ as function of
+proposed reparametrization. Replacing $\lambda$ and $\nu$ as functions of
 $\mu$ and $\phi$ in \autoref{eqn:pmf-cmp}, the reparametrized
@@
 \end{equation}
-where $\mu > 0$. We denote this distribution as
+where $\mu > 0$. We denote this reparameterized distribution as
 COM-Poisson$_\mu$. In \autoref{fig:pmf-cmp}, we show the shapes of
-COM-Poisson$_\mu$ distribution.
+the COM-Poisson$_\mu$ distribution.
 
@@
 restricted region is related to small underdispersed counts (averages
-smaller than 0.1 and $\nu>1$) a parameter region unexplored in practice.
+smaller than 0.1 and $\nu>1$) a parameter region unlikely to be of
+interest in practice.
 
@@
 $\text{ZI} > 0$, zero-deflation for $\text{ZI} < 0$ and no excess of
-zeros for $\text{ZI} = 0$. Finally, heavy-tail index indicates a
+zeros for $\text{ZI} = 0$. Finally, the heavy-tail index indicates a
 heavy-tail distribution for $\text{HT} \to 1$ when $y \to \infty$.
-
 These indexes are discussed by \citet{Bonat2017} to study the
 flexibility of Poisson-Tweedie distribution, and \citet{Puig2006} to
-describe count distributions. Regarding the COM-Poisson$_\mu$
-distribution, in \autoref{fig:indexes-plot} we present the relationship
-between (a) mean and variance, (b--c) the dispersion and zero-inflation
-indexes for different values of $\mu$ and $\phi$, and (d) heavy-tail
-index for $\mu=25$ and different values of $y$ and $\phi$.
+describe count distributions in general.
 
@@
 
-<<indexes-plot, fig.width=9.5, fig.height=4, fig.cap="Indexes for COM-Poisson distribution.
+<<indexes-plot, fig.width=9.5, fig.height=4, fig.cap="Indexes for the COM-Poisson distribution.
 (a) Mean and variance relationship, (b--d) dispersion, zero-inflation and heavy-tail indexes for different parameter values.
-Dotted lines represents the Poisson special case.">>=
-@
-
+Dotted lines represents the special case of the Poisson distribution.">>=
+@
+
+Regarding the COM-Poisson$_\mu$ distribution, in
+\autoref{fig:indexes-plot} we present the relationship between (a) mean
+and variance, (b--c) the dispersion and zero-inflation indexes for
+different values of $\mu$ and $\phi$, and (d) the heavy-tail index for
+$\mu=25$ and different values of $y$ and $\phi$.
 \autoref{fig:indexes-plot} shows that the indexes are slightly dependent
@@
 distribution is suitable to deal to dispersed counts, of course. For the
-parameter values evaluated the largest DI was \Sexpr{madi} and smallest
-was \Sexpr{midi}. \autoref{fig:indexes-plot}(c) shows the COM-Poisson
-can handle a limited amount of zero-inflation, in cases of
-overdispersion ($\phi < 0$). On the other hand, for $\phi > 0$
-(underdispersion) this distribution is suitable to deal with
-zero-deflated counts. Heavy-tail indexes in
-\autoref{fig:indexes-plot}(d) indicate the distribution is in general a
-light-tailed distribution, i.e. $\text{HT} \to 0$ for $y \to \infty$.
+parameter values evaluated the largest DI was $XXX$ and smallest was
+$XXX$. \autoref{fig:indexes-plot}(c) shows the COM-Poisson can handle
+a limited amount of zero-inflation, in cases of overdispersion
+($\phi < 0$). On the other hand, for $\phi > 0$ (underdispersion) this
+distribution is suitable to deal with zero-deflated counts. Heavy-tail
+indexes in \autoref{fig:indexes-plot}(d) indicate the distribution is in
+general a light-tailed distribution, i.e. $\text{HT} \to 0$ for
+$y \to \infty$.
 
@@
 
-In this section we describe the estimation and inference for the
-two forms of the COM-Poisson regression model based on the maximum
-likelihood method. The log-likelihood function for a set of independent
-observations $y_i$, $i=1,2,\ldots,n$ from the COM-Poisson$_\mu$
-distribution has the following form,
+In this section we describe estimation and inference for the two forms
+of the COM-Poisson regression model based on the maximum likelihood
+method. Inference can be done using the standard machinery of likelihood
+inference, including likelihood ratio tests for model comparison and
+Wald-tests for testing individual (or groups of) parameters. The
+log-likelihood function for a set of independent observations $y_i$,
+$i=1,2,\ldots,n$ from the COM-Poisson$_\mu$ distribution has the
+following form,
 \begin{equation}
@@
 \autoref{eqn:ll-rcmp}. Since the derivatives of $\ell$ cannot be
-obtained in closed forms, we obtain the Hessian matrix by central finite
-differences. We adopted the \texttt{BFGS} algorithm~\citep{Nocedal1995}
-as implemented in the function \texttt{optim()} for the statistical
-software \texttt{R} \citep{Rcore2017}. Standard errors for the
-regression coefficients are obtained based on the observed information
-matrix $\mathcal{I}(\bm{\theta})$, where
+obtained in closed forms, we use the \texttt{BFGS}
+algorithm~\citep{Nocedal1995} as implemented in the function
+\texttt{optim()} in \texttt{R} \citep{Rcore2017}. Standard errors for
+the regression coefficients are obtained based on the observed
+information matrix $\mathcal{I}(\bm{\theta})$, where
 $\mathcal{I}(\bm{\theta}) = -\mathcal{H}(\bm{\theta})$ (Hessian matrix)
-is computed numerically. Confidence intervals for $\hat{\mu}_i$ are
-obtained by using the delta method~\citep[p. 89]{Pawitan2001}.
-
-The parameter estimation for the COM-Poisson regression model in the
-original parametrization is analogous to the one presented for the
-COM-Poisson$_\mu$ distribution, however, it considers
-\autoref{eqn:ll-rcmp} in terms of $\lambda$. Even for the standard
+is computed numerically by central finite differences. Standard errors
+for $\hat{\mu}_i$ and hence confidence intervals are obtained by using
+the delta method~\citep[p. 89]{Pawitan2001}.
+
+Parameter estimation for the COM-Poisson regression model in the
+original parametrization is analogous to that presented for the
+COM-Poisson$_\mu$ distribution, however, it uses \autoref{eqn:ll-rcmp}
+re-expressed in terms of $\lambda$. Here, even for the standard
 COM-Poisson distribution, the dispersion parameter is taken on the log
@@
 
-In the applications we fitted the quasi-Poisson model
-\citep{Wedderburn1974} as a baseline model. This approach is based on a
-second-moment assumption that allows more flexibility to the model. In
-this case the variance of the response variable is fixed by an
-additional parameter $\sigma$, $\textrm{Var}(Y_i)=\sigma \mu_i$. These
-models are fitted in the \texttt{R} software using the function
-\texttt{glm(..., family = quasipoisson)}.
+For the applications we also fitted the quasi-Poisson model
+\citep{Wedderburn1974} as a baseline model. This approach is based only
+on a second-moment assumption and without specific underlying probablity
+model is less restrictive. In this model the variance of the response
+variable is specified by an additional dispersion parameter $\sigma$,
+with $\textrm{Var}(Y_i)=\sigma \mu_i$. These models are fitted in
+\texttt{R} using the function \texttt{glm(..., family = quasipoisson)}.
 
@@
 
-In this section we performed a simulation study to assess the properties
-of the maximum likelihood estimators and orthogonality of the
-reparametrized model as well as the flexibility of the COM-Poisson
-regression model to deal with non-equidispersed count data.
-
-We considered average counts varying from $3$ to $27$ according to a
+In this section we report a simulation study to assess the properties of
+the maximum likelihood estimators, the approximate parameter
+orthogonality of the reparametrized model, as well as the flexibility of
+the COM-Poisson regression model to deal with non-equidispersed count
+data.
+
+We considered average counts varying from $3$ to $27$ arising from a
 regression model with a continuous and a categorical covariate. The
-continuous covariate~($\bm{x}_1$) was generated as a sequence from $0$
-to $1$ and of length equal to the sample size.  Similarly, the
-categorical covariate~($\bm{x}_2$) was generated as a sequence of three
-values each one repeated $n/3$ times (rounding up when required), where
-$n$ denotes the sample size. The parameter $\mu$ of the reparametrized
-COM-Poisson random variable is given by
+continuous covariate~($\bm{x}_1$) was generated as a linearly increasing
+sequence from $0$ to $1$ with length equal to the sample size.
+Similarly, the categorical covariate~($\bm{x}_2$) was generated as a
+sequence of three values each one repeated $n/3$ times (rounding up when
+required), where $n$ denotes the sample size. The parameter $\mu$ of the
+reparametrized COM-Poisson random variable is given by
 $\bm{\mu} = \exp(\beta_0 + \beta_1 \bm{x}_1 + \beta_{21} \bm{x}_{21} +
 \beta_{22} \bm{x}_{22})$,
-where $\bm{x}_{21}$ and $\bm{x}_{22}$ are dummy representing the levels of
-$\bm{x}_2$.  The regression coefficients were fixed at the values,
+where $\bm{x}_{21}$ and $\bm{x}_{22}$ are dummy representing the levels
+of $\bm{x}_2$.  The regression coefficients were fixed at the values,
 $\beta_0 = 2$, $\beta_1 = 0.5$, $\beta_{21} = 0.8$ and
@@
 study.  These configurations allow us to assess the properties of the
-maximum likelihood estimators in extreme situations, such as high counts
-and low dispersion, and low counts and high dispersion, but also in the
-standard case of equidispersion.
+maximum likelihood estimators not only in extreme situations, such as
+high counts and low dispersion, and low counts and high dispersion, but
+also in the standard case of equidispersion.
 
@@
 data sets in each case. For sample sizes 50 and 100, we have 29 and 8
-simulations of the 1000 where algorithm did not converge. These non
-convergence situations occurred for $\phi=-1.6$. In
+simulations of the 1000 where the fitting algorithm did not
+converge. These non-convergence situations occurred for $\phi=-1.6$. In
 \autoref{fig:bias-plot}, we show the bias of the estimators for each
@@
 parameter and samples sizes) along with the confidence intervals
-calculated as average bias plus and minus $\Phi(0.975)$ times the
-average standard error. The scales are standardized for each parameter
-by dividing the average bias by the average standard error obtained for
-the sample of size $50$.
+calculated as average bias plus and minus $1.96$ times the average
+standard error. The scales are standardized for each parameter by
+dividing the average bias by the average standard error obtained for the
+sample of size $50$.
 
@@
 levels, both the average bias and standard errors tend to $0$ as the
-sample size increases. The estimators for the regression parameters are
-unbiased, consistent and their empirical distributions are
+sample size increases. Thus the estimators for the regression parameters
+are unbiased, consistent and their empirical distributions are
 symmetric. For the dispersion parameter, the estimator is asymptotically
@@
 become closer to the nominal level for large samples.  The worst
-scenario is when we have small sample size and strong overdispersed
+scenario is when we have a small sample size and strong overdispersed
 counts.
@@
 Cov$(\hat{\bm{\theta}}) = \mathcal{I}^{-1}(\bm{\theta})$.
-\autoref{fig:ortho-plot} shows the covariance between regression and
-dispersion parameter estimators for each simulation scenario, on the
-correlation scale. The correlations are close to zero in all cases
-suggesting the orthogonality property for the reparametrized
+\autoref{fig:ortho-plot} shows the correlation between the regression
+and dispersion parameter estimators for each simulation scenario, on the
+correlation scale. The correlations are gnerally close to zero in all
+cases suggesting the orthogonality property for the reparametrized
 model. Interestingly, results in the first panel show that
-cov($\hat{\beta}_{22}, \hat{\phi}$) is not very close to zero (values
+corr($\hat{\beta}_{22}, \hat{\phi}$) is not very close to zero (values
 between $-0.4$ and $0.2$) for strong overdispersion ($\phi = -1.6$).
@@
 To illustrate the orthogonality, \autoref{fig:ortho-surf} displays
-contour plots of the deviance surfaces for four simulated data set of
+contour plots of the deviance surfaces for four simulated data sets of
 size 1000 with $\mu=5$ and different values of the dispersion
@@
 the deviance function shape under strong overdispersion ($\phi=-1.6$) is
-not as well behaved as the others; this is due to the accurary of mean
-approximation \autoref{fig:approx-plot}. Although the reparametrized
-model is always valid, $\mu$ and $\phi$ are orthogonal only when $\mu$
-is the expectation of the distribution, in other words, when
-approximation $\lambda - (\nu-1)/(2\nu)$ is accurate. This also explains
-the results of Cov$(\hat{\beta}_{22}, \phi)$ in the first panel of
-\autoref{fig:ortho-plot}, since $\beta_{22}$ is negative and associated
-with low counts.
+not as well behaved as the others; this is due to the inaccuracy of the
+mean approximation \autoref{fig:approx-plot}(a). Although the
+reparametrized model is always valid, $\mu$ and $\phi$ are orthogonal
+only when $\mu$ is the expectation of the distribution, in other words,
+when the approximation $\lambda - (\nu-1)/(2\nu)$ is accurate. This also
+explains the results of corr$(\hat{\beta}_{22}, \phi)$ in the first
+panel of \autoref{fig:ortho-plot}, since $\beta_{22}$ is negative and
+associated with low counts.
 
@@
 
-In this section, we report three illustrative examples of count data
-analysis. We considered as alternative models for the analysis the
+In this section, we describe three illustrative examples of count data
+analysis. For the analyses as alternative models we considered the
 standard Poisson regression model, the COM-Poisson model in the two
-forms (original and new parametrization) and the quasi-Poisson
+forms (original and new parametrization), and the quasi-Poisson
 regression model.  The data sets and \texttt{R} code for their analysis
@@
 This example relates to cotton plants (\textit{Gossypium hirsutum})
-submitted to five levels of artificial defoliation (\texttt{des}) and
-crossed with five growth stages (\texttt{est}). The main goal of this
-study was to assess the effect of defoliation levels at different growth
-stages of cotton plants on the cotton production, expressed by the
-number of bolls produced. The study was conducted in a greenhouse and
-the experimental design was completely randomized with five
-replicates. This data set was analyzed by~\citet{Zeviani2014} using the
-Gamma-Count distribution.
+subjected to five levels of artificial defoliation (\texttt{des}) at
+each of five different growth stages (\texttt{est}). The main goal of
+this study was to assess the effect of defoliation levels at the
+different growth stages on the cotton production, measured by the number
+of bolls produced. The experimental study was conducted in a greenhouse
+and used a completely randomized design with five replicates. This data
+set was previously analyzed by~\citet{Zeviani2014} using the Gamma-Count
+distribution.
 
@@
 where $\eta_i = \log(\lambda_i)$ for the original COM-Poisson and
-$\eta_i = \log(\mu_{ij})$ for COM-Poisson$_\mu$. The parameter $\mu_i$
-is the (approximated) expected number of cotton bolls for the $i$-th
-defoliation level ($i=$ 1: 0\%, 2: 25\%, 3: 50\%, 4: 75\% e 5: 100\%)
-and $j$-th growth stage ($j$ = 1: vegetative, 2: flower bud, 3: blossom,
-4: boll, 5: boll open), i.e. we have a second order effect of
-defoliation in each growth stage. The parameters estimates and
-goodness-of-fit measures for the Poisson, COM-Poisson, COM-Poisson$_\mu$
-and quasi-Poisson regression models are presented in
+$\eta_i = \log(\mu_{ij})$ for COM-Poisson$_\mu$. The parameter
+$\mu_{ij}$ is the (approximated) expected number of cotton bolls for the
+$i$-th defoliation level ($i=$ 1: 0\%, 2: 25\%, 3: 50\%, 4: 75\% e 5:
+100\%) and $j$-th growth stage ($j$ = 1: vegetative, 2: flower bud, 3:
+blossom, 4: boll, 5: boll open), i.e. we have a different second order
+effect of defoliation at each growth stage. The parameter estimates and
+goodness-of-fit measures for the Poisson, COM-Poisson,
+COM-Poisson$_\mu$, and quasi-Poisson regression models are presented in
 \autoref{tab:coef-cotton}.
@@
 goodness-of-fit measures (log-likelihood, AIC and BIC) are quite similar
-for the COM-Poisson and COM-Poisson$_\mu$ models.  It suggests that the
-reparametrization does not change the model fit, as expected.  The
-Poisson model is clearly unsuitable, being overly conservative.  The
-difference in terms of log-likelihood value from the Poisson to the
-COM-Poisson$_\mu$ model was $\Sexpr{lrt.ncap[2, 4]}$, which in turn
-suggests the better fit of the COM-Poisson$_{\mu}$ model. A chi-square
-test also supports this statement.  Finally, the estimated value of the
-dispersion parameter $\\hat{\phi} = \Sexpr{co1[1]}$ suggests
-underdispersion.
+for the COM-Poisson and COM-Poisson$_\mu$ models. This suggests that the
+reparametrization does not change the model fit, as is to be expected.
+The Poisson model is clearly unsuitable, being overly conservative, due
+to the overestimated standard errors.  The difference in terms of
+$-2\times$log-likelihood value from the Poisson to the COM-Poisson$_\mu$
+model is $94.811$ for one additional parameter, which confirms the
+significantly improved fit of the COM-Poisson$_{\mu}$ model.  Finally,
+the estimated value of the dispersion parameter $\hat{\phi} = 1.582$
+suggests substantial underdispersion.
 
 Furthermore, results in \autoref{tab:coef-cotton} also show the
-advantage of the COM-Poisson$_\mu$ model, since the estimates are quite
-similar to the ones obtained by the Poisson model, whereas estimates
-obtained from the COM-Poisson model in the original parametrization are
-on a non interpretable scale.  The ratios between estimates and their
-respective standard errors for the COM-Poisson models are very close to
-ratios obtained by quasi-Poisson model. However, it is important to note
-that COM-Poisson model is a full parametric approach, i.e. there is a
-probability distribution associated to the counts. On the other hand,
-the quasi-Poisson model is a specification based only on second-moment
-assumptions.
+advantage of the COM-Poisson$_\mu$ model, since the regression parameter
+estimates are quite similar to those obtained for the Poisson model,
+whereas estimates from the original COM-Poisson model are on a different
+and not easily interpreted scale.  The ratios between estimates and
+their respective standard errors for the two COM-Poisson models are very
+close to ratios from the y quasi-Poisson model. However, it is important
+to note that COM-Poisson model is a full parametric approach, i.e. there
+is a probability distribution associated to the counts, while the
+quasi-Poisson model is based only on second-moment assumptions.
 
@@
 COM-Poisson$_{\mu}$ models, however, the confidence intervals are larger
-for the Poisson model because the equidispersion assumption.  The
-results from the COM-Poisson$_{\mu}$ model are consistent with those
-from the Gamma-Count model~\citep{Zeviani2014},
+for the Poisson model because of the inappropriate equidispersion
+assumption.  The results from the COM-Poisson$_{\mu}$ model are
+consistent with those from the Gamma-Count model~\citep{Zeviani2014},
 Poisson-Tweedie~\citep{Bonat2017} and the alternative parametrization of
-the COM-Poisson distribution proposed by~\citet{Huang2017}.  In all
-strategies the models indicated underdispersion and significant effects
-of defoliation for the vegetative, blossom and boll growth stages.
+the COM-Poisson distribution proposed by~\citet{Huang2017}.  All of
+these models indicated underdispersion and significant effects of
+defoliation at the vegetative, blossom and boll growth stages.
 
@@
 parameters, as computed by the asymptotic covariance matrix of the
-estimators, i.e. the inverse of the observed information.  The
-correlations are practically null considering the COM-Poisson$_\mu$. On
-the other hand, for the original parametrization such correlations are
-quite large, in particular for the parameter $\beta_0$ (due to effects
-parametrization in the linear predictor).  This result explain the
-better performance of the maximization algorithm in the new
-parametrization.  It is important to note that the initial values for
-the \texttt{BFGS} algorithm are provided by the Poisson model, then in
-the COM-Poisson$_{\mu}$ model the initial values are practically the
-maximum likelihood estimates and the effort of maximization is on the
-dispersion parameter $\phi$ only. To compare the computational times on
-the two parametrizations we repeat the fitting $50$ times. In
-this case COM-Poisson$_\mu$ fit was, on average, $38$\% faster than the
-original one.
+estimates, i.e. the inverse of the observed information.  The
+correlations are practically zero for the COM-Poisson$_\mu$, however,
+with the original parametrization some correlations are quite large, in
+particular for the parameter $\beta_0$ (due to the effects
+parametrization in the linear predictor).  This result explains the
+improved performance of the maximization algorithm in the new
+parametrization.  It is also important to note that as initial
+regression parameter values for the \texttt{BFGS} algorithm are provided
+by the Poisson model, then in the COM-Poisson$_{\mu}$ model the initial
+values are very close to the maximum likelihood estimates and
+maximization effort is essentially only on the dispersion parameter
+$\phi$. To compare the computational times for the two parametrizations
+we repeat the fitting $50$ times. In this example the COM-Poisson$_\mu$
+fit was, on average, $38$\% faster than for the original version.
 
@@
 
-The second example is a $5\times 3$ factorial experiment in a randomized
-complete block design. The aim of this study was to evaluate the effects
-of potassium doses (\texttt{K}) applied to soil (0, 0.3, 0.6, 1.2 and
-1.8 $\times$ 100mg dm$^{-3}$) and soil moisture (\texttt{umid}) levels
-(37.5, 50, 62.5\%) on  soybean (\emph{Glicine Max}) production. The
-experiment was carried out in a greenhouse, in pots with two plants, and
-the count variable measured was the number of bean seeds per
-pot~\citep{Serafim2012}.  \autoref{fig:desc-soy} (left) shows the number
-of bean seeds recorded for each combination of potassium dose and moisture
-level, it is important to note the indication of a quadratic effect of
-the potassium levels as shown by smoothing curves. Most points in the
-sample variance \textit{versus} sample means dispersion diagram (right)
-are above the identity line, suggesting overdispersion (block effect not
-yet removed).
+In this second example we consider a study of potassium doses and soil
+moisture levels on soybean (\emph{Glicine Max}) production. This was set
+up as a $5\times 3$ factorial experiment in a randomized complete block
+design with 5 replicates.  Five different potassium doses (\texttt{K})
+(0, 0.3, 0.6, 1.2 and 1.8 $\times$ 100mg dm$^{-3}$) were applied to the
+soil and soil moisture (\texttt{umid}) levels were controlled at (37.5,
+50, 62.5\%) . The experiment was carried out in a greenhouse and the
+experimental units were pots with two plants in each. The count response
+measured was the total number of bean seeds per pot~\citep{Serafim2012}.
+\autoref{fig:desc-soy} (left) shows the number of bean seeds recorded
+for each combination of potassium dose and moisture level; it is
+important to note the indication of a quadratic effect of the potassium
+levels, as indicated by the smoothers. Most points in the sample
+variance \textit{versus} sample means dispersion diagram (right) are
+above the identity line, suggesting overdispersion (block effect not yet
+removed).
 
@@
 <<desc-soy, fig.height=4, fig.width=8, fig.cap="Number of bean seeds per pot for each potassium dose and moisture level (left) and sample mean against sample variance of the five replicates for each experimental treatment (right).
-Solid lines are the smoothing curves on the left and the least of squares curve on the right.">>=
-@
-
-For the analysis of this data set based on the descriptive analysis
+Solid lines are smoothing curves (left) and the least-squares line (right).">>=
+@
+
+For the analysis of this data set, based on the descriptive analysis
 (\autoref{fig:desc-soy}), we proposed the following linear predictor
@@
 $$
-where $i=$1: block II, 2: block III, 3: block IV e 4: block V; $j=$1:
-50\% e 2: 62.5\%; and $k=$1: 0.0, 2: 0.3, 3: 0.6, 4: 1.2, 5: 1.8 100mg
-dm$^{-3}$, where $\gamma_i$ is the effect of $i$-th block ($i=$1: block
-II, 2: block III, 3: block IV and 4: block V), $\tau_j$ is the effect of
-$j$-th moisture level ($j=$1: 50\% and 2: 62.5\%) and $\beta_{3j}$ is
-interaction of the first order potassium effect (\texttt{K}) for the
-$j$-th moisture level (\texttt{umid}); The predictor $\eta_i$ is
-$\log(\lambda_i)$ for the original COM-Poisson and $\log(\mu_{ij})$ for
-COM-Poisson$_\mu$.\autoref{tab:coef-soy} presents the estimates, ratio
-between estimate and standard error and goodness-of-fit measures for the
-alternative models.
-
-The results in \autoref{tab:coef-soy} show that the two parametrization
-of COM-Poisson model presented very similar goodness-of-fit measures and
-better fit than the Poisson model. The difference between the
-log-likelihoods of the Poisson and COM-Poisson models was
-$\Sexpr{lrt.ngra[2, 4]}$, indicating that $\phi$ is significantly
-different from zero. The estimate of $\phi$ ($\Sexpr{co2[1]}$) indicates
-overdispersion, corroborating the descriptive analysis. Concerning to
-the regression parameters, the similarities between the models are
-analogous to the previous section. Both models indicate effects of
-block, potassium dose and moisture level, however the Poisson model
-indicates the effects with greater significance, because it does not
-take account of the extra variability.
+where $\gamma_i$ is the effect of $i$-th block ($i=$1: block II, 2:
+block III, 3: block IV and 4: block V), $\tau_j$ is the effect of $j$-th
+moisture level ($j=$1: 50\% and 2: 62.5\%), $\beta_{1}$ and $\beta_{2}$
+give the baseline quadratic response over potassium levels \texttt{K}
+($k=$1: 0.0, 2: 0.3, 3: 0.6, 4: 1.2, 5: 1.8 100mg dm$^{-3}$) and
+$\beta_{3j}$ is interaction of the first order potassium effect
+(\texttt{K}) for the $j$-th moisture level (\texttt{umid}); The
+predictor $\eta_i$ is $\log(\lambda_i)$ for the original COM-Poisson and
+$\log(\mu_{ij})$ for COM-Poisson$_\mu$. \autoref{tab:coef-soy} presents
+the estimates, ratio between estimate and standard error, and
+goodness-of-fit measures for the alternative models.
+
+The results in \autoref{tab:coef-soy} show that the two parametrizations
+of the COM-Poisson model presented very similar goodness-of-fit measures
+and a better fit than the Poisson model. The $-2\times$difference
+between the log-likelihoods of the Poisson and COM-Poisson models was
+$29.697$ on 1 degree of freedom, indicating that $\phi$ is significantly
+different from zero. The estimate of $\phi$ ($-0.782$) indicates
+overdispersion, corroborating the descriptive analysis. Concerning the
+regression parameters, the similarities between the models are analogous
+to those in the previous section. Both models show effects of block,
+potassium dose and moisture level, however the Poisson model indicates
+the effects with greater significance, because it does not take account
+of the extra variability and so underestimates standard errors.
 
@@
 the computation of the log-likelihood function is computationally
-expensive.  For the data set considered, the upper bound was fixed at
-$700$.  The \texttt{BFGS} algorithm evaluated the log-likelihood
-function $\Sexpr{c21[1]}$ and $\Sexpr{c22[2]}$ times to reach
-convergence, when using the original and new parametrization of the
-COM-Poisson distribution, respectively.  In terms of computational time,
-for 50 repetitions of fit, the proposed reparametrization was on average
-$110\%$ faster than the original one.  Probably, it is due to the better
-behaviour of the log-likelihood function as well as better initial
-values obtained from the Poisson fit. In \autoref{tab:corr-soy}, we
-present the empirical correlation between the regression and dispersion
-parameter estimates. The correlations are close to zero for the
+expensive.  For the data set considered here , the upper bound was fixed
+at $700$.  To reach convergence the \texttt{BFGS} algorithm evaluated
+the log-likelihood function 264 times when using the original
+parametrization of the COM-Poisson distribution and only 20 times with
+the new parametrization.  In terms of computational time, for 50
+repetitions of the fit, the proposed reparametrization was, on average,
+$110\%$ faster than the original one.  This is probably due to the
+better behaviour of the log-likelihood function as well as more relevant
+initial values from the Poisson fit. In \autoref{tab:corr-soy}, we
+present the empirical correlations between the regression and dispersion
+parameter estimates. Again, the correlations are close to zero for the
 COM-Poisson$_\mu$ model, indicating the empirical orthogonality between
@@
 same conclusions. On the other hand, confidence intervals for the
-Poisson model are smaller than the ones from the COM-Poisson$_{\mu}$,
-due to the equidispersion assumption underlying the Poisson model. The
+Poisson model are narrower than those from the COM-Poisson$_{\mu}$, due
+to the inappropriate equidispersion assumption of the Poisson model. The
 confidence intervals from the COM-Poisson$_{\mu}$ and quasi-Poisson
-models are really similar, which in turn shows the already highlighted
+models are also very similar, which again shows the already highlighted
 similarity between these approaches, however only the COM-Poisson
@@
 
-<<pred-soy, fig.height=3.8, fig.width=8, fig.cap="Dispersion diagrams of been seeds counts as function of potassium doses and humidity levels with fitted curves and confidence intervals (95\\%).">>=
+<<pred-soy, fig.height=3.8, fig.width=8, fig.cap="Observed bean seeds counts as functions of potassium doses and soil humidity levels with fitted curves and confidence intervals (95\\%).">>=
 @
@@
 
-The data set comes from an experiment to measure the reproductive
+This data set comes from an experiment to measure the reproductive
 toxicity of the herbicide, nitrofen, on a species of zooplankton
@@
 concentration of nitrofen ($0, 0.8, 1.6, 2.35$ and $3.10$
-$\mu$g$/10^2$litre) (\texttt{dose}). Then the number of live offspring
-was recorded.
+$\mu$g$/10^2$litre) (\texttt{dose}). Subsequently, the number of live
+offspring was recorded.
 
@@
 
-\footnotesize \raggedright np, number of parameters; diff $\ell$,
+\footnotesize \raggedright $\#p$, number of parameters; diff $\ell$,
 difference in log-likelihoods; QDev, quasi-deviance, F, F statistics
-based on quasi-deviances; diff np, difference in np.
+based on quasi-deviances; diff $\#p$, difference in number of parameters.
 \end{table}
@@
 indicate the significance of the cubic effect of the nitrofen
-concentration.  Considering this predictor, there is an evidence of
+concentration.  Considering this predictor, there is evidence of
 equidispersed counts, the $\phi$ estimate of the COM-Poisson is close to
-zero and $\sigma$ of quasi-Poisson is close to one.  It is interesting
-to note that if we omit the high order effects the models show evidence
-of overdispersion. This exemplifies the discussion on the causes of
-overdispersion made in \autoref{introduction}.  We can also note that
-the quasi-Poisson approach, although robust to equidispersion
-assumption, shows higher descriptive levels ($p$-values) than parametric
-models, that is, the tests under parametric models are more powerful
-than the ones under the quasi-Poisson model in the equidispersed case.
+zero and the $\sigma$ estimate of the quasi-Poisson is close to one.  It
+is interesting to note that if we omit the higher order effects the
+models show evidence of overdispersion --- this exemplifies the
+discussion on the possible causes of overdispersion in
+\autoref{introduction}.  We also note that the quasi-Poisson approach,
+although robust to the equidispersion assumption, shows higher
+descriptive levels ($p$-values) than parametric models, that is, the
+tests under parametric models are more powerful than the ones under the
+quasi-Poisson model in the equidispersed case.
 
 In \autoref{tab:coef-ovos}, we present the estimates of the regression
-parameters considering the cubic dose model. The interpretations are
-similar to others cases studies, however, in this case the Poisson model
-is also suitable for indicating the significance of the covariate
-effects.  In addition, note that the parameter estimates of the original
-COM-Poisson model are comparable to the others models. This occurs
-because we are in the particular case, where $\phi = 0$, which implies
-$\lambda = \mu$.
+parameters for the cubic dose models. The interpretations are similar to
+the other case studies, however, here the Poisson model is also suitable
+for indicating the significance of the covariate effects.  In addition,
+note that the parameter estimates of the original COM-Poisson model are
+comparable to the others models. This occurs because we are in the
+particular case, where $\phi \approx 0$, which implies
+$\lambda \approx \mu$.
 
@@
 \autoref{fig:pred-ovos} shows the number of live off-spring observed and
-fitted curves along with confidence intervals for all model strategies
-adopted.  The fitted values and confidence intervals are identical and
-have a complete overlap. It shows that the estimation of the extra
+fitted curves along with confidence intervals for the different cubic
+dose models.  The fitted values and confidence intervals are identical
+and have a complete overlap. This shows that the estimation of the extra
 dispersion parameter does not affect the estimation of the regression
@@
 
-<<pred-ovos, fig.height=3.5, fig.width=5.5, out.width="0.7\\textwidth", fig.cap="Number of live offsprings observed for each nitrofen concentration level with fitted curves and 95\\% confidence intervals.">>=
+<<pred-ovos, fig.height=3.5, fig.width=5.5, out.width="0.7\\textwidth", fig.cap="Number of live offspring observed for each nitrofen concentration level with fitted curves and 95\\% confidence intervals.">>=
 @
@@
 correlations are practically null. The computational times for fifty
-repetitions of fit are similar. The average time to fit for the
+repetitions of fit are similar; the average time to fit the
 COM-Poisson$_\mu$ and COM-Poisson models is 1.19 and 1.09 seconds,
@@
 
-In this paper, we presented and characterized a novel parametrization of
-the COM-Poisson distribution and the associated regression model. The
-novel parametrization was based on a simple asymptotic approximation for
-the expectation of the COM-Poisson distribution. The main advantage of
-the proposed reparametrization is the simple interpretation of the
+In this paper, we presented and characterized a novel reparametrization
+of the COM-Poisson distribution and associated regression model. The
+reparametrization was based on a simple asymptotic approximation for the
+expectation of the COM-Poisson distribution. The main advantage of the
+proposed reparametrization is the simple interpretation of the
 regression coefficients in terms of the expectation of the response
 variable as usual in the generalized linear models context. Thus, it is
-possible to compare the results of the COM-Poisson model with the ones
-from standard approaches as the Poisson and quasi-Poisson regression
-models. Furthermore, in the novel parametrization the COM-Poisson
-distribution is indexed by $\mu$ and an extra dispersion parameter
-$\phi$ which our data analysis suggest being orthogonal. This is similar
-to \Citeauthor{Huang2017}'s (\citeyear{Huang2017}) parametrization but
-is simpler, because the $\mu$ is obtained from simple algebra.
+possible to compare directly the results from the COM-Poisson model with
+those from standard approaches, such as the Poisson and quasi-Poisson
+regression models. Furthermore, in the new parametrization the
+COM-Poisson distribution is indexed by $\mu$ and an extra dispersion
+parameter $\phi$ which our studies suggest are approximately
+orthogonal. Overall the approach is similar to \Citeauthor{Huang2017}'s
+(\citeyear{Huang2017}) parametrization but ours is simpler, because the
+$\mu$ used here is obtained from simple algebra.
 
 The proposed parametrization is always valid, only the interpretation of
-$\mu$ parameter, as expectation of the distribution, depends on the
+$\mu$ parameter, as the expectation of the distribution, depends on the
 accuracy of the approximation. We evaluated the accuracy of the
@@
 showed that the approximation for the expectation is accurate for a
-large part of the parameter space, which in turn support our
-reparametrization. Moreover, we discuss the properties and flexibility
-of the distribution to deal with count data although dispersion,
-zero-inflation and heavy-tail indexes. We carried out a simulation study
-to assess the properties of the reparametrized COM-Poisson model to deal
-with different levels of dispersion as well as the properties of the
-maximum likelihood estimators. The results of our simulation study
-suggested that the maximum likelihood estimators of the regression and
-dispersion parameters are unbiased and consistent.  The empirical
-coverage rates of the confidence intervals computed based on the
-asymptotic distribution of the maximum likelihood estimators are close
-to the nominal level for sample size greater than $100$.  The worst
-scenario is when we have small sample sizes and strong overdispersed
-counts.  In general, we recommend the use of the asymptotic confidence
-intervals for computational simplicity.
-
-The data analyses have shown that the COM-Poisson regression model is a
-suitable choice to deal with dispersed count data.  The observed
-empirical correlation between the regression and dispersion parameter
-estimators and deviance surfaces suggest orthogonality between $\mu$ and
-$\phi$ in COM-Poisson$_\mu$ distribution. Thus, the computational
-procedure based on the proposed reparametrization is faster than in the
-original parametrization.
+large part of the parameter space, supporting our reparametrization. We
+discuss the properties and flexibility of the distribution to deal with
+count data through considerations of dispersion, zero-inflation and
+heavy-tail indexes. A simulation study was used to assess the properties
+of the reparametrized COM-Poisson model and its ability to deal with
+different levels of dispersion, as well as the properties of the maximum
+likelihood estimators. The results of our simulation study suggested
+that the maximum likelihood estimators of the regression and dispersion
+parameters are unbiased and consistent.  The empirical coverage rates of
+the confidence intervals computed based on the asymptotic distribution
+of the maximum likelihood estimators are close to the nominal level for
+sample sizes greater than $100$.  The worst case scenario is when we
+have a small sample size and strongly overdispersed counts.  In general,
+we recommend the use of the asymptotic confidence intervals for
+computational simplicity, although of course bootstrap intervals could
+be obtained, but would involve extensive computation.
+
+The three examples and data analyses have shown that the COM-Poisson
+regression model is a suitable choice to deal with generally dispersed
+count data, equi-, under- and over-dispersed.  The observed empirical
+correlation between the regression and dispersion parameter estimators
+and deviance surfaces suggest approximate orthogonality between $\mu$
+and $\phi$ in the COM-Poisson$_\mu$ distribution. Thus, the
+computational procedure based on the proposed reparametrization is
+faster than that for the original parametrization.
 
@@
 extension of the proposed model is to model both $\mu$ and $\phi$
-parameters as functions of covariate in a double generalized linear
-models framework.  Finally, the reparametrized version of the
-COM-Poisson model also encourages the specification of generalized
-linear mixed models using this distribution.
+parameters as functions of covariates in a double generalized linear
+model framework.  Finally, the reparametrized version of the COM-Poisson
+model would also be highly suitable for the specification of generalized
+linear mixed COM-Poisson models.
 

### 2018-09-13 codes/functions.R
Simplify functions to compute the pmf and dist moments

@@ -61,84 +61,53 @@ llcmp2 <- function (params, y, X, offset = NULL, sumto = NULL) {
 
 #=======================================================================
 # Probability mass function (mean parametrization)
-dcmp <- Vectorize(FUN = function(y, mu, phi, sumto = 100) {
-    exp(-llcmp2(c(phi, log(mu)), y = y, X = 1, sumto = sumto))
-}, vectorize.args = c("y", "mu", "phi"))
+dcmp <- function(y, mu, phi, sumto = 500L) {
+    nu <- exp(phi)
+    k0 <- nu * log(mu + (nu - 1)/(2 * nu))
+    j <- 1:sumto
+    Z <- 1 + sum(exp(j * k0 - nu * lfactorial(j)))
+    l <- y * k0 - nu * lfactorial(y) - log(Z)
+    return(exp(l))
+}
 
 #=======================================================================
 # Compute momentes mean and variance
-compute_mean <- function(mu, phi, sumto = NULL, tol = 1e-5) {
-    # Faz com que os parâmetros sejam vetores de mesmo tamanho.
+compute_mean <- function(mu, phi, sumto = NULL, upto = 500) {
+    y <- 0:upto
     names(mu) <- NULL
     names(phi) <- NULL
     pars <- data.frame(mu = mu, phi = phi)
-    # Calcula o ymax usando mu + 5 (sqrt(sigma)) aproximados
-    nu <- exp(phi)
-    sigma <- (1 / nu) * mu
-    ymax <- ceiling(max(mu + 5 * sqrt(sigma)))
-    # Agora verifica se a prob(ymax) é de fato pequena, se não, soma 1.
-    mumax <- max(pars$mu)
-    phimin <- min(pars$phi)
-    pmax <- dcmp(y = ymax, mu = mumax, phi = phimin, sumto = sumto)
-    while (pmax > tol) {
-        ymax <- ymax + 1
-        pmax <- dcmp(y = ymax, mumax, phimin, sumto = sumto)
-    }
-    # Define o vetor onde avaliar a densidade COM-Poisson.
-    y <- 1:ymax
-    estmean <- mapply(mu = pars$mu,
-                      phi = pars$phi,
-                      MoreArgs = list(y = y, sumto = sumto),
-                      FUN = function(mu, phi, y, sumto) {
-                          py <- dcmp(y, mu, phi, sumto)
-                          sum(y * py)
-                      },
[…]

### 2018-09-13 article.Rnw
Report the errors instead of quadratic errors and make minor corrections

@@
 (\autoref{eqn:mean-approx}), \autoref{fig:approx-plot} presents the
-squared errors for (a) expectation and (b) variance.  The squared errors
-are $[\mu - \text{E}(Y)]^2$ for the expectation and
-$[ \mu \exp(-\phi) - \text{Var}(Y)]^2$ for the variance. In both cases
-$\text{E}(Y)$ and Var$(Y)$ were computed numerically using 500
-terms\footnote{ E$(Y) \approx \sum_{y=0}^{500}yp(y)$ and
+errors for (a) expectation ($\mu - \text{E}(Y)$) and (b) variance
+($\mu \exp(-\phi) - \text{Var}(Y)$). In both cases $\text{E}(Y)$ and
+Var$(Y)$ were computed numerically using 500 terms\footnote{
+  E$(Y) \approx \sum_{y=0}^{500}yp(y)$ and
   E$(Y^2) \approx \sum_{y=0}^{500}y^2p(y)$.}. The dotted lines represent
@@
 
-<<approx-plot, fig.height=3, fig.width=6, out.width=".8\\textwidth", fig.cap="Squared errors for the approximation of the (a) expectation and (b) variance.
+<<approx-plot, fig.height=3, fig.width=6, out.width=".8\\textwidth", fig.cap="Errors for the approximation of the (a) expectation and (b) variance.
 Dotted lines represent the restriction for suitable approximations given by \\cite{Shmueli2005}.">>=
@@
 The results in \autoref{fig:approx-plot} show that the mean
-approximation is accurate, the largest squared error, for the parameter
-values evaluated, is $0.038$.  For the variance approximation, the
-largest squared error was $XXX$ and it occurs for negative values of
-$\phi$. Interestingly, the errors for the variance are larger for
-negative values of $\phi$ and present no clear relation with $\mu$, as
-opposed to the regions gives by \citet{Shmueli2005} ($\phi \leq 0$ and
-$\mu > 10 - \frac{\exp(\phi) - 1}{2\exp(\phi)}$).
+approximation is accurate, the largest (absolute) error, for the
+parameter values evaluated, is $-0.197$.  For the variance
+approximation, the largest error was $8.4$ and it occurs for negative
+values of $\phi$. Interestingly, the errors for the variance are larger
+for negative values of $\phi$ and present no clear relation with $\mu$,
+as opposed to the regions gives by \citet{Shmueli2005} ($\phi \leq 0$
+and $\mu > 10 - (\exp(\phi) - 1)/(2\exp(\phi))$).
 
@@
 distribution is suitable to deal to dispersed counts, of course. For the
-parameter values evaluated the largest DI was $XXX$ and smallest was
-$XXX$. \autoref{fig:indexes-plot}(c) shows the COM-Poisson can handle
+parameter values evaluated the largest DI was $4.21$ and smallest was
+$0.168$. \autoref{fig:indexes-plot}(c) shows the COM-Poisson can handle
 a limited amount of zero-inflation, in cases of overdispersion

### 2018-09-13 article.R

@@ -52,6 +52,7 @@ cols <- trellis.par.get("superpose.line")$col
 
 # Useful packages
 library(bbmle)
+library(methods)
 library(multcomp)
 library(plyr)
 library(tidyr)
@@ -83,7 +84,6 @@ grid$z <- apply(grid, 1, function(par) {
 })
 
 xt <- xtabs(z ~ nu + lambda, data = grid)
-
 
 ## ----results-Z
 caption <- paste(
@@ -119,53 +119,58 @@ print(xtable(xt2, digits = -2,
 # Mean and variance relationship
 aux <- expand.grid(
     mu = seq(3, 30, length.out = 55),
-    phi = seq(-1.5, 1.8, length.out = 50))
+    phi = seq(-1.5, 1.8, length.out = 50),
+    KEEP.OUT.ATTRS = FALSE)
 
-moments <- mapply(FUN = calc_moments,
-                  mu = aux$mu,
-                  phi = aux$phi,
-                  MoreArgs = list(sumto = 300),
-                  SIMPLIFY = FALSE)
-grid <- cbind(aux, t(do.call(cbind, moments)))
-grid <- transform(grid, va = mu / exp(phi))
+grid <- with(aux, t(calc_moments(mu, phi))) %>%
+    as.data.frame() %>%
+    cbind(aux, .) %>%
+    transform(va = mu / exp(phi)) %>%
+    transform(emu = (mu - mean),
+              eva = (va - var))
 
+maemu <- max(abs(grid$emu))
+maeva <- max(abs(grid$eva))
 
 ## ----approx-plot
 # Errors in approximations for E[Y] and V[Y]
-grid <- transform(grid,
-                  emu = (mu - mean)^2,
-                  eva = (va - var)^2)
 
-myreg <- colorRampPalette(c("gray90",  "gray20"))
-xy1 <- levelplot(emu ~ phi + mu, data = grid,
-                 aspect = "fill",
-                 col.regions = myreg,
-                 xlab = expression(phi),
-                 ylab = expression(mu),
-                 sub = "(a)",
-                 colorkey = list(space = "top"),
-                 par.settings = ps2,
-                 panel = function(x, y, z, ...) {
-                     panel.levelplot(x, y, z, ...)
-                     panel.curve(10 - (exp(x) - 1)/(2 * exp(x)),
-                                 lty = 2)
-                     panel.abline(v = 0, lty = 2)
-                 })
[…]

### 2018-09-13 article.Rnw
Make minor corrections

@@
 is computed numerically by central finite differences. Standard errors
-for $\hat{\mu}_i$ and hence confidence intervals are obtained by using
-the delta method~\citep[p. 89]{Pawitan2001}.
+for $\hat{\eta}_i = \log(\hat{\mu}_i)$ and hence confidence intervals
+are obtained by using the delta method~\citep[p. 89]{Pawitan2001}.
 
@@
 model. Interestingly, results in the first panel show that
-corr($\hat{\beta}_{22}, \hat{\phi}$) is not very close to zero (values
+Corr($\hat{\beta}_{22}, \hat{\phi}$) is not very close to zero (values
 between $-0.4$ and $0.2$) for strong overdispersion ($\phi = -1.6$).
@@
 when the approximation $\lambda - (\nu-1)/(2\nu)$ is accurate. This also
-explains the results of corr$(\hat{\beta}_{22}, \phi)$ in the first
+explains the results of Corr$(\hat{\beta}_{22}, \phi)$ in the first
 panel of \autoref{fig:ortho-plot}, since $\beta_{22}$ is negative and
@@
 Following~\citet{Zeviani2014}, the linear predictor is given by
-$$\eta_i = \beta_0 + \beta_{1j} \texttt{def}_i + \beta_{2j}
+$$\eta_{ij} = \beta_0 + \beta_{1j} \texttt{def}_i + \beta_{2j}
 \texttt{def}_i^2$$
-where $\eta_i = \log(\lambda_i)$ for the original COM-Poisson and
-$\eta_i = \log(\mu_{ij})$ for COM-Poisson$_\mu$. The parameter
+where $\eta_{ij} = \log(\lambda_{ij})$ for the original COM-Poisson and
+$\eta_{ij} = \log(\mu_{ij})$ for COM-Poisson$_\mu$. The parameter
 $\mu_{ij}$ is the (approximated) expected number of cotton bolls for the
@@
 $$
-\eta_i = \beta_0 + \gamma_i + \tau_j +
+\eta_{ijk} = \beta_0 + \gamma_i + \tau_j +
   \beta_{1}\texttt{K}_k + \beta_{2}\texttt{K}_k^2 +
@@
 (\texttt{K}) for the $j$-th moisture level (\texttt{umid}); The
-predictor $\eta_i$ is $\log(\lambda_i)$ for the original COM-Poisson and
-$\log(\mu_{ij})$ for COM-Poisson$_\mu$. \autoref{tab:coef-soy} presents
-the estimates, ratio between estimate and standard error, and
-goodness-of-fit measures for the alternative models.
+predictor $\eta_{ijk}$ is $\log(\lambda_{ijk})$ for the original
+COM-Poisson and $\log(\mu_{ijk})$ for
+COM-Poisson$_\mu$. \autoref{tab:coef-soy} presents the estimates, ratio
+between estimate and standard error, and goodness-of-fit measures for
+the alternative models.
 
@@
 
-For this data set we consider three models with linear predictors,
-\begin{center}
-\begin{minipage}{12cm}
-Linear: $\eta_i = \beta_0 + \beta_1 \texttt{dose}_i$\\
-Quadratic: $\eta_i = \beta_0 + \beta_1 \texttt{dose}_i +
-             \beta_2 \texttt{dose}_i^2$\\
-Cubic: $\eta_i = \beta_0 + \beta_1 \texttt{dose}_i +
-             \beta_2 \texttt{dose}_i^2 + \beta_3 \texttt{dose}_i^3$.
-\end{minipage}
-\end{center}
+For this data set we consider three models with linear
+predictors,\\[\parskip]
+\hspace*{1.5\parskip}
+\begin{tabular}{ll}
+  Linear:     & $\eta_i = \beta_0 + \beta_1 \texttt{dose}_i$,\\
+  Quadratic: & $\eta_i = \beta_0 + \beta_1 \texttt{dose}_i +
+               \beta_2 \texttt{dose}_i^2$ e\\
+  Cubic:     & $\eta_i = \beta_0 + \beta_1 \texttt{dose}_i +
+               \beta_2 \texttt{dose}_i^2 + \beta_3 \texttt{dose}_i^3$,
+\end{tabular}
+
 where $\eta_i = \log(\lambda_i)$ for the original COM-Poisson and
-$\eta_i = \log(\mu_{ij})$ for COM-Poisson$_\mu$.
+$\eta_i = \log(\mu_i)$ for COM-Poisson$_\mu$.
 
@@
   \toprule
- Poisson & np & $\ell$ & AIC & 2(diff $\ell$) & diff np & P($>\rchi^2$) & \\
+ Poisson & $\#p$ & $\ell$ & AIC & 2(diff $\ell$) & diff $\#p$ & P($>\rchi^2$) & \\
  \midrule
@@
 \specialrule{0em}{0.5em}{0em} %% Apenas para espaçamento
-  COM-Poisson & np & $\ell$ & AIC & 2(diff $\ell$) & diff np &
+  COM-Poisson & $\#p$ & $\ell$ & AIC & 2(diff $\ell$) & diff $\#p$ &
   P($>\rchi^2$) & $\hat{\phi}$ \\
@@
 \specialrule{0em}{0.5em}{0em} %% Apenas para espaçamento
-  COM-Poisson$_\mu$ & np & $\ell$ & AIC & 2(diff $\ell$) & diff np &
+  COM-Poisson$_\mu$ & $\#p$ & $\ell$ & AIC & 2(diff $\ell$) & diff $\#p$ &
   P($>\rchi^2$) & $\hat{\phi}$ \\
@@
 \specialrule{0em}{0.5em}{0em} %% Apenas para espaçamento
-  Quasi-Poisson & np & QDev & AIC & F & diff np & P($>F$) & $\hat{\sigma}$ \\
+  Quasi-Poisson & $\#p$ & QDev & AIC & F & diff $\#p$ & P($>F$) & $\hat{\sigma}$ \\
   \midrule

### 2018-09-13 review/review.Rnw
Make minor corrections on the point-by-point responses

@@
   \textbf{\underline{Response:}} The log-likelihood function cannot be
-  derived in relation ti the dispersion parameter using any
+  derived in relation to the dispersion parameter using any
   parametrization. We obtained the Hessian matrix by central finite
-  differences using the Richardson method implemented in
-  \texttt{R}. Although the expression of reparametrized log-likelihood
-  (Equation 3.1) seems more complicated, this function is better behaved
-  than the function expressed in the original form (see Figure
-  8). Consider the exponential family of distributions, with probability
-  function given by
+  differences using the Richardson method as implemented in R. Although
+  the expression of reparametrized log-likelihood (Equation 3.1) seems
+  more complicated, this function is better behaved than the function
+  expressed in the original form (see Figure 8). Consider the
+  exponential family of distributions, with probability function given
+  by
   $$f(y) = \exp\{\sigma[\theta y - b(\theta)] - c(y, \phi)\},$$
@@
   $b'(\theta) = \partial b(\theta)/\partial \theta$ in closed form, the
-  canonical link function has no important rule here. In this case, the
+  canonical link function has no important role here. In this case, the
   link function maps $\theta$ in the parameter space of $\lambda$.  In
-  the reparametrized version of the COM-Poisson, the log link function
-  maps the parameter space and provide simple interpretations (in terms
+  the reparametrized version of the COM-Poisson the log link function
+  maps the parameter space and provides simple interpretations (in terms
   of rate ratios).
+
 \item \textit{For model fitting, the authors used general purpose
@@
   below). This occurred only for strong overdispersion and small sample
-  sizes. We reported these situations in the updated version of the paper.
+  sizes. We reported these situations in the updated version of the
+  paper.
   \begin{table}[H]
@@
   more time to fit the models under equidispersion. This result is due
-  to choose of start values and normalizing constant upper bound. We
-  set, for all models, the upper bound 500 and we start the algorithm
-  from the real parameters.
+  to choice of starting values and normalizing constant upper bound. We
+  set, for all models, the upper bound as 500 and we start the algorithm
+  from the true parameters.
 
@@
 \end{minipage}
+\vspace{0.1cm}
 
@@
   cover a wide range of counts. The Figure below shows the variation of
-  the counts for $\mu=27$ for the different scenarios.
+  the counts for $\mu=27$ for the four different scenarios.
 
@@
   COM-Poisson distribution in terms of the dispersion, zero-inflation
-  and heavy-tail indexes that illustrate its applicability in the real
-  count data. These are the methodological novelties of the paper
-  besides the exploration of the results. However, in the updated
-  version of the paper, we provided a better discussion of the case
-  studies.
+  and heavy-tail indexes that illustrate its applicability to real count
+  data. These are the methodological novelties of the paper besides the
+  exploration of the results. However, in the updated version of the
+  paper, we provided a better discussion of the case studies.
 \end{enumerate}
@@
   behaved...'', p. 19, etc.). This needs to be addressed.}\\[0.3cm]
-\textbf{\underline{Response:}} We thanks the referee for exploring this
+\textbf{\underline{Response:}} We thank the referee for exploring this
 point. However, we didn't use the approximation of variance
 ($a\text{Var}(Y_i)$) to reparametrize or fit the COM-Poisson model. To
-compute the variances, that is showed in the Figure 3(a) and (b) we used
-$\text{Var}(Y)=\sum_{y=0}^{500} y^2p(y) - [\sum_{y=0}^{500} y
-p(y)]^2$.
+compute the variances, that are shown in the Figures 3(a) and (b) we
+used
+$\text{Var}(Y)=\sum_{y=0}^{500} y^2p(y) - [\sum_{y=0}^{500} y p(y)]^2$.
 As highlighted in the Figure 1(b), the approximation is not accurate,
@@
 setting) are orthogonal. We improved the discussion about this in the
-current version of the paper.  \vspace{0.4cm}
+current version of the paper.
+\vspace{0.4cm}
 
@@
 Figure 3(a) (and in referee's previous comment), the mean-variance
-relationship for COM-Poisson is linear. Therefore, is suitable to
-analyze data generated according to the COM-Poisson distribution, using
-the assumption $\text{Var}(Y_i)\propto \mu_i$. The advantage of the
-COM-Poisson approach is that it corresponds to a fully specified
-probability model allowing to compute the likelihood and some useful
-measures like deviance, AIC, BIC, LRT, etc.
-\vspace{0.4cm}
+relationship for COM-Poisson is linear. Therefore, it can be appropriate
+to analyze under-, equi-, and overdispersed data generated according to
+the COM-Poisson distribution, using the assumption
+$\text{Var}(Y_i)\propto \mu_i$. The advantage of the COM-Poisson
+approach is that it corresponds to a fully specified probability model
+allowing to compute the likelihood and some useful measures like
+deviance, AIC, BIC, LRT, etc.  \vspace{0.4cm}
 
@@
 \textbf{\underline{Response:}} Note that the $\lambda_i$ does not
-represent the expectation, nor approximately. The quadratic errors in
-Figure 1(a) is obtained by $[a\text{E(Y)} - \sum_{y=0}^{500}y p(y)]^2$.
-The models fitted in the case studies are
+represent the expectation, nor approximately. The errors in Figure 1(a)
+is obtained by $[a\text{E(Y)} - \sum_{y=0}^{500}y p(y)]^2$.  The models
+fitted in the case studies are
 $[\text{COMPo}(\lambda_i, \nu), \log(\lambda_i) = \beta^\top x_i]$ and
@@
   \textbf{\underline{Response:}} The constraint $\mu>0$ implies that
-  $\lambda > [(\nu-1)/(2\nu)]^\nu$. The Figure below show (a) the
+  $\lambda > [(\nu-1)/(2\nu)]^\nu$. The Figures below show (a) the
   constraint border in the parameter space; and (b) the expected values
-  (obtained by $\sum y p(y)$). The constraint lies on the very small
-  $\lambda$ and $\nu>1$ (underdispersion). This parameter region is
-  related to small expected values (smaller than 0.1) and
-  underdispersion. Therefore, although this constraint is undesirable,
-  it does not prejudice the application of the model. We thank the
-  referee for highlighting this point. We have discussed it in the
-  updated version.
+  (obtained by $\sum y p(y)$). The constraint applies to very small
+  values of $\lambda$ when $\nu>1$ (underdispersion). This infeasible
+  parameter region is related to small expected values (smaller than
+  0.1) and underdispersion. Therefore, although this constraint is
+  undesirable, it does not prejudice the application of the model. We
+  thank the referee for highlighting this point and we have discussed it
+  in the updated version.
 
@@
   given by Shmueli to reparametrize the COM-Poisson model. Moreover, as
-  discussed in Figure 1(a), the the errors are close to 0 for the
-  parameter grid evaluated and present no clear relation with regions
-  gives by \cite{Shmueli2005} ($\phi \leq 0$ and
-  $\mu > 10 - (\exp(\phi) - 1)/(2\exp(\phi))$. Except for
-  counts with small averages (close to 0), the reparametrized model fits
-  well. For examples presented in the paper, the parameter estimates
-  are summarizes below.
+  discussed in Figure 1(a), the errors are close to 0 for the parameter
+  grid evaluated and present no clear relation with regions gives by
+  \cite{Shmueli2005} ($\phi \leq 0$ and
+  $\mu > 10 - (\exp(\phi) - 1)/(2\exp(\phi))$. The paper proposal is a
+  parameterization, so it is always valid (except for the parameter
+  constraint imposed by the transformation). Only the interpretation of
+  the $\mu$ parameter depends on the approximation accuracy. For
+  examples presented in the paper, the parameter estimates (on the
+  original parametrization) are summarized below. The approximation is
+  good and therefore the $\beta$'s estimates for the Poisson model are
+  close to the $\beta$'s estimates for the reparametrized COM-Poisson
+  model.
 
-<<contraints-case-studies, results="asis", cache=TRUE>>=
-@
+% latex table generated in R 3.4.4 by xtable 1.8-2 package
+% Tue Aug 21 00:57:24 2018
+\begin{table}[H]
+\centering
+\begingroup\small
+\begin{tabular}{lrrrrr}
+  \toprule
+ & & & \multicolumn{3}{c}{$\hat{\lambda}_i$}\\
+ \cmidrule(lr){4-6}
+Case study & $\hat{\phi}$ & $\hat{\nu}$ & Minimun & Median & Maximum \\
+  \midrule
+  Cotton experiment   & 1.58 & 4.86 & 3.50 & 8.44 & 9.48 \\
+  Soybean experiment  & $-$0.78 & 0.46 &116.87 & 174.98 & 252.29 \\
+  Nitrofen experiment & 0.05 & 1.05 & 5.94 & 28.00 & 32.36 \\
+   \bottomrule
+\end{tabular}
+\endgroup
+\end{table}
 

### 2018-09-13 review/review.R

@@ -243,7 +243,7 @@ benchs <- readRDS("../codes/comparetimes.rds") %>%
     mutate(case = forcats::fct_inorder(case))
 xlabs <- expression("COMPo"(lambda[i], phi),
                     "COMPo"(mu[i], phi))
-bwplot(time/10e6 ~ expr | case,
+bwplot(time/10e8 ~ expr | case,
        ylab = "Time (seconds)",
        axis = axis.grid,
        scales = list(

### 2018-09-13 codes/00_explore-model.R
Rename the codes files and fix small bugs

@@ -7,8 +7,10 @@
 
 #-----------------------------------------------------------------------
 # Load package and functions
-source("functions.R")
-source("lattice-panels.R")
+source("helper01_general-functions.R")
+source("helper02_lattice-panels.R")
+
+library(dplyr)
 
 #-----------------------------------------------------------------------
 # Convergence of Z(lambda, nu) constant
@@ -39,54 +41,59 @@ xt
 
 # Parameters
 aux <- expand.grid(
-    mu = seq(2, 30, length.out = 50),
-    phi = seq(log(0.3), log(2.5), length.out = 50))
+    mu = seq(3, 30, length.out = 55),
+    phi = seq(-1.5, 1.8, length.out = 50),
+    KEEP.OUT.ATTRS = FALSE)
 
-# Compute approximately and numerically moments
-moments <- mapply(FUN = calc_moments,
-                  mu = aux$mu,
-                  phi = aux$phi,
-                  MoreArgs = list(sumto = 300),
-                  SIMPLIFY = FALSE)
-grid <- cbind(aux, t(do.call(cbind, moments)))
-grid <- transform(grid, va = mu / exp(phi))
+grid <- with(aux, t(calc_moments(mu, phi))) %>%
+    as.data.frame() %>%
+    cbind(aux, .) %>%
+    transform(va = mu / exp(phi)) %>%
+    transform(emu = (mu - mean),
+              eva = (va - var))
+
+max(abs(grid$emu))
+max(abs(grid$eva))
 
 #-------------------------------------------
 # Errors in approximations for E[Y] and V[Y]
-grid <- transform(grid,
-                  emu = (mu - mean)^2,
-                  eva = (va - var)^2)
 
-myreg <- colorRampPalette(c("gray90",  "gray20"))
-xy1 <- levelplot(emu ~ phi + mu, data = grid,
-                 aspect = "fill",
-                 col.regions = myreg,
-                 xlab = expression(phi),
-                 ylab = expression(mu),
-                 sub = "(a)",
-                 colorkey = list(space = "top"),
-                 par.settings = ps2,
-                 panel = function(x, y, z, ...) {
-                     panel.levelplot(x, y, z, ...)
-                     panel.curve(10 - ( exp(x) - 1)/(2 * exp(x)),
-                                 lty = 2)
-                     panel.abline(v = 0, lty = 2)
-                 })
+myreg <- colorRampPalette(c("gray20", "gray95"))
+xy1 <-
+    levelplot(emu ~ phi + mu, ,
[…]

### 2018-09-13 codes/01_simulation-study.R

@@ -7,8 +7,9 @@
 
 ##----------------------------------------------------------------------
 ## Load package and functions
-source("lattice-panels.R")
-source("functions.R")
+source("helper01_general-functions.R")
+source("helper02_lattice-panels.R")
+
 library(plyr)
 library(tidyr)
 library(bbmle)
@@ -25,7 +26,7 @@ rcmp <- Vectorize(FUN = function(n, mu, phi) {
 B <- 1000
 
 beta <- c("b0" = 2, "b1" = 0.5, "b21" = 0.8, "b22" = -0.8)
-phis <- c(0, -1.6, -1, 1.8)
+phis <- c(-1.6, -1, 0, 1.8)
 names(phis) <- sprintf("phi=%s", phis)
 
 sizes <- c(50, 100, 300, 1000)
@@ -59,6 +60,7 @@ xy1 <- xyplot(mu ~ x1, groups = x2,
 xy2 <- xyplot(di ~ x1 | phi, groups = x2,
               type = c("g", "l"),
               lwd = 2,
+              as.table = TRUE,
               # scales = "free",
               layout = c(2, 2),
               xlab = "Continous variable",
@@ -479,6 +481,7 @@ datime <-
     })
 saveRDS(datime, "simultimes.rds")
 
+datime <- readRDS("simultimes.rds")
 datime %>%
     mutate(size = factor(sizes[size], levels = sizes),
            phi = factor(phi, levels = names(phis)[c(2, 3, 1, 4)])) %>%

### 2018-09-13 codes/91_defoliation-analysis.R

@@ -1,6 +1,5 @@
 # =====================================================================
-# Analysis of number of grains in Soyabean under Soil Moisture and
-# Potassium Doses
+#  Analysis of Artificial Defoliation in Cotton Plants
 #                                                        Eduardo Junior
 #                                                    edujrrib@gmail.com
 #                                                            2017-08-05
@@ -8,49 +7,49 @@
 
 #-----------------------------------------------------------------------
 # Load package and functions
+source("helper01_general-functions.R")
+source("helper02_lattice-panels.R")
+
 library(bbmle)
 library(multcomp)
 library(dplyr)
-source("lattice-panels.R")
-source("functions.R")
 
 # Colors for legends
 cols <- trellis.par.get("superpose.line")$col
 
 #-----------------------------------------------------------------------
 # Load data
-# data(soyaBeans, package = "cmpreg")
-soyaBeans <- read.table("../data/soyaBeans.txt",
-                        header = TRUE, sep = "\t")
-soyaBeans$umid <- as.factor(soyaBeans$umid)
-soyaBeans <- soyaBeans[-74, ] ## Incorrect observation
-soyaBeans <- transform(soyaBeans, K = K / 100)
+# data(cottonBolls, package = "cmpreg")
+cottonBolls <- read.table("../data/cottonBolls.txt",
+                          header = TRUE, sep = "\t")
+str(cottonBolls)
 
 #-----------------------------------------------------------------------
 # Exploratory analysis
 
-# Scatter plot
-xy1 <- xyplot(ngra ~ K | umid,
-              data = soyaBeans,
-              xlab = "Potassium fertilization level",
-              ylab = "Number of grains per pot",
-              type = c("p", "g", "smooth"),
-              as.table =  TRUE,
-              layout = c(2, 2),
-              strip = strip.custom(
-                  strip.names = TRUE, var.name = "humidity",
-                  factor.levels = paste0(levels(soyaBeans$umid), "%")))
+# Scatter plot with a beewarm taste
+xy1 <- xyplot(ncap ~ des | est,
+              data = cottonBolls,
[…]

### 2018-09-13 codes/92_soybean-analysis.R

@@ -1,5 +1,6 @@
 # =====================================================================
-#  Analysis of Artificial Defoliation in Cotton Plants
+# Analysis of number of grains in Soyabean under Soil Moisture and
+# Potassium Doses
 #                                                        Eduardo Junior
 #                                                    edujrrib@gmail.com
 #                                                            2017-08-05
@@ -7,48 +8,50 @@
 
 #-----------------------------------------------------------------------
 # Load package and functions
+source("helper01_general-functions.R")
+source("helper02_lattice-panels.R")
+
+library(plyr)
+library(tidyr)
 library(bbmle)
-library(multcomp)
-library(dplyr)
-source("lattice-panels.R")
-source("functions.R")
 
 # Colors for legends
 cols <- trellis.par.get("superpose.line")$col
 
 #-----------------------------------------------------------------------
 # Load data
-# data(cottonBolls, package = "cmpreg")
-cottonBolls <- read.table("../data/cottonBolls.txt",
-                          header = TRUE, sep = "\t")
-str(cottonBolls)
+# data(soyaBeans, package = "cmpreg")
+soyaBeans <- read.table("../data/soyaBeans.txt",
+                        header = TRUE, sep = "\t")
+soyaBeans$umid <- as.factor(soyaBeans$umid)
+soyaBeans <- soyaBeans[-74, ] ## Incorrect observation
+soyaBeans <- transform(soyaBeans, K = K / 100)
 
 #-----------------------------------------------------------------------
 # Exploratory analysis
 
-# Scatter plot with a beewarm taste
-xy1 <- xyplot(ncap ~ des | est,
-              data = cottonBolls,
-              layout = c(2, 3),
-              as.table = TRUE,
-              grid = TRUE,
-              type = c("p", "smooth"),
-              xlab = "Artificial defoliation level",
-              ylab = "Number of bolls produced",
-              spread = 0.05,
-              panel = panel.beeswarm)
+# Scatter plot
+xy1 <- xyplot(ngra ~ K | umid,
+              data = soyaBeans,
+              xlab = "Potassium fertilization level",
[…]

### 2018-09-13 codes/93_nitrofen-analysis.R

@@ -8,11 +8,12 @@
 
 #-----------------------------------------------------------------------
 # Load package and functions
+source("helper01_general-functions.R")
+source("helper02_lattice-panels.R")
+
+library(plyr)
+library(tidyr)
 library(bbmle)
-library(multcomp)
-library(dplyr)
-source("lattice-panels.R")
-source("functions.R")
 
 # Colors for legends
 cols <- trellis.par.get("superpose.line")$col

### 2018-09-13 codes/94_timestofit-study.R

@@ -7,10 +7,11 @@
 
 #-----------------------------------------------------------------------
 # Load package and functions
+source("helper01_general-functions.R")
+source("helper02_lattice-panels.R")
+
 library(bbmle)
 library(microbenchmark)
-source("lattice-panels.R")
-source("functions.R")
 
 # Colors for legends
 cols <- trellis.par.get("superpose.line")$col
@@ -71,8 +72,8 @@ bench2 <- microbenchmark(
 
 # Times to the second case study
 bench3 <- microbenchmark(
-    "CMP  " = fitcm(form3, data = data3, model = "CP" , sumto = 700),
-    "CMPmu" = fitcm(form3, data = data3, model = "CP2", sumto = 700),
+    "CMP  " = fitcm(form3, data = data3, model = "CP" , sumto = 100),
+    "CMPmu" = fitcm(form3, data = data3, model = "CP2", sumto = 100),
     times = 50)
 
 #-------------------------------------------

### 2018-09-13 codes/papercomp.R
Add html paper companion

@@ -1,0 +1,2449 @@
+#' % Reparametrization of COM-Poisson Regression Models with Applications in the Analysis of Experimental Count Data
+#' % [Eduardo Elias Ribeiro Junior](http://www.leg.ufpr.br/~eduardojr) <br> [Walmes Marques Zeviani](http://www.leg.ufpr.br/~walmes) <br> [Wagner Hugo Bonat](http://www.leg.ufpr.br/~wagner) <br> [Clarice Garcia Borges Demétrio](http://www4.esalq.usp.br/pesquisa/node/25) <br> [John Hinde](http://www.nuigalway.ie/maths/jh/)
+#' % August 31, 2017
+
+#+ include=FALSE
+library(knitr)
+options(width = 110, digits = 4)
+opts_chunk$set(
+    warning = FALSE,
+    message = FALSE,
+    echo = TRUE,
+    cache = TRUE,
+    fig.width = 9,
+    fig.height = 5,
+    fig.align = "center",
+    dev.args = list(family = "Palatino")
+    )
+
+#' # 00_explore-model #
+# =====================================================================
+# Analysis of moments approximation
+#                                                        Eduardo Junior
+#                                                    edujrrib@gmail.com
+#                                                            2017-08-01
+# =====================================================================
+
+#-----------------------------------------------------------------------
+# Load package and functions
+source("helper01_general-functions.R")
+source("helper02_lattice-panels.R")
+
+library(dplyr)
+
+#-----------------------------------------------------------------------
+# Convergence of Z(lambda, nu) constant
+computeZ <- function(lambda, nu, maxit = 1e4, tol = 1e-5) {
+    z <- vector("numeric", maxit)
+    j = 1
+    z[j] <- exp(j * log(lambda) - nu * lfactorial(j))
+    while (abs(z[j] - 0) > tol && j <= maxit) {
+        j = j + 1
+        z[j] <- exp(j * log(lambda) - nu * lfactorial(j))
+    }
+    if (lambda == 1 & nu == 0) z <- Inf
+    return(sum(z, na.rm = TRUE) + 1)
+}
+
+grid <- expand.grid(
+    lambda = c(0.5, 1, 5, 10, 30, 50),
+    nu = seq(0, 1, length.out = 11))
+grid$z <- apply(grid, 1, function(par) {
+    computeZ(lambda = par[1], nu = par[2], maxit = 1e4)
+})
+
[…]

### 2018-10-30 article.R
Fix path

@@ -44,8 +44,8 @@ divide_matrix <- function(mat, divide) {
     return(list(mat1, mat2))
 }
 
-source("./codes/lattice-panels.R")
-source("./codes/functions.R")
+source("./codes/helper01_general-functions.R")
+source("./codes/helper02_lattice-panels.R")
 
 # Colors for legends
 cols <- trellis.par.get("superpose.line")$col

### 2018-10-30 presentations/2018mca/slides-mca2018.Rnw
Improve the directory names

--- 2018-08-27 presentations/esalq2018/slides-seminar.Rnw
@@
 \title[Reparametrization of COM-Poisson Regression Models]{
-  Reparametrization of COM-Poisson Regression Models for Analysis of
-  Count Data}
-
-\author[Eduardo E. Ribeiro Jr]{
+  Reparametrization of COM-Poisson Regression Models with Applications
+  in the Analysis of Experimental Count Data}
+
+\author[Clarice Garcia Borges Demétrio]{
   Eduardo Elias Ribeiro Junior \inst{1} \inst{2} \\
@@
 
+\begin{frame}{Count data}
+
+  Number of times an event occurs in the observation unit.\\[0.3cm]
+
+  Random variables that assume non-negative integer values.\\[0.3cm]
+
+  Let $Y$ be a counting random variable, so that $y=0,1,2,\ldots$\\[0.5cm]
+
+  Examples in experimental researches:
+
+  \begin{itemize}
+    \item number of grains produced by a plant;
+    \item number of fruits produced by a tree;
+    \item number of insects on a particular cell;
+    \item others.
+  \end{itemize}
+
+\end{frame}
+
 \begin{frame}{Poisson model and limitations}
-  \vspace{-0.3cm}
-  \begin{itemize}
-    \item[] {\bf Count data}
-    \item Number of times an event occurs in the observation unit;
-    \item Mean-variance relationship.
-      \vspace{0.3cm}
+
+  \begin{itemize}
     \item[] {\bf GLM framework \citep{Nelder1972}}
@@
   \begin{itemize}
-  \item The COM-Poisson \citep{Shmueli2005} belongs to the family of
-    weighted Poisson distributions with the weight function
-    $w(y,\nu) = (y!)^{1-\nu}$.
-    \vspace{0.3cm}
-  \item   The probability mass function of $Y$ a
-    COM-Poisson random variable is
+    \itemsep7.5pt\parskip0pt\parsep0pt
+  \item Probability mass function \citep{Shmueli2005} takes the form \\
+    \begin{equation*} \small
+    \label{eqn:pmf-cmp}
+    \Pr(Y = y \mid \lambda, \nu) =
+      \frac{\lambda^y}{(y!)^\nu Z(\lambda, \nu)}\,,
+    \qquad Z(\lambda, \nu) = \sum_{j=0}^\infty
+      \frac{\lambda^j}{(j!)^\nu}\,,
+  \end{equation*}
+    where $\lambda > 0$ and $\nu \geq 0$.
+  \item Moments are not available in closed form;
+  \item Expectation and variance can be closely approximated by
     \begin{equation*}
-      \label{eqn:pmf-compoisson}
-      \Pr(Y = y) = \frac{\lambda^y \exp(-\lambda)}{
-        (y!)^\nu \text{E}_\lambda[(Y!)^{1-\nu}]} =
-      \frac{\lambda^y}{(y!)^\nu \text{Z}(\lambda, \nu)}, \quad
-      \text{Z}(\lambda,\nu) = \sum_{j=0}^\infty \frac{\lambda^j}{(j!)^\nu},
+      \label{eqn:mean-aprox}
+      \text{E}(Y) \approx \lambda^{1/\nu} - \frac{\nu - 1}{2\nu} \qquad
+      \textrm{and} \qquad
+      \text{Var}(Y) \approx \frac{\lambda^{1/\nu}}{\nu}
     \end{equation*}
-    where $\nu$ is the dispersion parameter.
-    \vspace{0.3cm}
-  \item The moments of distribution are not obtained in closed forms.
-    There is approximations proposed by \citet{Shmueli2005,
-      Sellers2010}:
-    $$
-    \text{E}(Y) \approx \dot{\text{E}}(Y) = \lambda^{1/\nu} - \frac{\nu
-      - 1}{2\nu} \qquad \textrm{and} \qquad \text{Var}(Y) \approx
-    \frac{\lambda^{1/\nu}}{\nu}.
-    $$
-  \end{itemize}
+    with accurate approximations for $\nu \leq 1$ or $\lambda > 10^\nu$
+    \citep{Shmueli2005, Sellers2012}.
+  \end{itemize}
+
 \end{frame}
@@
 \begin{frame}{Orthogonality property of the MLEs}
-\vspace{-0.3cm}
 
@@
 
-\begin{frame}{Soybean data}
-\vspace{-0.2cm}
-<<desc-soy, fig.width=9, fig.height=5.8, fig.cap="">>=
-
-knitr::include_graphics("figure/desc-soy-1")
-
-@
-
-\end{frame}
-
 \begin{frame}{Model specification}
@@
 library(latticeExtra)
-source("../../codes/lattice-panels.R")
-
-benchs <- readRDS("../../codes/comparetimes.rds")
+source("../codes/lattice-panels.R")
+
+benchs <- readRDS("../codes/comparetimes.rds")
 xlabs <- expression("COM-Poisson", "COM-Poisson"[mu])
@@
   \vspace*{-0.5cm}
-  \bibliography{../../references.bib}
+  \bibliography{../references.bib}
 \end{frame}

### 2018-10-30 presentations/2018botucatu/slides-mca2018.Rnw
Add slides for the event in Botucatu

--- 2018-10-30 presentations/2018mca/slides-mca2018.Rnw
@@
 
-\date{\small \today \\[0.2cm]
-  \url{jreduardo@usp.br} | \url{edujrrib@gmail.com}}
+\date{\small 11th October, 2018 - UNESP/Botucatu \\[0.2cm]
+  \url{clarice.demetrio@usp.br} |\url{jreduardo@usp.br}}
 
@@
 
+\section{Motivating data sets}
+
+%-------------------------------------------
+\begin{frame}{Artificial defoliation in cotton phenology}
+
+  \begin{columns}
+    \begin{column}{0.4\textwidth}
+      \begin{figure}
+        \includegraphics[scale=0.25]{./images/defoliation}
+      \end{figure}
+    \end{column}
+    \begin{column}{0.6\textwidth}
+      {\bf Aim:} to assess the effects of five defoliation levels on the
+      bolls produced at five growth stages;\\[0.2cm]
+      {\bf Design:} factorial 5 $\times$ 5, with 5 replicates;\\[0.2cm]
+      {\bf Experimental unit:} a plot with 2 plants;\\[0.2cm]
+      {\bf Factors:}
+      \begin{itemize} \small
+        \item Artificial defoliation (\texttt{des}):
+        \item Growth stage (\texttt{est}):
+      \end{itemize}\vspace{0.1cm}
+      {\bf Response variable:} Total number of cotton bolls;\\
+    \end{column}
+  \end{columns}
+
+\end{frame}
+
+\begin{frame}{Artificial defoliation in cotton phenology}
+\vspace{-0.2cm}
+
+<<desc-cotton, fig.width=9, fig.height=5.8, fig.cap="(a) Perfis das contagens de ninfas observadas para cada bloco em cada uma das quatro cultivares BRS, as linhas em preto representam os perfis médios.
+(b) Dispersão das médias (eixo $x$) e variâncias (eixo $y$) amostrais, na escala logarítmica, calculadas sobre os blocos em cada tempo e cultivar para cada variável de interesse, a linha pontilhada representa a reta média $=$ variância (equidispersão) e a contínua representa um ajuste de mínimos quadrados.">>=
+include_graphics("figure/desc-cotton-1.pdf")
+@
+\end{frame}
+
+%-------------------------------------------
+\begin{frame}{Potassium fertilization on soybean culture}
+  \begin{columns}
+    \begin{column}{0.4\textwidth}
+      \begin{figure}
+        \includegraphics[width=4cm]{./images/soybean}
+      \end{figure}
+    \end{column}
+    \begin{column}{0.6\textwidth}
+      {\bf Aim:} evaluate the effects of potassium doses applied
+      to soil in different soil moisture levels;\\[0.2cm]
+      {\bf Design:} factorial 5 $\times$ 3 in a randomized complete
+      block design (5 blocks);\\[0.2cm]
+      {\bf Experimental unit:} a pot with a plant;\\[0.2cm]
+      {\bf Factors:}
+      \begin{itemize} \small
+        \item Potassium fertilization dose (\texttt{K}):
+        \item Soil moisture level (\texttt{umid}):
+      \end{itemize}\vspace{0.1cm}
+      {\bf Response variable:} Total number of bean seeds per pot;\\
+    \end{column}
+  \end{columns}
+\end{frame}
+
+\begin{frame}{Potassium fertilization on soybean culture}
+\vspace{-0.3cm}
+<<desc-soya, fig.width=9, fig.height=6, fig.cap="(a) Number of bean seeds per pot for each potassium dose and moisture level (left) and (b) sample mean against sample variance of the five replicates for each experimental treatment.
+Solid lines are smoothing curves and the least-squares line.">>=
+include_graphics("figure/desc-soy-1.pdf")
+@
+\end{frame}
+
+%-------------------------------------------
+\begin{frame}{Assessing toxicity of nitrofen in aquatic systems}
+  \begin{columns}
+    \begin{column}{0.4\textwidth}
+      \begin{figure}
+        \includegraphics[width=4cm]{./images/zooplankton}
+      \end{figure}
+    \end{column}
+    \begin{column}{0.6\textwidth}
+      {\bf Aim:} measure the reproductive toxicity of the herbicide
+      nitrofen on a species of zooplankton
+      (\textit{Ceriodaphnia dubia});\\[0.2cm]
+      {\bf Design:} completely randomized design, with 10
+      replicates;\\[0.2cm]
+      {\bf Experimental unit:} zooplankton animal;\\[0.2cm]
+      {\bf Factors:}
+      \begin{itemize} \small
+        \item herbicide nitrofen dose (\texttt{dose});
+      \end{itemize}\vspace{0.1cm}
+      {\bf Response variable:} Total number of live offspring;\\
+    \end{column}
+  \end{columns}
+\end{frame}
+
+\begin{frame}{Assessing toxicity of nitrofen in aquatic systems}
+<<desc-nitrofen, results="asis", fig.width=8, fig.height=3.5, fig.cap="(a) Number of live offsprings observed for each nitrofen concentration level and (b) scatterplot of the sample means against sample variances.">>=
+include_graphics("figure/desc-nitrofen-1")
+@
+\end{frame}
+
 \section{Background}
-
-\begin{frame}{Count data}
-
-  Number of times an event occurs in the observation unit.\\[0.3cm]
-
-  Random variables that assume non-negative integer values.\\[0.3cm]
-
-  Let $Y$ be a counting random variable, so that $y=0,1,2,\ldots$\\[0.5cm]
-
-  Examples in experimental researches:
-
-  \begin{itemize}
-    \item number of grains produced by a plant;
-    \item number of fruits produced by a tree;
-    \item number of insects on a particular cell;
-    \item others.
-  \end{itemize}
-
-\end{frame}
 
@@
 
-<<approx-plot, fig.height=3, fig.width=6, out.width=".9\\textwidth", fig.cap="Quadratic errors for the approximation of the (a) expectation and (b) variance.
+<<approx-plot, fig.height=3, fig.width=6, out.width=".9\\textwidth", fig.cap="Errors for the approximation of the (a) expectation and (b) variance.
 Dotted lines represent the restriction for suitable approximations given by \\cite{Shmueli2005}.">>=
@@
 
-\section{Case studies}
-
-\begin{frame}[t]{Motivating data sets and data analysis}
-
-  \begin{itemize}
-    \item Three illustrative examples of count data analysis are
-      reported.
-      \begin{itemize} \normalsize
-      \item Assessing toxicity of nitrofen in aquatic systems, an
-        equidispersed example;
-      \item Soil moisture and potassium doses on soybean culture, an
-        overdispersed example; and
-      \item Artificial defoliation in cotton phenology, an
-        underdispersed example.
-      \end{itemize}
-    \item In the data analysis, we consider the COM-Poisson model in the
-      two forms (original and new parametrization) and the quasi-Poisson
-      regression model as alternative models for the standard Poisson
-      regression model.
-  \end{itemize}
-\end{frame}
+\section{Data analysis results}
 
 \subsection{Artificial defoliation in cotton phenology}
-
-\begin{frame}{Cotton bolls data}
-
-  \begin{columns}
-    \begin{column}{0.4\textwidth}
-      \begin{figure}
-        \includegraphics[scale=0.25]{./images/defoliation}
-      \end{figure}
-    \end{column}
-    \begin{column}{0.6\textwidth}
-      {\bf Aim:} to assess the effects of five defoliation levels on the
-      bolls produced at five growth stages;\\[0.2cm]
-      {\bf Design:} factorial 5 $\times$ 5, with 5 replicates;\\[0.2cm]
-      {\bf Experimental unit:} a plot with 2 plants;\\[0.2cm]
-      {\bf Factors:}
-      \begin{itemize} \small
-        \item Artificial defoliation (\texttt{des}):
-        \item Growth stage (\texttt{est}):
-      \end{itemize}\vspace{0.1cm}
-      {\bf Response variable:} Total number of cotton bolls;\\
-    \end{column}
-  \end{columns}
-
-\end{frame}
 
@@
 
-\begin{frame}{Soybean data}
-
-  \begin{columns}
-    \begin{column}{0.4\textwidth}
-      \begin{figure}
-        \includegraphics[width=4cm]{./images/soybean}
-      \end{figure}
-    \end{column}
-    \begin{column}{0.6\textwidth}
-      {\bf Aim:} evaluate the effects of potassium doses applied
-      to soil in different soil moisture levels;\\[0.2cm]
-      {\bf Design:} factorial 5 $\times$ 3 in a randomized complete
-      block design (5 blocks);\\[0.2cm]
-      {\bf Experimental unit:} a pot with a plant;\\[0.2cm]
-      {\bf Factors:}
-      \begin{itemize} \small
-        \item Potassium fertilization dose (\texttt{K}):
-        \item Soil moisture level (\texttt{umid}):
-      \end{itemize}\vspace{0.1cm}
-      {\bf Response variable:} Total number of bean seeds per pot;\\
-    \end{column}
-  \end{columns}
-
-\end{frame}
-
 \begin{frame}{Model specification}
@@
 
-\begin{frame}{Nitrofen data}
-
-  \begin{columns}
-    \begin{column}{0.4\textwidth}
-      \begin{figure}
-        \includegraphics[width=4cm]{./images/zooplankton}
-      \end{figure}
-    \end{column}
-    \begin{column}{0.6\textwidth}
-      {\bf Aim:} measure the reproductive toxicity of the herbicide
-      nitrofen on a species of zooplankton
-      (\textit{Ceriodaphnia dubia});\\[0.2cm]
-      {\bf Design:} completely randomized design, with 10
-      replicates;\\[0.2cm]
-      {\bf Experimental unit:} zooplankton animal;\\[0.2cm]
-      {\bf Factors:}
-      \begin{itemize} \small
-        \item herbicide nitrofen dose (\texttt{dose});
-      \end{itemize}\vspace{0.1cm}
-      {\bf Response variable:} Total number of live offspring;\\
-    \end{column}
-  \end{columns}
-
-\end{frame}
-
 \begin{frame}{Model specification}
@@
 \begin{table}[!ht]
-\centering \footnotesize
+\centering \scriptsize
 \caption{Parameter estimates (Est) and ratio between estimate and
@@
 \end{table}
-  \vspace{-0.4cm}
-
-<<pred-ovos, fig.height=3.5, fig.width=5.5, out.width="0.8\\textwidth", fig.cap="Number of live offsprings observed for each nitrofen concentration level with fitted curves and 95\\% confidence intervals.">>=
+  \vspace{-0.3cm}
+
+<<pred-ovos, fig.height=3.5, fig.width=5.5, out.width="0.65\\textwidth", fig.cap="Number of live offsprings observed for each nitrofen concentration level with fitted curves and 95\\% confidence intervals.">>=
 include_graphics("figure/pred-ovos-1")
@@
 
-\subsection*{}
-
-\begin{frame}{Additional results}
+\subsection{Additional results}
+
+\begin{frame}{Computational times for fitting}
 
@@
 <<comp-times, fig.height=4, fig.width=8, fig.cap="Computational times to fit the models under original and reparametrized versions based on the fifty repetitions.">>=
-
-library(latticeExtra)
-source("../codes/lattice-panels.R")
-
-benchs <- readRDS("../codes/comparetimes.rds")
-xlabs <- expression("COM-Poisson", "COM-Poisson"[mu])
-bwplot(time ~ expr | case,
-       ylab = "Time (seconds)",
-       scales = list(
-           y = list(relation = "free"),
-           x = list(at = 1:2, labels = xlabs)
-       ),
-       data = benchs)
-
+include_graphics("figure/comp-times-1")
 @
@@
   \vspace*{-0.5cm}
-  \bibliography{../references.bib}
+  \bibliography{../../references.bib}
 \end{frame}

### 2019-01-30 article.Rnw
Add acknowledments and update submitted version
Typo

@@
 their respective standard errors for the two COM-Poisson models are very
-close to ratios from the y quasi-Poisson model. However, it is important
+close to ratios from the quasi-Poisson model. However, it is important
 to note that COM-Poisson model is a full parametric approach, i.e. there
@@
 
+\section*{Acknowledgments}
+
+The authors wish to thank the anonymous referees for many comments from
+which this paper has benefited. This work was partially supported by
+CNPq, a Brazilian science funding agency.
+
 %-----------------------------------------------------------------------
